### 2017-04-20 flsa.tex
first commit

\documentclass{article}
\input{pack.tex}
% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{Finite Time Analysis of Linear Stochastic Approximation}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
%  David S.~Hippocampus\thanks{Use footnote for providing further
%    information about author (webpage, alternative
%    address)---\emph{not} for acknowledging funding agencies.} \\
%  Department of Computer Science\\
%  Cranberry-Lemon University\\
%  Pittsburgh, PA 15213 \\
%  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle
\begin{abstract}

\end{abstract}

\section{Introduction}
\section{Notation}

\section{Problem Setup}
We consider linear stochastic approximation schemes given as
\begin{align}
\theta_t=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),
\end{align}
where we assume
\begin{assumption}
\item $A_t=A+M_t$ and $b_t=b+N_t$, where $\{M_t \in \R^{\ncn},\,t\geq 0\}$ and $\{N_t\in\R^n,\,t\geq 0\}$ are $i.i.d$ sequences such that $\E[M_t]=0$ and $\E[N_t]=0$.
\item There exists a $\ts\in \R^n$ such that $\ts=A^{-1}b$.
\item All the eigenvalues of $A$ have strictly positive real parts.
\item $\E[M_t^\top M_t\ts]=\sigma_1^2$ and $\E[\norm{N_t}^2]=\sigma_2^2$.
\end{assumption}

\begin{example}[Linear Least Squares]
\end{example}

\begin{example}[TD(0)]
\end{example}

\section{Definitions}
\begin{definition}[Weakly Admissible]
\end{definition}

\begin{definition}[Admissible]
\end{definition}

\begin{definition}[Strongly Admissible]
\end{definition}

\begin{lemma}\label{pdt}
Given a AS matrix $A$, there exists a non-singular matrix $U\in \C^{\ncn}$ such that $A=U\Lambda U^{-1}$ where $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
\end{lemma}
Note that in \cref{pdt} the transformation $U\Lambda U^{-1}$ is not unique.
\begin{example}
\end{example}

\begin{definition}
Define symmetric matrices
$
\SD \eqdef(\T_U(A)^\dag+\T_U(A))-\alpha \T_U(A)^\dag \T_U(A),\,
\SR \eqdef(\T_U(A)^\dag+\T_U(A))-\alpha \E[\T_U(A_1)^\dag \T_U(A_1)],\,
$
and constants
$\rhoD \eqdef \inf_{\norm{x}=1} \ip{x,\SD x},\, \rhoR \eqdef \inf_{\norm{x}=1} \ip{x,\SR x},\,$ and $\bu\eqdef \norm{U}^2\norm{U^{-1}}^2 \big(1+\frac2{\alpha\rhoD}\big)\frac{1}{\alpha \rhoR}$
\end{definition}
\begin{lemma}
Given any AS matrix $A$, there exists $\alpha_D>0,\,\alpha_R>0$ and a non-singluar matrix $U$ such that
\begin{enumerate}
\item $\rhoD>0,\,\forall \alpha \in (0,\alpha_D)$.
\item $\rhoR>0,\,\forall \alpha \in (0,\alpha_R)$.
\end{enumerate}
\end{lemma}
\begin{corollary}
There exists $\beta\eqdef\inf_{U : det(U)\neq 0,\alpha\in (0,\alpha_R)} \bu$.
\end{corollary}
\section{Main Results}
\begin{theorem}
\begin{align}
\E[\norm{\zb_t}^2]
\leq \beta
\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{e_0})}{t+1} \right)\,.
\end{align}
\end{theorem}

\section{Comparison With Prior Works}

\subsection{Stochastic Approximation}

\section{Interesting Special Cases}

\section{Application in Reinforcement Learning}
\subsection{RL Setting}

\subsection{TD(0) Algorithm}

\subsection{GTD}
\textbf{Flaws with GTD-MP:}

\textbf{Why is GTD slow:}

\subsection{iLSTD}

\subsection{Comparison of TD, GTD and iLSTD}

\section{Conclusion}

\appendix

\input{linal}

\end{document}

### 2017-04-20 linal.tex
first commit

\section{Linear Algebra Preliminaries}

We now present some useful results from linear algebra. In what follows, $A$ is an $\ncn$ matrix and $\I$ the $\ncn$.
\subsection{Results in Matrix Decomposition and Transformation}
Let $B$ be a $\ncn$ block diagonal matrix given by $B=\begin{bmatrix} B_1 &0 &0 &\ldots &0 \\ 0 &B_2 &0 &\ldots &0  \\ \vdots &\vdots &\vdots &\vdots &\vdots \\ 0 &\ldots &0 &0 &B_k \end{bmatrix}$, where $B_i$ is a $n_i \times n_i$ matrix such that $n_i<n$ (w.l.o.g) and $\sum_{i=1}^k n_i=n$. We also denote $B$ as
\begin{align*}
B=B_1 \op B_2 \op \ldots B_k
\end{align*}
\begin{lemma}
There exists a polynomial known as the \emph{characteristic} polynomial denoted by $\chi_A=(\lambda_1 \I-A)^{n_1}\ldots(\lambda_{k} \I-A)^{n_k}$, where $\lambda_i,\,i=1,\ldots,k (k\leq n)$ are the $k$ distinct eigenvalues of the matrix $A$ such that $\sum_{i=1}^k n_i=n$.
\end{lemma}
\begin{lemma}
There exists a polynomial known as the \emph{minimal} polynomial denoted by $m_A=(\lambda_1 \I-A)^{n'_1}\ldots(\lambda_k \I-A)^{n'_k}$, where $n'_i\leq n_i,\,i=1\ldots,k$.
\end{lemma}
\begin{lemma}
There exists a complex matrix $V\in \C^{\ncn}$ such that $A=V\Lambda V^{-1}$, where
$\tL=\tL_1\op\ldots\op\tL_k$, where each $\tL_i,\,i=1,\ldots,k$ can further be written as $\tL_i= {\tJ}^i_{d^i_1}\op \ldots \op {\tJ}^i_{d^i_{l(i)}}$
such that  $n'_i=\max\{d^i_1,\ldots, d^i_{l(i)} \}$ and $\sum_{j=1}^{l(i)} d^i_j =n_i$. Each of ${\tJ}^i_{d^i_j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
${\tJ}^i_{d_j}=\begin{bmatrix} \lambda_i &1 &0 &\ldots &0 &0\\ 0 &\lambda_i &1 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &1 \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.
\end{lemma}

\begin{lemma}
Given a AS matrix $A$, there exists a matrix $U_A\in \C^{\ncn}$ such that $A=U_A\Lambda_A U_A^{-1}$ where $\Lambda_A^\dag+\Lambda_A$ is a real symmetric positive definite matrix.
\end{lemma}
\begin{proof}
Consider the diagonal matrices $D^i_j=\begin{bmatrix} 1  &0 &0 &\ldots &0 &0\\ 0 &\re(\lambda_i) &0 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re(\lambda)^{d^i_j-1}_i &0 \\ 0 &\ldots &0 &0 &0 &\re(\lambda)^{d^i_j}_i \end{bmatrix},\,\forall j=1,\ldots,l(i)$, $D^i=D^i_1 \op\ldots\op D^i_{l(i)},\,\forall i=1,\ldots,k$ and $D=D^1 \op\ldots\op D^k$.
It follows that $A=(VD) \Lambda (VD)^{-1}$, where $\Lambda$ is a matrix such that
$\Lambda=\Lambda_1 \op \ldots \op \Lambda_k$, where each $\Lambda_i,\,i=1,\ldots,k$ can further be written as
$\Lambda_i=J^i_{d^i_1} \op \ldots \op J^i_{d^i_{l(i)}}$ such that  $n'_i=\max\{d^i_1,\ldots, d^i_{l(i)} \}$ and $\sum_{j=1}^{l(i)} d^i_j =n_i$. Each of $J^i_{d^i_j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
$J^i_{d_j}=\begin{bmatrix} \lambda_i &\re(\lambda_i) &0 &\ldots &0 &0\\ 0 &\lambda_i &\re(\lambda_i) &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &\re(\lambda_i) \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.

Now let $\frac{(\Lambda^\dag+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}(J^{i\dag}_{d_j}+J^i_{d_j})$, where $J^{i\dag}_{d_j}+J^i_{d_j}=\begin{bmatrix} \re(\lambda_i) &\frac{\re(\lambda_i)}{2} &0 &\ldots &0 &0\\ \frac{\re(\lambda_i)}{2} &\re(\lambda_i) &\frac{\re(\lambda_i)}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re(\lambda_i) &\frac{\re(\lambda_i)}{2} \\ 0 &\ldots &0 &0 &\frac{\re(\lambda_i)}{2} &\re(\lambda_i) \end{bmatrix} $. Then for any $x\in \C^n (\neq \mathbf{0})$, there exists a $b\in {-1,1}^n$, such that
\begin{align*}
x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\sum_{i=1}^n \ip{x_i,x_i}+\sum_{i=1}^{n-1}\ip{x_i,x_{i+1}}\\
&> (\sum_{i=1}^n b_i x_i)^2\\
&\geq 0
\end{align*}
\end{proof}

\section{Error Recursion}
\begin{assumption}\label{cmplxassmp}
\begin{enumerate}
\item $\Lt\eqdef \Lambda + L_t$ where $\Lambda$ is a positive definite matrix and $\{L_t \in C^{\ncn},\,t> 0\}$ is a sequence of $i.i.d$ matrices such that $\E[L_t]=\zero,\,\forall t>0$.
\item $\{\zeta_t,\,t\geq 0\in C^n\}$ is a sequence of noise vectors such that $\E[\zeta_t]=0,\,\forall t\geq 0$.
\item $\E[\norm{L_t \zeta_t}]=\hat{\sigma}_2^2$, $\E[\norm{\zeta_t}^2]=\hat{\sigma}_1^2$
\end{enumerate}
\end{assumption}

\begin{definition}
Define $\forall,\,i\geq j$, $F_{i,j}=(I-\alpha \L_i)\ldots (I-\alpha \L_j)$ and $\forall,\,i<j$ $F_{i,j}=\I$.
\end{definition}

\begin{align}
\theta_t&=\theta_{t-1}+\alpha\big(b_t-A_t\theta_{t-1}\big)\\
\theta_t-\ts&=\theta_{t-1}-\ts+\alpha\big(b_t-A_t(\theta_{t-1}-\ts+\ts)\big)\\
e_t&=(I-\alpha A_t)e_{t-1}+(b_t -b -(A_t-A)\ts)\\
U^{-1}e_t&=(I-\alpha U^{-1}A_t U) U^{-1}e_{t-1}+ U^{-1}(b_t -b -(A_t-A)\ts)\\
z_t&=(I-\alpha \Lt) z_{t-1}+ \zeta_t
\end{align}

\begin{align*}
z_t
& = (I-\alpha \Lambda_t) (I-\alpha \Lambda_{t-1}) z_{t-2}\\ &+ \alpha (I-\alpha \Lambda_t) \zeta_{t-1} +\alpha \zeta_t \\
& \quad \vdots\\
& = (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_1) e_0\\ &+ \alpha (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_2) \zeta_1 \\
& + \alpha (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_3) \zeta_2\\
&  \quad \vdots \\
&+ \alpha \zeta_t\,,
\end{align*}
which can be written compactly as
\begin{align}
\label{eq:etft}
z_t = F_{t,1} z_0 + \alpha (F_{t,2} \zeta_1 + \dots + F_{t,t+1} \zeta_t )\,,
\end{align}
\begin{align*}
\zb_t=\frac{1}{t+1}{\sum}_{i=0}^{t}z_i
=\frac{1}{t+1}&\Big\{{\sum}_{i=0}^{t} F_{i,1} z_0 \\
&+ \alpha \sum_{i=1}^{t} \left(\sum_{k=i}^{t} F_{k,i+1} \right)\zeta_i \Big\} ,
\end{align*}
where in the second sum we flipped the order of sums and swapped the names of the variables that the sum runs over.

It follows that \todoc{We should rather use $C$ instead of $H$ here?}
\begin{align*}
\E[\norm{\zb_t}^2]&=\E\ip{\zb_t,\zb_t}
=\frac{1}{(t+1)^2} \sum_{i,j=0}^t \E\ip{z_i,z_j}\,.
\end{align*}
Hence, we see that it suffices to bound $\EE{\ip{ e_i,  e_j }}$.
There are two cases depending on whether $i=j$. When $i< j$,
\begin{align*}
\E\ip{z_i,z_j}
&=\E \ip{z_i,\big[F_{j,i+1} e_i+\alpha\textstyle\sum_{k=i+1}^j F_{j,k+1}\zeta_{k}\big]}\\
&=\E\ip{z_i,F_{j,i+1} z_i}  \text{(from \cref{noisecancel})}\\
&=\E\ip{z_i, (I-\alpha \Lambda)^{j-i} z_i} \text{(from \cref{lem:unroll})}
\end{align*}
and therefore
\begin{align*}
\label{inter}
\sum_{i=0}^{t-1}\sum_{j=i+1}^t \E\ip{z_i,z_j}
&=\frac1{\alpha\rhoD} {\sum}_{i=0}^{t-1}\E\ip{z_i,z_i}\\
&\leq \frac2{\alpha\rhoD}{\sum}_{i=0}^{t}\E\ip{z_i,z_i}\,.
\end{align*}
Since $\sum_{i,j}\cdot{} = \sum_{i=j}\cdot{} + 2 \sum_i \sum_{j>i} \cdot{}$,
\begin{align*}
{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{z_i,z_j}&= \left(1+\frac2{\alpha\rhoD}\right){\sum}_{i=0}^{t}\E\ip{z_i,z_i}\,.
\end{align*}
Expanding $z_i$ using \eqref{eq:etft} and then using \cref{innerproduct} and \Cref{cmplxassmp}
\begin{align*}
\E\ip{z_i,z_i}&=\E\ip{F_{i,1}z_0,F_{i,1}z_0}+\alpha^2{\sum}_{j=1}^i\E\ip{ F_{i,j+1}\zeta_j, F_{i,j+1}\zeta_j}+\alpha\sum_{j=1}^i  \E\ip{F_{i,1} z_0, F_{i,j+1}\zeta_j}\\
&\leq (1-\alpha\rhoR)^i\norm{z_0}^2+ \alpha^2\frac{\hat{\sigma}_1^2}{\alpha \rhoR}+ \alpha \frac{\hat{\sigma}^2_2 \norm{z_0}}{\alpha\rhoR}\,,
\end{align*}
and so
\begin{align*}
{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{z_i,z_j}
&\leq \left(1+\frac2{\alpha\rhoD }\right)\, \frac1{\alpha\rhoR}\, (t(\alpha^2\hat{\sigma}_1^2+\alpha \hat{\sigma}^2_2\norm{z_0}) +\norm{z_0}^2)\,.
\end{align*}
Putting things together,
\begin{align}
\E[\norm{\zb_t}^2]
\leq \left(1+\frac2{\alpha\rhoD}\right)\, \frac1{\alpha\rhoR}\, \,
\left(\frac{\norm{z_0}^2}{(t+1)^2}+ \frac{\alpha^2\hat{\sigma}_1^2+\alpha \hat{\sigma}^2_2\norm{z_0}}{t+1} \right)\,.
\end{align}

Let $\F_t = \sigma( H_1,\dots,H_t, \zeta_1,\dots,\zeta_t ) $, with $\F_0$ the $\sigma$-field that holds all random variables.
\newcommand{\cF}{\F}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
Let $t>i\ge 1$, $x,y\in \R^n$ be $\F_{i}$-measurable random vectors. Then,
\begin{align*}
\E[x^\top F_{t,i+1}y|\F_i]=x^\top (I-\alpha H)^{t-i} y\,.
\end{align*}
\end{lemma}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proof}
By the definition of $F_{t,i+1}$,
and because $F_{t-1,i+1} = (I-\alpha H_{t-1}) \dots (I-\alpha H_{i+1})$ is $\cF_{t-1}$-measurable,
as are $x$ and $y$,
\begin{align*}
\EE{x^\top F_{t,i+1} y | \cF_{t-1} } &= x^\top \EE{ (I-\alpha H_t) | \cF_{t-1} } F_{t-1,i+1} y\\
&=x^\top  (I-\alpha H)  F_{t-1,i+1} y\,.
\end{align*}
By the tower-rule for conditional expectations and our measurability assumptions,
\begin{align*}
\EE{x^\top F_{t,i+1} y | \cF_{t-2} }
&=x^\top  (I-\alpha H)  \EE{F_{t-1,i+1} |\cF_{t-2}} y\\
&= x^\top (I-\alpha H)^2 F_{t-2,i+1} y\,.
\end{align*}
Continuing this way we get
\begin{align*}
\EE{x^\top F_{t,i+1} y | \cF_{t-j} }
= x^\top (I-\alpha H)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
\end{align*}
Specifically, for $j=t-i$ we get
\begin{align*}
\EE{x^\top F_{t,i+1} y | \cF_{i} }  = x^\top (I-\alpha H)^{t-i} y\,.
\end{align*}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{lemma}\label{noisecancel}
Let $t>i\ge 1$ and let $x$ be an $\F_{i-1}$-measurable random vector. Then,
$\E[x^\top F_{t,i+1}\zeta_{i}]=0$.
\end{lemma}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proof}
By \cref{lem:genunroll},
\begin{align*}
\EE{x^\top F_{t,i+1} \zeta_i | \cF_{i} }  = x^\top (I-\alpha H)^{t-i} \zeta_i\,.
\end{align*}
Using the tower rule,
\begin{align*}
\EE{x^\top F_{t,i+1} \zeta_i | \cF_{i-1} }
= x^\top (I-\alpha H)^{t-i}\EE{ \zeta_i | \cF_{i-1} }= 0\,.
\end{align*}
\end{proof}

\begin{lemma}\label{lem:unroll}
For all $t>i\ge 0$, $\E \ip{He_i,H F_{t,i+1} e_i}=\E\ip{He_i,H(I-\alpha H)^{t-i} e_i}$.
\end{lemma}
\begin{proof}
The lemma follows directly from \cref{lem:genunroll}. Indeed,
$\theta_i$ depends only on $H_1,\dots,H_{i},g_1,\dots,g_{i}$, $\theta_i$ and so is $e_i$ $\cF_i$-measurable.
Hence, the lemma is applicable and implies that
\begin{align*}
\EE{ \ip{He_i, HF_{t,i+1} e_i} | \cF_i } =
\EE{ \ip{He_i, H (I-\alpha H)^{t-i} e_i} | \cF_i }\,.
\end{align*}
Taking expectation of both sides gives the desired result.
\end{proof}

\begin{lemma}\label{innerproduct}
Let $i>j \ge 0$ and let $x\in \R^n$ be an $\F_j$-measurable random vector.
Then,
\begin{align*}
\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rho_\alpha)^{i-j}\E\norm{x}^2\,.
\end{align*}
\end{lemma}
\begin{proof}
Note that
$S_t\doteq \EE{ (I-\alpha H_t)^\top (I-\alpha H_t) | \F_{t-1} }
= I - \alpha (H^\top + H) + \alpha^2 \EE{ H_t^\top H_t | \F_{t-1} }$.
Since $(g_t,H_t)_t$ is an independent sequence, $\EE{ H_t^\top H_t|\F_{t-1}} = \EE{ H_1^\top H_1 }$.
Now, using the definition of $\rho_\alpha$ from \eqref{eq:rhodef},
$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (H^\top + H - \alpha \EE{H_1^\top H_1}) x
= 1-\alpha \rho_\alpha$.
Hence,
\begin{align*}
&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
&= \EE{x^\top F_{i-1,j+1}^\top (I-\alpha H_i)^\top (I-\alpha H_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
&=(x F_{i-1,j+1})^\top \, S_i \, F_{i-1,j+1} x\\
&\le (1-\alpha \rho_\alpha) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
& \le (1-\alpha \rho_\alpha)^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
& \quad \vdots \\
& \le (1-\alpha \rho_\alpha)^{i-j}\, \norm{x}^2\,.
%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
%&=(1-\alpha\rho_{\alpha}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
\end{align*}
\end{proof}

### 2017-04-22 flsa.tex
initial

@@
 where we assume
-\begin{assumption}
+\begin{assumption}\label{ass:lsa}
+\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
 \item $A_t=A+M_t$ and $b_t=b+N_t$, where $\{M_t \in \R^{\ncn},\,t\geq 0\}$ and $\{N_t\in\R^n,\,t\geq 0\}$ are $i.i.d$ sequences such that $\E[M_t]=0$ and $\E[N_t]=0$.
@@
 \item $\E[M_t^\top M_t\ts]=\sigma_1^2$ and $\E[\norm{N_t}^2]=\sigma_2^2$.
+\end{itemize}
 \end{assumption}
@@
 
+\section{Notation}
+Let $\R^{\ncn}$ and $\C^{\ncn}$ denote the set of $n\times n$ real and complex matrices respectively. Let $\gln\subset\C^{\ncn}$ denote the set of all invertible $\ncn$ complex matrices.
+
 \section{Definitions}
+\begin{definition}[Asymptotically Stable]
+An $\ncn$ matrix said to be \emph{asymptotically stable} iff all its eigenvalues have positive real parts.
+\end{definition}
+
+\begin{definition}\label{def:dist}[Distributions]
+Given a distribution $P$ over $\C^{\ncn}$ and let
+\begin{align*}
+A_P=\int M dP(M),\quad C_P=\int M^\dag M dP(M),
+\end{align*}
+and let $\P$ denote the set of probability distributions on $\C^{\ncn}$.
+\end{definition}
+
+\begin{definition}[Similarity Transformation]
+Given any $U\in \gln$, define $\T_U \colon \C^{\ncn}\rightarrow \C^{\ncn}$ be the linear map that takes $A\in \C^{\ncn}$ to $\T_U(A)=U^{-1}\,A\,U$.
+\end{definition}
+\begin{definition}[Transformed Measure]
+Given a distribution $P$ over $\C^{\ncn}$ define the distribution $P_U$ induced by the linear transformation $\T_U$ as follows:
+\begin{align*}
+\int_{S} dP_U(M) \eqdef \int_{\T_U^{-1}(S)} dP(M)
+\end{align*}
+\end{definition}
+
+\begin{comment}
+\begin{definition}[Induced Distribution]
+Let $P\in \P$ be any distribution over $\R^{\ncn}$ and let $p_c^U$ denote the probability distribution over $\C^{\ncn}$ induced by the linear map $\T_U$ such that the pullback of $p_c^U$ using $\T_U$ is $p$.
+\end{definition}
+\end{comment}
+
+\begin{definition}[Spectral Norms]
+Define $\rhod{P}\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{\bar{x},\big(A_P+A_P^\dag)-\alpha A_P^\dag A_P)x}$, $\rhos{P}\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{\bar{x},\big(A_P+A_P^\dag-\alpha C_P\big)x}$
+\end{definition}
+
 \begin{definition}[Weakly Admissible]
+A set of distributions $\P$ over $\C^{\ncn}$ is said to be \emph{weakly admissible} if for any given $P\in \P,\,\exists$ a matrix $U_P \in \gln$, real-valued constants $\alphapd>0$ and $\alphaps>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alphapd)$ and $\rhos{P_U}>0,\,\forall \alpha\in(0,\alphaps)$.
+\end{definition}
+\begin{definition}
+Given a weakly admissible class $\P$ and any $P\in\P$, $U\in \gln$ and $\alpha>0$ such that $\rhos{P_{U}}>0$ define
+\begin{align*}
+\bu&\eqdef \norm{U}^2\norm{U^{-1}}^2 \left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{1}{\alpha \rhos{P_U}}\\
+\beta(P)&\eqdef \inf_{\alpha,U\colon \bu>0} \bu
+\end{align*}
 \end{definition}
@@
 \begin{definition}[Admissible]
+A set of distributions $\P$ over $\C^{\ncn}$ is \emph{admissible} if  for any given $P\in \P,\,\exists$ a matrix $U_P\in \gln,\,$ and a real-valued constant $\alpha_0>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alpha_0)$ and $\rhos{P_{U_P}},\,\forall \alpha\in(0,\alpha_0)$.
 \end{definition}
-
+\begin{example}
+\end{example}
 \begin{definition}[Strongly Admissible]
+A set of distributions $\P$ over $\C^{\ncn}$ is \emph{strongly admissible} if  $\forall\,P\in \P,\,\exists U_P\in \gln,\, \alpha_0>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alpha_0]$ and $\rhos{P_{U_P}},\,\forall \alpha\in(0,\alpha_0]$, $\inf_{P\in \P} \rho_d(\alpha_0,P_{U_P})>0$ and $\inf_{P\in \P} \rho_s(\alpha_0,P_{U_P})>0$.
 \end{definition}
-
+\begin{example}
+\end{example}
+\begin{comment}
 \begin{lemma}\label{pdt}
@@
 \end{example}
-
-\begin{definition}
-Define symmetric matrices
-$
-\SD \eqdef(\T_U(A)^\dag+\T_U(A))-\alpha \T_U(A)^\dag \T_U(A),\,
-\SR \eqdef(\T_U(A)^\dag+\T_U(A))-\alpha \E[\T_U(A_1)^\dag \T_U(A_1)],\,
-$
-and constants
-$\rhoD \eqdef \inf_{\norm{x}=1} \ip{x,\SD x},\, \rhoR \eqdef \inf_{\norm{x}=1} \ip{x,\SR x},\,$ and $\bu\eqdef \norm{U}^2\norm{U^{-1}}^2 \big(1+\frac2{\alpha\rhoD}\big)\frac{1}{\alpha \rhoR}$
-\end{definition}
+\end{comment}
+\begin{comment}
 \begin{lemma}
@@
 \begin{enumerate}
-\item $\rhoD>0,\,\forall \alpha \in (0,\alpha_D)$.
-\item $\rhoR>0,\,\forall \alpha \in (0,\alpha_R)$.
+\item $\rhod{\Lambda_1}>0,\,\forall \alpha \in (0,\alpha_D)$.
+\item $\rhos{\Lambda_1}>0,\,\forall \alpha \in (0,\alpha_R)$.
 \end{enumerate}
@@
 \end{corollary}
+\end{comment}
 \section{Main Results}
+The following theorem states that weak admissibility implies stabilizability.
 \begin{theorem}
+Let $\P_{as}$ be a set of distributions supported on $\R^{\ncn}$ such that for any given $P\in \P$, all the eigenvalues of $A_{P}$ have positive real parts. $\P_{as}$ is \emph{weakly admissible}.
+\end{theorem}
+In short, so long as the matrices $A_t$ are distributed such that all the eigenvalues of $\E[A_t]$ have with positive real parts, then there always exists a positive step size $\alpha$ (albeit problem dependent) that results in \eqref{lsa} to converge in the mean-squared sense.
+
+In \Cref{thm:rate} that we state we below, we let $P_A$ denote the underlying distribution that generates $A_t, t\geq 0$, and in order to be consitent with \Cref{ass:lsa} and \Cref{def:dist} it is understood that $A_{P_A}=A$.
+\begin{theorem}\label{thm:rate}
 \begin{align}
 \E[\norm{\zb_t}^2]
-\leq \beta
+\leq \beta(P_A)
 \left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{e_0})}{t+1} \right)\,.
 \end{align}
+
 \end{theorem}

### 2017-04-22 linal.tex
initial

@@
 \begin{assumption}\label{cmplxassmp}
-\begin{enumerate}
-\item $\Lt\eqdef \Lambda + L_t$ where $\Lambda$ is a positive definite matrix and $\{L_t \in C^{\ncn},\,t> 0\}$ is a sequence of $i.i.d$ matrices such that $\E[L_t]=\zero,\,\forall t>0$.
-\item $\{\zeta_t,\,t\geq 0\in C^n\}$ is a sequence of noise vectors such that $\E[\zeta_t]=0,\,\forall t\geq 0$.
+\begin{enumerate}[leftmargin=*, before=\leavevmode\vspace{-\baselineskip}]
+\item $\Lt\eqdef \Lambda + L_t$ where $\Lambda$ is a positive definite matrix and $\{L_t \in \C^{\ncn},\,t> 0\}$ is a sequence of $i.i.d$ matrices such that $\E[L_t]=\zero,\,\forall t>0$.
+\item $\{\zeta_t,\,t\geq 0\in \C^n\}$ is a sequence of noise vectors such that $\E[\zeta_t]=0,\,\forall t\geq 0$.
 \item $\E[\norm{L_t \zeta_t}]=\hat{\sigma}_2^2$, $\E[\norm{\zeta_t}^2]=\hat{\sigma}_1^2$
@@
 \begin{align*}
-\zb_t=\frac{1}{t+1}{\sum}_{i=0}^{t}z_i
+\zh_t=\frac{1}{t+1}{\sum}_{i=0}^{t}z_i
 =\frac{1}{t+1}&\Big\{{\sum}_{i=0}^{t} F_{i,1} z_0 \\
@@
 \begin{align*}
-\E[\norm{\zb_t}^2]&=\E\ip{\zb_t,\zb_t}
+\E[\norm{\zh_t}^2]&=\E\ip{\zh_t,\zh_t}
 =\frac{1}{(t+1)^2} \sum_{i,j=0}^t \E\ip{z_i,z_j}\,.
@@
 \sum_{i=0}^{t-1}\sum_{j=i+1}^t \E\ip{z_i,z_j}
-&=\frac1{\alpha\rhoD} {\sum}_{i=0}^{t-1}\E\ip{z_i,z_i}\\
-&\leq \frac2{\alpha\rhoD}{\sum}_{i=0}^{t}\E\ip{z_i,z_i}\,.
+&=\frac1{\alpha\rhod{P}} {\sum}_{i=0}^{t-1}\E\ip{z_i,z_i}\\
+&\leq \frac2{\alpha\rhod{P}}{\sum}_{i=0}^{t}\E\ip{z_i,z_i}\,.
 \end{align*}
@@
 \begin{align*}
-{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{z_i,z_j}&= \left(1+\frac2{\alpha\rhoD}\right){\sum}_{i=0}^{t}\E\ip{z_i,z_i}\,.
+{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{z_i,z_j}&= \left(1+\frac2{\alpha\rhod{P}}\right){\sum}_{i=0}^{t}\E\ip{z_i,z_i}\,.
 \end{align*}
@@
 \E\ip{z_i,z_i}&=\E\ip{F_{i,1}z_0,F_{i,1}z_0}+\alpha^2{\sum}_{j=1}^i\E\ip{ F_{i,j+1}\zeta_j, F_{i,j+1}\zeta_j}+\alpha\sum_{j=1}^i  \E\ip{F_{i,1} z_0, F_{i,j+1}\zeta_j}\\
-&\leq (1-\alpha\rhoR)^i\norm{z_0}^2+ \alpha^2\frac{\hat{\sigma}_1^2}{\alpha \rhoR}+ \alpha \frac{\hat{\sigma}^2_2 \norm{z_0}}{\alpha\rhoR}\,,
+&\leq (1-\alpha\rhos{P})^i\norm{z_0}^2+ \alpha^2\frac{\hat{\sigma}_1^2}{\alpha \rhos{P}}+ \alpha \frac{\hat{\sigma}^2_2 \norm{z_0}}{\alpha\rhos{P}}\,,
 \end{align*}
@@
 {\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{z_i,z_j}
-&\leq \left(1+\frac2{\alpha\rhoD }\right)\, \frac1{\alpha\rhoR}\, (t(\alpha^2\hat{\sigma}_1^2+\alpha \hat{\sigma}^2_2\norm{z_0}) +\norm{z_0}^2)\,.
+&\leq \left(1+\frac2{\alpha\rhod{P} }\right)\, \frac1{\alpha\rhos{P}}\, (t(\alpha^2\hat{\sigma}_1^2+\alpha \hat{\sigma}^2_2\norm{z_0}) +\norm{z_0}^2)\,.
 \end{align*}
@@
 \begin{align}
-\E[\norm{\zb_t}^2]
-\leq \left(1+\frac2{\alpha\rhoD}\right)\, \frac1{\alpha\rhoR}\, \,
+\E[\norm{\zh_t}^2]
+\leq \left(1+\frac2{\alpha\rhod{P}}\right)\, \frac1{\alpha\rhos{P}}\, \,
 \left(\frac{\norm{z_0}^2}{(t+1)^2}+ \frac{\alpha^2\hat{\sigma}_1^2+\alpha \hat{\sigma}^2_2\norm{z_0}}{t+1} \right)\,.

### 2017-04-23 flsa.tex
rudimentary

@@
 \usepackage{microtype}      % microtypography
+%\usepackage{varioref}
+%\usepackage{thmtools}
+\usepackage{aliascnt}
+\usepackage{titlecaps,cleveref}
+\crefname{assumption}{Assumption}{Assumption}
 
@@
 \section{Problem Setup}
-We consider linear stochastic approximation schemes given as
+We consider linear stochastic approximation schemes with constant step size and PR-averaging given as below:
 \begin{align}
-\theta_t=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),
+\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
+\thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_s
 \end{align}
 where we assume
-\begin{assumption}\label{ass:lsa}
+\begin{assumption}\label{lsa}
 \begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
@@
 \end{assumption}
-
+\Cref{lsa}
 \begin{example}[Linear Least Squares]
@@
 
-\begin{definition}[Spectral Norms]
-Define $\rhod{P}\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{\bar{x},\big(A_P+A_P^\dag)-\alpha A_P^\dag A_P)x}$, $\rhos{P}\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{\bar{x},\big(A_P+A_P^\dag-\alpha C_P\big)x}$
+\begin{definition}\label{def:spectnorm}[Spectral Norms]
+Define $\rhod{P}\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{\bar{x},\left((A_P+A_P^\dag)-\alpha A_P^\dag A_P\right)x}$, $\rhos{P}\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{\bar{x},\left((A_P+A_P^\dag)-\alpha C_P\right)x}$
 \end{definition}
@@
 
-In \Cref{thm:rate} that we state we below, we let $P_A$ denote the underlying distribution that generates $A_t, t\geq 0$, and in order to be consitent with \Cref{ass:lsa} and \Cref{def:dist} it is understood that $A_{P_A}=A$.
+In \Cref{thm:rate} that we state we below, we let $P^A$ denote the underlying distribution that generates $A_t, t\geq 0$, and in order to be consistent with \Cref{lsa} and \Cref{def:dist} it is understood that $A_{P^A}=A$.
 \begin{theorem}\label{thm:rate}
@@
 \E[\norm{\zb_t}^2]
-\leq \beta(P_A)
+\leq \beta(P^A)
 \left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{e_0})}{t+1} \right)\,.

### 2017-04-23 linal.tex
rudimentary

@@
 \begin{lemma}
-There exists a complex matrix $V\in \C^{\ncn}$ such that $A=V\Lambda V^{-1}$, where
+There exists a complex matrix $V\in \C^{\ncn}$ such that $A=VJ V^{-1}$, where
 $\tL=\tL_1\op\ldots\op\tL_k$, where each $\tL_i,\,i=1,\ldots,k$ can further be written as $\tL_i= {\tJ}^i_{d^i_1}\op \ldots \op {\tJ}^i_{d^i_{l(i)}}$
@@
 \begin{lemma}
-Given a AS matrix $A$, there exists a matrix $U_A\in \C^{\ncn}$ such that $A=U_A\Lambda_A U_A^{-1}$ where $\Lambda_A^\dag+\Lambda_A$ is a real symmetric positive definite matrix.
+Given a AS matrix $A$, there exists a matrix $U_A\in \C^{\ncn}$ such that $A=U_AJ_A U_A^{-1}$ where $J_A^\dag+J_A$ is a real symmetric positive definite matrix.
 \end{lemma}
@@
 Consider the diagonal matrices $D^i_j=\begin{bmatrix} 1  &0 &0 &\ldots &0 &0\\ 0 &\re(\lambda_i) &0 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re(\lambda)^{d^i_j-1}_i &0 \\ 0 &\ldots &0 &0 &0 &\re(\lambda)^{d^i_j}_i \end{bmatrix},\,\forall j=1,\ldots,l(i)$, $D^i=D^i_1 \op\ldots\op D^i_{l(i)},\,\forall i=1,\ldots,k$ and $D=D^1 \op\ldots\op D^k$.
-It follows that $A=(VD) \Lambda (VD)^{-1}$, where $\Lambda$ is a matrix such that
-$\Lambda=\Lambda_1 \op \ldots \op \Lambda_k$, where each $\Lambda_i,\,i=1,\ldots,k$ can further be written as
-$\Lambda_i=J^i_{d^i_1} \op \ldots \op J^i_{d^i_{l(i)}}$ such that  $n'_i=\max\{d^i_1,\ldots, d^i_{l(i)} \}$ and $\sum_{j=1}^{l(i)} d^i_j =n_i$.
+It follows that $A=(VD) J (VD)^{-1}$, where $J$ is a matrix such that
+$J=J_1 \op \ldots \op J_k$, where each $J_i,\,i=1,\ldots,k$ can further be written as
+$J_i=J^i_{d^i_1} \op \ldots \op J^i_{d^i_{l(i)}}$ such that  $n'_i=\max\{d^i_1,\ldots, d^i_{l(i)} \}$ and $\sum_{j=1}^{l(i)} d^i_j =n_i$.
 Each of $J^i_{d^i_j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
@@
 
-Now let $\frac{(\Lambda^\dag+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}(J^{i\dag}_{d_j}+J^i_{d_j})$, where $J^{i\dag}_{d_j}+J^i_{d_j}=\begin{bmatrix} \re(\lambda_i) &\frac{\re(\lambda_i)}{2} &0 &\ldots &0 &0\\ \frac{\re(\lambda_i)}{2} &\re(\lambda_i) &\frac{\re(\lambda_i)}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re(\lambda_i) &\frac{\re(\lambda_i)}{2} \\ 0 &\ldots &0 &0 &\frac{\re(\lambda_i)}{2} &\re(\lambda_i) \end{bmatrix} $.
+Now let $\frac{(J^\dag+J)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}(J^{i\dag}_{d_j}+J^i_{d_j})$, where $J^{i\dag}_{d_j}+J^i_{d_j}=\begin{bmatrix} \re(\lambda_i) &\frac{\re(\lambda_i)}{2} &0 &\ldots &0 &0\\ \frac{\re(\lambda_i)}{2} &\re(\lambda_i) &\frac{\re(\lambda_i)}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re(\lambda_i) &\frac{\re(\lambda_i)}{2} \\ 0 &\ldots &0 &0 &\frac{\re(\lambda_i)}{2} &\re(\lambda_i) \end{bmatrix} $.
 Then for any $x\in \C^n (\neq \mathbf{0})$, there exists a $b\in {-1,1}^n$, such that
 \begin{align*}
-x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\sum_{i=1}^n \ip{x_i,x_i}+\sum_{i=1}^{n-1}\ip{x_i,x_{i+1}}\\
+x^\dag \frac{(J^\dag+J)}{2} x &=\sum_{i=1}^n \ip{x_i,x_i}+\sum_{i=1}^{n-1}\ip{x_i,x_{i+1}}\\
 &> (\sum_{i=1}^n b_i x_i)^2\\
@@
 \section{Error Recursion}
-\begin{assumption}\label{cmplxassmp}
-\begin{enumerate}[leftmargin=*, before=\leavevmode\vspace{-\baselineskip}]
-\item $\Lt\eqdef \Lambda + L_t$ where $\Lambda$ is a positive definite matrix and $\{L_t \in \C^{\ncn},\,t> 0\}$ is a sequence of $i.i.d$ matrices such that $\E[L_t]=\zero,\,\forall t>0$.
-\item $\{\zeta_t,\,t\geq 0\in \C^n\}$ is a sequence of noise vectors such that $\E[\zeta_t]=0,\,\forall t\geq 0$.
-\item $\E[\norm{L_t \zeta_t}]=\hat{\sigma}_2^2$, $\E[\norm{\zeta_t}^2]=\hat{\sigma}_1^2$
-\end{enumerate}
-\end{assumption}
-
+\subsection{Definition}
 \begin{definition}
@@
 \end{definition}
+
+\begin{assumption}\label{cmplxassmp}
+\begin{enumerate}[leftmargin=*, before=\leavevmode\vspace{-\baselineskip}]
+\item $J_t\eqdef J + G_t$ where $J$ is a positive definite matrix and $\{G_t \in \C^{\ncn},\,t> 0\}$ is a sequence of $i.i.d$ matrices such that $\E[G_t]=\zero,\,\forall t>0$.
+\item $\{H_t,\,t\geq 0\in \C^n\}$ is a sequence of noise vectors such that $\E[H_t]=0,\,\forall t\geq 0$.
+\item $\E[\norm{G_t H_t}]=\hat{\sigma}_2^2$, $\E[\norm{H_t}^2]=\hat{\sigma}_1^2$
+\end{enumerate}
+\end{assumption}
+Let $P^J$ be the distribution that generates $J_t$ and in accordance with \Cref{def:dist} we have $J=A_{P^J}$.
+In this light, the constants in \Cref{def:spectnorm} $\rhod{P^J}$ and $\rhos{P^J}$ can also be given by
+\begin{align*}
+\rhod{P^J}&={\inf}_{x\in\C^n\colon\norm{x}=1}\ip{\bar{x},\left((J+J^\dag)-\alpha J^\dag J\right)x}\\
+\rhos{P^J}&={\inf}_{x\in \C^n\colon\norm{x}=1}\ip{\bar{x},\left((J+J^\dag-\alpha \E_{P^J}[J_1^\dag J_1]\right)x}
+\end{align*}
 
@@
 U^{-1}e_t&=(I-\alpha U^{-1}A_t U) U^{-1}e_{t-1}+ U^{-1}(b_t -b -(A_t-A)\ts)\\
-z_t&=(I-\alpha \Lt) z_{t-1}+ \zeta_t
+z_t&=(I-\alpha J_t) z_{t-1}+ H_t
 \end{align}
@@
 z_t
-& = (I-\alpha \Lambda_t) (I-\alpha \Lambda_{t-1}) z_{t-2}\\ &+ \alpha (I-\alpha \Lambda_t) \zeta_{t-1} +\alpha \zeta_t \\
+& = (I-\alpha J_t) (I-\alpha J_{t-1}) z_{t-2}\\ &+ \alpha (I-\alpha J_t) \zeta_{t-1} +\alpha H_t \\
 & \quad \vdots\\
-& = (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_1) e_0\\ &+ \alpha (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_2) \zeta_1 \\
-& + \alpha (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_3) \zeta_2\\
+& = (I-\alpha J_t) \cdots (I-\alpha J_1) e_0\\ &+ \alpha (I-\alpha J_t) \cdots (I-\alpha J_2) \zeta_1 \\
+& + \alpha (I-\alpha J_t) \cdots (I-\alpha J_3) \zeta_2\\
 &  \quad \vdots \\
-&+ \alpha \zeta_t\,,
+&+ \alpha H_t\,,
 \end{align*}
@@
 \label{eq:etft}
-z_t = F_{t,1} z_0 + \alpha (F_{t,2} \zeta_1 + \dots + F_{t,t+1} \zeta_t )\,,
+z_t = F_{t,1} z_0 + \alpha (F_{t,2} \zeta_1 + \dots + F_{t,t+1} H_t )\,,
 \end{align}
@@
 &=\E\ip{z_i,F_{j,i+1} z_i}  \text{(from \cref{noisecancel})}\\
-&=\E\ip{z_i, (I-\alpha \Lambda)^{j-i} z_i} \text{(from \cref{lem:unroll})}
+&=\E\ip{z_i, (I-\alpha J)^{j-i} z_i} \text{(from \cref{lem:unroll})}
 \end{align*}
@@
 
-Let $\F_t = \sigma( H_1,\dots,H_t, \zeta_1,\dots,\zeta_t ) $, with $\F_0$ the $\sigma$-field that holds all random variables.
+Let $\F_t = \sigma( H_1,\dots,H_t, \zeta_1,\dots,H_t ) $, with $\F_0$ the $\sigma$-field that holds all random variables.
 \newcommand{\cF}{\F}

### 2017-04-27 flsa.tex
updated

@@
 \usepackage{titlecaps,cleveref}
+\usepackage{amsmath,amssymb,graphicx}
 \crefname{assumption}{Assumption}{Assumption}
@@
 \section{Introduction}
-\section{Notation}
 
+\section{Notations and Definitions}
+
+We denote the sets real and complex number by $\R$ and $\C$ respectively.
+For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively.
+We denote $n$-dimensional vector spaces over $\R$ and $\C$ by $\R^{n}$ and $\C^{n}$ respectively.
+We use $\R^{\ncn}$ and $\C^{\ncn}$ to denote $\ncn$ matrices with real and complex entries respectively.
+For any $x\in \C$ we denote the modulus of $x$ by $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$.
+For any $x\in \C^n$We denote complex conjugate of $x$ by $\bar{x}$.
+We denote the transpose of $x\in \C^{n}, A\in \C^{\ncn}$ by $x^\top$ and $A^\top$ respectively.
+Further, by $x^\dag$ denotes the conjugate transpose, i.e., $x^\dag=\bar{x}^\top$ (the same convention carries over to matrices as well).
+We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^\dag y$.
+Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
+The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^n:\norm{x}=1} \norm{Ax}$.
+For a positive real number $B>0$, we denote by $\C^{\ncn}_B=\{A\in \C^{\ncn}\mid \norm{A}\leq B\}$ and  $\R^{\ncn}_B=\{A\in \R^{\ncn}\mid \norm{A}\leq B,\,\forall i,j=1,\ldots,n\}$ the set of complex and real matrices whose norms are bounded $B$.
+%We denote by $\C^{\ncn}_B=\{A\in \C^{\ncn}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\ncn}_B=\{A\in \R^{\ncn}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
+%We use $\P$ to denote the set of probability distribution over $\C^{\ncn}$.
+We also denote the set of invertible $\ncn$ complex matrices by $\gln$.
+We use $A\sim P$ to denote the fact that the random variable $A$ is distributed according to distribution $P$.
+Let the random matrix $A\sim P$, and $U\in \gln$, we use $P_U$ to denote the distribution of the random matrix $U^{-1}A U$.
+We denote the identity matrix in $\C^{\ncn}$ by $\I$.
+%We denote the set of distributions over $\C_B^{\ncn}$ by $\P$.
+
+We use $A\succeq 0$ to denote that the
+square matrix $A$ is hermitian and positive semidefinite (HPD):
+$A = A^\dag$, $\inf_x x^\dag A x\ge 0$.
+For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
+We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
+We also use $\preceq$ and $\prec$ analogously. $\E$ denotes mathematical expectation.
+
+\begin{definition}
+Let $P$ be a distribution over $\C_B^{\ncn}$, then define
+\begin{align*}
+A_P=\int M dP(M),\quad C_P=\int M^\dag M dP(M),
+\end{align*}
+Note that $\norm{A_P}<B$ and $\norm{C_P}<B^2$.
+\end{definition}
+\begin{definition}\label{contract}[Contraction Factors]
+Define
+\begin{align*}
+\rhod{P}&\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha A_P^\dag A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}
+\end{align*}
+\end{definition}
+Note that $\rhod{P}>\rhos{P}$.
+
+\begin{definition}
+We define a matrix $A\in \C^{\ncn}$ to be \emph{positive definite} if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{n}$.
+We define a matix $A\in \C^{\ncn}$ to be \emph{asymptotically stable} if all eigenvalues of $A$ have positive real parts.
+\end{definition}
+\begin{definition}\label{distpd}
+Define a distribution $P\in \P$ to be \emph{positive definite} (PD)  if $A_P$ is positive definite.
+Define a distribution $P\in \P$ to be \emph{asymptotically stable} (AS) if $A_P$ is asymptotically stable.
+\end{definition}
+
+\section{Problem Setup}
+We consider the linear stochastic approximation scheme with constant step size and PR-averaging given as below:
+\begin{align}
+\label{conststep} \theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
+\label{iteravg} \thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_s,
+\end{align}
+where $\thh_t\in \C^{n}$ in \eqref{iteravg} is PR-average of the iterates $\theta_t\in \C^{n}$ of the constant step size stochastic recursion in \eqref{conststep}.
+For the quantities in \eqref{conststep,iteravg} we assume the following:
+\begin{assumption}\label{lsadist}
+$\{A_t,\, t\geq 0\}$ is an $i.i.d$ sequence of $\ncn$ matrices distributed according to $P\in \P$ and $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^n$ with mean $\E[b_t]=b$.
+\end{assumption}
+From \cref{lsadist} it follows that $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ are martingale difference sequences with respect to the $\sigma$-field given by $\F_{t-1}=\{\theta_0,M_1,\ldots, M_{t-1}\}$.
+Further, we assume that
+\begin{assumption}\label{matvar}
+%\begin{align*}\E\left[ M_t^\dag M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
+\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
+\end{assumption}
+
+Our objective is to analyse $\EE{\norm{\thh_t-\theta^*}^2}$.
+
+\begin{example}[TD(0)]
+\end{example}
+
+\begin{example}[Linear Prediction]
+\end{example}
+
+%\section{Error Recursion and Similarity Transformation}
+
+\section{Main Results}
+\begin{lemma}
+Let $P$ be a positive definite distribution over $\C_B^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+\end{lemma}
+\begin{theorem}
+Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\ncn}$. Then
+\begin{align}
+\EE{\norm{\thh_t-\theta^*}^2}
+\leq
+\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
+\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
+%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
+\end{align}
+
+\end{theorem}
+
+\begin{lemma}[Similarity Transformation]
+Let $P^{AS}\in\P$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
+\end{lemma}
+
+\begin{theorem}[Change of Basis]
+Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
+\begin{align}
+\EE{\norm{\gamma_t-\gamma^*}^2}
+\leq
+\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
+\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
+%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
+\end{align}
+where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
+\end{theorem}
+\begin{corollary}
+Let $P$ in \Cref{lsadist} be asymptotically stable. Then, there exists
+\begin{align}
+\EE{\norm{\thh_t-\theta^*}^2}
+\leq
+\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U}^2\norm{U^{-1}}^2}{\alpha \rhos{P_U}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
+\end{align}
+\end{corollary}
+
+\section{Uniform stepsizes}
+\begin{definition}
+Call a set of distributions $\P$ over $\C_B^{\ncn}$
+\emph{weakly admissible} if there exists $\alpha>0$ such that
+$\rhos{P}>0$ holds for all $P\in \P$.
+\end{definition}
+\begin{definition}
+Call a set of distributions $\P$ over $\C_B^{\ncn}$ \emph{admissible}
+if there exists some $\alpha>0$ such that $\inf_{P\in \P} \rhos{P}>0$.
+The value of $\alpha$ is called a witness.
+\end{definition}
+It is easy to see that $\alpha \mapsto \rhos{P}$ is decreasing,
+hence if $\alpha>0$ witnesses that $\P$ is (weakly) admissible
+then any $0<\alpha'\le \alpha$ is also witnessing this.
+
+If $\P$ is weakly admissible, then one can choose a stepsize $\alpha>0$ solely based on the knowledge of $\P$ and
+conclude that no matter the joint distribution of $(A_t,b_t)$, as long as the distribution of $A_t$ is an element of $\P$, the conclusions of \cref{maintheorem} hold.
+Further, if $\P$ is admissible then the error bound stated in  \cref{maintheorem} becomes independent of the instance.
+Hence, an interesting question to investigate is whether a given set $\P$ is (weakly) admissible.
+A hopeful set is the following:
+For $B>0$, let $\P_B$ is the set of distributions where for every $P\in \P_B$, $H_P+H_P^\top\succ 0$
+and if $H\sim P$ then $\norm{H}\le B$ holds with probability one.
+In what follows all distributions will be a subset of $\P_B$ with some $B>0$,
+Note that in most applications the knowledge of a suitable $B>0$ such that $\P\subset \P_B$ is available.
+By appropriately normalizing the matrices involved, we may as well assume that $B=1$.
+
+Our first result shows that we cannot be too ambitious:
+\begin{proposition}
+The set $\P_1$ is not weakly admissible.
+\end{proposition}
+\begin{proof}
+Fix an arbitrary $\alpha>0$. We show that there exists $P\in \P$ such that $\rho_\alpha(P)<0$.
+For $\epsilon \in (0,1/2)$ let $P$ be the distribution that is supported on $\{-I,I\}$ and takes on the value of $I$ with probability $1/2+\epsilon$.
+Then $H_P = 2\epsilon I \succ 0$, hence $P\in \P_1$.
+Further, $Q_P = I$.
+Hence, $\rho_\alpha(P) = 4\epsilon-\alpha$. Hence, if $\epsilon<\alpha/4$, $\rho_\alpha(P)<0$.
+\end{proof}
+Note that the result still holds if in the definition of $\P_B$ we replace the condition that $H_P + H_P^\top \succ 0$ with the
+stronger condition that $H_P\succ 0$. In words, positive definiteness of $H_P$ alone is not sufficient for weak admissibility.
+
+In linear least-squares problems (cf. \cref{ex:leastsquares}) the set $\P$ has more structure.
+In particular, the random matrices $\{H_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
+Define
+\begin{align*}
+\P_{\text{PSD},B} = \{ P \in \P_B \,:\,  \supp(P)\subset \text{PSD} \}\,.
+\end{align*}
+We have the following proposition:
+\begin{proposition}
+For any $B>0$, the set $\P_{\text{PSD},B}$ is weakly admissible
+and in particular any $0<\alpha < 2/B$ witnesses this.
+\end{proposition}
+\begin{proof}
+Take any $P\in \P_{\text{PSD}}$ and let $H\sim P$.
+Consider the SVD of $H$: $H = U \Lambda U^\top$ where $U$ is orthonormal and $\Lambda$ is diagonal with
+nonnegative elements. Note that $\Lambda \preceq B\, \I$ and thus $\Lambda^2 \preceq B \Lambda$.
+Then for any $x\in \R^n$, $x^\top H^\top H x = x^\top U \Lambda^2 U^\top x \le B x^\top U \Lambda U^\top x = B x^\top H x$.
+Taking expectations we find that $x^\top Q_P x \le B x^\top H_P x$.
+Hence, $\rho_\alpha(P) = 2 x^\top H_P x - \alpha x^\top Q_P x \ge (2- \alpha B ) \,x^\top H_P x $.
+Thus, for any $\alpha<2/B$, $\rho_\alpha(P)>0$.
+\end{proof}
+\todoc[inline]{So this altogether is weaker than what Bach founds. They prove strong admissibility.
+I am guessing that for this they use that the loss is special and also that the noise is special.
+This would be a good place to somehow incorporate this with a (short?) proof.
+}
+
+\begin{comment}
+\paragraph{Contraction over time $t$}
+The condition of contraction per time step is overly restrictive in the general case (other than the one presented in \Cref{rankonestep}).
+We can relax this condition and ask whether a weaker result in \Cref{multheorem} holds.
+\begin{lemma}\label{multheorem}
+Let $\P$ be the space of positive definite distributions supported on $\C_B^{\ncn}$.
+Then there exists an $\alpha_{\max}>0$ such that $\alpha\in(0,\alpha_{\max})$ implies $\us{\sup}{\norm{x}\leq 1,P\in\P,t\geq 0}\E_{P}[\norm{(I-\alpha H)^tx}]< \infty$.
+\end{lemma}
+\textbf{Possible Counter example for \Cref{multheorem} in $\R$}
+Consider $\H=[-1,1]$ and $\P$ supported on $\{-1,1\}$ with $P(1)=\frac{1}{2}+\epsilon$ and $P(-1)=\frac{1}{2}-\epsilon$. Note that
+\begin{align*}
+\E[\Pi_{i=1}^t (1-\alpha H_i)]&=\exp{\ous{\sum}{i=1}{t}\log(1-\alpha H_i)}\\
+&\approx \exp{t\E[\log(1-\alpha H_i)]}
+\end{align*}
+Now the condition $\E[\Pi_{i=1}^t (1-\alpha H_i)]<\infty,$ where $H_i\sim P$ loosely translates to the following
+\begin{align*}
+\E[\log(1-\alpha H_i)]<0,
+\end{align*}
+which can be verified by noting that
+\begin{align*}
+&\E[\log(1-\alpha H_i)]\\&=\E[(\frac{1}{2}+\epsilon)\log (1-\alpha)+(\frac{1}{2}-\epsilon)\log (1+\alpha)]\\
+&=\frac{1}{2}\log(1-\alpha^2)+\epsilon(\log(1-\alpha)-\log(1+\alpha))\\
+&<0
+\end{align*}
+\textbf{Possible Counter example for \Cref{multheorem} in $\R^2$}
+Given any $\alpha$ the aim is to choose a distribution $P$ over $\H$ such that $\E_{P}[\norm{(I-\alpha H)^tx}]$ grows with $t$.
+In other words, given an $\alpha>0$, we are looking for a distribution $P'$ supported in the $\alpha$ neighbourhood of $I$ denoted by $\F$ such that $\E_{P'}[F^t]$ grows with $t$ (where $F\sim P'$).
+We now look at a possible suggestion for such a $P'$.\par
+\begin{example}
+Pick $F_1=\begin{bmatrix} 0 & 1-\beta_2 \\ 1+\beta_1& 0\\\end{bmatrix}^{\frac{1}{p}}$, $F_2=\begin{bmatrix} 0 & 1+\beta_1 \\ 1-\beta_2& 0\\\end{bmatrix}^{\frac{1}{p}}$, where $\beta_2>\beta_1$ such that $1+\frac{1}{2}(\beta_1-\beta_2)$.
+Let $P'$ be supported on $\{F_1,F_2\}$ with $P'(F_1)=\frac{1}{2}$ and $P'(F_2)=\frac{1}{2}$, and  $p$ is sufficiently large such that $F_1,F_2$ belong to the $\alpha$-neighbourhood of $I$.
+A possible way the random product can blow up is when products $F_2^pF_1^p$ occur in a contiguous stream.
+To see this,
+\begin{align}
+F_2^pF_1^p=\begin{bmatrix} (1+\beta_1) &0  \\ 0& (1-\beta_2)^2\\\end{bmatrix}
+\end{align}
+Thus $\norm{F_2^p F_1^p}>1$.
+However, such (i.e., $F_2^p F_1^p$) sequences have probability of occurance $\frac{1}{2^{2p}}$ is a sequence of length $2p$.
+Thus probability of them occuring quite often in a given random product is small.
+\end{example}
+\todoc[inline]{Did not we also want to discuss the multiplicative ergodic theorem? I have not cleaned this part yet..}
+\end{comment}
+
+\appendix
+\input{linal}
+
+\begin{comment}
+\input{linal}
 \section{Problem Setup}
@@
 
-\begin{comment}
 \begin{definition}[Induced Distribution]
 Let $P\in \P$ be any distribution over $\R^{\ncn}$ and let $p_c^U$ denote the probability distribution over $\C^{\ncn}$ induced by the linear map $\T_U$ such that the pullback of $p_c^U$ using $\T_U$ is $p$.
-\end{definition}
-\end{comment}
-
-\begin{definition}\label{def:spectnorm}[Spectral Norms]
-Define $\rhod{P}\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{\bar{x},\left((A_P+A_P^\dag)-\alpha A_P^\dag A_P\right)x}$, $\rhos{P}\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{\bar{x},\left((A_P+A_P^\dag)-\alpha C_P\right)x}$
 \end{definition}
@@
 \end{example}
-\begin{comment}
+
 \begin{lemma}\label{pdt}
@@
 \end{example}
-\end{comment}
-\begin{comment}
+
 \begin{lemma}
@@
 \end{corollary}
-\end{comment}
+
 \section{Main Results}
@@
 
+\end{comment}
+
 \end{document}

### 2017-04-27 linal.tex
updated

--- 2017-04-22 linal.tex
@@
 \section{Linear Algebra Preliminaries}
-
-We now present some useful results from linear algebra. In what follows, $A$ is an $\ncn$ matrix and $\I$ the $\ncn$.
-\subsection{Results in Matrix Decomposition and Transformation}
+We now present some useful results from linear algebra.
+\subsection{Notation}
 Let $B$ be a $\ncn$ block diagonal matrix given by $B=\begin{bmatrix} B_1 &0 &0 &\ldots &0 \\ 0 &B_2 &0 &\ldots &0  \\ \vdots &\vdots &\vdots &\vdots &\vdots \\ 0 &\ldots &0 &0 &B_k \end{bmatrix}$, where $B_i$ is a $n_i \times n_i$ matrix such that $n_i<n$ (w.l.o.g) and $\sum_{i=1}^k n_i=n$.
@@
 \begin{align*}
-B=B_1 \op B_2 \op \ldots B_k
+B=B_1 \op B_2 \op \ldots B_k=\op_{i=1}^k B_i
 \end{align*}
+
+\subsection{Results in Matrix Decomposition and Transformation}
+We will now recall Jordon decomposition.
+
+\begin{comment}
 \begin{lemma}
-There exists a polynomial known as the \emph{characteristic} polynomial denoted by $\chi_A=(\lambda_1 \I-A)^{n_1}\ldots(\lambda_{k} \I-A)^{n_k}$, where $\lambda_i,\,i=1,\ldots,k (k\leq n)$ are the $k$ distinct eigenvalues of the matrix $A$ such that $\sum_{i=1}^k n_i=n$.
+Given a matrix $A\in \C^{\ncn} there exists a \emph{characteristic} polynomial denoted by $\chi_A=(\lambda_1 \I-A)^{n_1}\ldots(\lambda_{k} \I-A)^{n_k}$, where $\lambda_i,\,i=1,\ldots,k (k\leq n)$ are the $k$ distinct eigenvalues of the matrix $A$ such that $\sum_{i=1}^k n_i=n$.
 \end{lemma}
@@
 \end{lemma}
-\begin{lemma}
-There exists a complex matrix $V\in \C^{\ncn}$ such that $A=V\Lambda V^{-1}$, where
-$\tL=\tL_1\op\ldots\op\tL_k$, where each $\tL_i,\,i=1,\ldots,k$ can further be written as $\tL_i= {\tJ}^i_{d^i_1}\op \ldots \op {\tJ}^i_{d^i_{l(i)}}$
-such that  $n'_i=\max\{d^i_1,\ldots, d^i_{l(i)} \}$ and $\sum_{j=1}^{l(i)} d^i_j =n_i$. Each of ${\tJ}^i_{d^i_j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
-${\tJ}^i_{d_j}=\begin{bmatrix} \lambda_i &1 &0 &\ldots &0 &0\\ 0 &\lambda_i &1 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &1 \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.
+\end{comment}
+
+\begin{lemma}\label{jordon}
+Let $A\in \C^{\ncn}$ and $\{\lambda_i\in \C,i=1,\ldots,k\leq n \}$ denote its $k$ distinct eigenvalues.
+There exists a complex matrix $V\in \C^{\ncn}$ such that $A=V\tL V^{-1}$, where
+$\tL=\tL_1\op\ldots\op\tL_k$, where each $\tL_i,\,i=1,\ldots,k$ can further be written as $\tL_i= {\tL}^i_{1}\op \ldots \op {\tL}^i_{{l(i)}}$.
+Each of ${\tL}^i_{j},j=1,\ldots,l(i)$ is a $d^i_j\times d^i_j$ square matrix such that $\sum_{j=1}^{l(i)} d^i_j =n_i$ and has the special form given by
+${\tL}^i_{j}=\begin{bmatrix} \lambda_i &1 &0 &\ldots &0 &0\\ 0 &\lambda_i &1 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &1 \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.
 \end{lemma}
 
-\begin{lemma}
-Given a AS matrix $A$, there exists a matrix $U_A\in \C^{\ncn}$ such that $A=U_A\Lambda_A U_A^{-1}$ where $\Lambda_A^\dag+\Lambda_A$ is a real symmetric positive definite matrix.
+\begin{lemma}\label{simtran}
+Let $A\in \C^{\ncn}$ be asymptotically stable. There exists a matrix $U\in \gln$ such that $A=U\Lambda U^{-1}$ and $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
 \end{lemma}
 \begin{proof}
-Consider the diagonal matrices $D^i_j=\begin{bmatrix} 1  &0 &0 &\ldots &0 &0\\ 0 &\re(\lambda_i) &0 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re(\lambda)^{d^i_j-1}_i &0 \\ 0 &\ldots &0 &0 &0 &\re(\lambda)^{d^i_j}_i \end{bmatrix},\,\forall j=1,\ldots,l(i)$, $D^i=D^i_1 \op\ldots\op D^i_{l(i)},\,\forall i=1,\ldots,k$ and $D=D^1 \op\ldots\op D^k$.
-It follows that $A=(VD) \Lambda (VD)^{-1}$, where $\Lambda$ is a matrix such that
-$\Lambda=\Lambda_1 \op \ldots \op \Lambda_k$, where each $\Lambda_i,\,i=1,\ldots,k$ can further be written as
-$\Lambda_i=J^i_{d^i_1} \op \ldots \op J^i_{d^i_{l(i)}}$ such that  $n'_i=\max\{d^i_1,\ldots, d^i_{l(i)} \}$ and $\sum_{j=1}^{l(i)} d^i_j =n_i$.
-Each of $J^i_{d^i_j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
-$J^i_{d_j}=\begin{bmatrix} \lambda_i &\re(\lambda_i) &0 &\ldots &0 &0\\ 0 &\lambda_i &\re(\lambda_i) &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &\re(\lambda_i) \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.
+It is trivial to see that for any $\Lambda\in \C^{\ncn}$, $\left(\Lambda^\dag+\Lambda\right)$ is real symmetric.
+We will use the decomposition of $A=V \tL V^{-1}$ in \Cref{jordon} and also carry over the notations in \Cref{jordon}.
+Consider the diagonal matrices $D^i_j=\begin{bmatrix} 1  &0 &0 &\ldots &0 &0\\ 0 &\re{\lambda_i} &0 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda}^{d^i_j-1}_i &0 \\ 0 &\ldots &0 &0 &0 &\re{\lambda}^{d^i_j}_i \end{bmatrix},\,\forall j=1,\ldots,l(i)$, $D^i=D^i_1 \op\ldots\op D^i_{l(i)},\,\forall i=1,\ldots,k$ and $D=D^1 \op\ldots\op D^k$.
+It follows that $A=(VD) \Lambda (VD)^{-1}$, where $J$ is a matrix such that
+$\Lambda=A_1 \op \ldots \op A_k$, where each $A_i,\,i=1,\ldots,k$ can further be written as
+$A_i=\Lambda^i_{1} \op \ldots \op \Lambda^i_{{l(i)}}$. Each of $\Lambda^i_{j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
+$\Lambda^i_{j}=\begin{bmatrix} \lambda_i &\re{\lambda_i} &0 &\ldots &0 &0\\ 0 &\lambda_i &\re{\lambda_i} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &\re{\lambda_i} \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.
 
-Now let $\frac{(\Lambda^\dag+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}(J^{i\dag}_{d_j}+J^i_{d_j})$, where $J^{i\dag}_{d_j}+J^i_{d_j}=\begin{bmatrix} \re(\lambda_i) &\frac{\re(\lambda_i)}{2} &0 &\ldots &0 &0\\ \frac{\re(\lambda_i)}{2} &\re(\lambda_i) &\frac{\re(\lambda_i)}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re(\lambda_i) &\frac{\re(\lambda_i)}{2} \\ 0 &\ldots &0 &0 &\frac{\re(\lambda_i)}{2} &\re(\lambda_i) \end{bmatrix} $.
-Then for any $x\in \C^n (\neq \mathbf{0})$, there exists a $b\in {-1,1}^n$, such that
+Now let $\frac{(\Lambda^\dag+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}$, where $\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}=\begin{bmatrix} \re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 &0\\ \frac{\re{\lambda_i}}{2} &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} \\ 0 &\ldots &0 &0 &\frac{\re{\lambda_i}}{2} &\re{\lambda_i} \end{bmatrix} $.
+Then for any $x=(x(i),i=1,\ldots,n)\in \C^n (\neq \mathbf{0})$, there exists a $b\in \{-1,1\}^n$, such that
 \begin{align*}
-x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\sum_{i=1}^n \ip{x_i,x_i}+\sum_{i=1}^{n-1}\ip{x_i,x_{i+1}}\\
-&> (\sum_{i=1}^n b_i x_i)^2\\
-&\geq 0
+x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\re{\lambda_i} \left(\sum_{i=1}^n \bar{x}{(i)} x(i)+\sum_{i=1}^{n-1} \frac{\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)}{2}\right) \\
+&=\frac{\re{\lambda_i}}{2}\left(\md{x(1)}^2+ \md{x(n)}^2\right)+\frac{\re{\lambda_i}}{2}\left( \sum_{i=1}^{n-1} \md{x(i)}^2+\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)+\md{x(i+1)}^2 \right)\\
+&>\frac{\re{\lambda_i}}{2}\left(\sum_{i=1}^n \md{x(i)+x(i+1)}^2 \right)\\
+&> 0
 \end{align*}
@@
 \section{Error Recursion}
-\begin{assumption}\label{cmplxassmp}
-\begin{enumerate}[leftmargin=*, before=\leavevmode\vspace{-\baselineskip}]
-\item $\Lt\eqdef \Lambda + L_t$ where $\Lambda$ is a positive definite matrix and $\{L_t \in \C^{\ncn},\,t> 0\}$ is a sequence of $i.i.d$ matrices such that $\E[L_t]=\zero,\,\forall t>0$.
-\item $\{\zeta_t,\,t\geq 0\in \C^n\}$ is a sequence of noise vectors such that $\E[\zeta_t]=0,\,\forall t\geq 0$.
-\item $\E[\norm{L_t \zeta_t}]=\hat{\sigma}_2^2$, $\E[\norm{\zeta_t}^2]=\hat{\sigma}_1^2$
-\end{enumerate}
-\end{assumption}
 
-\begin{definition}
-Define $\forall,\,i\geq j$, $F_{i,j}=(I-\alpha \L_i)\ldots (I-\alpha \L_j)$ and $\forall,\,i<j$ $F_{i,j}=\I$.
-\end{definition}
+\begin{lemma}
+Let $P$ be a positive definite distribution over $\C_B^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+\end{lemma}
+\begin{proof}
+Since $P$ is positive definite from \Cref{distpd} it follows that $A_P$ is positive definite. Now
+\begin{align*}
+\rhos{P}&={\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}\\
+&\geq {\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)\right)x}-\alpha B^2\\
+&\geq \lambda_{\min}(A_P+A_P^\dag)-\alpha B^2
+\end{align*}
+By choosing $\alpha_P<\frac{\lambda_{\min}(A_P+A_P^\dag)}{B^2}$, it follows that $\rhos{P}>0,\forall \alpha \in (0,\alpha_P)$.
+Further, by noting that $C_P\succeq A_P^\dag A_P$, it is easy to check that $\rhod{P}>\rhos{P}>0,\,\forall \alpha \in(0,\alpha_P)$.
+\end{proof}
 
+\begin{theorem}\label{pdrate}
+Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\ncn}$. Then
 \begin{align}
+\EE{\norm{\thh_t-\theta^*}^2}
+\leq
+\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
+\end{align}
+\end{theorem}
+\begin{proof}
+
+\begin{align*}
 \theta_t&=\theta_{t-1}+\alpha\big(b_t-A_t\theta_{t-1}\big)\\
@@
 e_t&=(I-\alpha A_t)e_{t-1}+(b_t -b -(A_t-A)\ts)\\
-U^{-1}e_t&=(I-\alpha U^{-1}A_t U) U^{-1}e_{t-1}+ U^{-1}(b_t -b -(A_t-A)\ts)\\
-z_t&=(I-\alpha \Lt) z_{t-1}+ \zeta_t
-\end{align}
+\end{align*}
+Now define $\forall,\,i\geq j$, $F_{i,j}=(I-\alpha A_i)\ldots (I-\alpha A_j)$ and $\forall,\,i<j$ $F_{i,j}=\I$.
+Also let $\zeta_t\eqdef b_t-b-(A_t-A)\ts$ and $\EE{\norm{\zeta_t}^2}=\sigma_1^2=\sigma_P^2\norm{\ts}^2+\sigma_b^2$ and $\EE{\norm{M_t\zeta_t}}=\sigma_2^2=\sigma_P^2\norm{\ts}$.
 
 \begin{align*}
-z_t
-& = (I-\alpha \Lambda_t) (I-\alpha \Lambda_{t-1}) z_{t-2}\\ &+ \alpha (I-\alpha \Lambda_t) \zeta_{t-1} +\alpha \zeta_t \\
+e_t
+& = (I-\alpha A_t) (I-\alpha A_{t-1}) e_{t-2}\\ &+ \alpha (I-\alpha A_t) \zeta_{t-1} +\alpha \zeta_t \\
 & \quad \vdots\\
-& = (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_1) e_0\\ &+ \alpha (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_2) \zeta_1 \\
-& + \alpha (I-\alpha \Lambda_t) \cdots (I-\alpha \Lambda_3) \zeta_2\\
+& = (I-\alpha A_t) \cdots (I-\alpha A_1) e_0\\ &+ \alpha (I-\alpha A_t) \cdots (I-\alpha A_2) \zeta_1 \\
+& + \alpha (I-\alpha A_t) \cdots (I-\alpha A_3) \zeta_2\\
 &  \quad \vdots \\
@@
 \label{eq:etft}
-z_t = F_{t,1} z_0 + \alpha (F_{t,2} \zeta_1 + \dots + F_{t,t+1} \zeta_t )\,,
+e_t = F_{t,1} e_0 + \alpha (F_{t,2} \zeta_1 + \dots + F_{t,t+1} \zeta_t )\,,
 \end{align}
 \begin{align*}
-\zh_t=\frac{1}{t+1}{\sum}_{i=0}^{t}z_i
-=\frac{1}{t+1}&\Big\{{\sum}_{i=0}^{t} F_{i,1} z_0 \\
+\eh_t=\frac{1}{t+1}{\sum}_{i=0}^{t}e_i
+=\frac{1}{t+1}&\Big\{{\sum}_{i=0}^{t} F_{i,1} e_0 \\
 &+ \alpha \sum_{i=1}^{t} \left(\sum_{k=i}^{t} F_{k,i+1} \right)\zeta_i \Big\} ,
@@
 where in the second sum we flipped the order of sums and swapped the names of the variables that the sum runs over.
-
 It follows that \todoc{We should rather use $C$ instead of $H$ here?}
 \begin{align*}
-\E[\norm{\zh_t}^2]&=\E\ip{\zh_t,\zh_t}
-=\frac{1}{(t+1)^2} \sum_{i,j=0}^t \E\ip{z_i,z_j}\,.
+\E[\norm{\eh_t}^2]&=\E\ip{\eh_t,\eh_t}
+=\frac{1}{(t+1)^2} \sum_{i,j=0}^t \E\ip{e_i,e_j}\,.
 \end{align*}
@@
 \begin{align*}
-\E\ip{z_i,z_j}
-&=\E \ip{z_i,\big[F_{j,i+1} e_i+\alpha\textstyle\sum_{k=i+1}^j F_{j,k+1}\zeta_{k}\big]}\\
-&=\E\ip{z_i,F_{j,i+1} z_i}  \text{(from \cref{noisecancel})}\\
-&=\E\ip{z_i, (I-\alpha \Lambda)^{j-i} z_i} \text{(from \cref{lem:unroll})}
+\E\ip{e_i,e_j}
+&=\E \ip{e_i,\big[F_{j,i+1} e_i+\alpha\textstyle\sum_{k=i+1}^j F_{j,k+1}\zeta_{k}\big]}\\
+&=\E\ip{e_i,F_{j,i+1} e_i}  \text{(from \cref{noisecancel})}\\
+&=\E\ip{e_i, (I-\alpha A)^{j-i} e_i} \text{(from \cref{lem:unroll})}
 \end{align*}
@@
 \label{inter}
-\sum_{i=0}^{t-1}\sum_{j=i+1}^t \E\ip{z_i,z_j}
-&=\frac1{\alpha\rhod{P}} {\sum}_{i=0}^{t-1}\E\ip{z_i,z_i}\\
-&\leq \frac2{\alpha\rhod{P}}{\sum}_{i=0}^{t}\E\ip{z_i,z_i}\,.
+\sum_{i=0}^{t-1}\sum_{j=i+1}^t \E\ip{e_i,e_j}
+&=\frac1{\alpha\rhod{P}} {\sum}_{i=0}^{t-1}\E\ip{e_i,e_i}\\
+&\leq \frac2{\alpha\rhod{P}}{\sum}_{i=0}^{t}\E\ip{e_i,e_i}\,.
 \end{align*}
@@
 \begin{align*}
-{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{z_i,z_j}&= \left(1+\frac2{\alpha\rhod{P}}\right){\sum}_{i=0}^{t}\E\ip{z_i,z_i}\,.
+{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{e_i,e_j}&= \left(1+\frac2{\alpha\rhod{P}}\right){\sum}_{i=0}^{t}\E\ip{e_i,e_i}\,.
 \end{align*}
-Expanding $z_i$ using \eqref{eq:etft} and then using \cref{innerproduct} and \Cref{cmplxassmp}
+Expanding $e_i$ using \eqref{eq:etft} and then using \cref{innerproduct} and \Cref{cmplxassmp}
 \begin{align*}
-\E\ip{z_i,z_i}&=\E\ip{F_{i,1}z_0,F_{i,1}z_0}+\alpha^2{\sum}_{j=1}^i\E\ip{ F_{i,j+1}\zeta_j, F_{i,j+1}\zeta_j}+\alpha\sum_{j=1}^i  \E\ip{F_{i,1} z_0, F_{i,j+1}\zeta_j}\\
-&\leq (1-\alpha\rhos{P})^i\norm{z_0}^2+ \alpha^2\frac{\hat{\sigma}_1^2}{\alpha \rhos{P}}+ \alpha \frac{\hat{\sigma}^2_2 \norm{z_0}}{\alpha\rhos{P}}\,,
+\E\ip{e_i,e_i}&=\E\ip{F_{i,1}e_0,F_{i,1}e_0}+\alpha^2{\sum}_{j=1}^i\E\ip{ F_{i,j+1}\zeta_j, F_{i,j+1}\zeta_j}+\alpha\sum_{j=1}^i  \E\ip{F_{i,1} e_0, F_{i,j+1}\zeta_j}\\
+&\leq (1-\alpha\rhos{P})^i\norm{e_0}^2+ \alpha^2\frac{{\sigma}_1^2}{\alpha \rhos{P}}+ \alpha \frac{{\sigma}^2_2 \norm{e_0}}{\alpha\rhos{P}}\,,
 \end{align*}
@@
 \begin{align*}
-{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{z_i,z_j}
-&\leq \left(1+\frac2{\alpha\rhod{P} }\right)\, \frac1{\alpha\rhos{P}}\, (t(\alpha^2\hat{\sigma}_1^2+\alpha \hat{\sigma}^2_2\norm{z_0}) +\norm{z_0}^2)\,.
+{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{e_i,e_j}
+&\leq \left(1+\frac2{\alpha\rhod{P} }\right)\, \frac1{\alpha\rhos{P}}\, (t(\alpha^2{\sigma}_1^2+\alpha {\sigma}^2_2\norm{e_0}) +\norm{e_0}^2)\,.
 \end{align*}
@@
 \begin{align}
+\E[\norm{\eh_t}^2]
+\leq \left(1+\frac2{\alpha\rhod{P}}\right)\, \frac1{\alpha\rhos{P}}\, \,
+\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{e_0}}{t+1} \right)\,.
+\end{align}
+\end{proof}
+
+\begin{lemma}[Similarity Transformation]
+Let $P^{AS}\in\P$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
+\end{lemma}
+\begin{proof}
+Follows from \Cref{simtran}
+\end{proof}
+
+\begin{theorem}[Change of Basis]
+Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
+\begin{align}
+\EE{\norm{\gamma_t-\gamma^*}^2}
+\leq
+\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
+\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
+%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
+\end{align}
+where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
+\end{theorem}
+\begin{proof}
+From proof of \Cref{pdrate} we know that
+\begin{align*}
+e_t&=(I-\alpha A_t)e_{t-1}+\zeta_t\\
+U^{-1}e_t&=(I-\alpha U^{-1}A_t U) U^{-1}e_{t-1}+ U^{-1}\zeta_t\\
+z_t&=(I-\alpha \Lambda_t) z_{t-1}+ H_t,
+\end{align*}
+where $z_t=\gamma_t-\gamma^*$, $\Lambda_t=U^{-1}A_t U$ and $H_t=U^{-1}\zeta_t$.
+we know that $\EE{\norm{H_t}^2}\leq \norm{U^{-1}}^2\EE{\norm{\zeta}}$ and $\EE{\norm{\Lambda_t H_t}}=\EE{\norm{U^{-1}A_t UU^-1\zeta_t}}=\EE{\norm{U^{-1}A_t \zeta_t}}\leq \norm{U^-1}\EE{\norm{A_t\zeta_t}}=\norm{U^{-1}}\sigma_2^2$.
+We follow through the steps of \Cref{pdrate} to obtain
+\begin{align}
 \E[\norm{\zh_t}^2]
-\leq \left(1+\frac2{\alpha\rhod{P}}\right)\, \frac1{\alpha\rhos{P}}\, \,
-\left(\frac{\norm{z_0}^2}{(t+1)^2}+ \frac{\alpha^2\hat{\sigma}_1^2+\alpha \hat{\sigma}^2_2\norm{z_0}}{t+1} \right)\,.
+&\leq \left(1+\frac2{\alpha\rhod{P_U}}\right)\, \frac1{\alpha\rhos{P_U}}\, \,
+\left(\frac{\norm{z_0}^2}{(t+1)^2}+ \frac{\alpha^2(\norm{U^{-1}}^2\sigma_1^2)+\alpha (\norm{U^{-1}}\sigma_2^2)\norm{z_0}}{t+1} \right)\,\\
+&\leq \left(1+\frac2{\alpha\rhod{P_U}}\right)\, \frac1{\alpha\rhos{P_U}}\, \,
+\left(\frac{\norm{U^{-1}}^2\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2(\norm{U^{-1}}^2\sigma_1^2)+\alpha (\norm{U^{-1}}\sigma_2^2)\norm{U^{-1}}\norm{e_0}}{t+1} \right)\,
 \end{align}
 
+\end{proof}
+
+Let $\F_t = \sigma( A_1,\dots,A_t, \zeta_1,\dots,\zeta_t ) $, with $\F_0$ the $\sigma$-field that holds all random variables.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
+Let $t>i\ge 1$, $x,y\in \C^n$ be $\F_{i}$-measurable random vectors. Then,
+\begin{align*}
+\E[x^\dag F_{t,i+1}y|\F_i]=x^\dag (I-\alpha H)^{t-i} y\,.
+\end{align*}
+\end{lemma}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proof}
+By the definition of $F_{t,i+1}$,
+and because $F_{t-1,i+1} = (I-\alpha A_{t-1}) \dots (I-\alpha A_{i+1})$ is $\F_{t-1}$-measurable,
+as are $x$ and $y$,
+\begin{align*}
+\EE{x^\dag F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha A_t) | \F_{t-1} } F_{t-1,i+1} y\\
+&=x^\dag  (I-\alpha A_P)  F_{t-1,i+1} y\,.
+\end{align*}
+By the tower-rule for conditional expectations and our measurability assumptions,
+\begin{align*}
+\EE{x^\dag F_{t,i+1} y | \F_{t-2} }
+&=x^\dag  (I-\alpha A_P)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
+&= x^\dag (I-\alpha A_P)^2 F_{t-2,i+1} y\,.
+\end{align*}
+Continuing this way we get
+\begin{align*}
+\EE{x^\dag F_{t,i+1} y | \F_{t-j} }
+= x^\dag (I-\alpha A_P)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
+\end{align*}
+Specifically, for $j=t-i$ we get
+\begin{align*}
+\EE{x^\dag F_{t,i+1} y | \F_{i} }  = x^\dag (I-\alpha A_P)^{t-i} y\,.
+\end{align*}
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{lemma}\label{noisecancel}
+Let $t>i\ge 1$ and let $x$ be an $\F_{i-1}$-measurable random vector. Then,
+$\E[x^\dag F_{t,i+1}\zeta_{i}]=0$.
+\end{lemma}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proof}
+By \cref{lem:genunroll},
+\begin{align*}
+\EE{x^\dag F_{t,i+1} \zeta_i | \F_{i} }  = x^\dag (I-\alpha A_P)^{t-i} \zeta_i\,.
+\end{align*}
+Using the tower rule,
+\begin{align*}
+\EE{x^\dag F_{t,i+1} \zeta_i | \F_{i-1} }
+= x^\dag (I-\alpha A_P)^{t-i}\EE{ \zeta_i | \F_{i-1} }= 0\,.
+\end{align*}
+\end{proof}
+
+\begin{lemma}\label{lem:unroll}
+For all $t>i\ge 0$, $\E \ip{e_i,F_{t,i+1} e_i}=\E\ip{e_i,(I-\alpha A_P)^{t-i} e_i}$.
+\end{lemma}
+\begin{proof}
+The lemma follows directly from \cref{lem:genunroll}. Indeed,
+$\theta_i$ depends only on $A_1,\dots,A_{i},b_1,\dots,b_{i}$, $\theta_i$ and so is $e_i$ $\F_i$-measurable.
+Hence, the lemma is applicable and implies that
+\begin{align*}
+\EE{ \ip{e_i, F_{t,i+1} e_i} | \F_i } =
+\EE{ \ip{e_i, (I-\alpha A_P)^{t-i} e_i} | \F_i }\,.
+\end{align*}
+Taking expectation of both sides gives the desired result.
+\end{proof}
+
+\begin{lemma}\label{innerproduct}
+Let $i>j \ge 0$ and let $x\in \R^n$ be an $\F_j$-measurable random vector.
+Then,
+\begin{align*}
+\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rho_\alpha)^{i-j}\E\norm{x}^2\,.
+\end{align*}
+\end{lemma}
+\begin{proof}
+Note that
+$S_t\doteq \EE{ (I-\alpha A_t)^\dag (I-\alpha A_t) | \F_{t-1} }
+= I - \alpha (A_P^\dag + A_P) + \alpha^2 \EE{ A_t^\dag A_t | \F_{t-1} }$.
+Since $(b_t,A_t)_t$ is an independent sequence, $\EE{ A_t^\dag A_t|\F_{t-1}} = \EE{ A_1^\dag A_1 }$.
+Now, using the definition of $\rhos{P}$ from \eqref{contract},
+$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (A_P^\dag + A_P - \alpha \EE{A_1^\top A_1}) x
+= 1-\alpha \rhos{P}$.
+Hence,
+\begin{align*}
+&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
+&= \EE{x^\dag F_{i-1,j+1}^\top (I-\alpha A_i)^\dag (I-\alpha A_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
+&=(x F_{i-1,j+1})^\dag \, S_i \, F_{i-1,j+1} x\\
+&\le (1-\alpha \rhos{P}) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
+& \le (1-\alpha \rhos{P})^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
+& \quad \vdots \\
+& \le (1-\alpha \rho_\alpha)^{i-j}\, \norm{x}^2\,.
+%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
+%&=(1-\alpha\rhos{P}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
+\end{align*}
+\end{proof}
+
+\begin{comment}
 Let $\F_t = \sigma( H_1,\dots,H_t, \zeta_1,\dots,\zeta_t ) $, with $\F_0$ the $\sigma$-field that holds all random variables.
-\newcommand{\cF}{\F}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 By the definition of $F_{t,i+1}$,
-and because $F_{t-1,i+1} = (I-\alpha H_{t-1}) \dots (I-\alpha H_{i+1})$ is $\cF_{t-1}$-measurable,
+and because $F_{t-1,i+1} = (I-\alpha H_{t-1}) \dots (I-\alpha H_{i+1})$ is $\F_{t-1}$-measurable,
 as are $x$ and $y$,
 \begin{align*}
-\EE{x^\top F_{t,i+1} y | \cF_{t-1} } &= x^\top \EE{ (I-\alpha H_t) | \cF_{t-1} } F_{t-1,i+1} y\\
+\EE{x^\top F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha H_t) | \F_{t-1} } F_{t-1,i+1} y\\
 &=x^\top  (I-\alpha H)  F_{t-1,i+1} y\,.
@@
 \begin{align*}
-\EE{x^\top F_{t,i+1} y | \cF_{t-2} }
-&=x^\top  (I-\alpha H)  \EE{F_{t-1,i+1} |\cF_{t-2}} y\\
+\EE{x^\top F_{t,i+1} y | \F_{t-2} }
+&=x^\top  (I-\alpha H)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
 &= x^\top (I-\alpha H)^2 F_{t-2,i+1} y\,.
@@
 \begin{align*}
-\EE{x^\top F_{t,i+1} y | \cF_{t-j} }
+\EE{x^\top F_{t,i+1} y | \F_{t-j} }
 = x^\top (I-\alpha H)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
@@
 \begin{align*}
-\EE{x^\top F_{t,i+1} y | \cF_{i} }  = x^\top (I-\alpha H)^{t-i} y\,.
+\EE{x^\top F_{t,i+1} y | \F_{i} }  = x^\top (I-\alpha H)^{t-i} y\,.
 \end{align*}
@@
 \begin{align*}
-\EE{x^\top F_{t,i+1} \zeta_i | \cF_{i} }  = x^\top (I-\alpha H)^{t-i} \zeta_i\,.
+\EE{x^\top F_{t,i+1} \zeta_i | \F_{i} }  = x^\top (I-\alpha H)^{t-i} \zeta_i\,.
 \end{align*}
@@
 \begin{align*}
-\EE{x^\top F_{t,i+1} \zeta_i | \cF_{i-1} }
-= x^\top (I-\alpha H)^{t-i}\EE{ \zeta_i | \cF_{i-1} }= 0\,.
+\EE{x^\top F_{t,i+1} \zeta_i | \F_{i-1} }
+= x^\top (I-\alpha H)^{t-i}\EE{ \zeta_i | \F_{i-1} }= 0\,.
 \end{align*}
@@
 The lemma follows directly from \cref{lem:genunroll}. Indeed,
-$\theta_i$ depends only on $H_1,\dots,H_{i},g_1,\dots,g_{i}$, $\theta_i$ and so is $e_i$ $\cF_i$-measurable.
+$\theta_i$ depends only on $H_1,\dots,H_{i},g_1,\dots,g_{i}$, $\theta_i$ and so is $e_i$ $\F_i$-measurable.
 Hence, the lemma is applicable and implies that
 \begin{align*}
-\EE{ \ip{He_i, HF_{t,i+1} e_i} | \cF_i } =
-\EE{ \ip{He_i, H (I-\alpha H)^{t-i} e_i} | \cF_i }\,.
+\EE{ \ip{He_i, HF_{t,i+1} e_i} | \F_i } =
+\EE{ \ip{He_i, H (I-\alpha H)^{t-i} e_i} | \F_i }\,.
 \end{align*}
@@
 \end{proof}
+
+\begin{align*}
+U^{-1}e_t&=(I-\alpha U^{-1}A_t U) U^{-1}e_{t-1}+ U^{-1}(b_t -b -(A_t-A)\ts)\\
+e_t&=(I-\alpha J_t) e_{t-1}+ H_t
+\end{align*}
+
+\begin{align*}
+e_t
+& = (I-\alpha J_t) (I-\alpha J_{t-1}) e_{t-2}\\ &+ \alpha (I-\alpha J_t) \zeta_{t-1} +\alpha H_t \\
+& \quad \vdots\\
+& = (I-\alpha J_t) \cdots (I-\alpha J_1) e_0\\ &+ \alpha (I-\alpha J_t) \cdots (I-\alpha J_2) \zeta_1 \\
+& + \alpha (I-\alpha J_t) \cdots (I-\alpha J_3) \zeta_2\\
+&  \quad \vdots \\
+&+ \alpha H_t\,,
+\end{align*}
+which can be written compactly as
+\begin{align}
+\label{eq:etft}
+e_t = F_{t,1} e_0 + \alpha (F_{t,2} \zeta_1 + \dots + F_{t,t+1} H_t )\,,
+\end{align}
+\begin{align*}
+\zh_t=\frac{1}{t+1}{\sum}_{i=0}^{t}e_i
+=\frac{1}{t+1}&\Big\{{\sum}_{i=0}^{t} F_{i,1} e_0 \\
+&+ \alpha \sum_{i=1}^{t} \left(\sum_{k=i}^{t} F_{k,i+1} \right)\zeta_i \Big\} ,
+\end{align*}
+where in the second sum we flipped the order of sums and swapped the names of the variables that the sum runs over.
+
+It follows that \todoc{We should rather use $C$ instead of $H$ here?}
+\begin{align*}
+\E[\norm{\zh_t}^2]&=\E\ip{\zh_t,\zh_t}
+=\frac{1}{(t+1)^2} \sum_{i,j=0}^t \E\ip{e_i,e_j}\,.
+\end{align*}
+Hence, we see that it suffices to bound $\EE{\ip{ e_i,  e_j }}$.
+There are two cases depending on whether $i=j$. When $i< j$,
+\begin{align*}
+\E\ip{e_i,e_j}
+&=\E \ip{e_i,\big[F_{j,i+1} e_i+\alpha\textstyle\sum_{k=i+1}^j F_{j,k+1}\zeta_{k}\big]}\\
+&=\E\ip{e_i,F_{j,i+1} e_i}  \text{(from \cref{noisecancel})}\\
+&=\E\ip{e_i, (I-\alpha J)^{j-i} e_i} \text{(from \cref{lem:unroll})}
+\end{align*}
+and therefore
+\begin{align*}
+\label{inter}
+\sum_{i=0}^{t-1}\sum_{j=i+1}^t \E\ip{e_i,e_j}
+&=\frac1{\alpha\rhod{P}} {\sum}_{i=0}^{t-1}\E\ip{e_i,e_i}\\
+&\leq \frac2{\alpha\rhod{P}}{\sum}_{i=0}^{t}\E\ip{e_i,e_i}\,.
+\end{align*}
+Since $\sum_{i,j}\cdot{} = \sum_{i=j}\cdot{} + 2 \sum_i \sum_{j>i} \cdot{}$,
+\begin{align*}
+{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{e_i,e_j}&= \left(1+\frac2{\alpha\rhod{P}}\right){\sum}_{i=0}^{t}\E\ip{e_i,e_i}\,.
+\end{align*}
+Expanding $e_i$ using \eqref{eq:etft} and then using \cref{innerproduct} and \Cref{cmplxassmp}
+\begin{align*}
+\E\ip{e_i,e_i}&=\E\ip{F_{i,1}e_0,F_{i,1}e_0}+\alpha^2{\sum}_{j=1}^i\E\ip{ F_{i,j+1}\zeta_j, F_{i,j+1}\zeta_j}+\alpha\sum_{j=1}^i  \E\ip{F_{i,1} e_0, F_{i,j+1}\zeta_j}\\
+&\leq (1-\alpha\rhos{P})^i\norm{e_0}^2+ \alpha^2\frac{{\sigma}_1^2}{\alpha \rhos{P}}+ \alpha \frac{{\sigma}^2_2 \norm{e_0}}{\alpha\rhos{P}}\,,
+\end{align*}
+and so
+\begin{align*}
+{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{e_i,e_j}
+&\leq \left(1+\frac2{\alpha\rhod{P} }\right)\, \frac1{\alpha\rhos{P}}\, (t(\alpha^2{\sigma}_1^2+\alpha {\sigma}^2_2\norm{e_0}) +\norm{e_0}^2)\,.
+\end{align*}
+Putting things together,
+\begin{align}
+\E[\norm{\zh_t}^2]
+\leq \left(1+\frac2{\alpha\rhod{P}}\right)\, \frac1{\alpha\rhos{P}}\, \,
+\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+\alpha {\sigma}^2_2\norm{e_0}}{t+1} \right)\,.
+\end{align}
+
+\end{comment}

### 2017-04-30 flsa.tex
changes

@@
 We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
-We also use $\preceq$ and $\prec$ analogously. $\E$ denotes mathematical expectation.
+We also use $\preceq$ and $\prec$ analogously. For any $x\in\C^n$x, we denote the general quadratic norm with respect to a $C\succ 0$ by $\norm{x}^2_C\eqdef x^\dag\, C \,x$.
+$\E$ denotes mathematical expectation.
 
 \begin{definition}
-Let $P$ be a distribution over $\C_B^{\ncn}$, then define
+Let $P$ be a distribution over $\C^{\ncn}$, then define
 \begin{align*}
@@
 \end{align*}
-Note that $\norm{A_P}<B$ and $\norm{C_P}<B^2$.
 \end{definition}
@@
 \begin{definition}
-We define a matrix $A\in \C^{\ncn}$ to be \emph{positive definite} if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{n}$.
-We define a matix $A\in \C^{\ncn}$ to be \emph{asymptotically stable} if all eigenvalues of $A$ have positive real parts.
+We call a matrix $A\in \C^{\ncn}$ to be \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{n}$.
+We call a matix $A\in \C^{\ncn}$ to be \emph{asymptotically stable} (AS) if all eigenvalues of $A$ have positive real parts.
+We call a matrix $A\in \C^{\ncn}$ to be \emph{symmetric positive definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
 \end{definition}
 \begin{definition}\label{distpd}
-Define a distribution $P\in \P$ to be \emph{positive definite} (PD)  if $A_P$ is positive definite.
-Define a distribution $P\in \P$ to be \emph{asymptotically stable} (AS) if $A_P$ is asymptotically stable.
+We call a distribution $P$ to be PD/AS/SPD if $A_P$ is PD/AS/SPD.
 \end{definition}
@@
 \section{Problem Setup}
-We consider the linear stochastic approximation scheme with constant step size and PR-averaging given as below:
+We consider the linear stochastic approximation (LSA) algorithm with constant step size and PR-averaging of the iterates.
+\begin{subequations}\label{eq:lsa}
 \begin{align}
-\label{conststep} \theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
-\label{iteravg} \thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_s,
+\label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
+\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_i,
 \end{align}
-where $\thh_t\in \C^{n}$ in \eqref{iteravg} is PR-average of the iterates $\theta_t\in \C^{n}$ of the constant step size stochastic recursion in \eqref{conststep}.
-For the quantities in \eqref{conststep,iteravg} we assume the following:
-\begin{assumption}\label{lsadist}
-$\{A_t,\, t\geq 0\}$ is an $i.i.d$ sequence of $\ncn$ matrices distributed according to $P\in \P$ and $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^n$ with mean $\E[b_t]=b$.
-\end{assumption}
-From \cref{lsadist} it follows that $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ are martingale difference sequences with respect to the $\sigma$-field given by $\F_{t-1}=\{\theta_0,M_1,\ldots, M_{t-1}\}$.
-Further, we assume that
-\begin{assumption}\label{matvar}
+\end{subequations}
+where $\thh_t, \theta_t \in \C^{n}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\ncn}$ incrementally, in discrete time steps $t=1,2,\dots$
+based on data $b_t\in \C^n$, $A_t\in \C^{n\times n}$\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^n$ and $A_t\in \R^{\ncn}, b_t\in \R^n$.
+However, it serves well for our analysis to make the general assumption that these qunatites are complex valued.}.
+Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
+initial value $\theta_0$.
+The iterate $\theta_t$ is treated as an internal state of the algorithm, while $\th_t$ is the output at time step $t$.
+The update of $\theta_t$ alone is considered a form of constant stepsize LSA.
+Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(n)$ storage.
+
+Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$
+can be efficiently computed in $O(n)$ time.
+This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{} and linear prediction \cite{} will be presented in a later sections (\Cref{sec:rl,sec:lp}).
+The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
+can be implemented in $O(n)$ time and $O(n)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.
+We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa} In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.
+\begin{assumption}\label{assmp:lsa}
+\begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\item \label{dist} $\{A_t\in\C^{\ncn},\, t\geq 0\}$ is an $i.i.d$ sequence distributed according to $P$ which is either AS or PD or SPD.
+\item $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^n$ with mean $\E[b_t]=b$.
+\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ associated with $A_t$ and $b_t$ satisfy the following
 %\begin{align*}\E\left[ M_t^\dag M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
 \begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
+\item $A_P$ is invertible and there exists a $\ts=A_P^{-1}b$.
+\end{enumerate}
 \end{assumption}
-
-Our objective is to analyse $\EE{\norm{\thh_t-\theta^*}^2}$.
-
-\begin{example}[TD(0)]
-\end{example}
-
-\begin{example}[Linear Prediction]
-\end{example}
-
-%\section{Error Recursion and Similarity Transformation}
-
+\textbf{Performance Metric:}
+The ``aim'' of the updates in \eqref{eq:lsa} is to estimate $\ts$ based on the noisy versions $(A_t,b_t)$ of $(A_P,b)$.
+There are various ways of quantifying the error $\eh_t \doteq \thh_t- \ts$ of estimating $\ts$ using $\thh_t$, but the magnitude
+of the errors is usually measured in some weighted quadratic norm. The expected loss for time step $t$ then becomes
+\begin{align*}
+L_t \doteq \EE{ \norm{\eh_t}_{C}^2 }\,,
+\end{align*}
+where recall that for a SPD matrix $C$, $\norm{x}_C^2 = x^\dag C x$. Here, the choice of the matrix $C$ depends on the application.\\
+\textbf{Challenges:}
+The idea of constant step size and iterate averaging is not new and goes back to \cite{}, where it was first shown that $L_t\ra 0$ as $t\ra\infty$.
+Recently \cite{bach} revisited the idea in connection to the stochastic gradient descent (SGD) algorithm (which is also an LSA of the form in \eqref{eq:lsa}) to solve linear regression.
+Much similar to the spirit of \cite{bach}, our aim here is to understand the aspects that effect the behaviour of $L_t$, such as:
+\begin{itemize}
+\item The variances $\sigma_P^2$ and $\sigma_b^2$ of the noise terms $M_t$ and $b_t$ in the update equation.
+\item The baias $\norm{\theta_0-\ts}$ due to the initial condition $\theta_0$.
+\item The effect of the structural properties of the $A_P$, i.e., whether it is AS or PD or SPD.
+\end{itemize}
+Most resutls in the past including those that of \cite{bach} consider an SGD algorithm for quadratic loss minimization, where the matrix $A_P$ is SPD and the noise has additional strucutral properties (we discuss this briefly in \Cref{sec:lp}).
+As it turns out these nice properties are lost when $A_P$ is AS or PD (a property that is holds in RL settings).
+While from the standpoint of the tools and techniques used, we use techniques much similar to \cite{bach}, however, as mentioned, the fact that we no longer are dealing with gradient systems (that involve to SPD matrices) but general vector fields (that involve AS or PD matrices), poses important challenges.
+Many RL algorithms \cite{} are LSA algorithms that need only $O(n)$ updates.
+However, there is a lot of variability in their analyses.
+The analysis as well as the results in this paper are useful from the perspective that it presents a single unified way of understand the various LSA-RL algorithms.
+We now present and discuss the main results of this paper in the next section.
+The comparison with existing literature is done in two separate sections, one to compare our results with \cite{bach} in the linear predictions setting (\Cref{sec:lp}) and the other to compare our results with prior results in RL literature \cite{}.
+\todoch{Mention impotance of forgetting the bias etc. Mention that we really don't have control over $\norm{U}$}
 \section{Main Results}
 \begin{lemma}
-Let $P$ be a positive definite distribution over $\C_B^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+Let $P$ be a positive definite distribution over $\C^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
 \end{lemma}
 \begin{theorem}
-Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\ncn}$. Then
+Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C^{\ncn}$. Then
 \begin{align}
@@
 \begin{align}
-\EE{\norm{\gamma_t-\gamma^*}^2}
+\EE{\norm{\gh_t-\gamma^*}^2}
 \leq
 \left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
-\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
+\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
+%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
 %\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
@@
 \begin{definition}
-Call a set of distributions $\P$ over $\C_B^{\ncn}$
+Call a set of distributions $\P$ over $\C^{\ncn}$
 \emph{weakly admissible} if there exists $\alpha>0$ such that
@@
 \begin{definition}
-Call a set of distributions $\P$ over $\C_B^{\ncn}$ \emph{admissible}
+Call a set of distributions $\P$ over $\C^{\ncn}$ \emph{admissible}
 if there exists some $\alpha>0$ such that $\inf_{P\in \P} \rhos{P}>0$.
@@
 A hopeful set is the following:
-For $B>0$, let $\P_B$ is the set of distributions where for every $P\in \P_B$, $H_P+H_P^\top\succ 0$
-and if $H\sim P$ then $\norm{H}\le B$ holds with probability one.
+For $B>0$, let $\P_B$ be the set of distributions over $\C_B^{\ncn}$ where for every $P\in \P_B$ is positive definite.
 In what follows all distributions will be a subset of $\P_B$ with some $B>0$,
@@
 Our first result shows that we cannot be too ambitious:
-\begin{proposition}
+\begin{proposition}\label{notwad}
 The set $\P_1$ is not weakly admissible.
@@
 For $\epsilon \in (0,1/2)$ let $P$ be the distribution that is supported on $\{-I,I\}$ and takes on the value of $I$ with probability $1/2+\epsilon$.
-Then $H_P = 2\epsilon I \succ 0$, hence $P\in \P_1$.
+Then $A_P = 2\epsilon I \succ 0$, hence $P\in \P_1$.
 Further, $Q_P = I$.
-Hence, $\rho_\alpha(P) = 4\epsilon-\alpha$. Hence, if $\epsilon<\alpha/4$, $\rho_\alpha(P)<0$.
+Hence, $\rhos{P} = 4\epsilon-\alpha$. Hence, if $\epsilon<\alpha/4$, $\rho_\alpha(P)<0$.
 \end{proof}
-Note that the result still holds if in the definition of $\P_B$ we replace the condition that $H_P + H_P^\top \succ 0$ with the
-stronger condition that $H_P\succ 0$. In words, positive definiteness of $H_P$ alone is not sufficient for weak admissibility.
-
+Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
 In linear least-squares problems (cf. \cref{ex:leastsquares}) the set $\P$ has more structure.
-In particular, the random matrices $\{H_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
+In particular, the random matrices $\{A_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
 Define
@@
 }
+\subsection{Result Discussion}
+
+For more details see the discussion on related work section
+
+\section{Special Cases}
+\subsection{Additive Noise Case}
+
+\subsection{Symmetric Matrix}
+
+\subsection{Linear Prediction With Structured Noise}
+
+\section{Reinforcement Learning}
+
+\section{Stochastic Approximation and Mirror Prox}
 
@@
 \begin{lemma}\label{multheorem}
-Let $\P$ be the space of positive definite distributions supported on $\C_B^{\ncn}$.
+Let $\P$ be the space of positive definite distributions supported on $\C^{\ncn}$.
 Then there exists an $\alpha_{\max}>0$ such that $\alpha\in(0,\alpha_{\max})$ implies $\us{\sup}{\norm{x}\leq 1,P\in\P,t\geq 0}\E_{P}[\norm{(I-\alpha H)^tx}]< \infty$.
@@
 \end{corollary}
-
 \section{Main Results}
@@
 \end{align}
-
 \end{theorem}

### 2017-05-01 flsa.tex
splitting-files

--- 2017-04-23 flsa.tex
@@
 \section{Introduction}
-\section{Notation}
 
-\section{Problem Setup}
-We consider linear stochastic approximation schemes with constant step size and PR-averaging given as below:
-\begin{align}
-\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
-\thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_s
-\end{align}
-where we assume
-\begin{assumption}\label{lsa}
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item $A_t=A+M_t$ and $b_t=b+N_t$, where $\{M_t \in \R^{\ncn},\,t\geq 0\}$ and $\{N_t\in\R^n,\,t\geq 0\}$ are $i.i.d$ sequences such that $\E[M_t]=0$ and $\E[N_t]=0$.
-\item There exists a $\ts\in \R^n$ such that $\ts=A^{-1}b$.
-\item All the eigenvalues of $A$ have strictly positive real parts.
-\item $\E[M_t^\top M_t\ts]=\sigma_1^2$ and $\E[\norm{N_t}^2]=\sigma_2^2$.
-\end{itemize}
-\end{assumption}
-\Cref{lsa}
-\begin{example}[Linear Least Squares]
-\end{example}
+\input{not}
+\input{prob}
+\input{instance}
+\input{uniform}
 
-\begin{example}[TD(0)]
-\end{example}
+\section{Special Cases}
+\subsection{Additive Noise Case}
 
-\section{Notation}
-Let $\R^{\ncn}$ and $\C^{\ncn}$ denote the set of $n\times n$ real and complex matrices respectively. Let $\gln\subset\C^{\ncn}$ denote the set of all invertible $\ncn$ complex matrices.
+\subsection{Symmetric Matrix}
 
-\section{Definitions}
-\begin{definition}[Asymptotically Stable]
-An $\ncn$ matrix said to be \emph{asymptotically stable} iff all its eigenvalues have positive real parts.
-\end{definition}
+\subsection{Linear Prediction With Structured Noise}
 
-\begin{definition}\label{def:dist}[Distributions]
-Given a distribution $P$ over $\C^{\ncn}$ and let
-\begin{align*}
-A_P=\int M dP(M),\quad C_P=\int M^\dag M dP(M),
-\end{align*}
-and let $\P$ denote the set of probability distributions on $\C^{\ncn}$.
-\end{definition}
+\section{Reinforcement Learning}
 
-\begin{definition}[Similarity Transformation]
-Given any $U\in \gln$, define $\T_U \colon \C^{\ncn}\rightarrow \C^{\ncn}$ be the linear map that takes $A\in \C^{\ncn}$ to $\T_U(A)=U^{-1}\,A\,U$.
-\end{definition}
-\begin{definition}[Transformed Measure]
-Given a distribution $P$ over $\C^{\ncn}$ define the distribution $P_U$ induced by the linear transformation $\T_U$ as follows:
-\begin{align*}
-\int_{S} dP_U(M) \eqdef \int_{\T_U^{-1}(S)} dP(M)
-\end{align*}
-\end{definition}
-
-\begin{comment}
-\begin{definition}[Induced Distribution]
-Let $P\in \P$ be any distribution over $\R^{\ncn}$ and let $p_c^U$ denote the probability distribution over $\C^{\ncn}$ induced by the linear map $\T_U$ such that the pullback of $p_c^U$ using $\T_U$ is $p$.
-\end{definition}
-\end{comment}
-
-\begin{definition}\label{def:spectnorm}[Spectral Norms]
-Define $\rhod{P}\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{\bar{x},\left((A_P+A_P^\dag)-\alpha A_P^\dag A_P\right)x}$, $\rhos{P}\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{\bar{x},\left((A_P+A_P^\dag)-\alpha C_P\right)x}$
-\end{definition}
-
-\begin{definition}[Weakly Admissible]
-A set of distributions $\P$ over $\C^{\ncn}$ is said to be \emph{weakly admissible} if for any given $P\in \P,\,\exists$ a matrix $U_P \in \gln$, real-valued constants $\alphapd>0$ and $\alphaps>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alphapd)$ and $\rhos{P_U}>0,\,\forall \alpha\in(0,\alphaps)$.
-\end{definition}
-\begin{definition}
-Given a weakly admissible class $\P$ and any $P\in\P$, $U\in \gln$ and $\alpha>0$ such that $\rhos{P_{U}}>0$ define
-\begin{align*}
-\bu&\eqdef \norm{U}^2\norm{U^{-1}}^2 \left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{1}{\alpha \rhos{P_U}}\\
-\beta(P)&\eqdef \inf_{\alpha,U\colon \bu>0} \bu
-\end{align*}
-\end{definition}
-
-\begin{definition}[Admissible]
-A set of distributions $\P$ over $\C^{\ncn}$ is \emph{admissible} if  for any given $P\in \P,\,\exists$ a matrix $U_P\in \gln,\,$ and a real-valued constant $\alpha_0>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alpha_0)$ and $\rhos{P_{U_P}},\,\forall \alpha\in(0,\alpha_0)$.
-\end{definition}
-\begin{example}
-\end{example}
-\begin{definition}[Strongly Admissible]
-A set of distributions $\P$ over $\C^{\ncn}$ is \emph{strongly admissible} if  $\forall\,P\in \P,\,\exists U_P\in \gln,\, \alpha_0>0$ such that $\rhod{P_{U_P}}>0,\,\forall \alpha\in(0,\alpha_0]$ and $\rhos{P_{U_P}},\,\forall \alpha\in(0,\alpha_0]$, $\inf_{P\in \P} \rho_d(\alpha_0,P_{U_P})>0$ and $\inf_{P\in \P} \rho_s(\alpha_0,P_{U_P})>0$.
-\end{definition}
-\begin{example}
-\end{example}
-\begin{comment}
-\begin{lemma}\label{pdt}
-Given a AS matrix $A$, there exists a non-singular matrix $U\in \C^{\ncn}$ such that $A=U\Lambda U^{-1}$ where $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
-\end{lemma}
-Note that in \cref{pdt} the transformation $U\Lambda U^{-1}$ is not unique.
-\begin{example}
-\end{example}
-\end{comment}
-\begin{comment}
-\begin{lemma}
-Given any AS matrix $A$, there exists $\alpha_D>0,\,\alpha_R>0$ and a non-singluar matrix $U$ such that
-\begin{enumerate}
-\item $\rhod{\Lambda_1}>0,\,\forall \alpha \in (0,\alpha_D)$.
-\item $\rhos{\Lambda_1}>0,\,\forall \alpha \in (0,\alpha_R)$.
-\end{enumerate}
-\end{lemma}
-\begin{corollary}
-There exists $\beta\eqdef\inf_{U : det(U)\neq 0,\alpha\in (0,\alpha_R)} \bu$.
-\end{corollary}
-\end{comment}
-\section{Main Results}
-The following theorem states that weak admissibility implies stabilizability.
-\begin{theorem}
-Let $\P_{as}$ be a set of distributions supported on $\R^{\ncn}$ such that for any given $P\in \P$, all the eigenvalues of $A_{P}$ have positive real parts. $\P_{as}$ is \emph{weakly admissible}.
-\end{theorem}
-In short, so long as the matrices $A_t$ are distributed such that all the eigenvalues of $\E[A_t]$ have with positive real parts, then there always exists a positive step size $\alpha$ (albeit problem dependent) that results in \eqref{lsa} to converge in the mean-squared sense.
-
-In \Cref{thm:rate} that we state we below, we let $P^A$ denote the underlying distribution that generates $A_t, t\geq 0$, and in order to be consistent with \Cref{lsa} and \Cref{def:dist} it is understood that $A_{P^A}=A$.
-\begin{theorem}\label{thm:rate}
-\begin{align}
-\E[\norm{\zb_t}^2]
-\leq \beta(P^A)
-\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{e_0})}{t+1} \right)\,.
-\end{align}
-
-\end{theorem}
-
-\section{Comparison With Prior Works}
-
-\subsection{Stochastic Approximation}
-
-\section{Interesting Special Cases}
-
-\section{Application in Reinforcement Learning}
-\subsection{RL Setting}
-
-\subsection{TD(0) Algorithm}
-
-\subsection{GTD}
-\textbf{Flaws with GTD-MP:}
-
-\textbf{Why is GTD slow:}
-
-\subsection{iLSTD}
-
-\subsection{Comparison of TD, GTD and iLSTD}
-
-\section{Conclusion}
-
-\appendix
-
-\input{linal}
+\section{Stochastic Approximation and Mirror Prox}
 

### 2017-05-01 instance.tex
splitting-files

\section{Instance Dependent Bounds}
In this section, we derive instance dependent bounds for $L_t$ in \eqref{eq:exploss} which is valid for a fixed $P$ and in the next section, explore the question of deriving uniform bounds for all $P\in \P$ (where $\P$ is a class of distributions over $\C^{\ncn}$). We first present the results and then follow it up with a discussion.
\begin{lemma}\label{lm:pd}
Let $P$ be a positive definite distribution over $\C^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{theorem}\label{th:pdrate}
Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C^{\ncn}$. Then
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}

\end{theorem}

\begin{lemma}\label{lm:simtran}[Similarity Transformation]
Let $P^{AS}\in\P$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
\end{lemma}

\begin{theorem}\label{thm:simtran}[Change of Basis]
Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
\begin{align}
\EE{\norm{\gh_t-\gamma^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
\end{align}
where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
\end{theorem}
\begin{corollary}\label{cor:as}
Let $P$ in \Cref{lsadist} be asymptotically stable. Then, there exists
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U}^2\norm{U^{-1}}^2}{\alpha \rhos{P_U}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}
\end{corollary}
\subsection{Results Discussion}
\textbf{Error Recursion and Contraction Maps:}
The output $\thh_t$ of the algorithm \eqref{eq:lsa} is the average of the internal states at times $s=0,\ldots,t-1$. Hence, before we proceed to interpret the results in \Cref{lm:pd,thm:pdrate,lm:simtran,thm:simtran,cor:as}, it is worthwhile to spend some effort in trying to understand the dynamics of the internal states by looking at the behavior of $e_t\eqdef \theta_t-\ts$.
\begin{align}\label{eq:errec}
\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1})\nn\\
\theta_t-\ts&=\theta_{t-1}-\ts+\alpha\left(b_t-A_t(\theta_{t-1}-\ts+\ts)\right)\nn\\
e_t&=e_{t-1}+\alpha(b_t-A_t e_t -A_t\ts)\nn\\
e_t&=\underbrace{((I-\alpha A_t))}_{\text{Random-Matrix}} e_{t-1}+\underbrace{\alpha(N_t -(M_t)\ts)}_{\text{Noise}}
\end{align}
From \eqref{errec} it is clear that the error dynamics depends on the $i)$ properties of the random matrix, $ii)$ properties of the noise. In the absence of the noise term in \eqref{eq:errec}, we have $e_t=(I-\alpha A_t) e_{t-1}=\Pi_{s=1}(I-\alpha A_s) e_0$, i.e., the inital error is get multiplied by a product of random matrices. In such a scenario, we can guess that whether or not the algorithm forgets the bias $\norm{\theta_0-\ts}$, depends on whether the matrix product is contracting. One way to characterize this contracting property is to look at the spectral radius of the random matrix, which is given by
\begin{align}\label{eq:spectralrand}
\EE{\norm{I-\alpha A_t}^2}=1-\alpha \rhos{P}, {\norm{I-\alpha A_P}^2}=1-\alpha \rhod{P}
\end{align}
Notice that \eqref{eq:err} places our definitions of $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract}. Here the subscripts $s$ and $d$ stand for \emph{stochastic} and \emph{deterministic}, since they are associated with the spectral radius of stochastic and deterministic matrices respectively.\\
\textbf{Positive Definiteness:} The first result in \Cref{lm:pd} shows if $P$ is PD, then there exists a stepsize for which $\rhos{P}, \rhod{P}>0$ which in turn implies that the spectral radii in \eqref{eq:spectralrand} are less than $1$.
\textbf{Similarity Transformation:} 
\paragraph{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms. The first term is the bias term given by $\B=(1+4(\alpha\rho_{\alpha})^{-1}) (\alpha\rho_{\alpha})^{-1} \Big(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}\Big)$.  The \emph{Bias} term that captures the rate at which the initial condition $\norm{\theta_t-\ts}^2$ is forgotten. The second term is the \emph{Variance} given by $\V=(1+4(\alpha\rho_{\alpha})^{-1}) (\alpha\rho_{\alpha})^{-1} \Big(\frac{\alpha^2\sigma^2}{t+1} \Big)$. The variance term that captures the rate at which noise is rejected. The $\B$ and $\V$ terms capture two different sources of errors. To see this, note that in the absence of noise i.e., when $\zeta_t=0,~\forall t\geq 0$, we have $\E[\norm{H\eb_t}^2]=\B$ and in the presence of noise, i.e., $\zeta\neq 0,\forall t\geq 0$ and the perfect initial condition, i.e., $\theta_t=\ts$, we have $\E[\norm{H\eb_t}^2]=\V$.
\paragraph{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term. This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower. However, in this case the noise term does not blow up, a fact that can appreciated by looking at \eqref{lsergen} where $\alpha$ is seen to multiply the noise term $\zeta_t$. In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other. For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{\max}$, the bounds blow up again, due to the fact that $\rho_{\alpha}\ra 0$ in this case. This is due to the fact that the effect of both noise and initial conditions decay with the contraction factor of $F_{t,i}$, which gets closer to unity as $\alpha\ra\alpha_{\max}$.

### 2017-05-01 not.tex
splitting-files

\section{Notations and Definitions}\label{sec:def}

We denote the sets real and complex number by $\R$ and $\C$ respectively. For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively. We denote $n$-dimensional vector spaces over $\R$ and $\C$ by $\R^{n}$ and $\C^{n}$ respectively. We use $\R^{\ncn}$ and $\C^{\ncn}$ to denote $\ncn$ matrices with real and complex entries respectively. For any $x\in \C$ we denote the modulus of $x$ by $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$. For any $x\in \C^n$We denote complex conjugate of $x$ by $\bar{x}$. We denote the transpose of $x\in \C^{n}, A\in \C^{\ncn}$ by $x^\top$ and $A^\top$ respectively. Further, by $x^\dag$ denotes the conjugate transpose, i.e., $x^\dag=\bar{x}^\top$ (the same convention carries over to matrices as well). We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^\dag y$. Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^n:\norm{x}=1} \norm{Ax}$. For a positive real number $B>0$, we denote by $\C^{\ncn}_B=\{A\in \C^{\ncn}\mid \norm{A}\leq B\}$ and  $\R^{\ncn}_B=\{A\in \R^{\ncn}\mid \norm{A}\leq B,\,\forall i,j=1,\ldots,n\}$ the set of complex and real matrices whose norms are bounded $B$.
%We denote by $\C^{\ncn}_B=\{A\in \C^{\ncn}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\ncn}_B=\{A\in \R^{\ncn}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
%We use $\P$ to denote the set of probability distribution over $\C^{\ncn}$.
We also denote the set of invertible $\ncn$ complex matrices by $\gln$. We use $A\sim P$ to denote the fact that the random variable $A$ is distributed according to distribution $P$. Let the random matrix $A\sim P$, and $U\in \gln$, we use $P_U$ to denote the distribution of the random matrix $U^{-1}A U$.  We denote the identity matrix in $\C^{\ncn}$ by $\I$.
%We denote the set of distributions over $\C_B^{\ncn}$ by $\P$.

We use $A\succeq 0$ to denote that the
square matrix $A$ is hermitian and positive semidefinite (HPD):
$A = A^\dag$, $\inf_x x^\dag A x\ge 0$.
For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
We also use $\preceq$ and $\prec$ analogously. For any $x\in\C^n$x, we denote the general quadratic norm with respect to a $C\succ 0$ by $\norm{x}^2_C\eqdef x^\dag\, C \,x$.
$\E$ denotes mathematical expectation.

\begin{definition}
Let $P$ be a distribution over $\C^{\ncn}$, then define
\begin{align*}
A_P=\int M dP(M),\quad C_P=\int M^\dag M dP(M),
\end{align*}
\end{definition}
\begin{definition}\label{def:contract}[Contraction Factors]
Define
\begin{align*}
\rhod{P}&\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha A_P^\dag A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}
\end{align*}
\end{definition}
Note that $\rhod{P}>\rhos{P}$.

\begin{definition}
We call a matrix $A\in \C^{\ncn}$ to be \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{n}$. We call a matix $A\in \C^{\ncn}$ to be \emph{asymptotically stable} (AS) if all eigenvalues of $A$ have positive real parts. We call a matrix $A\in \C^{\ncn}$ to be \emph{symmetric positive definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
\end{definition}
\begin{definition}\label{distpd}
We call a distribution $P$ to be PD/AS/SPD if $A_P$ is PD/AS/SPD.
\end{definition}

### 2017-05-01 prob.tex
splitting-files

\section{Problem Setup}
We consider the linear stochastic approximation (LSA) algorithm with constant step size and PR-averaging of the iterates.
\begin{subequations}\label{eq:lsa}
\begin{align}
\label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{s=0}^{t}\theta_i,
\end{align}
\end{subequations}
where $\thh_t, \theta_t \in \C^{n}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\ncn}$ incrementally, in discrete time steps $t=1,2,\dots$
based on data $b_t\in \C^n$, $A_t\in \C^{n\times n}$\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^n$ and $A_t\in \R^{\ncn}, b_t\in \R^n$. However, it serves well for our analysis to make the general assumption that these qunatites are complex valued.}. Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
initial value $\theta_0$. The iterate $\theta_t$ is treated as an internal state of the algorithm, while $\th_t$ is the output at time step $t$. The update of $\theta_t$ alone is considered a form of constant stepsize LSA. Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(n)$ storage.

Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$
can be efficiently computed in $O(n)$ time.
This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{} and linear prediction \cite{} will be presented in a later sections (\Cref{sec:rl,sec:lp}). The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
can be implemented in $O(n)$ time and $O(n)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.
We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa} In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.
\begin{assumption}\label{assmp:lsa}
\begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item \label{dist} $\{A_t\in\C^{\ncn},\, t\geq 0\}$ is an $i.i.d$ sequence distributed according to $P$ which is either AS or PD or SPD.
\item $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^n$ with mean $\E[b_t]=b$.
\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ associated with $A_t$ and $b_t$ satisfy the following
%\begin{align*}\E\left[ M_t^\dag M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
\item $A_P$ is invertible and there exists a $\ts=A_P^{-1}b$.
\end{enumerate}
\end{assumption}
\textbf{Performance Metric:}
The ``aim'' of the updates in \eqref{eq:lsa} is to estimate $\ts$ based on the noisy versions $(A_t,b_t)$ of $(A_P,b)$.
There are various ways of quantifying the error $\eh_t \doteq \thh_t- \ts$ of estimating $\ts$ using $\thh_t$, but the magnitude
of the errors is usually measured in some weighted quadratic norm. The expected loss for time step $t$ then becomes
\begin{align}\label{eq:exploss}
L_t \doteq \EE{ \norm{\eh_t}_{C}^2 }\,,
\end{align}
where recall that for a SPD matrix $C$, $\norm{x}_C^2 = x^\dag C x$. Here, the choice of the matrix $C$ depends on the application.\\
\textbf{Challenges:}
The idea of constant step size and iterate averaging is not new and goes back to \cite{}, where it was first shown that $L_t\ra 0$ as $t\ra\infty$. Recently \cite{bach} revisited the idea in connection to the stochastic gradient descent (SGD) algorithm (which is also an LSA of the form in \eqref{eq:lsa}) to solve linear regression. Much similar to the spirit of \cite{bach}, our aim here is to understand the aspects that effect the behaviour of $L_t$, such as:
\begin{itemize}
\item The variances $\sigma_P^2$ and $\sigma_b^2$ of the noise terms $M_t$ and $b_t$ in the update equation.
\item The baias $\norm{\theta_0-\ts}$ due to the initial condition $\theta_0$.
\item The effect of the structural properties of the $A_P$, i.e., whether it is AS or PD or SPD.
\end{itemize}
Most resutls in the past including those that of \cite{bach} consider an SGD algorithm for quadratic loss minimization, where the matrix $A_P$ is SPD and the noise has additional strucutral properties (we discuss this briefly in \Cref{sec:lp}). As it turns out these nice properties are lost when $A_P$ is AS or PD (a property that is holds in RL settings). While from the standpoint of the tools and techniques used, we use techniques much similar to \cite{bach}, however, as mentioned, the fact that we no longer are dealing with gradient systems (that involve to SPD matrices) but general vector fields (that involve AS or PD matrices), poses important challenges.
Many RL algorithms \cite{} are LSA algorithms that need only $O(n)$ updates. However, there is a lot of variability in their analyses. The analysis as well as the results in this paper are useful from the perspective that it presents a single unified way of understand the various LSA-RL algorithms. We now present and discuss the main results of this paper in the next section. The comparison with existing literature is done in two separate sections, one to compare our results with \cite{bach} in the linear predictions setting (\Cref{sec:lp}) and the other to compare our results with prior results in RL literature \cite{}.
\todoch{Mention impotance of forgetting the bias etc. Mention that we really don't have control over $\norm{U}$}

### 2017-05-01 uniform.tex
splitting-files

\section{Uniform stepsizes}
\begin{definition}
Call a set of distributions $\P$ over $\C^{\ncn}$
\emph{weakly admissible} if there exists $\alpha>0$ such that
$\rhos{P}>0$ holds for all $P\in \P$.
\end{definition}
\begin{definition}
Call a set of distributions $\P$ over $\C^{\ncn}$ \emph{admissible}
if there exists some $\alpha>0$ such that $\inf_{P\in \P} \rhos{P}>0$.
The value of $\alpha$ is called a witness.
\end{definition}
It is easy to see that $\alpha \mapsto \rhos{P}$ is decreasing,
hence if $\alpha>0$ witnesses that $\P$ is (weakly) admissible
then any $0<\alpha'\le \alpha$ is also witnessing this.

If $\P$ is weakly admissible, then one can choose a stepsize $\alpha>0$ solely based on the knowledge of $\P$ and
conclude that no matter the joint distribution of $(A_t,b_t)$, as long as the distribution of $A_t$ is an element of $\P$, the conclusions of \cref{maintheorem} hold. Further, if $\P$ is admissible then the error bound stated in  \cref{maintheorem} becomes independent of the instance.
Hence, an interesting question to investigate is whether a given set $\P$ is (weakly) admissible.
A hopeful set is the following:
For $B>0$, let $\P_B$ be the set of distributions over $\C_B^{\ncn}$ where for every $P\in \P_B$ is positive definite.
In what follows all distributions will be a subset of $\P_B$ with some $B>0$,
Note that in most applications the knowledge of a suitable $B>0$ such that $\P\subset \P_B$ is available.
By appropriately normalizing the matrices involved, we may as well assume that $B=1$.

Our first result shows that we cannot be too ambitious:
\begin{proposition}\label{notwad}
The set $\P_1$ is not weakly admissible.
\end{proposition}
\begin{proof}
Fix an arbitrary $\alpha>0$. We show that there exists $P\in \P$ such that $\rho_\alpha(P)<0$.
For $\epsilon \in (0,1/2)$ let $P$ be the distribution that is supported on $\{-I,I\}$ and takes on the value of $I$ with probability $1/2+\epsilon$. Then $A_P = 2\epsilon I \succ 0$, hence $P\in \P_1$. Further, $Q_P = I$.
Hence, $\rhos{P} = 4\epsilon-\alpha$. Hence, if $\epsilon<\alpha/4$, $\rho_\alpha(P)<0$.
\end{proof}
Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
In linear least-squares problems (cf. \cref{ex:leastsquares}) the set $\P$ has more structure.
In particular, the random matrices $\{A_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
Define
\begin{align*}
\P_{\text{PSD},B} = \{ P \in \P_B \,:\,  \supp(P)\subset \text{PSD} \}\,.
\end{align*}
We have the following proposition:
\begin{proposition}
For any $B>0$, the set $\P_{\text{PSD},B}$ is weakly admissible
and in particular any $0<\alpha < 2/B$ witnesses this.
\end{proposition}
\begin{proof}
Take any $P\in \P_{\text{PSD}}$ and let $H\sim P$.
Consider the SVD of $H$: $H = U \Lambda U^\top$ where $U$ is orthonormal and $\Lambda$ is diagonal with
nonnegative elements. Note that $\Lambda \preceq B\, \I$ and thus $\Lambda^2 \preceq B \Lambda$.
Then for any $x\in \R^n$, $x^\top H^\top H x = x^\top U \Lambda^2 U^\top x \le B x^\top U \Lambda U^\top x = B x^\top H x$.
Taking expectations we find that $x^\top Q_P x \le B x^\top H_P x$.
Hence, $\rho_\alpha(P) = 2 x^\top H_P x - \alpha x^\top Q_P x \ge (2- \alpha B ) \,x^\top H_P x $.
Thus, for any $\alpha<2/B$, $\rho_\alpha(P)>0$.
\end{proof}
\todoc[inline]{So this altogether is weaker than what Bach founds. They prove strong admissibility.
I am guessing that for this they use that the loss is special and also that the noise is special.
This would be a good place to somehow incorporate this with a (short?) proof.
}

### 2017-05-01 prob.tex
latest

@@
 \item $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^n$ with mean $\E[b_t]=b$.
-\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ associated with $A_t$ and $b_t$ satisfy the following
+q\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ associated with $A_t$ and $b_t$ satisfy the following
 %\begin{align*}\E\left[ M_t^\dag M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}

### 2017-05-01 related.tex
latest

\section{Related Work}
\subsection{Linear Regression}
Let $(x_t,y_t)\in \R^n\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite. Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$. Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$. The stochastic gradient descent scheme with constant step size $\alpha>0$ to minimize $f(\theta)$ can be given as below:
\begin{align}\label{linreg}
\theta_t=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)
\end{align}
Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$. It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}. In this case $A_P$ is SPD.
In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
\begin{align*}
f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
\end{align*}

\subsection{Reinforcement Learning}
\subsection{GTD}
\todoch{Use LiHong Li's GTD results, choice of the basis $U$ etc}

### 2017-05-05 intro.tex
rough intro

\section{Introduction}
Stochastic Approximation (SA) algorithms can be expressed in the general form
\begin{align}\label{eq:sa}
\theta_t=\theta_{t-1}+\alpha_t f(\theta_{t-1},\zeta_t),
\end{align}
where $\theta_t\in \R^d$ are the iterates, and $f(\theta_{t-1},\zeta_t)$ of the function $g(\theta_{t-1})=\EE{f(\theta_{t-1})}_{\zeta_t}$, where the expectation is with respect to some noise process $\zeta_t$. At each time step, the algorithm updates the iterates in increments proportional to a step-size $\alpha_t>0$ and the idea is to eventually converge to $\ts\in \R^d$ which is a desired optimum or a fixed point. Due to their inherent online nature SA algorithms appear widely in machine learning applications \cite{konda,bhatnagar,borkar,gtd,td,bach}, where we have to learn from a noisy data stream.\par
The choice of the step-size $\alpha_t$ needs to trade-off learning rate versus filtering noise, so to speak. In general, one can guess that larger step-size lead to faster learning and at the same time smaller step-sizes are better to handle noise. A popular choice of step-size schedule that is known handle the noise $\alpha_t=O(\frac{1}{t})$. However, since the step size are diminishing, it can hurt the learning rate. An alternative suggestion is to use more aggressive step sizes without minding the noise in the iterates $\theta_t$ in \eqref{eq:sa}, however, introduce additional averaging (known as Polyak-Rupert averaging) to filter the noise. The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$, and an analysis of the PR-averaging technique is known from \cite{}.\par
A host of algorithms in machine learning and in particular, the temporal difference class of algorithms arising in reinforcement learning, turn out to be special cases of \eqref{eq:sa}, namley linear stochastic approximation (LSA) algorithms. LSA algorithms assume the general form
\begin{align}\label{eq:lsaintro}
\theta_t=\theta_{t-1}+\alpha_t (b_t-A_t \theta_{t-1}),
\end{align}
where $b_t\in \R^d$ and $A_t\in \R^{\dcd}$ are data dependent quantities that are preseneted to the learner in a sequential manner. Further, in most instances of LSA $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be computed in $O(n)$, and are desired for this low per iteration computational cost. Recently, \cite{bach} used \eqref{eq:lsaintro} with constant step-size, i.e., $\alpha_t=\alpha,\,\forall t\geq 0, \, \alpha >0$ and PR-averaging for the linear prediction problem with squared loss and $i.i.d$ sampling. A ``remarkable'' result by \cite{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, he expected squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\todoc{Dimension?}
In particular, as opposed to many earlier results available in the literature \todoc{Add citations}, $C$ does \emph{not} depend on the conditioning of the underlying system, while the total runtime of the algorithm scales linearly with both $t$ and $n$.\par
In this paper, we want to understand the performance of constant step size with PR-average of iterates for general class of LSAs (outside of the linear prediction setting considered by \cite{bach}). We are interested in the problem of linear value function approximation in reinforcement learning using temporal difference learning (TD) either from experience replay in a batch setting or solving linear systems using TD-style algorithms.
\todoc{TD-style, or just TD?} \todoc{Later we need to comment on TD with eligibility traces and other variations of TD, e.g., new variants.} We now list our contributions, and its implications to the TD class of algorithms, and the key differences between the techniques employed here in comparison to \cite{bach}. We now list our contributions, the implications of our results, and the challenges presented by the general setting in comparison to the linear prediction setting considered by \cite{bach}.
\paragraph{Contributions}:
\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item We show that in the case of general LSA, there exists instance dependent constant step size choice such that the mean squared error of the PR-average given by $\err_t=\EE{\norm{\thh_t-\ts}^2}$ is at most $O(\frac{1}{t})$. The error $\err_t$ further can be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
\item We also show (via counter example) that not all classes of general LSAs `admit' a constant uniform step size. Further, we also argue that the subclass of LSAs related to linear prediction problem does `admit' a problem independent constant step size.
\end{itemize}
\paragraph{Implications:}
\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item Our results hold under the weak assumption\footnote{When this weak assumption fails to hold, the LSA is unstable and hence the notion of convergence is void.} that all the eigen values of $\EE{A_t}$ have strictly real parts. This assumption holds for all known TD class of RL algorithms.
\item Two timescale LSA algorithms are those that use two set of iterates (namely slower and faster) and two separate step-size schedules. Our results also hold for two timescale LSAs as long as the two different step-sizes are held constant with respect to time. Thus our setting not only holds for TD(0) but also extends to GTD/GTD2 and TDC. This provides a unified view of the various temporal difference learning algorithms.
\item The intuition behind the error analysis can be explained in simple terms when the noise free case is considered (i.e., $A_t=A$ in \eqref{eq:lsaintro}). In this case, the LSA is just a linear time invariant dynamical system where the error $\theta_t-\ts=(I-\alpha A)^t (\theta_0-\ts)$ and $\norm{\theta_t-\ts}=\rho^t\norm{\theta_0-\ts}$ (for some $0<\rho<1$). Now the PR-average is given by $\frac{\norm{\theta_t-\ts}^2}{t^2}=\frac{\sum_{s=0}^{t-1}\rho^s}{t^2} \norm{\theta_0-\ts}^2$. Note that the numerator is just sum of terms of a geomteric series, which is summable as long as the matrix $(I-\alpha A)$ has eigen values within the unit circle.
\item The fact that the setting covers even two timescale LSAs with constant step size, coupled with the intuitive interpretation of the error dynamics together present a simple and unified view of the various temporal difference learning algroithms in literature.
\item One of the major bottlenecks in selecting step-size for general LSA is the lack of problem knoweldge. One way to handle this issue is to choose a diminishing step size schedule such as $\alpha_t=\frac{1}{t}$. However, such a strategy can fail when the underlying matrices are ill-conditioned (see \Cref{sec:stepsizes}). Another popular approach has been to choose $\alpha_t=\frac{c}{c+t}$, where initially the step-sizes are held almost constant and then as $t$ increases they decay at rate $O(\frac{1}{t})$. While this approach can mitigate the ill-conditioning problem, it can at the same time amplify the effect of noise. In this paper, we argue informally that the constant step-size with PR-averaging of iterates can be a promising way to mitigate these issues, by periodically checking for possible blow up of the norm of the iterates and then reducing the step-size based on this information (see \Cref{sec:stepsizes}).
\end{itemize}
\paragraph{Challenges:} The break up of error into bias and variance terms and their respective rates were previously worked out by \cite{bach} in the linear prediction setting, and here we extend the same to the general class of LSAs. Nevertheless, there are certain key differences between the general LSA setting that we consider to the setting considered by \cite{bach}, which are listed as under.
\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item In the linear least-squares problem is considered by \citet{bachharder} and the random matrices involved are known to be symmetric positive definite. This enables \citet{bachharder} to define operators that act between spaces of \emph{symmetric} matrices and carry out the computations making use of such operators. In our case, the matrices are known only to be (laxly) positive definite, i.e., they lack symmetry. \todoc{We should point out after our analysis in which step the symmetry helps \citet{bachharder}. We should also make this somehow clear here if possible.}
Thus, the above approach does not work, and instead we have resort to an analysis that makes use of only the expected norms of the random matrices involved.
\item Another significant difference is that in linear least-squares, the ``noise'' has a favorable ``structure''. Again, due to the lack of symmetry, this favorable structure does not apply. \todoc{Again, after the analysis, we should point out how the structured noise would have helped. I am wondering whether a more graceful degradation should happen? Is this really all or nothing?}
\end{itemize}

### 2017-05-05 intro.tex
updated intro

@@
 \section{Introduction}
+
 Stochastic Approximation (SA) algorithms can be expressed in the general form
@@
 The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$, and an analysis of the PR-averaging technique is known from \cite{}.\par
+
 A host of algorithms in machine learning and in particular, the temporal difference class of algorithms arising in reinforcement learning, turn out to be special cases of \eqref{eq:sa}, namley linear stochastic approximation (LSA) algorithms.
@@
 The error $\err_t$ further can be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
-\item We also show (via counter example) that not all classes of general LSAs `admit' a constant uniform step size.
+\item We also show (via counter example) that not all classes of general LSAs `admit' a constant step-size that is uniform across all the problems in the class.
 Further, we also argue that the subclass of LSAs related to linear prediction problem does `admit' a problem independent constant step size.
+\item The assumptions in our setting is comparable to those in the asymptotic analysis based on ordinary differential equation (ODE) method by \cite{borkarmeyn,borkarbook}.
+\item Two timescale LSA algorithms are those that use two set of iterates (namely slower and faster) and two separate step-size schedules.
+Our results also hold for two timescale LSAs as long as the two different step-sizes are held constant with respect to time.
+Thus our setting not only holds for TD(0) but also extends to GTD/GTD2 and TDC.
+\item On an intuitive level, the error dynamics of the constant step-size plus PR-averaging procedure is similar to the behavior of partial sums of a geometric series.
+This coupled with the fact that the result also encompasses two timescale LSAs presents a simple and unified view of the various TD algorithms.
+\item The break up of error into bias and variance terms and their respective rates were previously worked out by \cite{bach} in the linear prediction setting, and here we extend the same to the general class of LSAs.
+Nevertheless, there are certain key differences between the general LSA setting that we consider to the setting considered by \cite{bach}, which are listed as under.
 \end{itemize}
+\begin{comment}
 \paragraph{Implications:}
@@
 \end{itemize}
+\end{comment}

### 2017-05-06 instance.tex
initial version

@@
 \end{lemma}
+\begin{proof}
+\begin{align*}
+\rhos{P}&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^*\EE{A_t^* A_t} x\\
+&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x
+&\geq \lambda_{\min}(A^*+A)-\alpha \norm{A_P}^2-\sigma^2_P
+\end{align*}
+The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*+A)}{\norm{A_P}^2+\sigma^2_P}$
+\end{proof}
+
+\begin{lemma}\label{lm:simtran}[Similarity Transformation]
+Let $P^{AS}$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
+\end{lemma}
+
 \begin{theorem}\label{th:pdrate}
-Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C^{\ncn}$. Then
+Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\ncn}$. Then
 \begin{align}
@@
 
-\begin{lemma}\label{lm:simtran}[Similarity Transformation]
-Let $P^{AS}\in\P$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
-\end{lemma}
-
 \begin{theorem}\label{thm:simtran}[Change of Basis]
-Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
+Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be AS. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
 \begin{align}
@@
 \end{corollary}
-\subsection{Results Discussion}
-\textbf{Error Recursion and Contraction Maps:}
+\subsection{Proof Ingredients}
+\paragraph{Error Recursion and Contraction Maps:}
 The output $\thh_t$ of the algorithm \eqref{eq:lsa} is the average of the internal states at times $s=0,\ldots,t-1$.
@@
 \end{align}
-Notice that \eqref{eq:err} places our definitions of $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract}.
+Notice that \eqref{eq:spectralrand} places our definitions of $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract}.
 Here the subscripts $s$ and $d$ stand for \emph{stochastic} and \emph{deterministic}, since they are associated with the spectral radius of stochastic and deterministic matrices respectively.\\
-\textbf{Positive Definiteness:} The first result in \Cref{lm:pd} shows if $P$ is PD, then there exists a stepsize for which $\rhos{P}, \rhod{P}>0$ which in turn implies that the spectral radii in \eqref{eq:spectralrand} are less than $1$.
-\textbf{Similarity Transformation:} 
+\paragraph{Positive Definiteness:} By expanding \eqref{eq:spectralrand} we get
+\begin{align*}
+\EE{\norm{I-\alpha A_t}^2}=\sup_{x:\norm{x}=1} 1 -\alpha x^* (A_P+A_P^*)x +\alpha^2 \EE{A_t^* A_t},
+\end{align*}
+and thanks to \Cref{lm:pd} we know that when $A_P$ is positive definite it is possible to choose an $\alpha$ such that $\EE{\norm{I-\alpha A_t}}<1$.
+\paragraph{Similarity Transformation:} When $A_{P^{AS}}$ is known only to be AS, it is possible that there exists no choice of $\alpha>0$ for which $\rhos{P^{AS}}>0$ and $\rhod{P^{AS}}>0$ (see \Cref{ex:matrices}).
+However, it is potentially possible to re-write the recursion in a different basis via a similarity transformation $\gamma_t=U^{-1 }\theta_t$, and $z_t\eqdef U^{-1} e_t$
+\begin{align}\label{eq:errecsimtran}
+z_t&=\underbrace{(I-\alpha J_t)}_{\text{Transformed-Matrix}} z_{t-1}+\underbrace{\alpha U^{-1}(N_t -(M_t)\ts)}_{\text{Transformed-Noise}},
+\end{align}
+where $J_t\sim P_U$ and thanks to \Cref{lm:simtran} we know that $P_U$ is PD and from \Cref{lm:pd} we know that there exists $\alpha$ such that $\EE{\norm{I-\alpha J_t}}<1$.
+We note here that a similarity transformation does not result in \eqref{eq:errrecsimtran} being a diagonal system.
+However, \Cref{lm:simtran} side steps non-diagonizablility by taking refuge in Jordon decomposition theorem in linear algebra (see Proof of \Cref{lm:simtran} in the Appendix).
+\subsection{Result Discussion}
+We first discuss implications of \Cref{thm:pdrate}.
 \paragraph{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms.

### 2017-05-06 related.tex
initial version

@@
 \subsection{Reinforcement Learning}
-\subsection{GTD}
+
+\paragraph{RL Setting}
+We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^n\times\R^n\times \R$.
+Let $\gamma \in (0,1)$ be a fixed constant, we define $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$ the various temporal difference learning algorithms can be written as in \Cref{table:td}.
+In what follows, $\rho_t>0$ is an \emph{importance} sampling factor whose aim is to correct for mismatch in the (behavior) distribution with which the data was collected and the (target) distribution with respect to which one wants to learn.
+A factor $\rho_t=1,\,\forall t\geq 0$ will mean that no correction is required\footnote{This is known as the \emph{on-policy} case where the behavior is identical to the target.
+The general setting where $\rho_t>0$ is known as \emph{off-policy}.}.
+\begin{comment}
+\begin{table}
+
+\begin{tabular}{|c|c|c|c|c|}\hline
+Quantity& TD(0) & GTD & GTD2 & TDC \\ \hline
+$b_t$& $\phi_t r_t$ & $\begin{bmatrix}\phi_t r_t\\ \mathbf{0}_{n\times 1} \end{bmatrix}$ &$\begin{bmatrix}\phi_t r_t\\ \mathbf{0}_{n\times 1} \end{bmatrix}$ & $\begin{bmatrix}\phi_t r_t\\ \phi_t r_t \end{bmatrix}$
+\end{tabular}
+\end{table}
+\end{comment}
+\begin{table}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|c|c|c|}\hline
+Algorithm& Update &Stepsize &Rate \\ \hline
+TD(0)& $\begin{aligned}\theta_t&=\theta_{t-1}+\alpha_t (b_t -\Delta_t\theta_{t-1})\\ \thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t\end{aligned}$ &$\begin{aligned}\alpha_t&=O(\frac1{t})^\delta,\\ &\delta\in(1/2,1)\end{aligned}$ &$\EE{\norm{\eh_t}}\leq O(\frac{1}{\sqrt{t}})$ for $\delta\ra 1$ \\ \hline
+GTD& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\rho_t b_t -\rho_t \Delta_t \theta_{t-1}- Q_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\rho_t A^\top_t y_{t-1})\end{aligned}$ & $\begin{aligned}&\beta_t=\eta \alpha_t,\sum_{t\geq 0}\alpha_t=\infty, \\ &\sum_{t\geq 0}\alpha^2_t<\infty \end{aligned}$ & $e_t\ra 0$ w.p.1 \\\hline
+TDC& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\rho_t b_t -\rho_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\rho_t b_t-\rho_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$ & $\begin{aligned}&\beta_t,\alpha_t,\frac{\alpha_t}{\beta_t} \ra 0,~\text{as}~t\ra \infty,\\& \sum_{t\geq 0}\alpha_t=\infty, \sum_{t\geq 0}\beta_t=\infty, \\ &\sum_{t\geq 0}\alpha^2_t<\infty, \sum_{t\geq 0}\beta^2_t<\infty \end{aligned}$ & $e_t\ra 0$ w.p.1 \\\hline
+GTD& $\begin{aligned} y_{t}&=\Pi_Y \left(y_{t-1}+\alpha_t(\rho_t b_t -\rho_t \Delta_t \theta_{t-1}-Q_t y_{t-1})\right)\\ \theta_t &=\Pi_{\Theta}\left(\theta_{t-1}+\alpha_t (\rho_t A_t^\top y_{t-1})\right)\\ \thh_t&=\frac{\sum_{s=0}^t \alpha_s\theta_s}{\sum_{s=0}^t \alpha_s}\end{aligned}$ & $\alpha_t =O(\frac{1}{\sqrt{t}})$ & $\norm{A_P\eh_t-b}^2 \leq O(\frac{1}{\sqrt{t}})$ w.h.p\\\hline
+
+\end{tabular}
+}
+\end{table}
+
+1) No bias Variance Split in Korda
+2) Projection, and incorrectness in GTDMP
+3) TDC only SA result
+3) Recent results by LiHongLi (need to say what they say clearly)
+
 \todoch{Use LiHong Li's GTD results, choice of the basis $U$ etc}

### 2017-05-06 stepsizes.tex
initial version

\section{Choice of Step-sizes}
We now argue, in a heuristic manner, as to why constant step-size with PR-averaging might be a principled idea, especially in TD type algorithms. Consider the following LSA with time varying step-sizes $\alpha_t>0$
\begin{align}
\theta_t=\theta_{t-1}+\alpha_t(b_t-A_t\theta_{t-1}).
\end{align}
It is known from asymptotic results by \cite{borkarbook,kushner}\todoch{Check about Kushner} that in order to handle the noise the step-sizes have to be diminishing i.e., $\alpha_t\ra$, unsummable i.e., $\sum_{t \geq 0}\alpha_t=\infty$ and square summable $\sum_{t\geq 0}\alpha_t^2<\infty$. We now look at two specific diminishing step-size sequences, to illustrate the issues that the diminishing step-sizes presents us with. In what follows, we assume that the maximum eigevalue of $\EE{A_t}$ is less that $1$.
\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item \textbf{$\alpha_t=\frac{1}{t}$\footnote{While $\frac{c}{t}$, $c\geq 0$ is more general, it is enough to consider $\frac1t$ by accommodating the multiple $c$ as part of the matrix $A_t$}} is known yield a rate of $\frac{1}{t^{\lambda_{\min}}}$, where $\lambda_{\min}$ is the smallest real part of the eigen values of the matrix $\EE{A_t}$.
\item \textbf{$\alpha_t=\frac{c}{c+t}$} is another example, where the step size decays at a rate $\frac{1}{t/c}$. The rate of decay of error in this case turns out to be $\frac{c^2\sigma_b^2}{t}$, whenever $c>\frac{1}{\lambda_{\min}}$, which can be ensured only when $\lambda_{\min}$ is known.
\end{itemize}
In contrast, in the case of constant step-size with PR-averaging, not knowing the problem instance can result in instability, and when we manage to choose a step-size that results in stable behavior, the rates obtained in \Cref{th:pdrate,cor:as} hold for it. Using this observation, the constant-step size can be set in a more prinicipled manner as showing the pseudocode below.
\begin{algorithm}[H]
\caption*{Automatic Tuning of Constant Step-Size}
\begin{algorithmic}[1]
\STATE{Initialize: $\theta_0$, $\alpha=1$, $i=1$, $a_0=0$}
\FOR{$t=1,2,\ldots, $}
\STATE{$\theta_t=\theta_{t-1}+\alpha (b_t-A_t\theta_{t-1})$}
\IF{$mod(t,k)==0$}
\STATE{$a_{i}=\norm{\theta_t}$}
\IF{$Is-stable(a_0,\ldots,a_i)==0$}
\STATE{$\alpha=\alpha/2$}
\ENDIF
\STATE{$i=i+1$}
\ENDIF
\ENDFOR
\end{algorithmic}
\end{algorithm}

\begin{comment}
In order to calculate the rates we make use of the following asymptotic result by \cite{borkarbook}.
\begin{theorem}\label{th:borkar}
Let $\theta_t=\theta_{t-1}+\alpha_t(b_t-A_t\theta_{t-1})$, where $\theta_t,b_t\in\R^d$ and $A_t\in\R^{\dcd}$ that satisfy the conditions in \Cref{assmp:lsa} and $\sum_{t\geq 0}\alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$. Let $t_{ode}(k)=\eqdef \sum_{i=0}^{k-1}\alpha_n$ and let $\tilde{\theta}(t)$ be the trajectory of the ordinary differential equation $\dot{\theta}(t)=\EE{b_t}-A_P\theta(t)$. Then it follows that
\begin{align*}
\lim_{k\ra\infty}\sup\norm{\tilde{\theta_t(t_{ode}(k))}-\theta_t} \ra 0
\end{align*}
\end{theorem}

For this discussion, let us consider a simple example where $P$ is a distribution with all the mass concentrated at $A_P=\begin{bmatrix}\lambda_{\min} &0\\ 0 &\lambda_{\max}\end{bmatrix}$, $b_t$ is a zero mean $i.i.d$ sequence with variance $\sigma_2_b$.
\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item \textbf{$\alpha_t=\frac{1}{t}$:} In order to ensure stability, we want the eigenvalues of $I-\alpha_t A_P$ to be within unit circle. This restricts the choice of $c$ to be $0<c<\frac{2}{\lambda_{\max}}$. It is straighforward to show that a step-size choice of $\alpha_t=\frac{2}{\lambda_{\max}t}$ will yield a rate of $\frac{1}{t^{2\lambda_{\min}}}$, which can be very bad for small $\lambda_{\min}$.
\item \textbf{}
\end{itemize}
\end{comment}

### 2017-05-06 instance.tex
initial version

@@
 \section{Instance Dependent Bounds}
-In this section, we derive instance dependent bounds for $L_t$ in \eqref{eq:exploss} which is valid for a fixed $P$ and in the next section, explore the question of deriving uniform bounds for all $P\in \P$ (where $\P$ is a class of distributions over $\C^{\ncn}$).
+In this section, we derive instance dependent bounds for $L_t$ in \eqref{eq:exploss} which is valid for a fixed $P$ and in the next section, explore the question of deriving uniform bounds for all $P\in \P$ (where $\P$ is a class of distributions over $\C^{\dcd}$).
 We first present the results and then follow it up with a discussion.
 \begin{lemma}\label{lm:pd}
-Let $P$ be a positive definite distribution over $\C^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+Let $P$ be a positive definite distribution over $\C^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
 \end{lemma}
@@
 \begin{theorem}\label{th:pdrate}
-Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\ncn}$. Then
+Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
 \begin{align}

### 2017-05-06 linal.tex
initial version

@@
 \subsection{Notation}
-Let $B$ be a $\ncn$ block diagonal matrix given by $B=\begin{bmatrix} B_1 &0 &0 &\ldots &0 \\ 0 &B_2 &0 &\ldots &0  \\ \vdots &\vdots &\vdots &\vdots &\vdots \\ 0 &\ldots &0 &0 &B_k \end{bmatrix}$, where $B_i$ is a $n_i \times n_i$ matrix such that $n_i<n$ (w.l.o.g) and $\sum_{i=1}^k n_i=n$.
+Let $B$ be a $\dcd$ block diagonal matrix given by $B=\begin{bmatrix} B_1 &0 &0 &\ldots &0 \\ 0 &B_2 &0 &\ldots &0  \\ \vdots &\vdots &\vdots &\vdots &\vdots \\ 0 &\ldots &0 &0 &B_k \end{bmatrix}$, where $B_i$ is a $n_i \times n_i$ matrix such that $n_i<n$ (w.l.o.g) and $\sum_{i=1}^k n_i=n$.
 We also denote $B$ as
@@
 \begin{lemma}
-Given a matrix $A\in \C^{\ncn} there exists a \emph{characteristic} polynomial denoted by $\chi_A=(\lambda_1 \I-A)^{n_1}\ldots(\lambda_{k} \I-A)^{n_k}$, where $\lambda_i,\,i=1,\ldots,k (k\leq n)$ are the $k$ distinct eigenvalues of the matrix $A$ such that $\sum_{i=1}^k n_i=n$.
+Given a matrix $A\in \C^{\dcd} there exists a \emph{characteristic} polynomial denoted by $\chi_A=(\lambda_1 \I-A)^{n_1}\ldots(\lambda_{k} \I-A)^{n_k}$, where $\lambda_i,\,i=1,\ldots,k (k\leq n)$ are the $k$ distinct eigenvalues of the matrix $A$ such that $\sum_{i=1}^k n_i=n$.
 \end{lemma}
@@
 \begin{lemma}\label{jordon}
-Let $A\in \C^{\ncn}$ and $\{\lambda_i\in \C,i=1,\ldots,k\leq n \}$ denote its $k$ distinct eigenvalues.
-There exists a complex matrix $V\in \C^{\ncn}$ such that $A=V\tL V^{-1}$, where
+Let $A\in \C^{\dcd}$ and $\{\lambda_i\in \C,i=1,\ldots,k\leq n \}$ denote its $k$ distinct eigenvalues.
+There exists a complex matrix $V\in \C^{\dcd}$ such that $A=V\tL V^{-1}$, where
 $\tL=\tL_1\op\ldots\op\tL_k$, where each $\tL_i,\,i=1,\ldots,k$ can further be written as $\tL_i= {\tL}^i_{1}\op \ldots \op {\tL}^i_{{l(i)}}$.
@@
 \begin{lemma}\label{simtran}
-Let $A\in \C^{\ncn}$ be asymptotically stable. There exists a matrix $U\in \gln$ such that $A=U\Lambda U^{-1}$ and $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
+Let $A\in \C^{\dcd}$ be asymptotically stable. There exists a matrix $U\in \gln$ such that $A=U\Lambda U^{-1}$ and $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
 \end{lemma}
 \begin{proof}
-It is trivial to see that for any $\Lambda\in \C^{\ncn}$, $\left(\Lambda^\dag+\Lambda\right)$ is real symmetric.
+It is trivial to see that for any $\Lambda\in \C^{\dcd}$, $\left(\Lambda^\dag+\Lambda\right)$ is real symmetric.
 We will use the decomposition of $A=V \tL V^{-1}$ in \Cref{jordon} and also carry over the notations in \Cref{jordon}.
@@
 Now let $\frac{(\Lambda^\dag+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}$, where $\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}=\begin{bmatrix} \re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 &0\\ \frac{\re{\lambda_i}}{2} &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} \\ 0 &\ldots &0 &0 &\frac{\re{\lambda_i}}{2} &\re{\lambda_i} \end{bmatrix} $.
-Then for any $x=(x(i),i=1,\ldots,n)\in \C^n (\neq \mathbf{0})$, there exists a $b\in \{-1,1\}^n$, such that
+Then for any $x=(x(i),i=1,\ldots,n)\in \C^d (\neq \mathbf{0})$, there exists a $b\in \{-1,1\}^d$, such that
 \begin{align*}
-x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\re{\lambda_i} \left(\sum_{i=1}^n \bar{x}{(i)} x(i)+\sum_{i=1}^{n-1} \frac{\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)}{2}\right) \\
+x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\re{\lambda_i} \left(\sum_{i=1}^d \bar{x}{(i)} x(i)+\sum_{i=1}^{n-1} \frac{\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)}{2}\right) \\
 &=\frac{\re{\lambda_i}}{2}\left(\md{x(1)}^2+ \md{x(n)}^2\right)+\frac{\re{\lambda_i}}{2}\left( \sum_{i=1}^{n-1} \md{x(i)}^2+\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)+\md{x(i+1)}^2 \right)\\
-&>\frac{\re{\lambda_i}}{2}\left(\sum_{i=1}^n \md{x(i)+x(i+1)}^2 \right)\\
+&>\frac{\re{\lambda_i}}{2}\left(\sum_{i=1}^d \md{x(i)+x(i+1)}^2 \right)\\
 &> 0
@@
 \begin{lemma}
-Let $P$ be a positive definite distribution over $\C_B^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+Let $P$ be a positive definite distribution over $\C_B^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
 \end{lemma}
@@
 \begin{align*}
-\rhos{P}&={\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}\\
-&\geq {\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)\right)x}-\alpha B^2\\
+\rhos{P}&={\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}\\
+&\geq {\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)\right)x}-\alpha B^2\\
 &\geq \lambda_{\min}(A_P+A_P^\dag)-\alpha B^2
@@
 \begin{theorem}\label{pdrate}
-Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\ncn}$. Then
+Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\dcd}$. Then
 \begin{align}
@@
 \begin{lemma}[Product unroll lemma]\label{lem:genunroll}
-Let $t>i\ge 1$, $x,y\in \C^n$ be $\F_{i}$-measurable random vectors. Then,
+Let $t>i\ge 1$, $x,y\in \C^d$ be $\F_{i}$-measurable random vectors. Then,
 \begin{align*}
@@
 \begin{lemma}\label{innerproduct}
-Let $i>j \ge 0$ and let $x\in \R^n$ be an $\F_j$-measurable random vector.
+Let $i>j \ge 0$ and let $x\in \R^d$ be an $\F_j$-measurable random vector.
 Then,
@@
 \begin{lemma}[Product unroll lemma]\label{lem:genunroll}
-Let $t>i\ge 1$, $x,y\in \R^n$ be $\F_{i}$-measurable random vectors. Then,
+Let $t>i\ge 1$, $x,y\in \R^d$ be $\F_{i}$-measurable random vectors. Then,
 \begin{align*}

### 2017-05-06 not.tex
initial version

@@
 For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively.
-We denote $n$-dimensional vector spaces over $\R$ and $\C$ by $\R^{n}$ and $\C^{n}$ respectively.
-We use $\R^{\ncn}$ and $\C^{\ncn}$ to denote $\ncn$ matrices with real and complex entries respectively.
+We denote $n$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively.
+We use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively.
 For any $x\in \C$ we denote the modulus of $x$ by $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$.
-For any $x\in \C^n$We denote complex conjugate of $x$ by $\bar{x}$.
-We denote the transpose of $x\in \C^{n}, A\in \C^{\ncn}$ by $x^\top$ and $A^\top$ respectively.
-Further, by $x^\dag$ denotes the conjugate transpose, i.e., $x^\dag=\bar{x}^\top$ (the same convention carries over to matrices as well).
-We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^\dag y$.
+For any $x\in \C^d$We denote complex conjugate of $x$ by $\bar{x}$.
+We denote the transpose of $x\in \C^{d}, A\in \C^{\dcd}$ by $x^\top$ and $A^\top$ respectively.
+Further, by $x^*$ denotes the conjugate transpose, i.e., $x^*=\bar{x}^\top$ (the same convention carries over to matrices as well).
+We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^* y$.
 Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
-The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^n:\norm{x}=1} \norm{Ax}$.
-For a positive real number $B>0$, we denote by $\C^{\ncn}_B=\{A\in \C^{\ncn}\mid \norm{A}\leq B\}$ and  $\R^{\ncn}_B=\{A\in \R^{\ncn}\mid \norm{A}\leq B,\,\forall i,j=1,\ldots,n\}$ the set of complex and real matrices whose norms are bounded $B$.
-%We denote by $\C^{\ncn}_B=\{A\in \C^{\ncn}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\ncn}_B=\{A\in \R^{\ncn}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
-%We use $\P$ to denote the set of probability distribution over $\C^{\ncn}$.
-We also denote the set of invertible $\ncn$ complex matrices by $\gln$.
+The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^d:\norm{x}=1} \norm{Ax}$.
+For a positive real number $B>0$, we denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \norm{A}\leq B,\,\forall i,j=1,\ldots,n\}$ the set of complex and real matrices whose norms are bounded $B$.
+%We denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
+%We use $\P$ to denote the set of probability distribution over $\C^{\dcd}$.
+We also denote the set of invertible $\dcd$ complex matrices by $\gln$.
 We use $A\sim P$ to denote the fact that the random variable $A$ is distributed according to distribution $P$.
 Let the random matrix $A\sim P$, and $U\in \gln$, we use $P_U$ to denote the distribution of the random matrix $U^{-1}A U$.
-We denote the identity matrix in $\C^{\ncn}$ by $\I$.
-%We denote the set of distributions over $\C_B^{\ncn}$ by $\P$.
+We denote the identity matrix in $\C^{\dcd}$ by $\I$.
+%We denote the set of distributions over $\C_B^{\dcd}$ by $\P$.
 
@@
 square matrix $A$ is hermitian and positive semidefinite (HPD):
-$A = A^\dag$, $\inf_x x^\dag A x\ge 0$.
+$A = A^*$, $\inf_x x^* A x\ge 0$.
 For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
 We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
-We also use $\preceq$ and $\prec$ analogously. For any $x\in\C^n$x, we denote the general quadratic norm with respect to a $C\succ 0$ by $\norm{x}^2_C\eqdef x^\dag\, C \,x$.
+We also use $\preceq$ and $\prec$ analogously. For any $x\in\C^d$x, we denote the general quadratic norm with respect to a $C\succ 0$ by $\norm{x}^2_C\eqdef x^*\, C \,x$.
 $\E$ denotes mathematical expectation.
@@
 \begin{definition}
-Let $P$ be a distribution over $\C^{\ncn}$, then define
+Let $P$ be a distribution over $\C^{\dcd}$, then define
 \begin{align*}
-A_P=\int M dP(M),\quad C_P=\int M^\dag M dP(M),
+A_P=\int M dP(M),\quad C_P=\int M^* M dP(M),
 \end{align*}
@@
 \begin{align*}
-\rhod{P}&\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha A_P^\dag A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}
+\rhod{P}&\eqdef {\inf}_{x\in\C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha A_P^* A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha C_P\right)x}
 \end{align*}
@@
 \begin{definition}
-We call a matrix $A\in \C^{\ncn}$ to be \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{n}$.
-We call a matix $A\in \C^{\ncn}$ to be \emph{asymptotically stable} (AS) if all eigenvalues of $A$ have positive real parts.
-We call a matrix $A\in \C^{\ncn}$ to be \emph{symmetric positive definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
+We call a matrix $A\in \C^{\dcd}$ to be \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$.
+We call a matix $A\in \C^{\dcd}$ to be \emph{asymptotically stable} (AS) if all eigenvalues of $A$ have positive real parts.
+We call a matrix $A\in \C^{\dcd}$ to be \emph{symmetric positive definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
 \end{definition}
@@
 \end{definition}
+\begin{example}
+The matrices $\begin{bmatrix}0.1 &-1\\ 1 & 0.1\end{bmatrix}$, $\begin{bmatrix} 0.1 & 0.1 \\ 0 & 0.1\end{bmatrix}$ and $\begin{bmatrix}0.1 &0 \\ 0 & 0.1\end{bmatrix}$ are examples of AS, PD and SPD respectively.
+\end{example}

### 2017-05-06 prob.tex
initial version

@@
 \end{subequations}
-where $\thh_t, \theta_t \in \C^{n}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\ncn}$ incrementally, in discrete time steps $t=1,2,\dots$
-based on data $b_t\in \C^n$, $A_t\in \C^{n\times n}$\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^n$ and $A_t\in \R^{\ncn}, b_t\in \R^n$.
+where $\thh_t, \theta_t \in \C^{d}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\dcd}$ incrementally, in discrete time steps $t=1,2,\dots$
+based on data $b_t\in \C^d$, $A_t\in \C^{n\times n}$\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^d$ and $A_t\in \R^{\dcd}, b_t\in \R^d$.
 However, it serves well for our analysis to make the general assumption that these qunatites are complex valued.}.
@@
 \begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \label{dist} $\{A_t\in\C^{\ncn},\, t\geq 0\}$ is an $i.i.d$ sequence distributed according to $P$ which is either AS or PD or SPD.
-\item $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^n$ with mean $\E[b_t]=b$.
+\item \label{dist} $\{A_t\in\C^{\dcd},\, t\geq 0\}$ is an $i.i.d$ sequence distributed according to $P$ which is either AS or PD or SPD.
+\item $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^d$ with mean $\E[b_t]=b$.
 q\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ associated with $A_t$ and $b_t$ satisfy the following
-%\begin{align*}\E\left[ M_t^\dag M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
-\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
+%\begin{align*}\E\left[ M_t^* M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^* N_t]=\sigma_b^2.\end{align*}
+\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^* N_t]=\sigma_b^2.\end{align*}
 \item $A_P$ is invertible and there exists a $\ts=A_P^{-1}b$.
@@
 \end{align}
-where recall that for a SPD matrix $C$, $\norm{x}_C^2 = x^\dag C x$. Here, the choice of the matrix $C$ depends on the application.\\
+where recall that for a SPD matrix $C$, $\norm{x}_C^2 = x^* C x$. Here, the choice of the matrix $C$ depends on the application.\\
 \textbf{Challenges:}

### 2017-05-06 related.tex
initial version

@@
 \section{Related Work}
+\begin{comment}
 \subsection{Linear Regression}
-Let $(x_t,y_t)\in \R^n\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
+Let $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
 Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$.
@@
 \subsection{Reinforcement Learning}
+\end{comment}
 
 \paragraph{RL Setting}
-We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^n\times\R^n\times \R$.
+We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
 Let $\gamma \in (0,1)$ be a fixed constant, we define $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$ the various temporal difference learning algorithms can be written as in \Cref{table:td}.

### 2017-05-06 uniform.tex
initial version

@@
 \begin{definition}
-Call a set of distributions $\P$ over $\C^{\ncn}$
+Call a set of distributions $\P$ over $\C^{\dcd}$
 \emph{weakly admissible} if there exists $\alpha>0$ such that
@@
 \begin{definition}
-Call a set of distributions $\P$ over $\C^{\ncn}$ \emph{admissible}
+Call a set of distributions $\P$ over $\C^{\dcd}$ \emph{admissible}
 if there exists some $\alpha>0$ such that $\inf_{P\in \P} \rhos{P}>0$.
@@
 A hopeful set is the following:
-For $B>0$, let $\P_B$ be the set of distributions over $\C_B^{\ncn}$ where for every $P\in \P_B$ is positive definite.
+For $B>0$, let $\P_B$ be the set of distributions over $\C_B^{\dcd}$ where for every $P\in \P_B$ is positive definite.
 In what follows all distributions will be a subset of $\P_B$ with some $B>0$,
@@
 nonnegative elements. Note that $\Lambda \preceq B\, \I$ and thus $\Lambda^2 \preceq B \Lambda$.
-Then for any $x\in \R^n$, $x^\top H^\top H x = x^\top U \Lambda^2 U^\top x \le B x^\top U \Lambda U^\top x = B x^\top H x$.
+Then for any $x\in \R^d$, $x^\top H^\top H x = x^\top U \Lambda^2 U^\top x \le B x^\top U \Lambda U^\top x = B x^\top H x$.
 Taking expectations we find that $x^\top Q_P x \le B x^\top H_P x$.

### 2017-05-07 intro.tex
initial version
updated intro

@@
 \section{Introduction}
-
+\begin{comment}
 Stochastic Approximation (SA) algorithms can be expressed in the general form
@@
 The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$, and an analysis of the PR-averaging technique is known from \cite{}.\par
+\end{comment}
 
-A host of algorithms in machine learning and in particular, the temporal difference class of algorithms arising in reinforcement learning, turn out to be special cases of \eqref{eq:sa}, namley linear stochastic approximation (LSA) algorithms.
-LSA algorithms assume the general form
+A linear stochastic approximation (LSA) algorithm of the form given by
 \begin{align}\label{eq:lsaintro}
@@
 \end{align}
-where $b_t\in \R^d$ and $A_t\in \R^{\dcd}$ are data dependent quantities that are preseneted to the learner in a sequential manner.
+where the iterates $\theta_t\in \R^d$ and step-sizes $\alpha_t>0$ are quantities related to the algorithm, while $b_t\in \R^d$ and $A_t\in \R^{\dcd}$ is the data sequence presented to the algorithm.
+LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence $(b_t,A_t)$.
+Specific examples include the stochastic gradient descent algorithm for the problem of linear predition with quadratic loss, and the temporal difference class (TD) of learning algorithms in RL.
+An additional feature in these class of pplications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
+It is of interest to understand the effect of the step-sizes (a quantity that can be controlled by the algorithm designer) on the performance of the LSA.
+The choice of the step-sizes $\alpha_t$ needs to trade-off learning rate versus filtering noise, so to speak.
+In general, one can guess that larger step-sizes might lead to faster learning and at the same time smaller step-sizes are better at handling the noise.
+A technique in literature that goes back to \cite{polyak,ruppert} is to use a constant step-size (CS) $\alpha_t=\alpha>0,\forall t\geq 0$  in conjunction with averaging of iterates , known as Polyak-Rupert (PR) averaging to filter the noise.
+The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$.
+Under the CS-PR technique $\theta_t$ is the internal vairable and $\thh_t$ is the output of the algorithm.
+Intuitively, the CS-PR technique can be seen to address the learning vs filtering trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the filtering.
+\par
+Recently \cite{bach} considered stochastic gradient descent algorithm (SGD)\footnote{SGD is a special of the LSA algorithm in \eqref{eq:lsaintro}.} with CS-PR for the linear prediction problem with squared loss and $i.i.d$ sampling.
+A ``remarkable'' result by \cite{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the expected squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\todoc{Dimension?}
+\paragraph{Motivation and Challenges:} In this paper, we wish to study the effect of CS-PR in the case of general LSA algorithms of form \eqref{eq:lsaintro}.
+Our study is motivated by the problem of linear value function approximation from experience replay in a batch setting in RL using TD class of algorithms.
+An important difference between the linear prediction problem considered by \cite{bach} and the general LSA setting we consider is that in the former has certain special strucutres and the latter does not.
+One of the special strucutre exploited by \cite{bach} is the fact that $A_t$ are real symmetric matrices.
+The general LSA setting does not enjoy such an additional structure presents an interesting challenge.\par
+\paragraph{Contributions:} We now list our contributions in this paper
+\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\item We show that in the case of general LSA, there exists instance dependent constant step size choice such that the mean squared error (MSE) of the CS-PR procedure given by $\EE{\norm{\thh_t-\ts}^2}$ is at most $O(\frac{1}{t})$.
+The MSE further can be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
+ \item We also show (via counter example) that not all classes of general LSAs `admit' a constant step-size that is uniform across all the problems in the class.
+Further, we also argue that the subclass of LSAs related to linear prediction problem does `admit' a problem independent constant step size.
+\item We observe that when the constant step-size is chosen too big it results in instability, however, a conservative choice of the constant step-size affects only the constants in our error bounds without affecting the asymptotic rate (which might not be true with other popular choices of step-sizes).
+We aruge heuristically that CS-PR procedure can be deployed in a pricinpled manner by using an outer loop that automatically reduces that constant step-size by detecting detects the blow up of iterates.
+\end{itemize}
+\paragraph{Highlights:}
+ We only assume that $\EE{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts.
+This is the condition used in prior works by \cite{polyak-judisky,borkarbook}.
+Eventhough our proof technique resembles that by \cite{bach}, in addition, we make use of Jordon/Cyclic decomposition in a critical way to handle the general Hurwitz assumption.
+Our results are applicable to wide variety of TD algorithms such as TD(0), GTD/GDT2, TDC.
+Further, the error dynamics has a simple interpretation in terms of behaviour of the sum of terms in a geometric series, and hence offers a simple and unified understanding of the various TD algroithms.
+\begin{comment}
+A host of algorithms in machine learning and in particular, the temporal difference class of algorithms arising in reinforcement learning, turn out to be special cases of \eqref{eq:sa}, namley linear stochastic approximation (LSA) algorithms.
+LSA algorithms assume the general form
  Further, in most instances of LSA $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be computed in $O(n)$, and are desired for this low per iteration computational cost.
@@
 A ``remarkable'' result by \cite{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, he expected squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\todoc{Dimension?}
+
 In particular, as opposed to many earlier results available in the literature \todoc{Add citations}, $C$ does \emph{not} depend on the conditioning of the underlying system, while the total runtime of the algorithm scales linearly with both $t$ and $n$.\par
@@
 \end{itemize}
+\end{comment}
 \begin{comment}

### 2017-05-08 examples.tex
updated version

Our objective is to analyse $\EE{\norm{\thh_t-\theta^*}_C^2}$, where $C$ is a positive definite matrix.

\begin{example}\label{ex:tdzero}[Temporal Difference Learning-  (TD(0)) ]
The simplest temporal difference learning algorithm with a constant stepsize $\alpha>0$ is given by
\begin{align}
\label{tdzero}
\begin{split}
\theta_t
& = \theta_{t-1}+\alpha\, (R_t+ \gamma \ip{\theta_{t-1},\tilde{\phi}_t } - \ip{\theta_{t-1},\phi_t})\phi_t \\
& = \theta_{t-1}+\alpha\, \bigl( R_t\phi_t- (\phi_t \phi_t^\top-\gamma \phi_t \tilde{\phi}_t^\top)\theta_{t-1}\bigr)\,.
\end{split}
\end{align}
Here, $(\phi_t,\tilde{\phi}_t,R_t)\in \R^n\times \R^n \times \R$ is the data at time $t$.
Assuming $(\phi_t,\tilde{\phi}_t,R_t)$ is i.i.d. and say, bounded,
we see that with $A_t = \phi_t \phi_t^\top-\gamma \phi_t \tilde{\phi}_t^\top$, $b_t = R_t \phi_t$, \eqref{tdzero} is in the form of
\eqref{conststep}.
Note that the matrices $A_t$ are rank-1, hence the algorithm can indeed be implemented using $O(n)$ storage and
time (as is also shown in the first form of the update of $\theta_t$). In the so called `on-policy' setting, it is known that $A_P$ is positive definite \cite{}. Note that in general neither $A_t$ nor $A_P$ is symmetric.
\todoc[inline]{Finish: This is indeed satisfied in ``on policy'' TD, etc. Add at least one reference. Note that this is TD(0).}
In TD learning the error of a parameter $\theta$ is commonly measured using the expected squared
prediction error $L_t(\theta) = \EE{ (\phi_t^\top \theta - \phi_t^\top \ts)^2 } = \norm{\theta - \ts}_C^2$,
where $C = \EE{ \phi_t \phi_t^\top }$ is assumed to be PD.
\if0
\begin{align}\label{tdzero}
\theta_t=&\theta_{t-1}+\alpha \big(\phi(s_t) R(s_t)\nn\\&+ (\phi(s_t)\phi^\top(s_t)-\gamma \phi(s_t)\phi^\top(s'_t))\theta_{t-1}\big),
\end{align}
where are $s_t,s'_t\in S$ with $S$ being a finite set, $\phi(s)\in \R^n$ are the so called \emph{feature} vectors, $\gamma\in (0,1)$ is a given discount factor and $R\colon S\ra \R$ is map from S to reals. \todoc{Shoot, I wish we used $d$ in place of $n$..}
Here, $s_t$ are $i.i.d$ random variables distributed according to the law $s_t\sim \xi$ where $\xi$ is supported on $S$ and $s'_t\sim p(s_t,\cdot)$. The TD algorithm in \eqref{tdzero} can be cast in the general form as presented in \eqref{eq:lsa}, by letting $g_t=\phi(s_t)R(s_t)$ and $A_t=\phi(s_t)\phi^\top(s_t)-\gamma \phi(s_t)\phi^\top(s'_t)$.
\fi
\end{example}

\begin{example}[Temporal Difference Learning-  (GTD)) ]
The gradient temporal difference learning algorithm with constant stepsize $\alpha>0$ is given by
\begin{align}
\label{master} y_t=y_{t-1}+\alpha\, \bigl( b_t-A_t \theta_{t-1} -y_{t-1}\bigr)\\
\label{slave} \theta_t& = \theta_{t-1}+\beta\, \bigl(A_t^\top \theta_{t-1}\bigr),
\end{align}
where $A_t$ and $b_t$ are same as in \cref{ex:tdzero}.
In contrast to TD, the GTD algorithm has two recursions namely the \emph{master} that updates $y_t$ in \eqref{master} and the \emph{slave} that updates $\theta_t$ \eqref{slave}. Notice that the master-slave recursions can be written as a single recursion by lumping together the variables and properly taking care of the step sizes. This is accomplished below:
\begin{align}
\xi_{t}=\xi_{t-1}+\alpha (g_t -H_t),
\end{align}
where $g_t=\begin{bmatrix}b_t\\ 0\end{bmatix}$, and $H_t=\begin{bmatrix}-I & -A_t\\ \frac{\beta}{\alpha}A_t^\top \zero_{n\times n} \end{bmatrix}$
In TD learning the error of a parameter $\theta$ is commonly measured using the expected squared
prediction error $L_t(\theta) = \EE{ (\phi_t^\top \theta - \phi_t^\top \ts)^2 } = \norm{\theta - \ts}_C^2$,
where $C = \EE{ \phi_t \phi_t^\top }$ is assumed to be PD.
\if0
\begin{align}\label{tdzero}
\theta_t=&\theta_{t-1}+\alpha \big(\phi(s_t) R(s_t)\nn\\&+ (\phi(s_t)\phi^\top(s_t)-\gamma \phi(s_t)\phi^\top(s'_t))\theta_{t-1}\big),
\end{align}
where are $s_t,s'_t\in S$ with $S$ being a finite set, $\phi(s)\in \R^n$ are the so called \emph{feature} vectors, $\gamma\in (0,1)$ is a given discount factor and $R\colon S\ra \R$ is map from S to reals. \todoc{Shoot, I wish we used $d$ in place of $n$..}
Here, $s_t$ are $i.i.d$ random variables distributed according to the law $s_t\sim \xi$ where $\xi$ is supported on $S$ and $s'_t\sim p(s_t,\cdot)$. The TD algorithm in \eqref{tdzero} can be cast in the general form as presented in \eqref{eq:lsa}, by letting $g_t=\phi(s_t)R(s_t)$ and $H_t=\phi(s_t)\phi^\top(s_t)-\gamma \phi(s_t)\phi^\top(s'_t)$.
\fi
\end{example}

\begin{comment}
We consider the vanilla version of the linear regression problem, where `amazing' results are known. It serves us to contrast the results we obtain for general linear stochastic approximation algorithms to the results known for this setting.
\end{comment}
\begin{example}\label{ex:leastsquares}[Linear Regression]
Let $(x_t,y_t)\in \R^n\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite. Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$. Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$. The stochastic gradient descent scheme with constant step size $\alpha>0$ to minimize $f(\theta)$ can be given as below:
\begin{align}\label{linreg}
\theta_t=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)
\end{align}
Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$. It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}. In this case $A_P$ is PSD.
In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
\begin{align*}
f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
\end{align*}
\end{example}

### 2017-05-08 instance.tex
updated version

--- 2017-05-01 instance.tex
@@
-\section{Instance Dependent Bounds}
-In this section, we derive instance dependent bounds for $L_t$ in \eqref{eq:exploss} which is valid for a fixed $P$ and in the next section, explore the question of deriving uniform bounds for all $P\in \P$ (where $\P$ is a class of distributions over $\C^{\ncn}$).
+\section{Instance Dependent Bounds}\label{sec:mainresults}
+In this section, we derive instance dependent bounds  which is valid for a fixed $P$ and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions over $\C^{\dcd}$.
 We first present the results and then follow it up with a discussion.
 \begin{lemma}\label{lm:pd}
-Let $P$ be a positive definite distribution over $\C^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+Let $P$ be a positive definite distribution over $\C^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
 \end{lemma}
+\begin{comment}
+\begin{proof}
+\begin{align*}
+\rhos{P}&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^*\EE{A_t^* A_t} x\\
+&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x\\
+&\geq \lambda_{\min}(A^*_P+A_P)-\alpha \norm{A_P}^2-\sigma^2_P
+\end{align*}
+The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*_P+A_P)}{\norm{A_P}^2+\sigma^2_P}$
+\end{proof}
+\end{comment}
+\begin{lemma}\label{lm:simtran}[Similarity Transformation]
+Let $P^{AS}$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
+\end{lemma}
+
 \begin{theorem}\label{th:pdrate}
-Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C^{\ncn}$. Then
-\begin{align}
-\EE{\norm{\thh_t-\theta^*}^2}
+Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
+\begin{align*}
+\EE{\norm{\thh_t-\ts}^2}
 \leq
 \left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
-\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
-%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
-\end{align}
+\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
+%\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\ts})}{t+1} \right)\,.
+\end{align*}
 
@@
 
-\begin{lemma}\label{lm:simtran}[Similarity Transformation]
-Let $P^{AS}\in\P$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
-\end{lemma}
-
 \begin{theorem}\label{thm:simtran}[Change of Basis]
-Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
-\begin{align}
+Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be AS. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\ts$, and suppose if $P_U$ is positive definite, then
+\begin{align*}
 \EE{\norm{\gh_t-\gamma^*}^2}
@@
 %\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
-\end{align}
+\end{align*}
 where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
@@
 Let $P$ in \Cref{lsadist} be asymptotically stable. Then, there exists
-\begin{align}
-\EE{\norm{\thh_t-\theta^*}^2}
+\begin{align*}
+\EE{\norm{\thh_t-\ts}^2}
 \leq
-\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U}^2\norm{U^{-1}}^2}{\alpha \rhos{P_U}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
-\end{align}
+\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}
+\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
+\end{align*}
 \end{corollary}
-\subsection{Results Discussion}
-\textbf{Error Recursion and Contraction Maps:}
+\subsection{Result Discussion}
+\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} We can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{corss} terms of$\EE{\ip{e_i,e_j}},i\neq j$.
+The growth of the diagonal terms and the cross terms depends on the spectral norm of the matrices $I-\alpha A_t$ and $I-\alpha A_P$ respectively.
+These are given by
+\begin{comment}
 The output $\thh_t$ of the algorithm \eqref{eq:lsa} is the average of the internal states at times $s=0,\ldots,t-1$.
-Hence, before we proceed to interpret the results in \Cref{lm:pd,thm:pdrate,lm:simtran,thm:simtran,cor:as}, it is worthwhile to spend some effort in trying to understand the dynamics of the internal states by looking at the behavior of $e_t\eqdef \theta_t-\ts$.
+The error dynamics of the internal states by looking at the behavior of $e_t\eqdef \theta_t-\ts$.
 \begin{align}\label{eq:errec}
@@
 One way to characterize this contracting property is to look at the spectral radius of the random matrix, which is given by
+\end{comment}
 \begin{align}\label{eq:spectralrand}
@@
 \end{align}
-Notice that \eqref{eq:err} places our definitions of $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract}.
-Here the subscripts $s$ and $d$ stand for \emph{stochastic} and \emph{deterministic}, since they are associated with the spectral radius of stochastic and deterministic matrices respectively.\\
-\textbf{Positive Definiteness:} The first result in \Cref{lm:pd} shows if $P$ is PD, then there exists a stepsize for which $\rhos{P}, \rhod{P}>0$ which in turn implies that the spectral radii in \eqref{eq:spectralrand} are less than $1$.
-\textbf{Similarity Transformation:} 
-\paragraph{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms.
-The first term is the bias term given by $\B=(1+4(\alpha\rho_{\alpha})^{-1}) (\alpha\rho_{\alpha})^{-1} \Big(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}\Big)$.
-The \emph{Bias} term that captures the rate at which the initial condition $\norm{\theta_t-\ts}^2$ is forgotten.
-The second term is the \emph{Variance} given by $\V=(1+4(\alpha\rho_{\alpha})^{-1}) (\alpha\rho_{\alpha})^{-1} \Big(\frac{\alpha^2\sigma^2}{t+1} \Big)$.
+Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract} to the spectral norms of the aforementioned matrices.
+For the MSE to be bounded, we need the spectral norms to be less than unity, which in turn translates to the requirement that $\rhos{P}>0$ and $\rhod{P}>0$.
+\item \textbf{Role of $U$:} When $A_P$ is positive definite, it is straighforward to produce an $\alpha>0$ such that $\rhos{P}>0$ and $\rhod{P}>0$.
+However, $A_P$ is Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}) and a result an $\alpha>0$ such that $\rhos{P}>0$ and $\rhod{P}>0$.
+This poses a difficulty in applying results of \Cref{th:pdrate} in the case Hurwitz matrices.
+This can be remedied by a similiarity transformation using an appropriate $U$ that transforms $A_P$ to appropriate Jordon form $A_{P_U}$.
+\Cref{th:simtran} presents the rates for the transformed iterates   $\gamma_t=U^{-1 }\theta_t$.
+Note that \Cref{cor:as} is obtained from \Cref{th:simtran} by the fact that $\theta_t=U\gamma_t$.
+\begin{example}\label{ex:pdas}
+Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix}$ are not positive definite.
+However choosing $U_1=\begin{bmatrix} \frac{1}{\sqrt{2}}& -\frac{1}{\sqrt{2}}\\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}} \end{bmatrix}$, $U_2=\begin{bmatrix} \frac{\sqrt{2}}{1-i\sqrt{3}} & \frac{\sqrt{2}}{1+i\sqrt{3}} \\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}}\end{bmatrix}$ and $U_3=\begin{bmatrix} 1 & 0\\ 0 &\frac{1}{10} \end{bmatrix}$, we can see that matrices $J_i=U_i^{-1}A_i U_i,i=1,2,3$ are positive definite.
+\end{example}
+\item \textbf{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms.
+The first term is the bias term given by $\B=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}$.
+The \emph{bias} term that captures the rate at which the initial condition $\norm{\theta_0-\ts}^2$ is forgotten.
+The second term is the \emph{variance} given by $\V=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}} \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} $.
 The variance term that captures the rate at which noise is rejected.
-The $\B$ and $\V$ terms capture two different sources of errors.
-To see this, note that in the absence of noise i.e., when $\zeta_t=0,~\forall t\geq 0$, we have $\E[\norm{H\eb_t}^2]=\B$ and in the presence of noise, i.e., $\zeta\neq 0,\forall t\geq 0$ and the perfect initial condition, i.e., $\theta_t=\ts$, we have $\E[\norm{H\eb_t}^2]=\V$.
-\paragraph{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term.
+\item \textbf{$\alpha$, $\rhos{P}$ and $\rhod{P}$} only affect the constants in the rates and not the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
+\item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term.
 This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower.
-However, in this case the noise term does not blow up, a fact that can appreciated by looking at \eqref{lsergen} where $\alpha$ is seen to multiply the noise term $\zeta_t$.
+However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:pdrate,th:simtran,cor:as} where $\alpha$ is seen to multiply the variances $\sigma^2_P$ and $\sigma^2_b$.
 In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other.
-For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{\max}$, the bounds blow up again, due to the fact that $\rho_{\alpha}\ra 0$ in this case.
-This is due to the fact that the effect of both noise and initial conditions decay with the contraction factor of $F_{t,i}$, which gets closer to unity as $\alpha\ra\alpha_{\max}$.
+For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.
+\item \textbf{Geometric Series:} In simple terms the error dynamics is akin to sum of a geometric series with common factor $\alpha\rho$, a fact that is illustrated by the simple $1$-dimensional example in \Cref{ex:geo}.
+\begin{example}\label{ex:geo}
+Consider the one dimensional case where $A_t=\rho$ for some $0<\rho<1$, $b_t$ a bounded $i.i.d$ zero mean random variable with variance $\sigma^2$, and $\ts=0$.
+We know that $e_t=(1-\alpha\rho)^t e_0+\sum_{s=1}^t (1-\alpha \rho)^{t-s}\alpha b_s$ and the MSE is given by
+\begin{align*}
+\EE{\norm{\theta_t}^2}&=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2\\
+&\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)
+\end{align*}
+\end{example}
+\item \textbf{Lower Bounds:}
+Consider the LSA with $b_t=0$, and $P^0$ has all the mass concentrated on $A_{P^0}=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$.
+Note that in this example $\ts=0$.
+By choosing $\alpha<\frac2{\lambda_{\max}}$, in this case it is straightforward to write the expression for $\eh_t$ explicitly as below:
+\begin{align*}
+\eh_t&=\frac{1}{t+1}\sum_{s=0}^t e_t = \frac{1}{t+1}\sum_{s=0}^t (I-\alpha A_P)^{t-s} e_0\\
+&=\frac{1}{t+1}(\alpha A)^{-1}\left(I-(I-\alpha A)^{t+1}\right)e_0
+\end{align*}
+\begin{comment}
+\begin{align*}
+\norm{\eh_t}^2&=\frac{1}{(t+1)^2}\norm{(\alpha A_P)^{-1}\left(I-(I-\alpha A_P)^{t+1}\right)e_0}^2
+&\leq\frac{1}{(t+1)^2}\alpha^2\norm{A_P^{-1}}^2\norm{e_0}^2
+\end{align*}
+\end{comment}
+
+\begin{align*}
+\norm{\eh_t}^2\geq \eh^2_t(1)=\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)
+\end{align*}
+Note that in this example, $\rhos{P^0}=\rhod{P^0}=\lambda_{\min} -\alpha \lambda_{\min}^2=\lambda_{\min}(1-\alpha \lambda_{\min})$. Sandwiching the error between the lower and upper bounds we have
+\begin{align}
+\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
+\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
+\end{align}
+\begin{example}
+Consider distributions $P^1$, $P^2$ and $P^3$ which are concentrated on matrices $A_{P^1}=\begin{bmatrix} 10 &100\\ 0 &10\end{bmatrix}$, $A_{P^2}=\begin{bmatrix} 10 &10\\ 0 &10\end{bmatrix}$, $A_{P^3}=\begin{bmatrix} 10 &0 \\ 0 &10\end{bmatrix}$.
+Note that $P^1$, $P^2$ and $P^3$ are respectively AS, PD, SPD respectively.
+Consider the LSA for the three case of $P^1$, $P^2$ and $P^3$ with $b_t=0$ and $\ts=0$
+We now list the following observations:
+\begin{itemize}%[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\item In all the three instance, $0<\alpha<\frac2{10}$ ensures that the error $\norm{\eh_t}\ra 0$ as $t\ra\infty$. Thus the range of $\alpha$ is uniform across $P^i,i=1,2,3$.
+\item $\rhos{P^1}=\rhod{P^1}<0$, and hence one cannot apply results of \Cref{th:pdrate}.
+However, with  $U=\begin{bmatrix}1 &0 \\ 0 &0.2-\epsilon\end{bmatrix}$ for some $\epsilon>0$ we have $P^1_U$ to be the distribution concentrated on the matrix $A_{P^1_U}=\begin{bmatrix} 10 &20 \\ 0 &10\end{bmatrix}$ and one can apply the results of \Cref{thm:simtran,cor:as}.
+Notice that the condition number $\norm{U}{U^-1}\approx 5$.
+\end{itemize}
+We plot $L_t$ for the three cases for $t=1,\ldots,100$ and we can see that the condition number $\norm{U}{U^-1}\approx 5$, shows up in the AS case ($P^1$).
+\end{example}
+\end{itemize}
+\paragraph{Related Work}
+LSA with CS-PR was considered in \cite{polyak-judisky} under the additive noise assumption, wherein $A_t=A$.
+Recent work by \cite{bach} considers a stochastic gradient descent (SGD) algorithm for the problem of linear prediction given in \Cref{ex:linreg}.
+\begin{example}\label{ex:linreg}
+Let $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
+Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$.
+Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$.
+SGD with CS-PR to minimize $f(\theta)$ can be given as below:
+\begin{subequations}\label{eq:linreg}
+\begin{align}
+\theta_t&=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1}} -y_t\right)\\
+\thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t
+\end{align}
+\end{subequations}
+Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$.
+It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}.
+In this case $A_P$ is SPD.
+In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
+\begin{align}\label{eq:mselinreg}
+f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
+\end{align}
+\end{example}
+In this setting, \cite{bach} show that the variance term in the MSE is at most $O(\frac{C}{t})$, where $C$ does not depend on the condition number.
+This ``remarkable'' result is due to the following facts: $(i)$ $A_P$ in this case is symmetric positive definite, $(ii)$ the MSE in \eqref{eq:mselinreg} is measured with respect to $A_P$, $(iii)$ the noise $\epsilon_t=y_t-\ip{\ts,x_{t-1}}$ is structured i.e., $\EE{\epsilon_t^2 x_t x_t^\top}\leq R A_P$ for some constant $R>0$.
+At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.

### 2017-05-08 intro.tex
updated version

@@
 where the iterates $\theta_t\in \R^d$ and step-sizes $\alpha_t>0$ are quantities related to the algorithm, while $b_t\in \R^d$ and $A_t\in \R^{\dcd}$ is the data sequence presented to the algorithm.
-LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence $(b_t,A_t)$.
-Specific examples include the stochastic gradient descent algorithm for the problem of linear predition with quadratic loss, and the temporal difference class (TD) of learning algorithms in RL.
-An additional feature in these class of pplications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
+LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence $(b_t,A_t)$ using cheap per iteration computations.
+Specific examples include the stochastic gradient descent algorithm for the problem of linear predition with quadratic loss, and the temporal difference class (TD) of learning algorithms in RL.\par
+%An additional feature in these class of pplications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
 It is of interest to understand the effect of the step-sizes (a quantity that can be controlled by the algorithm designer) on the performance of the LSA.
@@
 Further, the error dynamics has a simple interpretation in terms of behaviour of the sum of terms in a geometric series, and hence offers a simple and unified understanding of the various TD algroithms.
+\paragraph{Organization:}
 \begin{comment}

### 2017-05-08 prob.tex
updated version

@@
 \section{Problem Setup}
-We consider the linear stochastic approximation (LSA) algorithm with constant step size and PR-averaging of the iterates.
+We consider the linear stochastic approximation (LSA) algorithm with constant step-size (CS) and Polyak-Ruppert (PR) averaging of the iterates.
 \begin{subequations}\label{eq:lsa}
@@
 \end{subequations}
-where $\thh_t, \theta_t \in \C^{d}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\dcd}$ incrementally, in discrete time steps $t=1,2,\dots$
-based on data $b_t\in \C^d$, $A_t\in \C^{n\times n}$\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^d$ and $A_t\in \R^{\dcd}, b_t\in \R^d$.
-However, it serves well for our analysis to make the general assumption that these qunatites are complex valued.}.
-Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
+where\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^d$ and $A_t\in \R^{\dcd}, b_t\in \R^d$.
+However, it serves well for our analysis to consider these qunatites to be complex valued.} $\thh_t, \theta_t \in \C^{d}$.
+The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\dcd}$ incrementally, in discrete time steps $t=1,2,\dots$
+based on data $b_t\in \C^d$, $A_t\in \C^{\dcd}$. Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
 initial value $\theta_0$.
-The iterate $\theta_t$ is treated as an internal state of the algorithm, while $\th_t$ is the output at time step $t$.
-The update of $\theta_t$ alone is considered a form of constant stepsize LSA.
-Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(n)$ storage.
-
-Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$
-can be efficiently computed in $O(n)$ time.
+The iterate $\theta_t$ is treated as an internal state of the algorithm, while $\thh_t$ is the output at time step $t$.
+The update of $\theta_t$ alone is considered a form of constant step-size LSA.\par
+Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(d)$ storage.
+Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$ can be efficiently computed in $O(d)$ time.
 This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{} and linear prediction \cite{} will be presented in a later sections (\Cref{sec:rl,sec:lp}).
 The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
-can be implemented in $O(n)$ time and $O(n)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.
-We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa} In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.
+can be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.\par
+We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa}.
+In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.
 \begin{assumption}\label{assmp:lsa}
@@
 \end{assumption}
-\textbf{Performance Metric:}
-The ``aim'' of the updates in \eqref{eq:lsa} is to estimate $\ts$ based on the noisy versions $(A_t,b_t)$ of $(A_P,b)$.
-There are various ways of quantifying the error $\eh_t \doteq \thh_t- \ts$ of estimating $\ts$ using $\thh_t$, but the magnitude
-of the errors is usually measured in some weighted quadratic norm. The expected loss for time step $t$ then becomes
-\begin{align}\label{eq:exploss}
-L_t \doteq \EE{ \norm{\eh_t}_{C}^2 }\,,
-\end{align}
-where recall that for a SPD matrix $C$, $\norm{x}_C^2 = x^* C x$. Here, the choice of the matrix $C$ depends on the application.\\
+%For the rest of the paper we denote the errors of the internal and the output variables of \eqref{eq:lsa} at time $t$ by $e_t\eqdef \theta_t-\ts$ and $\eh_t\eqdef \thh_t-\ts$ respectively.
+\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ We note here that $\EE{\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$.
+However, unless $C$ has some special structure, it is not possible to exploit this generality in our bounds.
+Further, when data is bounded $\norm{C}^2$ is also bounded and it is straightforward to see that $\EE{\norm{\thh_t-\ts}}^2_C\leq \norm{C}^2\EE{\norm{\thh_t-\ts}^2}$.
+So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in the linear regression setting, where its usage is relevant.} by $\EE{\norm{\thh_t-\ts}^2}$.
+We wish to understand how the MSE is affected by the choice of $\alpha$ and the spectral properties of $A_P$.
+Another aspect of interest is to understand the contribution of the \emph{bias} (caused by the initial condition $\theta_0$) and the \emph{variance} (caused by $\sigma^2_P,\sigma^2_b$) to the MSE.\par
+\begin{comment}
 \textbf{Challenges:}
@@
 \todoch{Mention impotance of forgetting the bias etc. Mention that we really don't have control over $\norm{U}$}
+\end{comment}

### 2017-05-08 related.tex
updated version

--- 2017-05-01 related.tex
@@
-\section{Related Work}
-\subsection{Linear Regression}
-Let $(x_t,y_t)\in \R^n\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
-Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$.
-Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$.
-The stochastic gradient descent scheme with constant step size $\alpha>0$ to minimize $f(\theta)$ can be given as below:
-\begin{align}\label{linreg}
-\theta_t=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)
-\end{align}
-Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$.
-It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}.
-In this case $A_P$ is SPD.
-In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
-\begin{align*}
-f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
-\end{align*}
+\section{Applications in Reinforcement Learning}
+\paragraph{Definitions:} We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
+Let $\gamma \in (0,1)$ be a fixed constant, we define $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$.
+In what follows, $\mu_t>0$ is an \emph{importance} sampling factor whose aim is to correct for mismatch in the (behavior) distribution with which the data was collected and the (target) distribution with respect to which one wants to learn.
+A factor $\mu_t=1,\,\forall t\geq 0$ will mean that no correction is required\footnote{This is known as the \emph{on-policy} case where the behavior is identical to the target.
+The general setting where $\mu_t>0$ is known as \emph{off-policy}.}.
+\begin{table}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|c|c|c|}\hline
+Algorithm& Update &Stepsize &Rate \\ \hline
+TD(0)& $\begin{aligned}\theta_t&=\theta_{t-1}+\alpha_t (b_t -\Delta_t\theta_{t-1})\\ \thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t\end{aligned}$ &$\begin{aligned}\alpha_t&=O(\frac1{t})^\delta,\\ &\delta\in(1/2,1)\end{aligned}$ &$\EE{\norm{\eh_t}}\leq O(\frac{1}{\sqrt{t}})$ for $\delta\ra 1$ \\ \hline
+GTD& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}- Q_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t A^\top_t y_{t-1})\end{aligned}$ & $\begin{aligned}&\beta_t=\eta \alpha_t,\sum_{t\geq 0}\alpha_t=\infty, \\ &\sum_{t\geq 0}\alpha^2_t<\infty \end{aligned}$ & $e_t\ra 0$ w.p.1 \\\hline
+TDC& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t b_t-\mu_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$ & $\begin{aligned}&\beta_t,\alpha_t,\frac{\alpha_t}{\beta_t} \ra 0,~\text{as}~t\ra \infty,\\& \sum_{t\geq 0}\alpha_t=\infty, \sum_{t\geq 0}\beta_t=\infty, \\ &\sum_{t\geq 0}\alpha^2_t<\infty, \sum_{t\geq 0}\beta^2_t<\infty \end{aligned}$ & $e_t\ra 0$ w.p.1 \\\hline
+GTD& $\begin{aligned} y_{t}&=\Pi_Y \left(y_{t-1}+\alpha_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-Q_t y_{t-1})\right)\\ \theta_t &=\Pi_{\Theta}\left(\theta_{t-1}+\alpha_t (\mu_t A_t^\top y_{t-1})\right)\\ \thh_t&=\frac{\sum_{s=0}^t \alpha_s\theta_s}{\sum_{s=0}^t \alpha_s}\end{aligned}$ & $\alpha_t =O(\frac{1}{\sqrt{t}})$ & $\norm{A_P\eh_t-b}^2 \leq O(\frac{1}{\sqrt{t}})$ w.h.p\\\hline
 
-\subsection{Reinforcement Learning}
-\subsection{GTD}
-\todoch{Use LiHong Li's GTD results, choice of the basis $U$ etc}
+\end{tabular}
+}
+\label{tb:rates}
+\end{table}
+\paragraph{Related Work in RL} The various temporal difference learning algorithms can be written as in \Cref{table:td}.
+The TD(0) algorithm is an on-policy algorithm ($\mu_t=1$) and is unstable in the general {off-policy} setting.
+The gradient temporal difference (GTD) algorithm \cite{} is stable in off-policy setting as well.
+Recent work by \cite{korda-prashanth} derive finite time bounds for the TD(0) for two different cases of step-sizes one without PR averaging and other with PR averaging.
+In the case without PR averaging, \cite{korda-prashanth} consider step-sizes of the form $\frac{c_0c}{c+t}$, where $c_0$ and $c$ are problem dependent positive constants, to dervie a rate of $O(\frac{1}{t})$ for the MSE.
+In the case with PR averaging (row $1$ of \Cref{tb:rates}), \cite{korda-prashanth} consider step-size of the from $c_0\left(\frac{c}{c+t}\right)^\delta,\,\delta \in (1/2,1)$ where $c_0, c$ are problem independent positive constants, and show that this choice yeilds a rate of $O(\frac{1}{t})$ for $\delta \ra 1$ and $t\geq t_0$ (where $t_0$ is a problem dependent time instant).
+A possible limiting feature of the results by \cite{korda-prashanth} is that either the problem knowledge is required or the bounds hold only after a problem dependent time.
+A centered TD(0) variant is also considered by \cite{korda-prashanth}, however, a projection operator is required to keep the iterates bounded.
+Also, the bounds by \cite{korda-prashanth} don't have an explicit bias variance decomposition.
+The works on GTD and TDC by \cite{gtdfirst,tdc} show only asymptotic convergence w.p.1.
+Recently, the work by \cite{dtdmp} shows a rate of $\frac{1}{\sqrt{t}}$ for GTD with use of projection operation, and also proposed a variant namely GTD-Mirror-prox based on stochastic mirror-prox ideas which achieves $\frac{1}{\sqrt{t}}$ rate without use of projection.
+In comparison, we don't use of projections, our bounds hold for all time $t$, and the bounds are dependent on the problem instance only to the extent of choosing a step-size that leads to stable behavior, in which case we show that $O\frac{1}{t}$ rate holds (the scaling factor hidden in the $O$ expression is problem dependent).
+Further, our results provide explicit bias variance decomposition.

### 2017-05-08 stepsizes.tex
updated version

@@
-\section{Choice of Step-sizes}
+\section{Automatic Tuning of Step-Sizes}
+\begin{comment}
 We now argue, in a heuristic manner, as to why constant step-size with PR-averaging might be a principled idea, especially in TD type algorithms.
@@
 Using this observation, the constant-step size can be set in a more prinicipled manner as showing the pseudocode below.
+\end{comment}
+In \Cref{sec:uniform} we showed a negative result that for the general class of LSAs, it is not possible to pick a fixed uniform step-size that can be used for any instance of the class. Further,
+recall from our discussion in \Cref{sec:mainresults} that as $\alpha\ra\alpha_P$ (or $\alpha\ra\alpha_{P_U}$), the MSE blows up and for $\alpha>\alpha_P$ the LSA is unstable because the spectral norm of the random matrix $I-\alpha A_t$ is no longer less that unity.
+Combining these facts, we conclude that if the step-size is incorrect for a particular instance then we would observe instability, a fact that can be used to reset the step-size to a smaller value, and this procedure can be continued in an iterative fashion to find the right choice of step-size for a given problem.
+We formalize this idea in the form of a step-size tuning algorithm in \Cref{alg:tuning}
+
 \begin{algorithm}[H]
@@
 \begin{algorithmic}[1]
-\STATE{Initialize: $\theta_0$, $\alpha=1$, $i=1$, $a_0=0$}
+\STATE{Initialize: $\theta_0$, $\alpha=1$, $i=1$, $\rho_0=0$}
 \FOR{$t=1,2,\ldots, $}
 \STATE{$\theta_t=\theta_{t-1}+\alpha (b_t-A_t\theta_{t-1})$}
-\IF{$mod(t,k)==0$}
-\STATE{$a_{i}=\norm{\theta_t}$}
-\IF{$Is-stable(a_0,\ldots,a_i)==0$}
+\IF{$t (mod k)==0$}
+\STATE{$\rho_{i}=\norm{\theta_t}$}
+\IF{IsUnstable($\{\rho_j,j=1,\ldots,i\}$)==1}
 \STATE{$\alpha=\alpha/2$}
@@
 \end{algorithmic}
+\label{alg:tuning}
 \end{algorithm}
-
+In \Cref{alg:tuning} IsUnstable$(\rho_j,j=1,\ldots,i\}$) is a routine whose input is $\norm{\theta_t}$ measured at time instances $k$ apart.
+It is easy to show from Gronwall type arguments that the $\norm{\theta_t}\leq (1-exp{\beta t})$, where $\beta>0$ implies stability and $\beta<0$ imples instability.
+From this Gronwall based inequality, we observe that when the iterates are stable the ratio $\frac{\rho_{k+1}}{\rho_k},k=1,\ldots,i-1$ should be `roughly' decreasing (because there is the iterates are persistently corrupted by noise) and should converge to $1$.
+This condition on the sequence $\frac{\rho_{k+1}}{\rho_k}$ can be checked in practice with appropriate tolerances.
 \begin{comment}

### 2017-05-09 instance.tex
rearraged

@@
-\section{Instance Dependent Bounds}\label{sec:mainresults}
+\section{Main Results}\label{sec:mainresults}
 In this section, we derive instance dependent bounds  which is valid for a fixed $P$ and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions over $\C^{\dcd}$.
@@
 \end{theorem}
-
 \begin{theorem}\label{thm:simtran}[Change of Basis]
@@
 \end{corollary}
+\begin{theorem}[Lower Bound]
+There exists a $P$ such that
+\begin{align}
+\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
+\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
+\end{align}
+
+\end{theorem}
+\begin{comment}
+Consider the LSA with $b_t=0$, and $P^0$ has all the mass concentrated on $A_{P^0}=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$.
+Note that in this example $\ts=0$.
+By choosing $\alpha<\frac2{\lambda_{\max}}$, in this case it is straightforward to write the expression for $\eh_t$ explicitly as below:
+\begin{align*}
+\eh_t&=\frac{1}{t+1}\sum_{s=0}^t e_t = \frac{1}{t+1}\sum_{s=0}^t (I-\alpha A_P)^{t-s} e_0\\
+&=\frac{1}{t+1}(\alpha A)^{-1}\left(I-(I-\alpha A)^{t+1}\right)e_0
+\end{align*}
+
+\begin{align*}
+\norm{\eh_t}^2&=\frac{1}{(t+1)^2}\norm{(\alpha A_P)^{-1}\left(I-(I-\alpha A_P)^{t+1}\right)e_0}^2
+&\leq\frac{1}{(t+1)^2}\alpha^2\norm{A_P^{-1}}^2\norm{e_0}^2
+\end{align*}
+
+\begin{align*}
+\norm{\eh_t}^2\geq \eh^2_t(1)=\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)
+\end{align*}
+Note that in this example, $\rhos{P^0}=\rhod{P^0}=\lambda_{\min} -\alpha \lambda_{\min}^2=\lambda_{\min}(1-\alpha \lambda_{\min})$. Sandwiching the error between the lower and upper bounds we have
+\begin{align}
+\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
+\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
+\end{align}
+\end{comment}
+
 \subsection{Result Discussion}
@@
 Note that \Cref{cor:as} is obtained from \Cref{th:simtran} by the fact that $\theta_t=U\gamma_t$.
-\begin{example}\label{ex:pdas}
-Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix}$ are not positive definite.
-However choosing $U_1=\begin{bmatrix} \frac{1}{\sqrt{2}}& -\frac{1}{\sqrt{2}}\\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}} \end{bmatrix}$, $U_2=\begin{bmatrix} \frac{\sqrt{2}}{1-i\sqrt{3}} & \frac{\sqrt{2}}{1+i\sqrt{3}} \\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}}\end{bmatrix}$ and $U_3=\begin{bmatrix} 1 & 0\\ 0 &\frac{1}{10} \end{bmatrix}$, we can see that matrices $J_i=U_i^{-1}A_i U_i,i=1,2,3$ are positive definite.
-\end{example}
 \item \textbf{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms.
@@
 For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.
-\item \textbf{Geometric Series:} In simple terms the error dynamics is akin to sum of a geometric series with common factor $\alpha\rho$, a fact that is illustrated by the simple $1$-dimensional example in \Cref{ex:geo}.
+
+\subsection{Illustrative Examples}
+\begin{example}\label{ex:pdas}
+Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix}$ are not positive definite.
+However choosing $U_1=\begin{bmatrix} \frac{1}{\sqrt{2}}& -\frac{1}{\sqrt{2}}\\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}} \end{bmatrix}$, $U_2=\begin{bmatrix} \frac{\sqrt{2}}{1-i\sqrt{3}} & \frac{\sqrt{2}}{1+i\sqrt{3}} \\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}}\end{bmatrix}$ and $U_3=\begin{bmatrix} 1 & 0\\ 0 &\frac{1}{10} \end{bmatrix}$, we can see that matrices $J_i=U_i^{-1}A_i U_i,i=1,2,3$ are positive definite.
+\end{example}
+\paragraph{Remark}
+
 \begin{example}\label{ex:geo}
@@
 \end{example}
-\item \textbf{Lower Bounds:}
-Consider the LSA with $b_t=0$, and $P^0$ has all the mass concentrated on $A_{P^0}=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$.
-Note that in this example $\ts=0$.
-By choosing $\alpha<\frac2{\lambda_{\max}}$, in this case it is straightforward to write the expression for $\eh_t$ explicitly as below:
-\begin{align*}
-\eh_t&=\frac{1}{t+1}\sum_{s=0}^t e_t = \frac{1}{t+1}\sum_{s=0}^t (I-\alpha A_P)^{t-s} e_0\\
-&=\frac{1}{t+1}(\alpha A)^{-1}\left(I-(I-\alpha A)^{t+1}\right)e_0
-\end{align*}
-\begin{comment}
-\begin{align*}
-\norm{\eh_t}^2&=\frac{1}{(t+1)^2}\norm{(\alpha A_P)^{-1}\left(I-(I-\alpha A_P)^{t+1}\right)e_0}^2
-&\leq\frac{1}{(t+1)^2}\alpha^2\norm{A_P^{-1}}^2\norm{e_0}^2
-\end{align*}
-\end{comment}
-
-\begin{align*}
-\norm{\eh_t}^2\geq \eh^2_t(1)=\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)
-\end{align*}
-Note that in this example, $\rhos{P^0}=\rhod{P^0}=\lambda_{\min} -\alpha \lambda_{\min}^2=\lambda_{\min}(1-\alpha \lambda_{\min})$. Sandwiching the error between the lower and upper bounds we have
-\begin{align}
-\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
-\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
-\end{align}
+\paragraph{Geometric Series:} In simple terms the error dynamics is akin to sum of a geometric series with common factor $\alpha\rho$, a fact that is illustrated by the simple $1$-dimensional example in \Cref{ex:geo}.
 \begin{example}
@@
 \end{itemize}
-\paragraph{Related Work}
+\subsection{Related Work}
 LSA with CS-PR was considered in \cite{polyak-judisky} under the additive noise assumption, wherein $A_t=A$.

### 2017-05-09 intro.tex
rearraged

@@
 In general, one can guess that larger step-sizes might lead to faster learning and at the same time smaller step-sizes are better at handling the noise.
-A technique in literature that goes back to \cite{polyak,ruppert} is to use a constant step-size (CS) $\alpha_t=\alpha>0,\forall t\geq 0$  in conjunction with averaging of iterates , known as Polyak-Rupert (PR) averaging to filter the noise.
+A technique in literature that goes back to \cite{polyak-judisky,ruppert} is to use a constant step-size (CS) $\alpha_t=\alpha>0,\forall t\geq 0$  in conjunction with averaging of iterates , known as Polyak-Rupert (PR) averaging to filter the noise.
 The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$.
@@
  We only assume that $\EE{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts.
-This is the condition used in prior works by \cite{polyak-judisky,borkarbook}.
+This assumption is quite common in literatures and appreas in the works by \cite{polyak-judisky,borkarbook}.
 Eventhough our proof technique resembles that by \cite{bach}, in addition, we make use of Jordon/Cyclic decomposition in a critical way to handle the general Hurwitz assumption.

### 2017-05-09 prob.tex
rearraged

@@
 Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$ can be efficiently computed in $O(d)$ time.
-This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{} and linear prediction \cite{} will be presented in a later sections (\Cref{sec:rl,sec:lp}).
+This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp} and linear prediction \cite{bach,bachaistats} will be presented in a later sections (\Cref{sec:rl,sec:lp}).
 The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
@@
 \textbf{Challenges:}
-The idea of constant step size and iterate averaging is not new and goes back to \cite{}, where it was first shown that $L_t\ra 0$ as $t\ra\infty$.
+The idea of constant step size and iterate averaging is not new and goes back to \cite{polyak-judisky}, where it was first shown that $L_t\ra 0$ as $t\ra\infty$.
 Recently \cite{bach} revisited the idea in connection to the stochastic gradient descent (SGD) algorithm (which is also an LSA of the form in \eqref{eq:lsa}) to solve linear regression.

### 2017-05-09 related.tex
rearraged

@@
 The TD(0) algorithm is an on-policy algorithm ($\mu_t=1$) and is unstable in the general {off-policy} setting.
-The gradient temporal difference (GTD) algorithm \cite{} is stable in off-policy setting as well.
+The gradient temporal difference (GTD) algorithm \cite{gtd,gtd2,gtdmp} is stable in off-policy setting as well.
 Recent work by \cite{korda-prashanth} derive finite time bounds for the TD(0) for two different cases of step-sizes one without PR averaging and other with PR averaging.
@@
 Also, the bounds by \cite{korda-prashanth} don't have an explicit bias variance decomposition.
-The works on GTD and TDC by \cite{gtdfirst,tdc} show only asymptotic convergence w.p.1.
-Recently, the work by \cite{dtdmp} shows a rate of $\frac{1}{\sqrt{t}}$ for GTD with use of projection operation, and also proposed a variant namely GTD-Mirror-prox based on stochastic mirror-prox ideas which achieves $\frac{1}{\sqrt{t}}$ rate without use of projection.
+The works on GTD and TDC by \cite{gtd,gtd2} show only asymptotic convergence w.p.1.
+Recently, the work by \cite{gtdmp} shows a rate of $\frac{1}{\sqrt{t}}$ for GTD with use of projection operation, and also proposed a variant namely GTD-Mirror-prox based on stochastic mirror-prox ideas which achieves $\frac{1}{\sqrt{t}}$ rate without use of projection.
 In comparison, we don't use of projections, our bounds hold for all time $t$, and the bounds are dependent on the problem instance only to the extent of choosing a step-size that leads to stable behavior, in which case we show that $O\frac{1}{t}$ rate holds (the scaling factor hidden in the $O$ expression is problem dependent).

### 2017-05-10 intro.tex
intro backup before removing comments

--- 2017-05-07 intro.tex
@@
 
-A linear stochastic approximation (LSA) algorithm of the form given by
+A linear stochastic approximation (LSA) algorithm is given by
 \begin{align}\label{eq:lsaintro}
@@
 \end{align}
-where the iterates $\theta_t\in \R^d$ and step-sizes $\alpha_t>0$ are quantities related to the algorithm, while $b_t\in \R^d$ and $A_t\in \R^{\dcd}$ is the data sequence presented to the algorithm.
-LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence $(b_t,A_t)$.
-Specific examples include the stochastic gradient descent algorithm for the problem of linear predition with quadratic loss, and the temporal difference class (TD) of learning algorithms in RL.
-An additional feature in these class of pplications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
-It is of interest to understand the effect of the step-sizes (a quantity that can be controlled by the algorithm designer) on the performance of the LSA.
-The choice of the step-sizes $\alpha_t$ needs to trade-off learning rate versus filtering noise, so to speak.
-In general, one can guess that larger step-sizes might lead to faster learning and at the same time smaller step-sizes are better at handling the noise.
-A technique in literature that goes back to \cite{polyak,ruppert} is to use a constant step-size (CS) $\alpha_t=\alpha>0,\forall t\geq 0$  in conjunction with averaging of iterates , known as Polyak-Rupert (PR) averaging to filter the noise.
-The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$.
-Under the CS-PR technique $\theta_t$ is the internal vairable and $\thh_t$ is the output of the algorithm.
-Intuitively, the CS-PR technique can be seen to address the learning vs filtering trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the filtering.
+where the iterates, $\theta_t\in \R^d$, and step-sizes, $\alpha_t>0$, are algorithmic quantities, and $(b_t,A_t)\in \R^d\times \R^{\dcd}$ is the noisy data sequence presented to the algorithm.
+LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence, using $O(d)$ updates per iteration.
+Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear prediction with quadratic loss \cite{bach,bachaistats}, and the temporal difference class (TD) of learning algorithms in RL \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.\par
+%An additional feature in these class of applications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
+It is of interest to understand the effect of the step-sizes $\alpha_t$ (a quantity that can be controlled by the algorithm designer) on the performance of the LSA in \eqref{eq:lsaintro}.
+The choice of the step-sizes dictates learning rate, noise rejection and stability.
+Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster learning, while at the same time too large step-sizes might result in instability of \eqref{eq:lsaintro}.
+Diminishing and square-summable step-sizes have been a popular choice for the step-sizes, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$ \cite{gtd2,borkarbook,konda-tsitsiklis}.
+An alternative technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$.
+In LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
+Averaging in stochastic approximation to improve noise rejection is a procedure that goes back to  Polyak and Ruppert \cite{polyak-judisky,ruppert}.
+%In what follows, when we say CS-PR for LSA or LSA with CS-PR, we allude to the case where \eqref{eq:lsaintro} is a constant step-size LSA with $\alpha_t=\alpha>0,\forall t\geq 0$, $\theta_t$ in \eqref{eq:lsaintro} is considered an internal variable and $\thh_t$ is the final output of the LSA with CS-PR.
+%In LSA with CS-PR, \eqref{eq:lsaintro} is a constant step-size LSA with $\alpha_t=\alpha>0,\forall t\geq 0$, $\theta_t$ in \eqref{eq:lsaintro} is considered an internal variable and $\thh_t$ is the final output of the LSA with CS-PR.
+Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection.
 \par
-Recently \cite{bach} considered stochastic gradient descent algorithm (SGD)\footnote{SGD is a special of the LSA algorithm in \eqref{eq:lsaintro}.} with CS-PR for the linear prediction problem with squared loss and $i.i.d$ sampling.
-A ``remarkable'' result by \cite{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the expected squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\todoc{Dimension?}
-\paragraph{Motivation and Challenges:} In this paper, we wish to study the effect of CS-PR in the case of general LSA algorithms of form \eqref{eq:lsaintro}.
-Our study is motivated by the problem of linear value function approximation from experience replay in a batch setting in RL using TD class of algorithms.
-An important difference between the linear prediction problem considered by \cite{bach} and the general LSA setting we consider is that in the former has certain special strucutres and the latter does not.
-One of the special strucutre exploited by \cite{bach} is the fact that $A_t$ are real symmetric matrices.
-The general LSA setting does not enjoy such an additional structure presents an interesting challenge.\par
-\paragraph{Contributions:} We now list our contributions in this paper
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item We show that in the case of general LSA, there exists instance dependent constant step size choice such that the mean squared error (MSE) of the CS-PR procedure given by $\EE{\norm{\thh_t-\ts}^2}$ is at most $O(\frac{1}{t})$.
+Recently, \cite{bach} considered SGD\footnote{SGD is a special of the LSA algorithm in \eqref{eq:lsaintro}.} with CS-PR for the linear prediction problem with quadratic loss and $i.i.d$ sampling.
+A ``remarkable'' result by \cite{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the expected squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
+\paragraph{Motivation and Contribution:} In this paper, we wish to study LSA with CS-PR for $i.i.d$ data with bounded variance.
+An additional restriction we place is that the matrix $\EEP{A_t}$ (where $A_t\sim P$ ) is Hurwitz, i.e., all its eigenvalues have positive real parts.
+For a given $P$, we wish to measure the performance in terms of the mean square error (MSE) $\EEP{\norm{\thh_t-\ts}^2}$.
+Further, for a given class of distributions $\P$, we also would like to know whether it is possible to ensure uniform performance $\forall P\in \P$.
+Our study is motivated by the problem \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp} of algorithms.
+We now list our contributions in this paper
+%While asymptotically diminishing step-sizes are common in RL literature \cite{ilstd,gtdmp,korda-prashanth}, the constant step-size choice has also been used in the past \cite{gtd2}.
+\begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
+\item \textbf{Instance Dependent Bounds} We show that for a fixed $P$, there exists a $\alpha_P>0$ such that the mean squared error (MSE) of the LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$.
 The MSE further can be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
- \item We also show (via counter example) that not all classes of general LSAs `admit' a constant step-size that is uniform across all the problems in the class.
-Further, we also argue that the subclass of LSAs related to linear prediction problem does `admit' a problem independent constant step size.
-\item We observe that when the constant step-size is chosen too big it results in instability, however, a conservative choice of the constant step-size affects only the constants in our error bounds without affecting the asymptotic rate (which might not be true with other popular choices of step-sizes).
-We aruge heuristically that CS-PR procedure can be deployed in a pricinpled manner by using an outer loop that automatically reduces that constant step-size by detecting detects the blow up of iterates.
+\item \textbf{Uniform Step-Sizes:} We also show (via counter example) that there is a class $\P$ that does not `admit' a constant step-size choice that guarantees uniform performance $\forall P\in \P$.
+We also argue that the subclass $\P_{LR}$ related to SGD for linear regression does `admit' a problem independent constant step size, thereby making a connection between our framework and that considered by \cite{bach}.
+\item \textbf{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion, and the performance of CS-PR is comparable to other popular step-size choices used for LSA in RL literature.
 \end{itemize}
+\begin{comment}
 \paragraph{Highlights:}
- We only assume that $\EE{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts.
-This is the condition used in prior works by \cite{polyak-judisky,borkarbook}.
-Eventhough our proof technique resembles that by \cite{bach}, in addition, we make use of Jordon/Cyclic decomposition in a critical way to handle the general Hurwitz assumption.
+The performance of CS-PR for general LSA has been previously studied under additive noise assumptions where $A_t=A$ for some fixed $A\in\R^{\dcd}$ and the noise is present only in $b_t$ \cite{polyak-judisky}.
+The setting we consider in this paper is known as the multiplicative noise setting, whose analysis requires tools that exploit the linear structure of the LSA \cite{bach}.
+However,
+ in the absence of special structures (such as symmetry of $A_t$),  we make use of Jordan/Cyclic decomposition in a critical way to handle the general Hurwitz assumption.
 Our results are applicable to wide variety of TD algorithms such as TD(0), GTD/GDT2, TDC.
-Further, the error dynamics has a simple interpretation in terms of behaviour of the sum of terms in a geometric series, and hence offers a simple and unified understanding of the various TD algroithms.
+Further, the error dynamics of the CS-PR technique for LSAs has a simple interpretation in terms of behaviour of the sum of terms in a geometric series, and hence offers a simple and unified understanding of its use in the various TD algorithms.
+An important difference between the linear prediction problem considered by \cite{bach} and the general LSA setting we consider is that in the former has certain special structures and the latter does not.
+One of the special structure exploited by \cite{bach} is the fact that $A_t$ are real symmetric matrices.
+The general LSA setting does not enjoy such an additional structure presents an interesting challenge.
+Further, CS-PR for general LSAs has been studied previous only under the additive noise setting \cite{polyal-judisky}.\par
+\end{comment}
+\paragraph{Organization:}
 \begin{comment}
-A host of algorithms in machine learning and in particular, the temporal difference class of algorithms arising in reinforcement learning, turn out to be special cases of \eqref{eq:sa}, namley linear stochastic approximation (LSA) algorithms.
+A host of algorithms in machine learning and in particular, the temporal difference class of algorithms arising in reinforcement learning, turn out to be special cases of \eqref{eq:sa}, namely linear stochastic approximation (LSA) algorithms.
 LSA algorithms assume the general form
@@
 Now the PR-average is given by $\frac{\norm{\theta_t-\ts}^2}{t^2}=\frac{\sum_{s=0}^{t-1}\rho^s}{t^2} \norm{\theta_0-\ts}^2$.
-Note that the numerator is just sum of terms of a geomteric series, which is summable as long as the matrix $(I-\alpha A)$ has eigen values within the unit circle.
-\item The fact that the setting covers even two timescale LSAs with constant step size, coupled with the intuitive interpretation of the error dynamics together present a simple and unified view of the various temporal difference learning algroithms in literature.
-\item One of the major bottlenecks in selecting step-size for general LSA is the lack of problem knoweldge.
+Note that the numerator is just sum of terms of a geometric series, which is summable as long as the matrix $(I-\alpha A)$ has eigen values within the unit circle.
+\item The fact that the setting covers even two timescale LSAs with constant step size, coupled with the intuitive interpretation of the error dynamics together present a simple and unified view of the various temporal difference learning algorithms in literature.
+\item One of the major bottlenecks in selecting step-size for general LSA is the lack of problem knowledge.
 One way to handle this issue is to choose a diminishing step size schedule such as $\alpha_t=\frac{1}{t}$.

### 2017-05-10 intro.tex
pruned the intro

@@
 \section{Introduction}
-\begin{comment}
-Stochastic Approximation (SA) algorithms can be expressed in the general form
-\begin{align}\label{eq:sa}
-\theta_t=\theta_{t-1}+\alpha_t f(\theta_{t-1},\zeta_t),
-\end{align}
-where $\theta_t\in \R^d$ are the iterates, and $f(\theta_{t-1},\zeta_t)$ of the function $g(\theta_{t-1})=\EE{f(\theta_{t-1})}_{\zeta_t}$, where the expectation is with respect to some noise process $\zeta_t$.
-At each time step, the algorithm updates the iterates in increments proportional to a step-size $\alpha_t>0$ and the idea is to eventually converge to $\ts\in \R^d$ which is a desired optimum or a fixed point.
-Due to their inherent online nature SA algorithms appear widely in machine learning applications \cite{konda,bhatnagar,borkar,gtd,td,bach}, where we have to learn from a noisy data stream.\par
-The choice of the step-size $\alpha_t$ needs to trade-off learning rate versus filtering noise, so to speak.
-In general, one can guess that larger step-size lead to faster learning and at the same time smaller step-sizes are better to handle noise.
-A popular choice of step-size schedule that is known handle the noise $\alpha_t=O(\frac{1}{t})$.
-However, since the step size are diminishing, it can hurt the learning rate.
-An alternative suggestion is to use more aggressive step sizes without minding the noise in the iterates $\theta_t$ in \eqref{eq:sa}, however, introduce additional averaging (known as Polyak-Rupert averaging) to filter the noise.
-The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$, and an analysis of the PR-averaging technique is known from \cite{}.\par
-\end{comment}
-
 A linear stochastic approximation (LSA) algorithm is given by
@@
 Diminishing and square-summable step-sizes have been a popular choice for the step-sizes, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$ \cite{gtd2,borkarbook,konda-tsitsiklis}.
-An alternative technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$.
+An alternative technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{s=0}^t \theta_t$.
 In LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
@@
 Our study is motivated by the problem \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp} of algorithms.
+A result on the performance of LSA with CS-PR can be directly translated for the TD class of algorithms that are also LSA algorithms.
 We now list our contributions in this paper
@@
 \item \textbf{Uniform Step-Sizes:} We also show (via counter example) that there is a class $\P$ that does not `admit' a constant step-size choice that guarantees uniform performance $\forall P\in \P$.
-We also argue that the subclass $\P_{LR}$ related to SGD for linear regression does `admit' a problem independent constant step size, thereby making a connection between our framework and that considered by \cite{bach}.
+This is a negative result in the context of RL.
+However,
+we also argue that the subclass $\P_{LR}$ related to SGD for linear regression does `admit' a problem independent constant step size, thereby making a connection between our framework and the one considered by \cite{bach}.
 \item \textbf{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion, and the performance of CS-PR is comparable to other popular step-size choices used for LSA in RL literature.
 \end{itemize}
-\begin{comment}
-\paragraph{Highlights:}
-The performance of CS-PR for general LSA has been previously studied under additive noise assumptions where $A_t=A$ for some fixed $A\in\R^{\dcd}$ and the noise is present only in $b_t$ \cite{polyak-judisky}.
-The setting we consider in this paper is known as the multiplicative noise setting, whose analysis requires tools that exploit the linear structure of the LSA \cite{bach}.
-However,
- in the absence of special structures (such as symmetry of $A_t$),  we make use of Jordan/Cyclic decomposition in a critical way to handle the general Hurwitz assumption.
-Our results are applicable to wide variety of TD algorithms such as TD(0), GTD/GDT2, TDC.
-Further, the error dynamics of the CS-PR technique for LSAs has a simple interpretation in terms of behaviour of the sum of terms in a geometric series, and hence offers a simple and unified understanding of its use in the various TD algorithms.
-An important difference between the linear prediction problem considered by \cite{bach} and the general LSA setting we consider is that in the former has certain special structures and the latter does not.
-One of the special structure exploited by \cite{bach} is the fact that $A_t$ are real symmetric matrices.
-The general LSA setting does not enjoy such an additional structure presents an interesting challenge.
-Further, CS-PR for general LSAs has been studied previous only under the additive noise setting \cite{polyal-judisky}.\par
-\end{comment}
 \paragraph{Organization:}
-\begin{comment}
-A host of algorithms in machine learning and in particular, the temporal difference class of algorithms arising in reinforcement learning, turn out to be special cases of \eqref{eq:sa}, namely linear stochastic approximation (LSA) algorithms.
-LSA algorithms assume the general form
- Further, in most instances of LSA $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be computed in $O(n)$, and are desired for this low per iteration computational cost.
-Recently, \cite{bach} used \eqref{eq:lsaintro} with constant step-size, i.e., $\alpha_t=\alpha,\,\forall t\geq 0, \, \alpha >0$ and PR-averaging for the linear prediction problem with squared loss and $i.i.d$ sampling.
-A ``remarkable'' result by \cite{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, he expected squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\todoc{Dimension?}
-
-In particular, as opposed to many earlier results available in the literature \todoc{Add citations}, $C$ does \emph{not} depend on the conditioning of the underlying system, while the total runtime of the algorithm scales linearly with both $t$ and $n$.\par
-In this paper, we want to understand the performance of constant step size with PR-average of iterates for general class of LSAs (outside of the linear prediction setting considered by \cite{bach}).
-We are interested in the problem of linear value function approximation in reinforcement learning using temporal difference learning (TD) either from experience replay in a batch setting or solving linear systems using TD-style algorithms.
-\todoc{TD-style, or just TD?} \todoc{Later we need to comment on TD with eligibility traces and other variations of TD, e.g., new variants.} We now list our contributions, and its implications to the TD class of algorithms, and the key differences between the techniques employed here in comparison to \cite{bach}.
-We now list our contributions, the implications of our results, and the challenges presented by the general setting in comparison to the linear prediction setting considered by \cite{bach}.
-\paragraph{Contributions}:
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item We show that in the case of general LSA, there exists instance dependent constant step size choice such that the mean squared error of the PR-average given by $\err_t=\EE{\norm{\thh_t-\ts}^2}$ is at most $O(\frac{1}{t})$.
-The error $\err_t$ further can be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
-\item We also show (via counter example) that not all classes of general LSAs `admit' a constant step-size that is uniform across all the problems in the class.
-Further, we also argue that the subclass of LSAs related to linear prediction problem does `admit' a problem independent constant step size.
-\item The assumptions in our setting is comparable to those in the asymptotic analysis based on ordinary differential equation (ODE) method by \cite{borkarmeyn,borkarbook}.
-\item Two timescale LSA algorithms are those that use two set of iterates (namely slower and faster) and two separate step-size schedules.
-Our results also hold for two timescale LSAs as long as the two different step-sizes are held constant with respect to time.
-Thus our setting not only holds for TD(0) but also extends to GTD/GTD2 and TDC.
-\item On an intuitive level, the error dynamics of the constant step-size plus PR-averaging procedure is similar to the behavior of partial sums of a geometric series.
-This coupled with the fact that the result also encompasses two timescale LSAs presents a simple and unified view of the various TD algorithms.
-\item The break up of error into bias and variance terms and their respective rates were previously worked out by \cite{bach} in the linear prediction setting, and here we extend the same to the general class of LSAs.
-Nevertheless, there are certain key differences between the general LSA setting that we consider to the setting considered by \cite{bach}, which are listed as under.
-\end{itemize}
-\end{comment}
-\begin{comment}
-\paragraph{Implications:}
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item Our results hold under the weak assumption\footnote{When this weak assumption fails to hold, the LSA is unstable and hence the notion of convergence is void.} that all the eigen values of $\EE{A_t}$ have strictly real parts.
-This assumption holds for all known TD class of RL algorithms.
-\item Two timescale LSA algorithms are those that use two set of iterates (namely slower and faster) and two separate step-size schedules.
-Our results also hold for two timescale LSAs as long as the two different step-sizes are held constant with respect to time.
-Thus our setting not only holds for TD(0) but also extends to GTD/GTD2 and TDC.
-This provides a unified view of the various temporal difference learning algorithms.
-\item The intuition behind the error analysis can be explained in simple terms when the noise free case is considered (i.e., $A_t=A$ in \eqref{eq:lsaintro}).
-In this case, the LSA is just a linear time invariant dynamical system where the error $\theta_t-\ts=(I-\alpha A)^t (\theta_0-\ts)$ and $\norm{\theta_t-\ts}=\rho^t\norm{\theta_0-\ts}$ (for some $0<\rho<1$).
-Now the PR-average is given by $\frac{\norm{\theta_t-\ts}^2}{t^2}=\frac{\sum_{s=0}^{t-1}\rho^s}{t^2} \norm{\theta_0-\ts}^2$.
-Note that the numerator is just sum of terms of a geometric series, which is summable as long as the matrix $(I-\alpha A)$ has eigen values within the unit circle.
-\item The fact that the setting covers even two timescale LSAs with constant step size, coupled with the intuitive interpretation of the error dynamics together present a simple and unified view of the various temporal difference learning algorithms in literature.
-\item One of the major bottlenecks in selecting step-size for general LSA is the lack of problem knowledge.
-One way to handle this issue is to choose a diminishing step size schedule such as $\alpha_t=\frac{1}{t}$.
-However, such a strategy can fail when the underlying matrices are ill-conditioned (see \Cref{sec:stepsizes}).
-Another popular approach has been to choose $\alpha_t=\frac{c}{c+t}$, where initially the step-sizes are held almost constant and then as $t$ increases they decay at rate $O(\frac{1}{t})$.
-While this approach can mitigate the ill-conditioning problem, it can at the same time amplify the effect of noise.
-In this paper, we argue informally that the constant step-size with PR-averaging of iterates can be a promising way to mitigate these issues, by periodically checking for possible blow up of the norm of the iterates and then reducing the step-size based on this information (see \Cref{sec:stepsizes}).
-\end{itemize}
-\paragraph{Challenges:} The break up of error into bias and variance terms and their respective rates were previously worked out by \cite{bach} in the linear prediction setting, and here we extend the same to the general class of LSAs.
-Nevertheless, there are certain key differences between the general LSA setting that we consider to the setting considered by \cite{bach}, which are listed as under.
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item In the linear least-squares problem is considered by \citet{bachharder} and the random matrices involved are known to be symmetric positive definite.
-This enables \citet{bachharder} to define operators that act between spaces of \emph{symmetric} matrices and carry out the computations making use of such operators.
-In our case, the matrices are known only to be (laxly) positive definite, i.e., they lack symmetry.
-\todoc{We should point out after our analysis in which step the symmetry helps \citet{bachharder}.
-We should also make this somehow clear here if possible.}
-Thus, the above approach does not work, and instead we have resort to an analysis that makes use of only the expected norms of the random matrices involved.
-\item Another significant difference is that in linear least-squares, the ``noise'' has a favorable ``structure''.
-Again, due to the lack of symmetry, this favorable structure does not apply.
-\todoc{Again, after the analysis, we should point out how the structured noise would have helped.
-I am wondering whether a more graceful degradation should happen?
-Is this really all or nothing?}
-\end{itemize}
-\end{comment}

### 2017-05-11 illustrate.tex
revisions

--- 2017-05-06 instance.tex
@@
-\section{Instance Dependent Bounds}
-In this section, we derive instance dependent bounds for $L_t$ in \eqref{eq:exploss} which is valid for a fixed $P$ and in the next section, explore the question of deriving uniform bounds for all $P\in \P$ (where $\P$ is a class of distributions over $\C^{\dcd}$).
-We first present the results and then follow it up with a discussion.
-\begin{lemma}\label{lm:pd}
-Let $P$ be a positive definite distribution over $\C^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
-\end{lemma}
-\begin{proof}
+\subsection{Illustrative Examples}
+In this section, we present some illustrative examples and make some useful remarks about the same.
+\begin{example}\label{ex:pdas}
+Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix},0<\epsilon<0.1$ are not positive definite.
+However choosing $U_1=\begin{bmatrix} \frac{1}{\sqrt{2}}& -\frac{1}{\sqrt{2}}\\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}} \end{bmatrix}$, $U_2=\begin{bmatrix} \frac{\sqrt{2}}{1-i\sqrt{3}} & \frac{\sqrt{2}}{1+i\sqrt{3}} \\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}}\end{bmatrix}$ and $U_3=\begin{bmatrix} 1 & 0\\ 0 &\frac{1}{10} \end{bmatrix}$, we can see that matrices $J_i=U_i^{-1}A_i U_i,i=1,2,3$ are positive definite.
+\end{example}
+The above example \Cref{ex:pdas} illustrates the use of appropriate $U_i,i=1,2,3$ matrices to transform Hurwitz problems into positive definite problems.
+In particular, problem $P^2$ involves the use of a complex matrix $U_2$.
+\begin{example}\label{ex:geo}
+Consider the one dimensional case where $A_t=\rho$ for some $0<\rho<1$, $b_t$ a bounded $i.i.d$ zero mean random variable with variance $\sigma^2$, and $\ts=0$.
+We know that $e_t=(1-\alpha\rho)^t e_0+\sum_{s=1}^t (1-\alpha \rho)^{t-s}\alpha b_s$ and the MSE is given by
 \begin{align*}
-\rhos{P}&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^*\EE{A_t^* A_t} x\\
-&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x
-&\geq \lambda_{\min}(A^*+A)-\alpha \norm{A_P}^2-\sigma^2_P
+\EE{\norm{\thh_t}^2}&=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2\\
+&\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)
 \end{align*}
-The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*+A)}{\norm{A_P}^2+\sigma^2_P}$
-\end{proof}
-
-\begin{lemma}\label{lm:simtran}[Similarity Transformation]
-Let $P^{AS}$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
-\end{lemma}
-
-\begin{theorem}\label{th:pdrate}
-Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
-\begin{align}
-\EE{\norm{\thh_t-\theta^*}^2}
-\leq
-\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
-\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
-%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
-\end{align}
-
-\end{theorem}
-
-\begin{theorem}\label{thm:simtran}[Change of Basis]
-Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be AS. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
-\begin{align}
-\EE{\norm{\gh_t-\gamma^*}^2}
-\leq
-\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
-\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
-%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
-%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
-\end{align}
-where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
-\end{theorem}
-\begin{corollary}\label{cor:as}
-Let $P$ in \Cref{lsadist} be asymptotically stable. Then, there exists
-\begin{align}
-\EE{\norm{\thh_t-\theta^*}^2}
-\leq
-\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U}^2\norm{U^{-1}}^2}{\alpha \rhos{P_U}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
-\end{align}
-\end{corollary}
-\subsection{Proof Ingredients}
-\paragraph{Error Recursion and Contraction Maps:}
-The output $\thh_t$ of the algorithm \eqref{eq:lsa} is the average of the internal states at times $s=0,\ldots,t-1$.
-Hence, before we proceed to interpret the results in \Cref{lm:pd,thm:pdrate,lm:simtran,thm:simtran,cor:as}, it is worthwhile to spend some effort in trying to understand the dynamics of the internal states by looking at the behavior of $e_t\eqdef \theta_t-\ts$.
-\begin{align}\label{eq:errec}
-\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1})\nn\\
-\theta_t-\ts&=\theta_{t-1}-\ts+\alpha\left(b_t-A_t(\theta_{t-1}-\ts+\ts)\right)\nn\\
-e_t&=e_{t-1}+\alpha(b_t-A_t e_t -A_t\ts)\nn\\
-e_t&=\underbrace{((I-\alpha A_t))}_{\text{Random-Matrix}} e_{t-1}+\underbrace{\alpha(N_t -(M_t)\ts)}_{\text{Noise}}
-\end{align}
-From \eqref{errec} it is clear that the error dynamics depends on the $i)$ properties of the random matrix, $ii)$ properties of the noise.
-In the absence of the noise term in \eqref{eq:errec}, we have $e_t=(I-\alpha A_t) e_{t-1}=\Pi_{s=1}(I-\alpha A_s) e_0$, i.e., the inital error is get multiplied by a product of random matrices.
-In such a scenario, we can guess that whether or not the algorithm forgets the bias $\norm{\theta_0-\ts}$, depends on whether the matrix product is contracting.
-One way to characterize this contracting property is to look at the spectral radius of the random matrix, which is given by
-\begin{align}\label{eq:spectralrand}
-\EE{\norm{I-\alpha A_t}^2}=1-\alpha \rhos{P}, {\norm{I-\alpha A_P}^2}=1-\alpha \rhod{P}
-\end{align}
-Notice that \eqref{eq:spectralrand} places our definitions of $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract}.
-Here the subscripts $s$ and $d$ stand for \emph{stochastic} and \emph{deterministic}, since they are associated with the spectral radius of stochastic and deterministic matrices respectively.\\
-\paragraph{Positive Definiteness:} By expanding \eqref{eq:spectralrand} we get
-\begin{align*}
-\EE{\norm{I-\alpha A_t}^2}=\sup_{x:\norm{x}=1} 1 -\alpha x^* (A_P+A_P^*)x +\alpha^2 \EE{A_t^* A_t},
-\end{align*}
-and thanks to \Cref{lm:pd} we know that when $A_P$ is positive definite it is possible to choose an $\alpha$ such that $\EE{\norm{I-\alpha A_t}}<1$.
-\paragraph{Similarity Transformation:} When $A_{P^{AS}}$ is known only to be AS, it is possible that there exists no choice of $\alpha>0$ for which $\rhos{P^{AS}}>0$ and $\rhod{P^{AS}}>0$ (see \Cref{ex:matrices}).
-However, it is potentially possible to re-write the recursion in a different basis via a similarity transformation $\gamma_t=U^{-1 }\theta_t$, and $z_t\eqdef U^{-1} e_t$
-\begin{align}\label{eq:errecsimtran}
-z_t&=\underbrace{(I-\alpha J_t)}_{\text{Transformed-Matrix}} z_{t-1}+\underbrace{\alpha U^{-1}(N_t -(M_t)\ts)}_{\text{Transformed-Noise}},
-\end{align}
-where $J_t\sim P_U$ and thanks to \Cref{lm:simtran} we know that $P_U$ is PD and from \Cref{lm:pd} we know that there exists $\alpha$ such that $\EE{\norm{I-\alpha J_t}}<1$.
-We note here that a similarity transformation does not result in \eqref{eq:errrecsimtran} being a diagonal system.
-However, \Cref{lm:simtran} side steps non-diagonizablility by taking refuge in Jordon decomposition theorem in linear algebra (see Proof of \Cref{lm:simtran} in the Appendix).
-\subsection{Result Discussion}
-We first discuss implications of \Cref{thm:pdrate}.
-\paragraph{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms.
-The first term is the bias term given by $\B=(1+4(\alpha\rho_{\alpha})^{-1}) (\alpha\rho_{\alpha})^{-1} \Big(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}\Big)$.
-The \emph{Bias} term that captures the rate at which the initial condition $\norm{\theta_t-\ts}^2$ is forgotten.
-The second term is the \emph{Variance} given by $\V=(1+4(\alpha\rho_{\alpha})^{-1}) (\alpha\rho_{\alpha})^{-1} \Big(\frac{\alpha^2\sigma^2}{t+1} \Big)$.
-The variance term that captures the rate at which noise is rejected.
-The $\B$ and $\V$ terms capture two different sources of errors.
-To see this, note that in the absence of noise i.e., when $\zeta_t=0,~\forall t\geq 0$, we have $\E[\norm{H\eb_t}^2]=\B$ and in the presence of noise, i.e., $\zeta\neq 0,\forall t\geq 0$ and the perfect initial condition, i.e., $\theta_t=\ts$, we have $\E[\norm{H\eb_t}^2]=\V$.
-\paragraph{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term.
-This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower.
-However, in this case the noise term does not blow up, a fact that can appreciated by looking at \eqref{lsergen} where $\alpha$ is seen to multiply the noise term $\zeta_t$.
-In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other.
-For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{\max}$, the bounds blow up again, due to the fact that $\rho_{\alpha}\ra 0$ in this case.
-This is due to the fact that the effect of both noise and initial conditions decay with the contraction factor of $F_{t,i}$, which gets closer to unity as $\alpha\ra\alpha_{\max}$.
+\end{example}
+The above example helps to interpret the error dynamics of \Cref{eq:lsa} in terms of sum of a geometric series with common factor $\alpha\rho$.
+\begin{example}
+Consider deterministic problems $P^1$, $P^2$ and $P^3$ with $A_{P^1}=\begin{bmatrix} 10 &100\\ 0 &10\end{bmatrix}$, $A_{P^2}=\begin{bmatrix} 10 &9.99\\ 0 &10\end{bmatrix}$, $A_{P^3}=\begin{bmatrix} 10 &0 \\ 0 &10\end{bmatrix}$, $\sigma^2_{A_{P^i}}=\sigma_{b_{P^i}}=0,\,i=1,2,3$.
+Note that $P^1$, $P^2$ and $P^3$ are respectively H, PD, SPD respectively.
+\end{example}
+\begin{comment}
+We now list the following observations:
+\begin{itemize}%[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\item In all the three instance, $0<\alpha<\frac2{10}$ ensures that the error $\norm{\eh_t}\ra 0$ as $t\ra\infty$. Thus the range of $\alpha$ is uniform across $P^i,i=1,2,3$.
+\item $\rhos{P^1}=\rhod{P^1}<0$, and hence one cannot apply results of \Cref{th:pdrate}.
+However, with  $U=\begin{bmatrix}1 &0 \\ 0 &0.2-\epsilon\end{bmatrix}$ for some $\epsilon>0$ we have $P^1_U$ to be the distribution concentrated on the matrix $A_{P^1_U}=\begin{bmatrix} 10 &20 \\ 0 &10\end{bmatrix}$ and one can apply the results of \Cref{thm:simtran,cor:as}.
+Notice that the condition number $\norm{U}{U^-1}\approx 5$.
+\end{itemize}
+\end{comment}
+\input{plts}

### 2017-05-11 instance.tex
revisions

@@
 In this section, we derive instance dependent bounds  which is valid for a fixed $P$ and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions over $\C^{\dcd}$.
-We first present the results and then follow it up with a discussion.
+Here, we only present the results followed by a discussion and the detailed proofs can be found in \Cref{sec:appendix}.
 \begin{lemma}\label{lm:pd}
-Let $P$ be a positive definite distribution over $\C^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+Let $P$ be positive definite, then there exists an $\alpha_P>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_P)$.
+\end{lemma}
+\begin{lemma}\label{lm:simtran}[Similarity Transformation]
+Let $P$ be Hurwitz, then there exists a $U\in \gln$ such that $P_U$ is positive definite.
 \end{lemma}
@@
 \end{comment}
-\begin{lemma}\label{lm:simtran}[Similarity Transformation]
-Let $P^{AS}$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
-\end{lemma}
 
+\begin{comment}
 \begin{theorem}\label{th:pdrate}
-Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
+Let $P^A$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
 \begin{align*}
@@
 \end{theorem}
-\begin{corollary}\label{cor:as}
-Let $P$ in \Cref{lsadist} be asymptotically stable. Then, there exists
+\end{comment}
+
+\begin{theorem}\label{th:rate}
+Let $P$ be Hurwitz, then there exists $U\in\gln$ and $\alpha_{P_U}>0$ such that
 \begin{align*}
@@
 \end{align*}
-\end{corollary}
+\end{theorem}
+
 \begin{theorem}[Lower Bound]
-There exists a $P$ such that
+There exists a $P$ Hurwitz, $\alpha_P>0$, such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in (0,\alpha_P)$ and
+\begin{align*}
+\EE{\norm{\eh_t}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big( \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \alpha^2 \sigma^2_b \Big)
+\end{align*}
+
+\begin{comment}
 \begin{align}
@@
 \end{align}
+\end{comment}
+\end{theorem}
 
-\end{theorem}
 \begin{comment}
+
 Consider the LSA with $b_t=0$, and $P^0$ has all the mass concentrated on $A_{P^0}=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$.
@@
 \begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} We can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{corss} terms of$\EE{\ip{e_i,e_j}},i\neq j$.
-The growth of the diagonal terms and the cross terms depends on the spectral norm of the matrices $I-\alpha A_t$ and $I-\alpha A_P$ respectively.
+\item \textbf{Role of $U$:} When the given problem $P$ is positive definite, it is straighforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$.
+However, the problem $P$ being Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}).
+This can be remedied by a similiarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$.
+To see this, let $\gamma_t\eqdef U^{-1 }\theta_t$ be a basis transformation, and let us re-write \eqref{conststep} in $\gamma_t$ by pre-multiplying \eqref{constste[} by $U^{-1}$ as shown below:
+\begin{align*}
+U^{-1}\theta_t&=U^{-1}\theta_{t-1}+\alpha(U^{-1} b_t- U^{-1}A_t U U^{-1}\theta_{t-1}),\\
+\gamma_t&=\gamma_{t-1}+\alpha(U^{-1} b_t- J_t\gamma_{t-1}),
+\end{align*}
+This completes the transformation of a Hurwirz problem $P$ to a positive definite problem $P_U$.
+One can revert to $\theta_t$ by using $\theta_t=U\gamma_t$, and this explains the reason for the condition number $\cond{U}$ in \Cref{th:rate}.
+We would like to add that when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}.
+We would like to mention that such similarity transformations have been in the past for analyzing LSA algorithms in RL \cite{lihong} (see \Cref{sec:related} for more details).
+\item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} When $P$ is positive definite, we can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{cross} terms of$\EE{\ip{e_i,e_j}},i\neq j$.
+The growth of the diagonal terms and the cross terms depends on the spectral norm of the random matrix $H_t=I-\alpha A_t$ and the deterministic matrix $H_P=I-\alpha A_P$ respectively.
 These are given by
@@
 \end{align}
-Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract} to the spectral norms of the aforementioned matrices.
+Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:dist} to the spectral norms matrices $H_t$ and $H_P$.
 For the MSE to be bounded, we need the spectral norms to be less than unity, which in turn translates to the requirement that $\rhos{P}>0$ and $\rhod{P}>0$.
-\item \textbf{Role of $U$:} When $A_P$ is positive definite, it is straighforward to produce an $\alpha>0$ such that $\rhos{P}>0$ and $\rhod{P}>0$.
-However, $A_P$ is Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}) and a result an $\alpha>0$ such that $\rhos{P}>0$ and $\rhod{P}>0$.
-This poses a difficulty in applying results of \Cref{th:pdrate} in the case Hurwitz matrices.
-This can be remedied by a similiarity transformation using an appropriate $U$ that transforms $A_P$ to appropriate Jordon form $A_{P_U}$.
-\Cref{th:simtran} presents the rates for the transformed iterates   $\gamma_t=U^{-1 }\theta_t$.
-Note that \Cref{cor:as} is obtained from \Cref{th:simtran} by the fact that $\theta_t=U\gamma_t$.
+If $P$ is Hurwitz, we can argue on similar lines by first transforming it into a positive definite problem $P_U$ and replacing $\rhos{P}$ and $\rhod{P}$ by $\rhos{P_U}$ and $\rhod{P_U}$, and introducing $\cond{U}$ to account for the forward ($\gamma=U^{-1}\theta$) and reverse ($\theta=U\gamma$) transformations using $U^{-1}$ and $U$ respectively.
+\item \textbf{Constants} {$\alpha$, $\rhos{P}$ and $\rhod{P}$} do not affect the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
 \item \textbf{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms.
@@
 The variance term that captures the rate at which noise is rejected.
-\item \textbf{$\alpha$, $\rhos{P}$ and $\rhod{P}$} only affect the constants in the rates and not the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
 \item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term.
 This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower.
-However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:pdrate,th:simtran,cor:as} where $\alpha$ is seen to multiply the variances $\sigma^2_P$ and $\sigma^2_b$.
+However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:rate} where $\alpha$ is seen to multiply the variances $\sigma^2_P$ and $\sigma^2_b$.
 In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other.
 For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.
-
-\subsection{Illustrative Examples}
-\begin{example}\label{ex:pdas}
-Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix}$ are not positive definite.
-However choosing $U_1=\begin{bmatrix} \frac{1}{\sqrt{2}}& -\frac{1}{\sqrt{2}}\\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}} \end{bmatrix}$, $U_2=\begin{bmatrix} \frac{\sqrt{2}}{1-i\sqrt{3}} & \frac{\sqrt{2}}{1+i\sqrt{3}} \\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}}\end{bmatrix}$ and $U_3=\begin{bmatrix} 1 & 0\\ 0 &\frac{1}{10} \end{bmatrix}$, we can see that matrices $J_i=U_i^{-1}A_i U_i,i=1,2,3$ are positive definite.
-\end{example}
-\paragraph{Remark}
-
-\begin{example}\label{ex:geo}
-Consider the one dimensional case where $A_t=\rho$ for some $0<\rho<1$, $b_t$ a bounded $i.i.d$ zero mean random variable with variance $\sigma^2$, and $\ts=0$.
-We know that $e_t=(1-\alpha\rho)^t e_0+\sum_{s=1}^t (1-\alpha \rho)^{t-s}\alpha b_s$ and the MSE is given by
-\begin{align*}
-\EE{\norm{\theta_t}^2}&=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2\\
-&\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)
-\end{align*}
-\end{example}
-\paragraph{Geometric Series:} In simple terms the error dynamics is akin to sum of a geometric series with common factor $\alpha\rho$, a fact that is illustrated by the simple $1$-dimensional example in \Cref{ex:geo}.
-\begin{example}
-Consider distributions $P^1$, $P^2$ and $P^3$ which are concentrated on matrices $A_{P^1}=\begin{bmatrix} 10 &100\\ 0 &10\end{bmatrix}$, $A_{P^2}=\begin{bmatrix} 10 &10\\ 0 &10\end{bmatrix}$, $A_{P^3}=\begin{bmatrix} 10 &0 \\ 0 &10\end{bmatrix}$.
-Note that $P^1$, $P^2$ and $P^3$ are respectively AS, PD, SPD respectively.
-Consider the LSA for the three case of $P^1$, $P^2$ and $P^3$ with $b_t=0$ and $\ts=0$
-We now list the following observations:
-\begin{itemize}%[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item In all the three instance, $0<\alpha<\frac2{10}$ ensures that the error $\norm{\eh_t}\ra 0$ as $t\ra\infty$. Thus the range of $\alpha$ is uniform across $P^i,i=1,2,3$.
-\item $\rhos{P^1}=\rhod{P^1}<0$, and hence one cannot apply results of \Cref{th:pdrate}.
-However, with  $U=\begin{bmatrix}1 &0 \\ 0 &0.2-\epsilon\end{bmatrix}$ for some $\epsilon>0$ we have $P^1_U$ to be the distribution concentrated on the matrix $A_{P^1_U}=\begin{bmatrix} 10 &20 \\ 0 &10\end{bmatrix}$ and one can apply the results of \Cref{thm:simtran,cor:as}.
-Notice that the condition number $\norm{U}{U^-1}\approx 5$.
+\item \textbf{Lower Bound} in \Cref{th:lb} shows that
 \end{itemize}
-We plot $L_t$ for the three cases for $t=1,\ldots,100$ and we can see that the condition number $\norm{U}{U^-1}\approx 5$, shows up in the AS case ($P^1$).
-\end{example}
-\end{itemize}
-\subsection{Related Work}
-LSA with CS-PR was considered in \cite{polyak-judisky} under the additive noise assumption, wherein $A_t=A$.
-Recent work by \cite{bach} considers a stochastic gradient descent (SGD) algorithm for the problem of linear prediction given in \Cref{ex:linreg}.
-\begin{example}\label{ex:linreg}
-Let $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
-Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$.
-Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$.
-SGD with CS-PR to minimize $f(\theta)$ can be given as below:
-\begin{subequations}\label{eq:linreg}
-\begin{align}
-\theta_t&=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1}} -y_t\right)\\
-\thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t
-\end{align}
-\end{subequations}
-Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$.
-It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}.
-In this case $A_P$ is SPD.
-In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
-\begin{align}\label{eq:mselinreg}
-f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
-\end{align}
-\end{example}
-In this setting, \cite{bach} show that the variance term in the MSE is at most $O(\frac{C}{t})$, where $C$ does not depend on the condition number.
-This ``remarkable'' result is due to the following facts: $(i)$ $A_P$ in this case is symmetric positive definite, $(ii)$ the MSE in \eqref{eq:mselinreg} is measured with respect to $A_P$, $(iii)$ the noise $\epsilon_t=y_t-\ip{\ts,x_{t-1}}$ is structured i.e., $\EE{\epsilon_t^2 x_t x_t^\top}\leq R A_P$ for some constant $R>0$.
-At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.
+\input{illustrate}

### 2017-05-11 intro.tex
revisions

@@
 where the iterates, $\theta_t\in \R^d$, and step-sizes, $\alpha_t>0$, are algorithmic quantities, and $(b_t,A_t)\in \R^d\times \R^{\dcd}$ is the noisy data sequence presented to the algorithm.
-LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence, using $O(d)$ updates per iteration.
-Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear prediction with quadratic loss \cite{bach,bachaistats}, and the temporal difference class (TD) of learning algorithms in RL \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.\par
+LSA algorithms (or simply LSAs) are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using only the noisy data sequence and $O(d)$ updates per iteration.
+Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear regression with quadratic loss \cite{bach,bachaistats}, and the temporal difference class (TD) of learning algorithms in RL \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.\par
 %An additional feature in these class of applications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
-It is of interest to understand the effect of the step-sizes $\alpha_t$ (a quantity that can be controlled by the algorithm designer) on the performance of the LSA in \eqref{eq:lsaintro}.
-The choice of the step-sizes dictates learning rate, noise rejection and stability.
+
+The choice of the step-sizes $\alpha_t$ is critical for the performance of the LSA in \eqref{eq:lsaintro}.
 Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster learning, while at the same time too large step-sizes might result in instability of \eqref{eq:lsaintro}.
-Diminishing and square-summable step-sizes have been a popular choice for the step-sizes, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$ \cite{gtd2,borkarbook,konda-tsitsiklis}.
-An alternative technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{s=0}^t \theta_t$.
+A useful choice in RL has been the diminishing and square-summable step-sizes \cite{gtd2,borkarbook,konda-tsitsiklis}, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$.
+An alternate technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{s=0}^t \theta_t$.
 In LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
-Averaging in stochastic approximation to improve noise rejection is a procedure that goes back to  Polyak and Ruppert \cite{polyak-judisky,ruppert}.
-%In what follows, when we say CS-PR for LSA or LSA with CS-PR, we allude to the case where \eqref{eq:lsaintro} is a constant step-size LSA with $\alpha_t=\alpha>0,\forall t\geq 0$, $\theta_t$ in \eqref{eq:lsaintro} is considered an internal variable and $\thh_t$ is the final output of the LSA with CS-PR.
-%In LSA with CS-PR, \eqref{eq:lsaintro} is a constant step-size LSA with $\alpha_t=\alpha>0,\forall t\geq 0$, $\theta_t$ in \eqref{eq:lsaintro} is considered an internal variable and $\thh_t$ is the final output of the LSA with CS-PR.
+The idea of averaging in stochastic approximation to improve noise rejection is a technique that goes back to  \citet{ruppert,polyak-judisky}.
 Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection.
 \par
-Recently, \cite{bach} considered SGD\footnote{SGD is a special of the LSA algorithm in \eqref{eq:lsaintro}.} with CS-PR for the linear prediction problem with quadratic loss and $i.i.d$ sampling.
-A ``remarkable'' result by \cite{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the expected squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
-\paragraph{Motivation and Contribution:} In this paper, we wish to study LSA with CS-PR for $i.i.d$ data with bounded variance.
-An additional restriction we place is that the matrix $\EEP{A_t}$ (where $A_t\sim P$ ) is Hurwitz, i.e., all its eigenvalues have positive real parts.
-For a given $P$, we wish to measure the performance in terms of the mean square error (MSE) $\EEP{\norm{\thh_t-\ts}^2}$.
-Further, for a given class of distributions $\P$, we also would like to know whether it is possible to ensure uniform performance $\forall P\in \P$.
-Our study is motivated by the problem \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp} of algorithms.
-A result on the performance of LSA with CS-PR can be directly translated for the TD class of algorithms that are also LSA algorithms.
+Recently, \citet{bach} considered SGD\footnote{SGD is a special of the LSA algorithm in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and $i.i.d$ sampling.
+A ``remarkable'' result by \citet{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the expected squared regression error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
+
+\paragraph{Motivation and Contribution:} In this paper, we wish to study LSA with CS-PR where data $(b_t,A_t)$ is $i.i.d$ with some underlying distribution $P$ such that variance is bounded.
+An additional restriction we place on $P$ is that the matrix $\EEP{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts.
+For a given $P$, we wish to measure the performance of LSA with CS-PR in terms of the mean square error (MSE) $\EEP{\norm{\thh_t-\ts}^2}$.
+Further, for a given class of distributions $\P$, we would like to know whether it is possible to ensure uniform performance $\forall P\in \P$.
+Our study is motivated by the problem \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class  of algorithms \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.
+Here, we wish to understand the effectiveness of the CS-PR technique for the TD class of LSA algorithms used in RL.
+
 We now list our contributions in this paper
 %While asymptotically diminishing step-sizes are common in RL literature \cite{ilstd,gtdmp,korda-prashanth}, the constant step-size choice has also been used in the past \cite{gtd2}.
-\begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{Instance Dependent Bounds} We show that for a fixed $P$, there exists a $\alpha_P>0$ such that the mean squared error (MSE) of the LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$.
-The MSE further can be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
-\item \textbf{Uniform Step-Sizes:} We also show (via counter example) that there is a class $\P$ that does not `admit' a constant step-size choice that guarantees uniform performance $\forall P\in \P$.
+%\begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
+ \begin{center}Algorithm\end{center}
+\paragraph{Instance Dependent Bounds} We show that for a fixed $P$, there exists a $\alpha_P>0$ such that the mean squared error (MSE) of the LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$.
+The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
+\begin{center} Structure in Data\end{center}
+\paragraph{Uniform Step-Sizes:} We also show (via counter example) that there is a class $\P$ that does not `admit' a single constant step-size $\alpha_{\P}$ that guarantees uniform performance $\forall P\in \P$.
 This is a negative result in the context of RL.
 However,
-we also argue that the subclass $\P_{LR}$ related to SGD for linear regression does `admit' a problem independent constant step size, thereby making a connection between our framework and the one considered by \cite{bach}.
-\item \textbf{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion, and the performance of CS-PR is comparable to other popular step-size choices used for LSA in RL literature.
-\end{itemize}
+we also argue that the subclass $\P_{LR}$ related to SGD for linear regression does `admit' a problem independent constant step-size, thereby recovering a part of the result by \citet{bach}.
+\begin{center} Meta-Parameter $\alpha$\end{center}
+\paragraph{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion, and the performance of CS-PR is comparable to other popular step-size choices used for LSA in RL literature.
 \paragraph{Organization:}

### 2017-05-11 not.tex
revisions

@@
 \section{Notations and Definitions}\label{sec:def}
+We denote the sets real and complex numbers by $\R$ and $\C$ respectively. For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively.
+For any $x\in \C$ we denote its modulus $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$ and its complex conjugate by $\bar{x}=a-ib$.
+We denote $d$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively.
+We denote the transpose of $x\in \C^{d}$ by $x^\top$ and the conjugate transpose by $x^*={\bar{x}}^\top$.
+We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^* y$.
+%We denote the transpose of $x\in \C^{d}, A\in \C^{\dcd}$ by $x^\top$ and $A^\top$ respectively.
+%Further, by $x^*$ denotes the conjugate transpose, i.e., $x^*=\bar{x}^\top$ (the same convention carries over to matrices as well).
 
-We denote the sets real and complex number by $\R$ and $\C$ respectively.
-For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively.
-We denote $n$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively.
-We use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively.
-For any $x\in \C$ we denote the modulus of $x$ by $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$.
-For any $x\in \C^d$We denote complex conjugate of $x$ by $\bar{x}$.
-We denote the transpose of $x\in \C^{d}, A\in \C^{\dcd}$ by $x^\top$ and $A^\top$ respectively.
-Further, by $x^*$ denotes the conjugate transpose, i.e., $x^*=\bar{x}^\top$ (the same convention carries over to matrices as well).
-We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^* y$.
-Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
+ %Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
+We use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively. The notation for (conjugage) transpose carries over from vectors.
 The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^d:\norm{x}=1} \norm{Ax}$.
-For a positive real number $B>0$, we denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \norm{A}\leq B,\,\forall i,j=1,\ldots,n\}$ the set of complex and real matrices whose norms are bounded $B$.
-%We denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
-%We use $\P$ to denote the set of probability distribution over $\C^{\dcd}$.
-We also denote the set of invertible $\dcd$ complex matrices by $\gln$.
-We use $A\sim P$ to denote the fact that the random variable $A$ is distributed according to distribution $P$.
-Let the random matrix $A\sim P$, and $U\in \gln$, we use $P_U$ to denote the distribution of the random matrix $U^{-1}A U$.
-We denote the identity matrix in $\C^{\dcd}$ by $\I$.
-%We denote the set of distributions over $\C_B^{\dcd}$ by $\P$.
-
+For any $x\in\C^d$, we denote the general quadratic norm with respect to a hermitian matrix $C$ by $\norm{x}^2_C\eqdef x^*\, C \,x$.
+We use $\cond{A}=\norm{A}\norm{A^{-1}}$ to denote the condition number of the matrix $A$.
+We denote the identity matrix in $\C^{\dcd}$ by $\I$ and the set of invertible $\dcd$ complex matrices by $\gln$.
+\begin{comment}
 We use $A\succeq 0$ to denote that the
@@
 We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
-We also use $\preceq$ and $\prec$ analogously. For any $x\in\C^d$x, we denote the general quadratic norm with respect to a $C\succ 0$ by $\norm{x}^2_C\eqdef x^*\, C \,x$.
+We also use $\preceq$ and $\prec$ analogously.
+\end{comment}
+For a positive real number $B>0$, we denote by $\C^{d}_B=\{b\in \C^d\mid \norm{b}\leq B\}$ and by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ the set of complex vectors and matrices whose norms are bounded $B$.
+%We denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
+%We use $\P$ to denote the set of probability distribution over $\C^{\dcd}$.
+%We denote the set of distributions over $\C_B^{\dcd}$ by $\P$.
+Let $P$ be a distribution over $C^{\dcd}$, we use $A\sim P$ to denote $A$ is distributed according to $P$.
+For $U\in \gln$, we denote by $P_U$ the distribution of $U^{-1}\,A\,U$, where $A\sim P$.
 $\E$ denotes mathematical expectation.
-
-\begin{definition}
-Let $P$ be a distribution over $\C^{\dcd}$, then define
+\begin{comment}
+ We use $A\sim P$ to denote the fact that the random variable $A$ is distributed according to distribution $P$.
+Let the random matrix $A\sim P$, and $U\in \gln$, we use $P_U$ to denote the distribution of the random matrix $U^{-1}A U$.
+\end{comment}
+\begin{definition}\label{def:dist}
+Let $P=(P^V,P^M)$ denote a $2$-tuple of probability distributions; $P^V$ over $\C^{d}$ and $P^M$ over $\C^{\dcd}$. Define
 \begin{align*}
-A_P=\int M dP(M),\quad C_P=\int M^* M dP(M),
-\end{align*}
-\end{definition}
-\begin{definition}\label{def:contract}[Contraction Factors]
-Define
-\begin{align*}
+A_P&=\int M dP^M(M),\quad C_P=\int M^* M dP^M(M), \quad b_P=\int V dP^V(V)\\
 \rhod{P}&\eqdef {\inf}_{x\in\C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha A_P^* A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha C_P\right)x}
@@
 Note that $\rhod{P}>\rhos{P}$.
+\begin{definition}\label{def:simdist}
+Let $P=(P^V,P^M)$ as in \Cref{def:dist}; $b\sim P^V$ and $A\sim \P^M$ be random variables distributed according to $P^V$ and $P^M$.
+For $U\in \gln$ define $P_U=(P_U^V,P_U^M)$ to be the distributions of $U^{-1}b$ and $U^{-1}AU$.
+\end{definition}
+\begin{definition}
+We call a matix $A\in \C^{\dcd}$ to be \emph{Hurwitz} (H) if all eigenvalues of $A$ have positive real parts.
+We call a matrix $A\in \C^{\dcd}$ to be \emph{Positive Definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$.
+We call a matrix $A\in \C^{\dcd}$ to be \emph{Symmetric Positive Definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
+\end{definition}
+\begin{definition}\label{distpd}
+We call a distribution $P$ to be H/PD/SPD if $A_P$ is H/PD/SPD.
+\end{definition}
+\begin{example}
+The matrices $\begin{bmatrix}0.1 &-1\\ 1 & 0.1\end{bmatrix}$, $\begin{bmatrix} 0.1 & 0.1 \\ 0 & 0.1\end{bmatrix}$ and $\begin{bmatrix}0.1 &0 \\ 0 & 0.1\end{bmatrix}$ are examples of H, PD and SPD respectively.
+\end{example}
 
 \begin{definition}
-We call a matrix $A\in \C^{\dcd}$ to be \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$.
-We call a matix $A\in \C^{\dcd}$ to be \emph{asymptotically stable} (AS) if all eigenvalues of $A$ have positive real parts.
-We call a matrix $A\in \C^{\dcd}$ to be \emph{symmetric positive definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
+Call a set of distributions $\P=(\P^V,\P^M)$ over $\C^{d}\times \C^{\dcd}$
+\emph{weakly admissible} if there exists $\alpha_{\P}>0$ such that
+$\rhos{P}>0,\forall \alpha\in(0,\alpha_{\P})$ holds for all $P\in \P$.
 \end{definition}
-\begin{definition}\label{distpd}
-We call a distribution $P$ to be PD/AS/SPD if $A_P$ is PD/AS/SPD.
+\begin{definition}
+Call a set of distributions $\P=(\P^V,\P^M)$ over $\C^{d}\times \C^{\dcd}$ \emph{admissible}
+if there exists some $\alpha_{\P}>0$ such that $\inf_{P\in \P} \rhos{P}>0,\,\forall \alpha\in(0,\alpha_{\P})$.
+The value of $\alpha_{\P}$ is called a witness.
 \end{definition}
-\begin{example}
-The matrices $\begin{bmatrix}0.1 &-1\\ 1 & 0.1\end{bmatrix}$, $\begin{bmatrix} 0.1 & 0.1 \\ 0 & 0.1\end{bmatrix}$ and $\begin{bmatrix}0.1 &0 \\ 0 & 0.1\end{bmatrix}$ are examples of AS, PD and SPD respectively.
-\end{example}
+
+It is easy to see that $\alpha \mapsto \rhos{P}$ is decreasing,
+hence if $\alpha_{\P}>0$ witnesses that $\P$ is (weakly) admissible
+then any $0<\alpha'\le \alpha_{\P}$ is also witnessing this.

### 2017-05-11 plts.tex
revisions

\FloatBarrier
\begin{figure}[htp]
\resizebox{\columnwidth}{!}{
\begin{tabular}{ccc}
\begin{tikzpicture}[scale=0.3,font=\Large]
    \begin{axis}[
        xlabel=$t$,
         ylabel=$\norm{\thh_t-\ts}^2$, legend style={at={(0.5,-0.1)},anchor=north}
]
    \addplot[only marks,mark=square,red] plot file {./exp/as};
    \end{axis}
    \end{tikzpicture}
&
\begin{tikzpicture}[scale=0.3,font=\Large,]
\begin{axis}[
xlabel=$t$,
ylabel=$\norm{\thh_t-\ts}^2$, legend style={at={(0.5,-0.1)},anchor=north}
]
\addplot[only marks,mark=square,red] plot file {./exp/pd};
\end{axis}
\end{tikzpicture}

&
\begin{tikzpicture}[scale=0.3,font=\Large]
\begin{axis}[
xlabel=$t$,
ylabel=$\norm{\thh_t-\ts}^2$, legend style={at={(0.5,-0.1)},anchor=north}
]
\addplot[only marks,mark=square,red] plot file {./exp/spd};
\end{axis}
\end{tikzpicture}
\end{tabular}
}
\end{figure}

### 2017-05-11 prob.tex
revisions

@@
 \section{Problem Setup}
-We consider the linear stochastic approximation (LSA) algorithm with constant step-size (CS) and Polyak-Ruppert (PR) averaging of the iterates.
+We consider the linear stochastic approximation (LSA) algorithm with constant step-size (CS) and Polyak-Ruppert (PR) averaging of the iterates given as below:
 \begin{subequations}\label{eq:lsa}
@@
 The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
-can be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.\par
+can then be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.\par
 We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa}.
-In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.
+In what follows, we make use of the $\sigma$-field given by $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.
 \begin{assumption}\label{assmp:lsa}
 \begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \label{dist} $\{A_t\in\C^{\dcd},\, t\geq 0\}$ is an $i.i.d$ sequence distributed according to $P$ which is either AS or PD or SPD.
-\item $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^d$ with mean $\E[b_t]=b$.
-q\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ associated with $A_t$ and $b_t$ satisfy the following
-%\begin{align*}\E\left[ M_t^* M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^* N_t]=\sigma_b^2.\end{align*}
-\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^* N_t]=\sigma_b^2.\end{align*}
-\item $A_P$ is invertible and there exists a $\ts=A_P^{-1}b$.
+\item \label{dist} $(b_t, A_t)\sim (P^b,P^A)$ is an $i.i.d$ sequence, where $P^b$ is a distribution over $\C^d$ and $P^A$ is a Hurwitz distribution over $\C^{\dcd}$.
+\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following
+\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_{A_P}, \, \E[N_t^* N_t]=\sigma^2_{b_P}.\end{align*}
+\item $A_P$ is invertible and there exists a $\ts=A^{-1}_Pb_P$.
 \end{enumerate}
@@
 %For the rest of the paper we denote the errors of the internal and the output variables of \eqref{eq:lsa} at time $t$ by $e_t\eqdef \theta_t-\ts$ and $\eh_t\eqdef \thh_t-\ts$ respectively.
-\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ We note here that $\EE{\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$.
+\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ $\EE{\norm{\thh_t-\ts}_C^2}$, the MSE with respect a hermitian matrix $C$ is desirable, we cannot exploit the use of $C$ unless it is connected to $P$ in a special way.
+For more discussion, see \Cref{sec:related}.}
+ by $\EE{\norm{\thh_t-\ts}^2}$.
+\begin{comment}
+\footnote{ We note here that $\EE{\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$.
 However, unless $C$ has some special structure, it is not possible to exploit this generality in our bounds.
 Further, when data is bounded $\norm{C}^2$ is also bounded and it is straightforward to see that $\EE{\norm{\thh_t-\ts}}^2_C\leq \norm{C}^2\EE{\norm{\thh_t-\ts}^2}$.
-So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in the linear regression setting, where its usage is relevant.} by $\EE{\norm{\thh_t-\ts}^2}$.
-We wish to understand how the MSE is affected by the choice of $\alpha$ and the spectral properties of $A_P$.
+So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in the linear regression setting, where its usage is relevant.}
+\end{comment}
+%We wish to understand how the MSE is affected by the choice of $\alpha$ and the spectral properties of $A_P$.
 Another aspect of interest is to understand the contribution of the \emph{bias} (caused by the initial condition $\theta_0$) and the \emph{variance} (caused by $\sigma^2_P,\sigma^2_b$) to the MSE.\par

### 2017-05-11 related.tex
revisions

@@
 \section{Applications in Reinforcement Learning}
+\subsection{Related Work}
+Make this more detailed by splitting into various parts.
+\begin{comment}
+LSA with CS-PR was considered in \cite{polyak-judisky} under the additive noise assumption, wherein $A_t=A$.
+Recent work by \cite{bach} considers a stochastic gradient descent (SGD) algorithm for the problem of linear prediction given in \Cref{ex:linreg}.
+\begin{example}\label{ex:linreg}
+Let $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
+Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$.
+Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$.
+SGD with CS-PR to minimize $f(\theta)$ can be given as below:
+\begin{subequations}\label{eq:linreg}
+\begin{align}
+\theta_t&=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1}} -y_t\right)\\
+\thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t
+\end{align}
+\end{subequations}
+Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$.
+It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}.
+In this case $A_P$ is SPD.
+In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
+\begin{align}\label{eq:mselinreg}
+f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
+\end{align}
+\end{example}
+In this setting, \cite{bach} show that the variance term in the MSE is at most $O(\frac{C}{t})$, where $C$ does not depend on the condition number.
+This ``remarkable'' result is due to the following facts: $(i)$ $A_P$ in this case is symmetric positive definite, $(ii)$ the MSE in \eqref{eq:mselinreg} is measured with respect to $A_P$, $(iii)$ the noise $\epsilon_t=y_t-\ip{\ts,x_{t-1}}$ is structured i.e., $\EE{\epsilon_t^2 x_t x_t^\top}\leq R A_P$ for some constant $R>0$.
+At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.
+\end{comment}
 \paragraph{Definitions:} We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
@@
 \end{table}
+
+\begin{itemize}
+\item It follows that for each instance of blah blah
+\end{itemize}
+
 \paragraph{Related Work in RL} The various temporal difference learning algorithms can be written as in \Cref{table:td}.

### 2017-05-11 uniform.tex
revisions

@@
 \section{Uniform stepsizes}
-\begin{definition}
-Call a set of distributions $\P$ over $\C^{\dcd}$
-\emph{weakly admissible} if there exists $\alpha>0$ such that
-$\rhos{P}>0$ holds for all $P\in \P$.
-\end{definition}
-\begin{definition}
-Call a set of distributions $\P$ over $\C^{\dcd}$ \emph{admissible}
-if there exists some $\alpha>0$ such that $\inf_{P\in \P} \rhos{P}>0$.
-The value of $\alpha$ is called a witness.
-\end{definition}
-It is easy to see that $\alpha \mapsto \rhos{P}$ is decreasing,
-hence if $\alpha>0$ witnesses that $\P$ is (weakly) admissible
-then any $0<\alpha'\le \alpha$ is also witnessing this.
-
-If $\P$ is weakly admissible, then one can choose a stepsize $\alpha>0$ solely based on the knowledge of $\P$ and
-conclude that no matter the joint distribution of $(A_t,b_t)$, as long as the distribution of $A_t$ is an element of $\P$, the conclusions of \cref{maintheorem} hold.
+If $\P$ is weakly admissible, then one can choose a stepsize $\alpha_{\P}>0$ solely based on the knowledge of $\P$ and
+conclude that no matter the joint distribution $P$ of $(b_t,A_t)$, as long as the distribution of $P$ is an element of $\P$, the conclusions of \Cref{th:rate} hold.
 Further, if $\P$ is admissible then the error bound stated in  \cref{maintheorem} becomes independent of the instance.
 Hence, an interesting question to investigate is whether a given set $\P$ is (weakly) admissible.
-A hopeful set is the following:
-For $B>0$, let $\P_B$ be the set of distributions over $\C_B^{\dcd}$ where for every $P\in \P_B$ is positive definite.
-In what follows all distributions will be a subset of $\P_B$ with some $B>0$,
-Note that in most applications the knowledge of a suitable $B>0$ such that $\P\subset \P_B$ is available.
-By appropriately normalizing the matrices involved, we may as well assume that $B=1$.
+We consider some reasonable classes of distributions and test whether or not they are (weakly) admissible.
+\begin{definition}
+\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\item For $B>0$, let $\P_B$ be the set of distributions over $\C_B^{d}\times \C_B^{\dcd}$ where for every $P\in \P_B$ is positive definite.
+\item  $\P_{\text{PSD},B}$ be such that  $\P^M_{\text{PSD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PSD} \}\,$ , where PSD denotes the set of positive semi-definite matrices.
+\end{itemize}
+\end{definition}
+Note that all distributions are a subset of $\P_B$ and in most applications the knowledge of a suitable $B>0$ is available,and by appropriately normalizing the vectors and matrices involved, we may as well assume that $B=1$ without any loss of generality.
+\begin{proposition}\label{notwad}
+The set $\P_B$ is not weakly admissible.
+\end{proposition}
+\begin{proposition}
+For any $B>0$, the set $\P_{\text{PSD},B}$ is weakly admissible
+and in particular any $0<\alpha < 2/B$ witnesses this.
+\end{proposition}
+\begin{proposition}
+For any $B>0$, the set $\P_{\text{PSD},B}$ is not admissible.
+\end{proposition}
 
-Our first result shows that we cannot be too ambitious:
-\begin{proposition}\label{notwad}
-The set $\P_1$ is not weakly admissible.
-\end{proposition}
+\begin{comment}
 \begin{proof}
 Fix an arbitrary $\alpha>0$. We show that there exists $P\in \P$ such that $\rho_\alpha(P)<0$.
-For $\epsilon \in (0,1/2)$ let $P$ be the distribution that is supported on $\{-I,I\}$ and takes on the value of $I$ with probability $1/2+\epsilon$.
+For $\epsilon \in (0,1/2)$ let $P=(P^V,P^M)$ be the distribution such that $P^M$ is supported on $\{-I,I\}$ and takes on the value of $I$ with probability $1/2+\epsilon$.
 Then $A_P = 2\epsilon I \succ 0$, hence $P\in \P_1$.
@@
 \end{proof}
-Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
-In linear least-squares problems (cf. \cref{ex:leastsquares}) the set $\P$ has more structure.
-In particular, the random matrices $\{A_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
-Define
-\begin{align*}
-\P_{\text{PSD},B} = \{ P \in \P_B \,:\,  \supp(P)\subset \text{PSD} \}\,.
-\end{align*}
-We have the following proposition:
-\begin{proposition}
-For any $B>0$, the set $\P_{\text{PSD},B}$ is weakly admissible
-and in particular any $0<\alpha < 2/B$ witnesses this.
-\end{proposition}
+\end{comment}
+
+\begin{comment}
 \begin{proof}
@@
 \end{proof}
+\end{comment}
 \todoc[inline]{So this altogether is weaker than what Bach founds. They prove strong admissibility.
@@
 }
+Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility. When we impost
+In linear least-squares problems (cf. \cref{ex:leastsquares}) the set $\P$ has more structure.
+In particular, the random matrices $\{A_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
+We have the following proposition:

### 2017-05-13 linreg.tex
refreshing

--- 2017-05-11 related.tex
@@
-\section{Applications in Reinforcement Learning}
-\subsection{Related Work}
-Make this more detailed by splitting into various parts.
 \begin{comment}
@@
 \end{comment}
-\paragraph{Definitions:} We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
-Let $\gamma \in (0,1)$ be a fixed constant, we define $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$.
-In what follows, $\mu_t>0$ is an \emph{importance} sampling factor whose aim is to correct for mismatch in the (behavior) distribution with which the data was collected and the (target) distribution with respect to which one wants to learn.
-A factor $\mu_t=1,\,\forall t\geq 0$ will mean that no correction is required\footnote{This is known as the \emph{on-policy} case where the behavior is identical to the target.
-The general setting where $\mu_t>0$ is known as \emph{off-policy}.}.
-\begin{table}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|c|c|c|}\hline
-Algorithm& Update &Stepsize &Rate \\ \hline
-TD(0)& $\begin{aligned}\theta_t&=\theta_{t-1}+\alpha_t (b_t -\Delta_t\theta_{t-1})\\ \thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t\end{aligned}$ &$\begin{aligned}\alpha_t&=O(\frac1{t})^\delta,\\ &\delta\in(1/2,1)\end{aligned}$ &$\EE{\norm{\eh_t}}\leq O(\frac{1}{\sqrt{t}})$ for $\delta\ra 1$ \\ \hline
-GTD& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}- Q_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t A^\top_t y_{t-1})\end{aligned}$ & $\begin{aligned}&\beta_t=\eta \alpha_t,\sum_{t\geq 0}\alpha_t=\infty, \\ &\sum_{t\geq 0}\alpha^2_t<\infty \end{aligned}$ & $e_t\ra 0$ w.p.1 \\\hline
-TDC& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t b_t-\mu_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$ & $\begin{aligned}&\beta_t,\alpha_t,\frac{\alpha_t}{\beta_t} \ra 0,~\text{as}~t\ra \infty,\\& \sum_{t\geq 0}\alpha_t=\infty, \sum_{t\geq 0}\beta_t=\infty, \\ &\sum_{t\geq 0}\alpha^2_t<\infty, \sum_{t\geq 0}\beta^2_t<\infty \end{aligned}$ & $e_t\ra 0$ w.p.1 \\\hline
-GTD& $\begin{aligned} y_{t}&=\Pi_Y \left(y_{t-1}+\alpha_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-Q_t y_{t-1})\right)\\ \theta_t &=\Pi_{\Theta}\left(\theta_{t-1}+\alpha_t (\mu_t A_t^\top y_{t-1})\right)\\ \thh_t&=\frac{\sum_{s=0}^t \alpha_s\theta_s}{\sum_{s=0}^t \alpha_s}\end{aligned}$ & $\alpha_t =O(\frac{1}{\sqrt{t}})$ & $\norm{A_P\eh_t-b}^2 \leq O(\frac{1}{\sqrt{t}})$ w.h.p\\\hline
-
-\end{tabular}
-}
-\label{tb:rates}
-\end{table}
-
-\begin{itemize}
-\item It follows that for each instance of blah blah
-\end{itemize}
-
-\paragraph{Related Work in RL} The various temporal difference learning algorithms can be written as in \Cref{table:td}.
-The TD(0) algorithm is an on-policy algorithm ($\mu_t=1$) and is unstable in the general {off-policy} setting.
-The gradient temporal difference (GTD) algorithm \cite{gtd,gtd2,gtdmp} is stable in off-policy setting as well.
-Recent work by \cite{korda-prashanth} derive finite time bounds for the TD(0) for two different cases of step-sizes one without PR averaging and other with PR averaging.
-In the case without PR averaging, \cite{korda-prashanth} consider step-sizes of the form $\frac{c_0c}{c+t}$, where $c_0$ and $c$ are problem dependent positive constants, to dervie a rate of $O(\frac{1}{t})$ for the MSE.
-In the case with PR averaging (row $1$ of \Cref{tb:rates}), \cite{korda-prashanth} consider step-size of the from $c_0\left(\frac{c}{c+t}\right)^\delta,\,\delta \in (1/2,1)$ where $c_0, c$ are problem independent positive constants, and show that this choice yeilds a rate of $O(\frac{1}{t})$ for $\delta \ra 1$ and $t\geq t_0$ (where $t_0$ is a problem dependent time instant).
-A possible limiting feature of the results by \cite{korda-prashanth} is that either the problem knowledge is required or the bounds hold only after a problem dependent time.
-A centered TD(0) variant is also considered by \cite{korda-prashanth}, however, a projection operator is required to keep the iterates bounded.
-Also, the bounds by \cite{korda-prashanth} don't have an explicit bias variance decomposition.
-The works on GTD and TDC by \cite{gtd,gtd2} show only asymptotic convergence w.p.1.
-Recently, the work by \cite{gtdmp} shows a rate of $\frac{1}{\sqrt{t}}$ for GTD with use of projection operation, and also proposed a variant namely GTD-Mirror-prox based on stochastic mirror-prox ideas which achieves $\frac{1}{\sqrt{t}}$ rate without use of projection.
-In comparison, we don't use of projections, our bounds hold for all time $t$, and the bounds are dependent on the problem instance only to the extent of choosing a step-size that leads to stable behavior, in which case we show that $O\frac{1}{t}$ rate holds (the scaling factor hidden in the $O$ expression is problem dependent).
-Further, our results provide explicit bias variance decomposition.

### 2017-05-13 related.tex
refreshing

@@
-\section{Applications in Reinforcement Learning}
-\subsection{Related Work}
-Make this more detailed by splitting into various parts.
+\section{Related Work}
+We first discuss the related work outside of RL setting, followed by related work in the RL setting. In both cases, we highlight the new insights that follows from the results in this paper.\par
+\paragraph{General LSA:} Analysis of LSA with CS-PR goes back to the work by \citet{polyak-judisky}, wherein they considered the additive noise setting i.e., $A_t=A$ for some deterministic Hurwitz matrix $A\in \R^{\dcd}$.
+A key improvement in this paper is that we consider the `multiplicative' noise case, i.e., $A_t$ is a random matrix.
+\paragraph{SGD for Linear Regression} As mentioned in \Cref{sec:intro}, we were motivated by the ``remarkable'' results obtained by \citet{bach}, where the authors analyse the performance of SGD for the problem of linear regression with quadratic loss.
 \begin{comment}
-LSA with CS-PR was considered in \cite{polyak-judisky} under the additive noise assumption, wherein $A_t=A$.
-Recent work by \cite{bach} considers a stochastic gradient descent (SGD) algorithm for the problem of linear prediction given in \Cref{ex:linreg}.
-\begin{example}\label{ex:linreg}
-Let $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
-Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$.
-Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$.
+Here, $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ is an $i.i.d$ a bounded (by $B$) data sequence such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
+$x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$.
 SGD with CS-PR to minimize $f(\theta)$ can be given as below:
@@
 It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}.
-In this case $A_P$ is SPD.
-In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
-\begin{align}\label{eq:mselinreg}
-f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
-\end{align}
-\end{example}
-In this setting, \cite{bach} show that the variance term in the MSE is at most $O(\frac{C}{t})$, where $C$ does not depend on the condition number.
-This ``remarkable'' result is due to the following facts: $(i)$ $A_P$ in this case is symmetric positive definite, $(ii)$ the MSE in \eqref{eq:mselinreg} is measured with respect to $A_P$, $(iii)$ the noise $\epsilon_t=y_t-\ip{\ts,x_{t-1}}$ is structured i.e., $\EE{\epsilon_t^2 x_t x_t^\top}\leq R A_P$ for some constant $R>0$.
+%In this case $A_P$ is SPD.
+In the linear regression setting, the prediction error is given by $\EE{\norm{\theta-\ts}^2_{A_P}}$.
+%$(ii)$ the MSE in \eqref{eq:mselinreg} is measured with respect to $A_P$, $(iii)$ the noise $%\epsilon_t=y_t-\ip{\ts,x_{t-1}}$ is structured i.e., $\EE{\epsilon_t^2 x_t x_t^\top}\leq R A_P$ for some constant $R>0$.
+\end{comment}
+The problem class considered by them belongs to $\P_{\text{PSD},B}$ (see \Cref{def:pclass} ) and hence is weakly admissible with a uniform step-size choice.
+In order to produce a rate of $\frac{C}{t}$ (where $C>0$ is a problem independent constant) for the bias term \citet{bach} do not rely on admissibility, but instead use the fact that the MSE $\EE{\norm{\thh_t-\ts}^2_{A_P}}$ for the linear prediction problem is measured in the general quadratic norm induced by $A_P$ and the fact that the noise is `structured' i.e., the noise is bounded by $R\,A_P$ for some constant $R>0$.
 At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.
-\end{comment}
-\paragraph{Definitions:} We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
+\subsection{Reinforcement Learning}
+\paragraph{Notation:} We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
 Let $\gamma \in (0,1)$ be a fixed constant, we define $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$.
@@
 The general setting where $\mu_t>0$ is known as \emph{off-policy}.}.
+\FloatBarrier
+\begin{table}[h]
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|l|p{60mm}|}\hline
+Algorithm& Update & Remark\\ \hline
+TD(0)& $\begin{aligned}\theta_t&=\theta_{t-1}+\alpha_t (b_t -\Delta_t\theta_{t-1})\end{aligned}$ & \cite{korda-prashanth}:$\alpha_t=O(\frac{1}{t})^\beta,\beta\in(0,1)$; PR-avg; $\EE{\norm{\eh_t}}= O(\frac{1}{\sqrt{t}}).$ \\ \hline
+GTD/GTD2& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}- Q_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t A^\top_t y_{t-1})\end{aligned}$
+& \cite{gtd2}: $\beta_t=\eta \alpha_t,\sum_{t\geq 0}\alpha_t=\infty, \sum_{t\geq 0}\alpha^2_t<\infty$; $e_t\ra 0$ as $t\ra\infty$ w.p.1. \cite{gtdmp}:
+$\alpha_t=\beta_t=O(\frac{1}{\sqrt{t}})$; Projection+PR; $\,e_t =O(\frac{1}{{t}^{\frac{1}{4}}})$ w.h.p.\\\hline
+TDC& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t b_t-\mu_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$ & \cite{gtd2}: $\beta_t=\eta \alpha_t,\sum_{t\geq 0}\alpha_t=\infty, \sum_{t\geq 0}\alpha^2_t<\infty$; $e_t\ra 0$ as $t\ra\infty$ w.p.1.
+\\\hline
+\end{tabular}
+}
+\label{tb:rates}
+\caption{Some popular TD algorithms with rates known in literature \cite{korda-prashanth,gtd2,gtdmp}. }
+\end{table}
+\begin{comment}
 \begin{table}
@@
 \end{table}
-
+\end{comment}
+Implications
 \begin{itemize}
 \item It follows that for each instance of blah blah
+\item Bias vairance trade off
 \end{itemize}
-
+\begin{comment}
 \paragraph{Related Work in RL} The various temporal difference learning algorithms can be written as in \Cref{table:td}.
@@
 Further, our results provide explicit bias variance decomposition.
+\end{comment}
+An interesting fact to note is while theoretical analysis for GTD/GTD2/TDC are presented for the case of diminishing step-size, the experimental results are reported for constant step-size choices \cite{gtd2,gtdmp}.

### 2017-05-13 stepsizes.tex
refreshing

@@
 \end{comment}
+\begin{comment}
 In \Cref{sec:uniform} we showed a negative result that for the general class of LSAs, it is not possible to pick a fixed uniform step-size that can be used for any instance of the class. Further,
@@
 We formalize this idea in the form of a step-size tuning algorithm in \Cref{alg:tuning}
-
+\end{comment}
 \begin{algorithm}[H]
@@
 \end{algorithm}
+\begin{comment}
 In \Cref{alg:tuning} IsUnstable$(\rho_j,j=1,\ldots,i\}$) is a routine whose input is $\norm{\theta_t}$ measured at time instances $k$ apart.
@@
 This condition on the sequence $\frac{\rho_{k+1}}{\rho_k}$ can be checked in practice with appropriate tolerances.
+\end{comment}
 \begin{comment}

### 2017-05-13 uniform.tex
refreshing

@@
 We consider some reasonable classes of distributions and test whether or not they are (weakly) admissible.
-\begin{definition}
+\begin{definition}\label{def:pclass}
 \begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
@@
 \item  $\P_{\text{PSD},B}$ be such that  $\P^M_{\text{PSD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PSD} \}\,$ , where PSD denotes the set of positive semi-definite matrices.
+\item  $\P_{\text{PD},B}$ be such that  $\P^M_{\text{SD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PD} \}\,$ , where PD denotes the set of positive definite matrices.
 \end{itemize}
@@
 }
-Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility. When we impost
-In linear least-squares problems (cf. \cref{ex:leastsquares}) the set $\P$ has more structure.
-In particular, the random matrices $\{A_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
-We have the following proposition:
+Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
+However, when we impose additional structure in the data in the form of positive semidefiniteness,  we then observe weak admissibility in \Cref{wad}, and at the same time, it is not enough to guarantee admissiblity in \Cref{notad}.
+Finally, when we impose strucutre in the form of positive definiteness we observe admissibility.
+\FloatBarrier
+\begin{table}[H]
+\begin{tabular}{|c|c|c|c|}\hline
+Problem Class & Weak Admissible & Admissible &Example \\ \hline
+$\P_B$ &\ding{53} &\ding{53} & RL\\\hline
+$\P_{\text{PSD},B}$ &\ding{51} &\ding{53} & SGD for Linear Regression\\\hline
+$\P_{\text{PD},B}$ &\ding{51} &\ding{51} & SGD for Linear Regression with regularization\\\hline
+\end{tabular}
+\end{table}

### 2017-05-14 related.tex
recent updates

@@
 We first discuss the related work outside of RL setting, followed by related work in the RL setting. In both cases, we highlight the new insights that follows from the results in this paper.\par
-\paragraph{General LSA:} Analysis of LSA with CS-PR goes back to the work by \citet{polyak-judisky}, wherein they considered the additive noise setting i.e., $A_t=A$ for some deterministic Hurwitz matrix $A\in \R^{\dcd}$.
-A key improvement in this paper is that we consider the `multiplicative' noise case, i.e., $A_t$ is a random matrix.
 \paragraph{SGD for Linear Regression} As mentioned in \Cref{sec:intro}, we were motivated by the ``remarkable'' results obtained by \citet{bach}, where the authors analyse the performance of SGD for the problem of linear regression with quadratic loss.
-\begin{comment}
-Here, $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ is an $i.i.d$ a bounded (by $B$) data sequence such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite.
-$x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$.
-SGD with CS-PR to minimize $f(\theta)$ can be given as below:
-\begin{subequations}\label{eq:linreg}
-\begin{align}
-\theta_t&=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1}} -y_t\right)\\
-\thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t
-\end{align}
-\end{subequations}
-Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$.
-It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}.
-%In this case $A_P$ is SPD.
-In the linear regression setting, the prediction error is given by $\EE{\norm{\theta-\ts}^2_{A_P}}$.
-%$(ii)$ the MSE in \eqref{eq:mselinreg} is measured with respect to $A_P$, $(iii)$ the noise $%\epsilon_t=y_t-\ip{\ts,x_{t-1}}$ is structured i.e., $\EE{\epsilon_t^2 x_t x_t^\top}\leq R A_P$ for some constant $R>0$.
-\end{comment}
 The problem class considered by them belongs to $\P_{\text{PSD},B}$ (see \Cref{def:pclass} ) and hence is weakly admissible with a uniform step-size choice.
@@
 At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.
+\paragraph{General LSA:} Analysis of LSA with CS-PR goes back to the work by \citet{polyak-judisky}, wherein they considered the additive noise setting i.e., $A_t=A$ for some deterministic Hurwitz matrix $A\in \R^{\dcd}$.
+\paragraph{Our Results:} A key improvement in this paper is that we consider the `multiplicative' noise case, i.e., $A_t$ is a random matrix.
+To tackle the multiplicative noise we use newer analysis introduced by \citet{bach}.
+However, since the general LSA setting does not enjoy special structures of the SGD setting, we make use of Jordan decomposition and similarity transformations in a critical way to prove our results.
 \subsection{Reinforcement Learning}
@@
 \end{comment}
-Implications
-\begin{itemize}
-\item It follows that for each instance of blah blah
-\item Bias vairance trade off
-\end{itemize}
+The \emph{temporal difference} learning algorithm or simply known as TD(0), is the most basic of the class of TD algorithms.
+An important shortcoming of TD(0) was its instability in the \emph{off-policy} case, and was successfully mitigated by the \emph{gradient temporal difference} learing GTD \cite{gtd2}.
+The GTD was first proposed by \citet{gtd} and its variants namely GTD2 and TDC were proposed later by \citet{gtd2}.
+However, the initial convergence analysis for GTD/GTD2/TDC were only asymptotic in nature \cite{gtd,gtd2}.\par
+The most relevant to our results are those by \citet{korda-prashanth} in TD(0) and by \citet{gtdmp} in GTD.
+For the TD(0) case,  diminishing step-sizes $alpha_t=O(\frac{1}{t})^\beta,\beta \in(0,1)$ with PR averaging is showed to exhibit a rate of $O(\frac{1}{t})$ decay for the MSE \cite{korda-prashanth}.
+In the case of GTD/GTD2 diminishing step-sizes $\alpha_t=O(\frac{1}{\sqrt{t}})$, projection of iterates and PR-averaging leads to a rate of $O(\frac{1}{\sqrt{t}})$ is achieved for the prediction error $\norm{A_P\thh_t-b_P}^2$ with high probability.
+\citet{gtdmp} also suggests a new version of GTD, based on stochastic mirror prox ideas, called the GTD-Mirror-Prox which also achieves $O(\frac{1}{\sqrt{t}})$ rate for $\norm{\eh_t}$ with high probability under similar step-size choice that was used by them for the GTD.
+\paragraph{Our Results} \Cref{th:rate} applies to TD(0)/TDC/GTD/GTD2/GTDMP with CS-PR.
+Finite time performance of TDC was unkown in prior literature and our work provides a first step by understanding TDC with CS-PR.
+Further, our results show that the error in the GTD class of algorithms decay at $O(\frac{1}{t})$ rate instead of $O(\frac{1}{\sqrt{t}})$, (without use of projection or mirror maps) which is an improvement.
+In comparison to the TD(0) results by \citet{korda-prashanth}, \Cref{th:rate} is better in that it provides the bias/variance decomposition.
+However, it is important to note that \citet{korda-prashanth} handle the Markov noise case which is not dealt with in this paper.
 \begin{comment}
@@
 \end{comment}
-An interesting fact to note is while theoretical analysis for GTD/GTD2/TDC are presented for the case of diminishing step-size, the experimental results are reported for constant step-size choices \cite{gtd2,gtdmp}.

### 2017-05-14 stepsizes.tex
recent updates

@@
 \section{Automatic Tuning of Step-Sizes}
+It is straightforwatd to see from \eqref{eq:lsaintro} that $\alpha_t$ cannot be asymptotically increasing.
+We now present some heuristic arguments in favour of a constant step-size over asymptotically diminishing step-sizes in \eqref{eq:lsaintro}.
+It has been observed that when the step-sizes of form $\alpha_t=\frac{c}{t}$ or $\alpha_t=\frac{c}{c+t}$ (for some $c>0$) are used, the MSE, $\EE{\norm{\theta_t-\ts}^2}$, is not robust to the choice of $c>0$ \cite{korda-prashanth,bach-moulines}.
+In particular only a $O(\frac{1}{t^{\mu c}/2})$ decay can be achieved for the MSE, where $\mu$ is the smallest positive part of the $A_P$ \cite{bach-moulines}.
+Note that, in the case of LSA with CS-PR, \Cref{th:rate} guarantees a $O(\frac{1}{t})$ rate of decay for the MSE and the problem dependent quantities affect only the constants and not the exponent.
+Also, in the case of important TD algorithms such as GTD/GTD2/TDC, while the theoretical analysis uses diminishing step-sizes, the experimental results are with a constant step-size or with CS and PR averaging \cite{gtd2,gtdmp}.
+Independently, \citet{dann} also observe in their experiments that a constant step-size is better than diminishing step-sizes.\par
+Motivated by \Cref{th:rate} and also by the usage of constant step-size in practice \cite{dann,gtd2,gtdmp}, we suggest a heuristic algorithm to tune the constant step-size (see \Cref{alg:tuning}).
+
 \begin{comment}
@@
 \begin{algorithm}[H]
-\caption*{Automatic Tuning of Constant Step-Size}
+\caption{Automatic Tuning of Constant Step-Size}
 \begin{algorithmic}[1]
-\STATE{Initialize: $\theta_0$, $\alpha=1$, $i=1$, $\rho_0=0$}
+\STATE{Initialize: $\theta_0$, $\alpha=\alpha_{\max}$, $k$, $T$ }
 \FOR{$t=1,2,\ldots, $}
-\STATE{$\theta_t=\theta_{t-1}+\alpha (b_t-A_t\theta_{t-1})$}
-\IF{$t (mod k)==0$}
-\STATE{$\rho_{i}=\norm{\theta_t}$}
-\IF{IsUnstable($\{\rho_j,j=1,\ldots,i\}$)==1}
+\STATE{$\theta_t=\theta_{t-1}+\alpha (b_t-A_t\theta_{t-1}), \thh_t=\thh_{t-1}+\frac{1}{t+1}(\theta_t-\thh_{t-1})$}
+\IF{$IsUnstable(\norm{\theta_t},\ldots,\norm{\theta_{(t-kT)\wedge 0}})=1$}
 \STATE{$\alpha=\alpha/2$}
-\ENDIF
-\STATE{$i=i+1$}
 \ENDIF
@@
 \end{algorithm}
+In \Cref{alg:tuning} $T>0$ is a time epoch and $k$ is a given integer and $\alpha_{\max}>0$ is the maximum step-size that is allowable.
+It is easy to show using a Gronwall argument that in \Cref{alg:tuning} $\norm{\theta_t}\leq (1-e^{\beta t})$, where ($\beta>0)$) $\beta<0$ implies (in)stability of iterates.
+Using this fact, we observe that the sequence $r_i=\frac{\norm{\theta_{(t-kT+iT)\wedge 0}}}{\norm{\theta_{(t-kT+(i-1)T)\wedge 0}}},i=1,\ldots,k$ should be `roughly' (making allowance for the presistent noise) decreasing and converge to $1$.
+The $IsUnstable()$ routine in \Cref{alg:tuning} can take this sequence of $r_i$s as input, and in the case of instability, the step-size can be scaled down by a factor of $2$ (or any constant factor).
 \begin{comment}

### 2017-05-15 abstract.tex
updates

\begin{abstract}
We consider $d$ dimensional linear stochastic algorithms with constant step-size and averaging, also known as Polyak-Ruppert (PR) averaging) of iterates.  In most applications, the aim of the LSA is to compute some $\ts\in \R^d$ using noisy data and $O(d)$ updates per iteration. We are motivated by the problem policy evaluation from experience replay using the \emph{temporal difference} (TD) class of learning algorithms arising in reinforcement learning (RL) that are also LSA algorithms. Under a constant step-size, and PR averaging, we provide bounds for the mean squared error after of the LSA  after $t$ iterations. The restrictions we place are that the data has finite variance and that the expected dynamics of the LSA is Hurwitz. We observe that under the said assumptions the MSE error in LSA with constant step-size and PR averaging decays as $O(\frac{1}{t})$.\par
We also examine  the conditions under which the constant step-size can be chosen in a data independent fashion for a given LSA, and show that not all problem classes admit such a uniform constant step-size.
We also suggest a heuristic step-size tuning algorithm that choose a constant step-size for each given problem instance. We compare our results with related work and also discuss the implication of our results in the context of TD algorithms arising in RL.
\end{abstract}

### 2017-05-15 illustrate.tex
updates

@@
 The above example helps to interpret the error dynamics of \Cref{eq:lsa} in terms of sum of a geometric series with common factor $\alpha\rho$.
+\begin{comment}
 \begin{example}
@@
 \end{example}
-\begin{comment}
-We now list the following observations:
-\begin{itemize}%[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item In all the three instance, $0<\alpha<\frac2{10}$ ensures that the error $\norm{\eh_t}\ra 0$ as $t\ra\infty$. Thus the range of $\alpha$ is uniform across $P^i,i=1,2,3$.
-\item $\rhos{P^1}=\rhod{P^1}<0$, and hence one cannot apply results of \Cref{th:pdrate}.
-However, with  $U=\begin{bmatrix}1 &0 \\ 0 &0.2-\epsilon\end{bmatrix}$ for some $\epsilon>0$ we have $P^1_U$ to be the distribution concentrated on the matrix $A_{P^1_U}=\begin{bmatrix} 10 &20 \\ 0 &10\end{bmatrix}$ and one can apply the results of \Cref{thm:simtran,cor:as}.
-Notice that the condition number $\norm{U}{U^-1}\approx 5$.
-\end{itemize}
+\input{plts}
 \end{comment}
-\input{plts}

### 2017-05-15 instance.tex
updates

@@
 \section{Main Results}\label{sec:mainresults}
-In this section, we derive instance dependent bounds  which is valid for a fixed $P$ and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions over $\C^{\dcd}$.
-Here, we only present the results followed by a discussion and the detailed proofs can be found in \Cref{sec:appendix}.
+In this section, we derive instance dependent bounds that are valid for a given problem $P$ (satisfying \Cref{assmp:lsa}) and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions (problems).
+Here, we only present the main results followed by a discussion and the detailed proofs can be found in \Cref{sec:appendix}.
 \begin{lemma}\label{lm:pd}
@@
 \begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{Role of $U$:} When the given problem $P$ is positive definite, it is straighforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$.
+\item \textbf{Role of $U$:} When the given problem $P$ is positive definite, it is straightforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$.
 However, the problem $P$ being Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}).
-This can be remedied by a similiarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$.
+This can be remedied by a similarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$.
 To see this, let $\gamma_t\eqdef U^{-1 }\theta_t$ be a basis transformation, and let us re-write \eqref{conststep} in $\gamma_t$ by pre-multiplying \eqref{constste[} by $U^{-1}$ as shown below:
@@
 \end{align*}
-This completes the transformation of a Hurwirz problem $P$ to a positive definite problem $P_U$.
+This completes the transformation of a Hurwitz problem $P$ to a positive definite problem $P_U$.
 One can revert to $\theta_t$ by using $\theta_t=U\gamma_t$, and this explains the reason for the condition number $\cond{U}$ in \Cref{th:rate}.
@@
 From \eqref{errec} it is clear that the error dynamics depends on the $i)$ properties of the random matrix, $ii)$ properties of the noise.
-In the absence of the noise term in \eqref{eq:errec}, we have $e_t=(I-\alpha A_t) e_{t-1}=\Pi_{s=1}(I-\alpha A_s) e_0$, i.e., the inital error is get multiplied by a product of random matrices.
+In the absence of the noise term in \eqref{eq:errec}, we have $e_t=(I-\alpha A_t) e_{t-1}=\Pi_{s=1}(I-\alpha A_s) e_0$, i.e., the initial error is get multiplied by a product of random matrices.
 In such a scenario, we can guess that whether or not the algorithm forgets the bias $\norm{\theta_0-\ts}$, depends on whether the matrix product is contracting.
@@
 The variance term that captures the rate at which noise is rejected.
-\item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term.
+\item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of step-size, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term.
 This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower.
@@
 For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.
-\item \textbf{Lower Bound} in \Cref{th:lb} shows that
+\item \textbf{Lower Bound} in \Cref{th:lb} shows that the factor $\rhos{P}\rhod{P}$ appearing in the upper bound in \Cref{th:rate} is inevitable for general LSA with CS-PR.
+However, there are specific examples such as SGD for the linear regression problem considered by \citet{bach} where this factor can be avoided (for further remarks see \Cref{sec:related}).
 \end{itemize}

### 2017-05-15 not.tex
updates

@@
  %Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
-We use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively. The notation for (conjugage) transpose carries over from vectors.
+We use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively. The notation for (conjugate) transpose carries over from vectors.
 The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^d:\norm{x}=1} \norm{Ax}$.
@@
 \begin{definition}
-We call a matix $A\in \C^{\dcd}$ to be \emph{Hurwitz} (H) if all eigenvalues of $A$ have positive real parts.
+We call a matrix $A\in \C^{\dcd}$ to be \emph{Hurwitz} (H) if all eigenvalues of $A$ have positive real parts.
 We call a matrix $A\in \C^{\dcd}$ to be \emph{Positive Definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$.

### 2017-05-15 prob.tex
updates

@@
 The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
-can then be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.\par
+can then be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $d$ is in the range of thousands or more.\par
 We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa}.
@@
 \item The variances $\sigma_P^2$ and $\sigma_b^2$ of the noise terms $M_t$ and $b_t$ in the update equation.
-\item The baias $\norm{\theta_0-\ts}$ due to the initial condition $\theta_0$.
+\item The bias $\norm{\theta_0-\ts}$ due to the initial condition $\theta_0$.
 \item The effect of the structural properties of the $A_P$, i.e., whether it is AS or PD or SPD.
 \end{itemize}
-Most resutls in the past including those that of \cite{bach} consider an SGD algorithm for quadratic loss minimization, where the matrix $A_P$ is SPD and the noise has additional strucutral properties (we discuss this briefly in \Cref{sec:lp}).
+Most results in the past including those that of \cite{bach} consider an SGD algorithm for quadratic loss minimization, where the matrix $A_P$ is SPD and the noise has additional structural properties (we discuss this briefly in \Cref{sec:lp}).
 As it turns out these nice properties are lost when $A_P$ is AS or PD (a property that is holds in RL settings).
@@
 The comparison with existing literature is done in two separate sections, one to compare our results with \cite{bach} in the linear predictions setting (\Cref{sec:lp}) and the other to compare our results with prior results in RL literature \cite{}.
-\todoch{Mention impotance of forgetting the bias etc. Mention that we really don't have control over $\norm{U}$}
+\todoch{Mention importance of forgetting the bias etc. Mention that we really don't have control over $\norm{U}$}
 \end{comment}

### 2017-05-15 related.tex
updates

@@
 We first discuss the related work outside of RL setting, followed by related work in the RL setting. In both cases, we highlight the new insights that follows from the results in this paper.\par
-\paragraph{SGD for Linear Regression} As mentioned in \Cref{sec:intro}, we were motivated by the ``remarkable'' results obtained by \citet{bach}, where the authors analyse the performance of SGD for the problem of linear regression with quadratic loss.
+\paragraph{SGD for Linear Regression} As mentioned in \Cref{sec:intro}, we were motivated by the ``remarkable'' results obtained by \citet{bach}, where the authors analyze the performance of SGD for the problem of linear regression with quadratic loss.
 The problem class considered by them belongs to $\P_{\text{PSD},B}$ (see \Cref{def:pclass} ) and hence is weakly admissible with a uniform step-size choice.
-In order to produce a rate of $\frac{C}{t}$ (where $C>0$ is a problem independent constant) for the bias term \citet{bach} do not rely on admissibility, but instead use the fact that the MSE $\EE{\norm{\thh_t-\ts}^2_{A_P}}$ for the linear prediction problem is measured in the general quadratic norm induced by $A_P$ and the fact that the noise is `structured' i.e., the noise is bounded by $R\,A_P$ for some constant $R>0$.
+In order to produce a rate of $\frac{C}{t}$ (where $C>0$ is a problem independent constant) for the bias term \citet{bach} do not rely on admissibility, but instead use the fact that the MSE, $\EE{\norm{\thh_t-\ts}^2_{A_P}}$ for the linear prediction problem is measured in the general quadratic norm with respect to the symmetric positive definite matrix $A_P$ and the fact that the noise is `structured' i.e., the variance is bounded by $R\,A_P$ for some constant $R>0$.
 At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.
-\paragraph{General LSA:} Analysis of LSA with CS-PR goes back to the work by \citet{polyak-judisky}, wherein they considered the additive noise setting i.e., $A_t=A$ for some deterministic Hurwitz matrix $A\in \R^{\dcd}$.
-\paragraph{Our Results:} A key improvement in this paper is that we consider the `multiplicative' noise case, i.e., $A_t$ is a random matrix.
+\paragraph{Our Results:} Analysis of LSA with CS-PR goes back to the work by \citet{polyak-judisky}, wherein they considered the additive noise setting i.e., $A_t=A$ for some deterministic Hurwitz matrix $A\in \R^{\dcd}$.
+A key improvement in our paper is that we consider the `multiplicative' noise case, i.e., $A_t$ is a random matrix.
 To tackle the multiplicative noise we use newer analysis introduced by \citet{bach}.
-However, since the general LSA setting does not enjoy special structures of the SGD setting, we make use of Jordan decomposition and similarity transformations in a critical way to prove our results.
-\subsection{Reinforcement Learning}
-\paragraph{Notation:} We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
-Let $\gamma \in (0,1)$ be a fixed constant, we define $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$.
+However, since the general LSA setting (with Hurwitz assumption) does not enjoy special structures of the SGD setting of \citet{bach}, we make use of Jordan decomposition and similarity transformations in a critical way to prove our results.
+\paragraph{Notation for RL:} We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
+For a fixed constant $\gamma \in (0,1)$ define  $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$.
 In what follows, $\mu_t>0$ is an \emph{importance} sampling factor whose aim is to correct for mismatch in the (behavior) distribution with which the data was collected and the (target) distribution with respect to which one wants to learn.
@@
 The general setting where $\mu_t>0$ is known as \emph{off-policy}.}.
+The various TD class of algorithms that can be cast as LSAs are given in \Cref{tb:tdalgo}
 \FloatBarrier
@@
 }
-\label{tb:rates}
+\label{tb:tdalgo}
 \caption{Some popular TD algorithms with rates known in literature \cite{korda-prashanth,gtd2,gtdmp}. }
@@
 \end{comment}
-The \emph{temporal difference} learning algorithm or simply known as TD(0), is the most basic of the class of TD algorithms.
-An important shortcoming of TD(0) was its instability in the \emph{off-policy} case, and was successfully mitigated by the \emph{gradient temporal difference} learing GTD \cite{gtd2}.
+\paragraph{Related Work in RL:} The \emph{temporal difference} learning algorithm or simply known as TD(0), is the most basic of the class of TD algorithms.
+An important shortcoming of TD(0) was its instability in the \emph{off-policy} case, and was successfully mitigated by the \emph{gradient temporal difference} learning GTD algorithm\cite{gtd2}.
 The GTD was first proposed by \citet{gtd} and its variants namely GTD2 and TDC were proposed later by \citet{gtd2}.
-However, the initial convergence analysis for GTD/GTD2/TDC were only asymptotic in nature \cite{gtd,gtd2}.\par
+However, the initial convergence analysis for GTD/GTD2/TDC were only asymptotic in nature \cite{gtd,gtd2} with diminishing step-sizes.\par
 The most relevant to our results are those by \citet{korda-prashanth} in TD(0) and by \citet{gtdmp} in GTD.
-For the TD(0) case,  diminishing step-sizes $alpha_t=O(\frac{1}{t})^\beta,\beta \in(0,1)$ with PR averaging is showed to exhibit a rate of $O(\frac{1}{t})$ decay for the MSE \cite{korda-prashanth}.
-In the case of GTD/GTD2 diminishing step-sizes $\alpha_t=O(\frac{1}{\sqrt{t}})$, projection of iterates and PR-averaging leads to a rate of $O(\frac{1}{\sqrt{t}})$ is achieved for the prediction error $\norm{A_P\thh_t-b_P}^2$ with high probability.
-\citet{gtdmp} also suggests a new version of GTD, based on stochastic mirror prox ideas, called the GTD-Mirror-Prox which also achieves $O(\frac{1}{\sqrt{t}})$ rate for $\norm{\eh_t}$ with high probability under similar step-size choice that was used by them for the GTD.
-\paragraph{Our Results} \Cref{th:rate} applies to TD(0)/TDC/GTD/GTD2/GTDMP with CS-PR.
+For the TD(0) case,  diminishing step-sizes $\alpha_t=O(\frac{1}{t})^\beta,\beta \in(0,1)$ with PR averaging is showed to exhibit a rate of $O(\frac{1}{t})$ decay for the MSE when $\beta\ra 1$ \cite{korda-prashanth}.
+In the case of GTD/GTD2 diminishing step-sizes $\alpha_t=O(\frac{1}{\sqrt{t}})$, projection of iterates and PR-averaging leads to a rate of $O(\frac{1}{\sqrt{t}})$ is achieved for the prediction error $\norm{A_P\thh_t-b_P}^2$ with high probability \cite{gtdmp}.
+\citet{gtdmp} also suggest a new version of GTD based on stochastic mirror prox ideas, called the GTD-Mirror-Prox which also achieves $O(\frac{1}{\sqrt{t}})$ rate for $\norm{A_P\thh_t-b_P}^2$ with high probability under similar step-size choice that was used by them for the GTD.
+\paragraph{Our Results} \Cref{th:rate} applies to TD(0)/TDC/GTD/GTD2/GTDMP with CS-PR since all the TD class of algorithms satisfy the Hurwitz assumption.
 Finite time performance of TDC was unkown in prior literature and our work provides a first step by understanding TDC with CS-PR.
-Further, our results show that the error in the GTD class of algorithms decay at $O(\frac{1}{t})$ rate instead of $O(\frac{1}{\sqrt{t}})$, (without use of projection or mirror maps) which is an improvement.
+Further, our results show that the error in the GTD class of algorithms decay at $O(\frac{1}{t})$ rate (even without use of projection or mirror maps) instead of $O(\frac{1}{\sqrt{t}})$,  which is an improvement.
 In comparison to the TD(0) results by \citet{korda-prashanth}, \Cref{th:rate} is better in that it provides the bias/variance decomposition.
+Further, their results are specific to the TD(0) where $A_P$ is known to be positive definite, and hence cannot be extended in a straightforward manner to GTD class or TDC.
 However, it is important to note that \citet{korda-prashanth} handle the Markov noise case which is not dealt with in this paper.

### 2017-05-15 stepsizes.tex
updates

@@
 \section{Automatic Tuning of Step-Sizes}
-It is straightforwatd to see from \eqref{eq:lsaintro} that $\alpha_t$ cannot be asymptotically increasing.
+It is straightforward to see from \eqref{eq:lsaintro} that $\alpha_t$ cannot be asymptotically increasing.
 We now present some heuristic arguments in favour of a constant step-size over asymptotically diminishing step-sizes in \eqref{eq:lsaintro}.
@@
 Independently, \citet{dann} also observe in their experiments that a constant step-size is better than diminishing step-sizes.\par
-Motivated by \Cref{th:rate} and also by the usage of constant step-size in practice \cite{dann,gtd2,gtdmp}, we suggest a heuristic algorithm to tune the constant step-size (see \Cref{alg:tuning}).
-
-\begin{comment}
-We now argue, in a heuristic manner, as to why constant step-size with PR-averaging might be a principled idea, especially in TD type algorithms.
-Consider the following LSA with time varying step-sizes $\alpha_t>0$
-\begin{align}
-\theta_t=\theta_{t-1}+\alpha_t(b_t-A_t\theta_{t-1}).
-\end{align}
-It is known from asymptotic results by \cite{borkarbook,kushner}\todoch{Check about Kushner} that in order to handle the noise the step-sizes have to be diminishing i.e., $\alpha_t\ra$, unsummable i.e., $\sum_{t \geq 0}\alpha_t=\infty$ and square summable $\sum_{t\geq 0}\alpha_t^2<\infty$.
-We now look at two specific diminishing step-size sequences, to illustrate the issues that the diminishing step-sizes presents us with.
-In what follows, we assume that the maximum eigevalue of $\EE{A_t}$ is less that $1$.
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{$\alpha_t=\frac{1}{t}$\footnote{While $\frac{c}{t}$, $c\geq 0$ is more general, it is enough to consider $\frac1t$ by accommodating the multiple $c$ as part of the matrix $A_t$}} is known yield a rate of $\frac{1}{t^{\lambda_{\min}}}$, where $\lambda_{\min}$ is the smallest real part of the eigen values of the matrix $\EE{A_t}$.
-\item \textbf{$\alpha_t=\frac{c}{c+t}$} is another example, where the step size decays at a rate $\frac{1}{t/c}$.
-The rate of decay of error in this case turns out to be $\frac{c^2\sigma_b^2}{t}$, whenever $c>\frac{1}{\lambda_{\min}}$, which can be ensured only when $\lambda_{\min}$ is known.
-\end{itemize}
-In contrast, in the case of constant step-size with PR-averaging, not knowing the problem instance can result in instability, and when we manage to choose a step-size that results in stable behavior, the rates obtained in \Cref{th:pdrate,cor:as} hold for it.
-Using this observation, the constant-step size can be set in a more prinicipled manner as showing the pseudocode below.
-\end{comment}
-\begin{comment}
-In \Cref{sec:uniform} we showed a negative result that for the general class of LSAs, it is not possible to pick a fixed uniform step-size that can be used for any instance of the class. Further,
-recall from our discussion in \Cref{sec:mainresults} that as $\alpha\ra\alpha_P$ (or $\alpha\ra\alpha_{P_U}$), the MSE blows up and for $\alpha>\alpha_P$ the LSA is unstable because the spectral norm of the random matrix $I-\alpha A_t$ is no longer less that unity.
-Combining these facts, we conclude that if the step-size is incorrect for a particular instance then we would observe instability, a fact that can be used to reset the step-size to a smaller value, and this procedure can be continued in an iterative fashion to find the right choice of step-size for a given problem.
-We formalize this idea in the form of a step-size tuning algorithm in \Cref{alg:tuning}
-\end{comment}
+We would like to remind that in \Cref{sec:uniform} we showed that weak admissibility might not hold for all problem classes, and hence a uniform choice for the constant step-size might not be possible, However, motivated by \Cref{th:rate} and also by the usage of constant step-size in practice \cite{dann,gtd2,gtdmp}, we suggest a heuristic algorithm to tune the constant step-size (see \Cref{alg:tuning}).
 \begin{algorithm}[H]
@@
 It is easy to show using a Gronwall argument that in \Cref{alg:tuning} $\norm{\theta_t}\leq (1-e^{\beta t})$, where ($\beta>0)$) $\beta<0$ implies (in)stability of iterates.
-Using this fact, we observe that the sequence $r_i=\frac{\norm{\theta_{(t-kT+iT)\wedge 0}}}{\norm{\theta_{(t-kT+(i-1)T)\wedge 0}}},i=1,\ldots,k$ should be `roughly' (making allowance for the presistent noise) decreasing and converge to $1$.
+Using this fact, we observe that the sequence $r_i=\frac{\norm{\theta_{(t-kT+iT)\wedge 0}}}{\norm{\theta_{(t-kT+(i-1)T)\wedge 0}}},i=1,\ldots,k$ should be `roughly' (making allowance for the persistent noise) decreasing and converge to $1$.
 The $IsUnstable()$ routine in \Cref{alg:tuning} can take this sequence of $r_i$s as input, and in the case of instability, the step-size can be scaled down by a factor of $2$ (or any constant factor).
-\begin{comment}
-In \Cref{alg:tuning} IsUnstable$(\rho_j,j=1,\ldots,i\}$) is a routine whose input is $\norm{\theta_t}$ measured at time instances $k$ apart.
-It is easy to show from Gronwall type arguments that the $\norm{\theta_t}\leq (1-exp{\beta t})$, where $\beta>0$ implies stability and $\beta<0$ imples instability.
-From this Gronwall based inequality, we observe that when the iterates are stable the ratio $\frac{\rho_{k+1}}{\rho_k},k=1,\ldots,i-1$ should be `roughly' decreasing (because there is the iterates are persistently corrupted by noise) and should converge to $1$.
-This condition on the sequence $\frac{\rho_{k+1}}{\rho_k}$ can be checked in practice with appropriate tolerances.
-\end{comment}
-\begin{comment}
-In order to calculate the rates we make use of the following asymptotic result by \cite{borkarbook}.
-\begin{theorem}\label{th:borkar}
-Let $\theta_t=\theta_{t-1}+\alpha_t(b_t-A_t\theta_{t-1})$, where $\theta_t,b_t\in\R^d$ and $A_t\in\R^{\dcd}$ that satisfy the conditions in \Cref{assmp:lsa} and $\sum_{t\geq 0}\alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$.
-Let $t_{ode}(k)=\eqdef \sum_{i=0}^{k-1}\alpha_n$ and let $\tilde{\theta}(t)$ be the trajectory of the ordinary differential equation $\dot{\theta}(t)=\EE{b_t}-A_P\theta(t)$.
-Then it follows that
-\begin{align*}
-\lim_{k\ra\infty}\sup\norm{\tilde{\theta_t(t_{ode}(k))}-\theta_t} \ra 0
-\end{align*}
-\end{theorem}
-
-For this discussion, let us consider a simple example where $P$ is a distribution with all the mass concentrated at $A_P=\begin{bmatrix}\lambda_{\min} &0\\ 0 &\lambda_{\max}\end{bmatrix}$, $b_t$ is a zero mean $i.i.d$ sequence with variance $\sigma_2_b$.
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{$\alpha_t=\frac{1}{t}$:} In order to ensure stability, we want the eigenvalues of $I-\alpha_t A_P$ to be within unit circle.
-This restricts the choice of $c$ to be $0<c<\frac{2}{\lambda_{\max}}$.
-It is straighforward to show that a step-size choice of $\alpha_t=\frac{2}{\lambda_{\max}t}$ will yield a rate of $\frac{1}{t^{2\lambda_{\min}}}$, which can be very bad for small $\lambda_{\min}$.
-\item \textbf{}
-\end{itemize}
-\end{comment}

### 2017-05-15 uniform.tex
recent updates
updates

@@
 \section{Uniform stepsizes}
-If $\P$ is weakly admissible, then one can choose a stepsize $\alpha_{\P}>0$ solely based on the knowledge of $\P$ and
+If $\P$ is weakly admissible, then one can choose a step-size $\alpha_{\P}>0$ solely based on the knowledge of $\P$ and
 conclude that no matter the joint distribution $P$ of $(b_t,A_t)$, as long as the distribution of $P$ is an element of $\P$, the conclusions of \Cref{th:rate} hold.
@@
 Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
-However, when we impose additional structure in the data in the form of positive semidefiniteness,  we then observe weak admissibility in \Cref{wad}, and at the same time, it is not enough to guarantee admissiblity in \Cref{notad}.
-Finally, when we impose strucutre in the form of positive definiteness we observe admissibility.
+However, when we impose additional structure in the data in the form of positive semi-definiteness,  we then observe weak admissibility in \Cref{wad}, and at the same time, it is not enough to guarantee admissibility in \Cref{notad}.
+Finally, when we impose structure in the form of positive definiteness we observe admissibility.
+\begin{comment}
 \FloatBarrier
@@
 \end{table}
+\end{comment}

### 2017-05-17 abstract.tex
latest

@@
 \begin{abstract}
-We consider $d$ dimensional linear stochastic algorithms with constant step-size and averaging, also known as Polyak-Ruppert (PR) averaging) of iterates.
-In most applications, the aim of the LSA is to compute some $\ts\in \R^d$ using noisy data and $O(d)$ updates per iteration.
-We are motivated by the problem policy evaluation from experience replay using the \emph{temporal difference} (TD) class of learning algorithms arising in reinforcement learning (RL) that are also LSA algorithms.
-Under a constant step-size, and PR averaging, we provide bounds for the mean squared error after of the LSA  after $t$ iterations.
-The restrictions we place are that the data has finite variance and that the expected dynamics of the LSA is Hurwitz.
-We observe that under the said assumptions the MSE error in LSA with constant step-size and PR averaging decays as $O(\frac{1}{t})$.\par
-We also examine  the conditions under which the constant step-size can be chosen in a data independent fashion for a given LSA, and show that not all problem classes admit such a uniform constant step-size.
-We also suggest a heuristic step-size tuning algorithm that choose a constant step-size for each given problem instance.
-We compare our results with related work and also discuss the implication of our results in the context of TD algorithms arising in RL.
+We consider $d$-dimensional linear stochastic approximation algorithms (LSAs) with a constant step-size and the so called Polyak-Ruppert (PR) averaging of iterates.
+LSAs are widely applied in machine learning and reinforcement learning (RL), where the aim is to compute an appropriate $\ts\in \R^d$ (that is an optimum or a fixed point) using noisy data and $O(d)$ updates per iteration.
+In this paper, we are motivated by the problem (in RL) of policy evaluation from experience replay using the \emph{temporal difference} (TD) class of learning algorithms that are also LSAs.
+For LSAs with a constant step-size, and PR averaging, we provide bounds for the mean squared error (MSE) after $t$ iterations.
+We assume that data is $i.i.d$ with finite variance (underlying distribution being $P$) and that the expected dynamics is Hurwitz.
+For a given LSA with PR averaging, and data distribution $P$ satisfying the said assumptions, we show that there exists a range of constant step-sizes such that its MSE decays as $O(\frac{1}{t})$.\par
+We examine the conditions under which a constant step-size can be chosen uniformly for a class of data distributions $\P$, and show that not all data distributions `admit' such a uniform constant step-size.
+We also suggest a heuristic step-size tuning algorithm to choose a constant step-size of a given LSA for a given data distribution $P$.
+We compare our results with related work and also discuss the implication of our results in the context of LSA-TD algorithms.
 \end{abstract}

### 2017-05-17 instance.tex
latest

@@
 \section{Main Results}\label{sec:mainresults}
-In this section, we derive instance dependent bounds that are valid for a given problem $P$ (satisfying \Cref{assmp:lsa}) and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions (problems).
-Here, we only present the main results followed by a discussion and the detailed proofs can be found in \Cref{sec:appendix}.
-\begin{lemma}\label{lm:pd}
-Let $P$ be positive definite, then there exists an $\alpha_P>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_P)$.
-\end{lemma}
-\begin{lemma}\label{lm:simtran}[Similarity Transformation]
-Let $P$ be Hurwitz, then there exists a $U\in \gln$ such that $P_U$ is positive definite.
+In this section, we derive instance dependent bounds that are valid for a given problem $P$ (satisfying \Cref{assmp:lsa}) and in the \Cref{sec:uniform}, we address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions (problems).
+Here, we only present the main results followed by a discussion and the detailed proofs can be found in \Cref{sec:proofs}.
+In what follows, for the sake of brevity, we drop the subscript $P$ in the quantities $\EEP{\cdot}$, $\sigma^2_{A_P}$ and $\sigma^2_{b_P}$.
+%\begin{lemma}\label{lm:simtran}[Similarity Transformation]
+%Let $P$ be Hurwitz, then there exists a $U\in \gln$ such that $P_U$ is positive definite.
+%\end{lemma}
+\begin{lemma}\label{lm:hur}
+Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, then there exists an $\alpha_{P_U}>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_{P_U})$.
 \end{lemma}
@@
 \begin{theorem}\label{th:rate}
-Let $P$ be Hurwitz, then there exists $U\in\gln$ and $\alpha_{P_U}>0$ such that
+Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, then there exists $U\in\gln$ and $\alpha_{P_U}>0$ such that
 \begin{align*}
@@
 \left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}
-\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
+\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,,
 \end{align*}
 \end{theorem}
-
+$\forall \alpha\in (0,\alpha_{P_U})$.
 \begin{theorem}[Lower Bound]
-There exists a $P$ Hurwitz, $\alpha_P>0$, such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in (0,\alpha_P)$ and
+There exists a distribution $P$ over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, such that $\alpha_P>0$, such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in (0,\alpha_P)$ and
 \begin{align*}
-\EE{\norm{\eh_t}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big( \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \alpha^2 \sigma^2_b \Big)
+\EE{\norm{\thh_t-\ts}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big( \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \big(\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}\big) \Big)
 \end{align*}
@@
 This can be remedied by a similarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$.
-To see this, let $\gamma_t\eqdef U^{-1 }\theta_t$ be a basis transformation, and let us re-write \eqref{conststep} in $\gamma_t$ by pre-multiplying \eqref{constste[} by $U^{-1}$ as shown below:
+To see this, let $\gamma_t\eqdef U^{-1 }\theta_t$ be a basis transformation, and let us re-write \eqref{conststep} in terms of $\gamma_t$ by pre-multiplying by $U^{-1}$ as shown below:
 \begin{align*}
@@
 The \emph{bias} term that captures the rate at which the initial condition $\norm{\theta_0-\ts}^2$ is forgotten.
-The second term is the \emph{variance} given by $\V=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}} \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} $.
+The second term is the \emph{variance} given by $\V=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}} \frac{\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} $.
 The variance term that captures the rate at which noise is rejected.
@@
 This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower.
-However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:rate} where $\alpha$ is seen to multiply the variances $\sigma^2_P$ and $\sigma^2_b$.
+However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:rate} where $\alpha$ is seen to multiply the variances $\sigma^2_A$ and $\sigma^2_b$.
 In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other.

### 2017-05-17 intro.tex
recent updates
latest

@@
 \section{Introduction}
-A linear stochastic approximation (LSA) algorithm is given by
+We consider linear stochastic approximation algorithms (LSA) of the general form given by
 \begin{align}\label{eq:lsaintro}
@@
 \end{align}
-where the iterates, $\theta_t\in \R^d$, and step-sizes, $\alpha_t>0$, are algorithmic quantities, and $(b_t,A_t)\in \R^d\times \R^{\dcd}$ is the noisy data sequence presented to the algorithm.
-LSA algorithms (or simply LSAs) are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using only the noisy data sequence and $O(d)$ updates per iteration.
-Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear regression with quadratic loss \cite{bach,bachaistats}, and the temporal difference class (TD) of learning algorithms in RL \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.\par
+where the iterates, $\theta_t\in \R^d$, and step-sizes, $\alpha_t>0$, are algorithmic quantities, and $(b_t,A_t), t\geq 0\in \R^d\times \R^{\dcd}$ is the noisy data sequence presented to the algorithm.
+LSAs are widely applied in machine learning and reinforcement learning (RL), where the aim is to compute an appropriate $\ts\in \R^d$ using only the noisy data sequence and $O(d)$ updates per iteration.
+Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear regression with quadratic loss \cite{bach,bachaistats}, and the \emph{temporal difference} (TD) class of learning algorithms in RL \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.\par
 %An additional feature in these class of applications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
 
-The choice of the step-sizes $\alpha_t$ is critical for the performance of the LSA in \eqref{eq:lsaintro}.
-Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster learning, while at the same time too large step-sizes might result in instability of \eqref{eq:lsaintro}.
-A useful choice in RL has been the diminishing and square-summable step-sizes \cite{gtd2,borkarbook,konda-tsitsiklis}, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$.
-An alternate technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{s=0}^t \theta_t$.
+The choice of the step-sizes $\alpha_t$ is critical for the performance of LSAs such as \eqref{eq:lsaintro}.
+Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster learning, and at the same time, step-sizes that are too large might result in instability of \eqref{eq:lsaintro}.
+A useful choice in RL has been the diminishing step-sizes \cite{gtd2,gtdmp,borkarbook,konda-tsitsiklis}, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$.
+An alternate technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{i=0}^t \theta_i$.
 In LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
-The idea of averaging in stochastic approximation to improve noise rejection is a technique that goes back to  \citet{ruppert,polyak-judisky}.
+The idea of averaging in stochastic approximation to improve noise rejection is a technique that goes back to  \citet{ruppert} and \citet{polyak-judisky}.
 Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection.
-\par
-Recently, \citet{bach} considered SGD\footnote{SGD is a special of the LSA algorithm in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and $i.i.d$ sampling.
-A ``remarkable'' result by \citet{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the expected squared regression error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
-
-\paragraph{Motivation and Contribution:} In this paper, we wish to study LSA with CS-PR where data $(b_t,A_t)$ is $i.i.d$ with some underlying distribution $P$ such that variance is bounded.
+\paragraph{Motivation and Contribution:} Recently, \citet{bach} considered SGD\footnote{SGD is an LSA of the form in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and $i.i.d$ sampling.
+A ``remarkable'' result by \citet{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the mean squared (regression) error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
+In this paper, we study LSA with CS-PR (thereby extending prior work by \citet{bach} from SGD to general LSA) where data $(b_t,A_t),t\geq 0$ is $i.i.d$ with some underlying distribution $P$ such that the variance is bounded.
 An additional restriction we place on $P$ is that the matrix $\EEP{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts.
-For a given $P$, we wish to measure the performance of LSA with CS-PR in terms of the mean square error (MSE) $\EEP{\norm{\thh_t-\ts}^2}$.
-Further, for a given class of distributions $\P$, we would like to know whether it is possible to ensure uniform performance $\forall P\in \P$.
-Our study is motivated by the problem \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class  of algorithms \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.
+Our study is motivated by the problem of \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class  of algorithms \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.
 Here, we wish to understand the effectiveness of the CS-PR technique for the TD class of LSA algorithms used in RL.
-
-We now list our contributions in this paper
+We now list our contributions in this paper as follows:
 %While asymptotically diminishing step-sizes are common in RL literature \cite{ilstd,gtdmp,korda-prashanth}, the constant step-size choice has also been used in the past \cite{gtd2}.
-%\begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
- \begin{center}Algorithm\end{center}
-\paragraph{Instance Dependent Bounds} We show that for a fixed $P$, there exists a $\alpha_P>0$ such that the mean squared error (MSE) of the LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$.
+\begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
+\item \textbf{Instance Dependent Bounds} For a given $P$, we  measure the performance of LSA with CS-PR in terms of the mean square error (MSE) given by $\EEP{\norm{\thh_t-\ts}^2}$.
+We show there exists an $\alpha_P>0$ such that the mean squared error (MSE) of LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$.
 The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
-\begin{center} Structure in Data\end{center}
-\paragraph{Uniform Step-Sizes:} We also show (via counter example) that there is a class $\P$ that does not `admit' a single constant step-size $\alpha_{\P}$ that guarantees uniform performance $\forall P\in \P$.
+\item \textbf{Uniform Step-Sizes:} It is of interest to know, for a given class $\P$ of distributions, whether it is possible to ensure uniform performance $\forall P\in \P$.
+We also show (via an example) that there is a class $\P$ that does not `admit' a constant step-size $\alpha_{\P}$ that guarantees uniform performance $\forall P\in \P$.
 This is a negative result in the context of RL.
-However,
-we also argue that the subclass $\P_{LR}$ related to SGD for linear regression does `admit' a problem independent constant step-size, thereby recovering a part of the result by \citet{bach}.
-\begin{center} Meta-Parameter $\alpha$\end{center}
-\paragraph{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion, and the performance of CS-PR is comparable to other popular step-size choices used for LSA in RL literature.
+However, we also argue that the subclass (which we denote by $\P_{\text{SPD},B}$) related to SGD for linear regression does `admit' a uniform constant step-size, thereby recovering a part of the result by \citet{bach}.
+\item \textbf{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion.
+This way we argue that the performance LSA with CS-PR is comparable to other popular step-size choices used for LSAs arising in RL.
+\end{itemize}
+\paragraph{Implications:}
 \paragraph{Organization:}

### 2017-05-17 linal.tex
updates
latest

@@
-\section{Linear Algebra Preliminaries}
+\appendix
+\section{Linear Algebra Preliminaries}\label{sec:appendix}
+\subsection{Additional Notations}\label{sec:addnot}
+For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively. Given a $x\in \C^d$, for $1\leq i \leq d$, $x(i)$ denotes the $i^{th}$ component of $x$.
+For any $x\in \C$ we denote its modulus $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$ and its complex conjugate by $\bar{x}=a-ib$.
+We use $A\succeq 0$ to denote that the
+square matrix $A$ is Hermitian and positive semidefinite (HPSD):
+$A = A^*$, $\inf_x x^* A x\ge 0$. We use $A\succ 0$ to denote that the square matrix $A$ is Hermitian and positive definite (HPD): $A=A^*$, $\inf_x x^* A x > 0$.
+For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
+We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
+We also use $\preceq$ and $\prec$ analogously. We denote the smallest eigen value of a real symmetric positive definite matrix $A$ by $\lambda_{\min}(A)$.\par
 We now present some useful results from linear algebra.
-\subsection{Notation}
-Let $B$ be a $\dcd$ block diagonal matrix given by $B=\begin{bmatrix} B_1 &0 &0 &\ldots &0 \\ 0 &B_2 &0 &\ldots &0  \\ \vdots &\vdots &\vdots &\vdots &\vdots \\ 0 &\ldots &0 &0 &B_k \end{bmatrix}$, where $B_i$ is a $n_i \times n_i$ matrix such that $n_i<n$ (w.l.o.g) and $\sum_{i=1}^k n_i=n$.
+
+Let $B$ be a $\dcd$ block diagonal matrix given by $B=\begin{bmatrix} B_1 &0 &0 &\ldots &0 \\ 0 &B_2 &0 &\ldots &0  \\ \vdots &\vdots &\vdots &\vdots &\vdots \\ 0 &\ldots &0 &0 &B_k \end{bmatrix}$, where $B_i$ is a $d_i \times d_i$ matrix such that $d_i<d,\,\forall i=1,\ldots,k$ (w.l.o.g) and $\sum_{i=1}^k d_i=d$.
 We also denote $B$ as
@@
 We will now recall Jordon decomposition.
-
-\begin{comment}
-\begin{lemma}
-Given a matrix $A\in \C^{\dcd} there exists a \emph{characteristic} polynomial denoted by $\chi_A=(\lambda_1 \I-A)^{n_1}\ldots(\lambda_{k} \I-A)^{n_k}$, where $\lambda_i,\,i=1,\ldots,k (k\leq n)$ are the $k$ distinct eigenvalues of the matrix $A$ such that $\sum_{i=1}^k n_i=n$.
-\end{lemma}
-\begin{lemma}
-There exists a polynomial known as the \emph{minimal} polynomial denoted by $m_A=(\lambda_1 \I-A)^{n'_1}\ldots(\lambda_k \I-A)^{n'_k}$, where $n'_i\leq n_i,\,i=1\ldots,k$.
-\end{lemma}
-\end{comment}
-
 \begin{lemma}\label{jordon}
-Let $A\in \C^{\dcd}$ and $\{\lambda_i\in \C,i=1,\ldots,k\leq n \}$ denote its $k$ distinct eigenvalues.
+Let $A\in \C^{\dcd}$ and $\{\lambda_i\in \C,i=1,\ldots,k\leq d \}$ denote its $k$ distinct eigenvalues.
 There exists a complex matrix $V\in \C^{\dcd}$ such that $A=V\tL V^{-1}$, where
 $\tL=\tL_1\op\ldots\op\tL_k$, where each $\tL_i,\,i=1,\ldots,k$ can further be written as $\tL_i= {\tL}^i_{1}\op \ldots \op {\tL}^i_{{l(i)}}$.
-Each of ${\tL}^i_{j},j=1,\ldots,l(i)$ is a $d^i_j\times d^i_j$ square matrix such that $\sum_{j=1}^{l(i)} d^i_j =n_i$ and has the special form given by
+Each of ${\tL}^i_{j},j=1,\ldots,l(i)$ is a $d^i_j\times d^i_j$ square matrix such that $\sum_{j=1}^{l(i)} d^i_j =d_i$ and has the special form given by
 ${\tL}^i_{j}=\begin{bmatrix} \lambda_i &1 &0 &\ldots &0 &0\\ 0 &\lambda_i &1 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\lambda_i &1 \\ 0 &\ldots &0 &0 &0 &\lambda_i \end{bmatrix}$.
@@
 
-\begin{lemma}\label{simtran}
-Let $A\in \C^{\dcd}$ be asymptotically stable. There exists a matrix $U\in \gln$ such that $A=U\Lambda U^{-1}$ and $\Lambda^\dag+\Lambda$ is a real symmetric positive definite matrix.
+\begin{lemma}\label{lm:simtran}
+Let $A\in \C^{\dcd}$ be a Hurwitz matrix. There exists a matrix $U\in \gln$ such that $A=U\Lambda U^{-1}$ and $\Lambda^*+\Lambda$ is a real symmetric positive definite matrix.
 \end{lemma}
 \begin{proof}
-It is trivial to see that for any $\Lambda\in \C^{\dcd}$, $\left(\Lambda^\dag+\Lambda\right)$ is real symmetric.
+It is trivial to see that for any $\Lambda\in \C^{\dcd}$, $\left(\Lambda^*+\Lambda\right)$ is Hermitian.
 We will use the decomposition of $A=V \tL V^{-1}$ in \Cref{jordon} and also carry over the notations in \Cref{jordon}.
-Consider the diagonal matrices $D^i_j=\begin{bmatrix} 1  &0 &0 &\ldots &0 &0\\ 0 &\re{\lambda_i} &0 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda}^{d^i_j-1}_i &0 \\ 0 &\ldots &0 &0 &0 &\re{\lambda}^{d^i_j}_i \end{bmatrix},\,\forall j=1,\ldots,l(i)$, $D^i=D^i_1 \op\ldots\op D^i_{l(i)},\,\forall i=1,\ldots,k$ and $D=D^1 \op\ldots\op D^k$.
-It follows that $A=(VD) \Lambda (VD)^{-1}$, where $J$ is a matrix such that
-$\Lambda=A_1 \op \ldots \op A_k$, where each $A_i,\,i=1,\ldots,k$ can further be written as
+Consider the diagonal matrices $D^i_j=\begin{bmatrix} 1  &0 &0 &\ldots &0 &0\\ 0 &\re{\lambda_i} &0 &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda_i}^{d^i_j-1} &0 \\ 0 &\ldots &0 &0 &0 &\re{\lambda_i}^{d^i_j} \end{bmatrix},\,\forall j=1,\ldots,l(i)$, $D^i=D^i_1 \op\ldots\op D^i_{l(i)},\,\forall i=1,\ldots,k$ and $D=D^1 \op\ldots\op D^k$.
+It follows that $A=(VD) \Lambda (VD)^{-1}$, where $\Lambda$ is a matrix such that
+$\Lambda=\Lambda_1 \op \ldots \op \Lambda_k$, where each $\Lambda_i,\,i=1,\ldots,k$ can further be written as
 $A_i=\Lambda^i_{1} \op \ldots \op \Lambda^i_{{l(i)}}$. Each of $\Lambda^i_{j}$ is a $d^i_j\times d^i_j$ square matrix with the special form given by
@@
 
-Now let $\frac{(\Lambda^\dag+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}$, where $\frac{\Lambda^{i\dag}_{j}+\Lambda^i_{j}}{2}=\begin{bmatrix} \re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 &0\\ \frac{\re{\lambda_i}}{2} &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} \\ 0 &\ldots &0 &0 &\frac{\re{\lambda_i}}{2} &\re{\lambda_i} \end{bmatrix} $.
-Then for any $x=(x(i),i=1,\ldots,n)\in \C^d (\neq \mathbf{0})$, there exists a $b\in \{-1,1\}^d$, such that
+Now we have $\frac{(\Lambda^*+\Lambda)}{2}=\op_{i=1}^k \op_{j=1}^{l(i)}\frac{\Lambda^{i*}_{j}+\Lambda^i_{j}}{2}$, where $\frac{\Lambda^{i*}_{j}+\Lambda^i_{j}}{2}=\begin{bmatrix} \re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 &0\\ \frac{\re{\lambda_i}}{2} &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} &0 &\ldots &0 \\ 0 &\vdots &\vdots &0 &\re{\lambda_i} &\frac{\re{\lambda_i}}{2} \\ 0 &\ldots &0 &0 &\frac{\re{\lambda_i}}{2} &\re{\lambda_i} \end{bmatrix} $.
+Then for any $x=(x(i),i=1,\ldots,d)\in \C^d (\neq \mathbf{0})$, we have %there exists a $b\in \{-1,1\}^d$, such that
 \begin{align*}
-x^\dag \frac{(\Lambda^\dag+\Lambda)}{2} x &=\re{\lambda_i} \left(\sum_{i=1}^d \bar{x}{(i)} x(i)+\sum_{i=1}^{n-1} \frac{\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)}{2}\right) \\
-&=\frac{\re{\lambda_i}}{2}\left(\md{x(1)}^2+ \md{x(n)}^2\right)+\frac{\re{\lambda_i}}{2}\left( \sum_{i=1}^{n-1} \md{x(i)}^2+\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)+\md{x(i+1)}^2 \right)\\
+x^* \frac{(\Lambda^*+\Lambda)}{2} x &=\re{\lambda_i} \left(\sum_{i=1}^d \bar{x}{(i)} x(i)+\sum_{i=1}^{d-1} \frac{\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)}{2}\right) \\
+&=\frac{\re{\lambda_i}}{2}\left(\md{x(1)}^2+ \md{x(d)}^2\right)+\frac{\re{\lambda_i}}{2}\left( \sum_{i=1}^{d-1} \md{x(i)}^2+\bar{x}(i) x(i+1) + x(i)\bar{x}(i+1)+\md{x(i+1)}^2 \right)\\
 &>\frac{\re{\lambda_i}}{2}\left(\sum_{i=1}^d \md{x(i)+x(i+1)}^2 \right)\\
@@
 \end{proof}
+\section{Proofs}\label{sec:proofs}
+\subsection{LSA with CS-PR for Positive Definite Distributions}
+%In this subsection, we assume that $P$ satisfies the assumptions in \Cref{assmp:lsa} and in addition $P$ is \emph{positive definite}.
+In this subsection, we re-write \eqref{eq:lsa} and \Cref{assmp:lsa} to accomodate complex number computations and in addition assume that $P$ is \emph{positive definite}. To this end,
+\begin{subequations}\label{eq:lsacmplx}
+\begin{align}
+\label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
+\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_i,
+\end{align}
+\end{subequations}
+where $\thh_t, \theta_t \in \C^{d}$. We now assume,
+\begin{assumption}\label{assmp:lsacmplx}
+\begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\item \label{dist} $(b_t, A_t)\sim (P^b,P^A), t\geq 0$ is an $i.i.d$ sequence, where $P^b$ is a distribution over $\C^d$ and $P^A$ is a distribution over $\C^{\dcd}$.
+We assume that $P$ is positive definite.
+\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following
+\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_{A_P}, \, \E[N_t^* N_t]=\sigma^2_{b_P}.\end{align*}
+\item $A_P$ is invertible and there exists a $\ts=A^{-1}_Pb_P$.
+\end{enumerate}
+\end{assumption}
 
-\section{Error Recursion}
+We now define the error variables and present the recurison for the error dynamics. In what follows, definitions in \Cref{sec:def} and \Cref{sec:prob} continue to hold.
+\begin{definition}\label{def:err}
+\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\item Define error variables $e_t\eqdef \theta_t-\ts$ and $\eh_t\eqdef \thh_t-\ts$.
+\item Define $\forall\, t\geq 0$ random vectors $\zeta_t\eqdef b_t-b-(A_t-A_P)\ts$.
+%\item Define constants $\EE{\norm{\zeta_t}^2}\eqdef\sigma_1^2=\sigma_A^2\norm{\ts}^2+\sigma_b^2$ and $\EE{\norm{M_t\zeta_t}}\eqdef \sigma_2^2=\sigma_A^2\norm{\ts}$.
+\item Define constants $\EE{\norm{\zeta_t}^2}\eqdef\sigma_1^2$ and $\EE{\norm{M_t\zeta_t}}\eqdef \sigma_2^2$.
+\item Define $\forall\,i\geq j$, the random matrices $F_{i,j}=(I-\alpha A_i)\ldots (I-\alpha A_j)$ and $\forall,\,i<j$ $F_{i,j}=\I$.
+\end{itemize}
+\end{definition}
 
+\paragraph{Error Recursion} Let us now look at the dynamics of the error terms defined by
+\begin{align}\label{eq:errrec}
+\begin{split}
+\theta_t&=\theta_{t-1}+\alpha\big(b_t-A_t\theta_{t-1}\big)\\
+\theta_t-\ts&=\theta_{t-1}-\ts+\alpha\big(b_t-A_t(\theta_{t-1}-\ts+\ts)\big)\\
+e_t&=(I-\alpha A_t)e_{t-1}+\alpha(b_t -b -(A_t-A)\ts)\\
+e_t&=(I-\alpha A_t)e_{t-1}+\alpha\zeta_t
+\end{split}
+\end{align}
+
+\begin{lemma}\label{lm:pd}
+Let $P$ be a distribution over $\C^d\times \C^{\dcd}$ satisfying \Cref{assmp:lsacmplx}, then there exists an $\alpha_P>0$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+\end{lemma}
+\begin{proof}
+\begin{align*}
+\rhos{P}&\stackrel{(a)}{=}\inf_{x:\norm{x}=1}x^* (A_P^*+A_P)x -\alpha x^*\EE{A_t^* A_t} x\\
+&\stackrel{(b)}{=}\inf_{x:\norm{x}=1}x^* (A_P^*+A_P)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x\\
+&\stackrel{(c)}{\geq} \lambda_{\min}(A^*_P+A_P)-\alpha \norm{A_P}^2-\sigma^2_A
+\end{align*}
+The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*_P+A_P)}{\norm{A_P}^2+\sigma^2_A}$.
+Here $(a)$ follows from definition of $\rhos{P}$ in \Cref{def:dist}, $(b)$ follows from the fact that $M_t$ is a martingale difference term (see \Cref{assmp:lsacmplx}) and $(c)$ follows from the fact that for a real symmetric matrix $M$ the smallest eigen value is given by $\lambda_{\min}=\inf_{x:\norm{x}=1} x^* M x $.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
+Let $t>i\ge 1$, $x,y\in \C^d$ be $\F_{i}$-measurable random vectors. Then,
+\begin{align*}
+\E[x^* F_{t,i+1}y|\F_i]=x^* (I-\alpha A_P)^{t-i} y\,.
+\end{align*}
+\end{lemma}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proof}
+By the definition of $F_{t,i+1}$,
+and because $F_{t-1,i+1} = (I-\alpha A_{t-1}) \dots (I-\alpha A_{i+1})$ is $\F_{t-1}$-measurable,
+as are $x$ and $y$,
+\begin{align*}
+\EE{x^* F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha A_t) | \F_{t-1} } F_{t-1,i+1} y\\
+&=x^*  (I-\alpha A_P)  F_{t-1,i+1} y\,.
+\end{align*}
+By the tower-rule for conditional expectations and our measurability assumptions,
+\begin{align*}
+\EE{x^* F_{t,i+1} y | \F_{t-2} }
+&=x^*  (I-\alpha A_P)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
+&= x^* (I-\alpha A_P)^2 F_{t-2,i+1} y\,.
+\end{align*}
+Continuing this way we get
+\begin{align*}
+\EE{x^* F_{t,i+1} y | \F_{t-j} }
+= x^* (I-\alpha A_P)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
+\end{align*}
+Specifically, for $j=t-i$ we get
+\begin{align*}
+\EE{x^* F_{t,i+1} y | \F_{i} }  = x^* (I-\alpha A_P)^{t-i} y\,.
+\end{align*}
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{lemma}\label{noisecancel}
+Let $t>i\ge 1$ and let $x\in \C^d$ be a $\F_{i-1}$-measurable random vector. Then,
+$\E[x^* F_{t,i+1}\zeta_{i}]=0$.
+\end{lemma}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{proof}
+By \Cref{lem:genunroll},
+\begin{align*}
+\EE{x^* F_{t,i+1} \zeta_i | \F_{i} }  = x^* (I-\alpha A_P)^{t-i} \zeta_i\,.
+\end{align*}
+Using the tower rule,
+\begin{align*}
+\EE{x^* F_{t,i+1} \zeta_i | \F_{i-1} }
+= x^* (I-\alpha A_P)^{t-i}\EE{ \zeta_i | \F_{i-1} }= 0\,.
+\end{align*}
+\end{proof}
+
+\begin{lemma}\label{lem:unroll}
+For all $t>i\ge 0$, $\E \ip{e_i,F_{t,i+1} e_i}=\E\ip{e_i,(I-\alpha A_P)^{t-i} e_i}$.
+\end{lemma}
+\begin{proof}
+The lemma follows directly from \Cref{lem:genunroll}. Indeed,
+$\theta_i$ depends only on $A_1,\dots,A_{i},b_1,\dots,b_{i}$, $\theta_i$ and so is $e_i$ $\F_i$-measurable.
+Hence, the lemma is applicable and implies that
+\begin{align*}
+\EE{ \ip{e_i, F_{t,i+1} e_i} | \F_i } =
+\EE{ \ip{e_i, (I-\alpha A_P)^{t-i} e_i} | \F_i }\,.
+\end{align*}
+Taking expectation of both sides gives the desired result.
+\end{proof}
+
+\begin{lemma}\label{innerproduct}
+Let $i>j \ge 0$ and let $x\in \R^d$ be an $\F_j$-measurable random vector.
+Then,
+\begin{align*}
+\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rhos{P})^{i-j}\E\norm{x}^2\,.
+\end{align*}
+\end{lemma}
+\begin{proof}
+Note that
+$S_t\doteq \EE{ (I-\alpha A_t)^* (I-\alpha A_t) | \F_{t-1} }
+= I - \alpha (A_P^* + A_P) + \alpha^2 \EE{ A_t^* A_t | \F_{t-1} }$.
+Since $(b_t,A_t)_t$ is an independent sequence, $\EE{ A_t^* A_t|\F_{t-1}} = \EE{ A_1^* A_1 }$.
+Now, using the definition of $\rhos{P}$ from \Cref{def:dist}
+$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (A_P^* + A_P - \alpha \EE{A_1^\top A_1}) x
+= 1-\alpha \rhos{P}$.
+Hence,
+\begin{align*}
+&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
+&= \EE{x^* F_{i-1,j+1}^\top (I-\alpha A_i)^* (I-\alpha A_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
+&=(x F_{i-1,j+1})^* \, S_i \, F_{i-1,j+1} x\\
+&\le (1-\alpha \rhos{P}) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
+& \le (1-\alpha \rhos{P})^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
+& \quad \vdots \\
+& \le (1-\alpha \rhos{P})^{i-j}\, \norm{x}^2\,.
+%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
+%&=(1-\alpha\rhos{P}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
+\end{align*}
+\end{proof}
+
+\begin{comment}
 \begin{lemma}
-Let $P$ be a positive definite distribution over $\C_B^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
+Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
 \end{lemma}
@@
 \begin{align*}
-\rhos{P}&={\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}\\
-&\geq {\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)\right)x}-\alpha B^2\\
-&\geq \lambda_{\min}(A_P+A_P^\dag)-\alpha B^2
+\rhos{P}&={\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha C_P\right)x}\\
+&\geq {\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)\right)x}-\alpha B^2\\
+&\geq \lambda_{\min}(A_P+A_P^*)-\alpha B^2
 \end{align*}
-By choosing $\alpha_P<\frac{\lambda_{\min}(A_P+A_P^\dag)}{B^2}$, it follows that $\rhos{P}>0,\forall \alpha \in (0,\alpha_P)$.
-Further, by noting that $C_P\succeq A_P^\dag A_P$, it is easy to check that $\rhod{P}>\rhos{P}>0,\,\forall \alpha \in(0,\alpha_P)$.
+By choosing $\alpha_P<\frac{\lambda_{\min}(A_P+A_P^*)}{B^2}$, it follows that $\rhos{P}>0,\forall \alpha \in (0,\alpha_P)$.
+Further, by noting that $C_P\succeq A_P^* A_P$, it is easy to check that $\rhod{P}>\rhos{P}>0,\,\forall \alpha \in(0,\alpha_P)$.
 \end{proof}
+\end{comment}
 
-\begin{theorem}\label{pdrate}
-Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\dcd}$. Then
+\begin{theorem}\label{th:pdrate}
+Let $\eh_t$ be as in \Cref{def:err}. %Let $P$ be a distribution over $\C^d\times\C^{\dcd}$ satisfying \Cref{assmp:lsacmplx}.
+Then
 \begin{align}
-\EE{\norm{\thh_t-\theta^*}^2}
-\leq
-\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
+%\EE{\norm{\eh_t}^2}
+%\leq
+%\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}} \left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2\norm{\ts}+\alpha \norm{e_0})}{t+1} \right)\,.
+\E[\norm{\eh_t}^2]
+\leq \left(1+\frac2{\alpha\rhod{P}}\right)\, \frac1{\alpha\rhos{P}}\, \,
+%\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{e_0}}{t+1} \right)\,.
+\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_1^2)+\alpha \sigma_2^2\norm{e_0}}{t+1} \right)\,.
 \end{align}
@@
 \begin{proof}
-
-\begin{align*}
-\theta_t&=\theta_{t-1}+\alpha\big(b_t-A_t\theta_{t-1}\big)\\
-\theta_t-\ts&=\theta_{t-1}-\ts+\alpha\big(b_t-A_t(\theta_{t-1}-\ts+\ts)\big)\\
-e_t&=(I-\alpha A_t)e_{t-1}+(b_t -b -(A_t-A)\ts)\\
-\end{align*}
-Now define $\forall,\,i\geq j$, $F_{i,j}=(I-\alpha A_i)\ldots (I-\alpha A_j)$ and $\forall,\,i<j$ $F_{i,j}=\I$.
-Also let $\zeta_t\eqdef b_t-b-(A_t-A)\ts$ and $\EE{\norm{\zeta_t}^2}=\sigma_1^2=\sigma_P^2\norm{\ts}^2+\sigma_b^2$ and $\EE{\norm{M_t\zeta_t}}=\sigma_2^2=\sigma_P^2\norm{\ts}$.
 
@@
 &=\E \ip{e_i,\big[F_{j,i+1} e_i+\alpha\textstyle\sum_{k=i+1}^j F_{j,k+1}\zeta_{k}\big]}\\
-&=\E\ip{e_i,F_{j,i+1} e_i}  \text{(from \cref{noisecancel})}\\
-&=\E\ip{e_i, (I-\alpha A)^{j-i} e_i} \text{(from \cref{lem:unroll})}
+&=\E\ip{e_i,F_{j,i+1} e_i}  \text{(from \Cref{noisecancel})}\\
+&=\E\ip{e_i, (I-\alpha A)^{j-i} e_i} \text{(from \Cref{lem:unroll})}
 \end{align*}
@@
 \end{align*}
-Expanding $e_i$ using \eqref{eq:etft} and then using \cref{innerproduct} and \Cref{cmplxassmp}
+Expanding $e_i$ using \eqref{eq:etft} and then using \Cref{innerproduct} and \Cref{assmp:lsacmplx}
 \begin{align*}
@@
 \leq \left(1+\frac2{\alpha\rhod{P}}\right)\, \frac1{\alpha\rhos{P}}\, \,
-\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{e_0}}{t+1} \right)\,.
+%\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{e_0}}{t+1} \right)\,.
+\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_1^2)+\alpha \sigma_2^2\norm{e_0}}{t+1} \right)\,.
 \end{align}
@@
 
-\begin{lemma}[Similarity Transformation]
-Let $P^{AS}\in\P$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
+\paragraph{Proof of \Cref{lm:hur}}
+\begin{lemma}
+Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, then there exists an $\alpha_{P_U}>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_P)$.
 \end{lemma}
 \begin{proof}
-Follows from \Cref{simtran}
+We know that $A_P$ is Hurwitz and from  \Cref{lm:simtran} it follows that there exists an $U\in \gld$ such that  $\Lambda=U^{-1} A_P U$ and $(\Lambda^*+\Lambda)$ is real symmetric and positive definite.
+Using \Cref{def:simdist}, we have $A_{P_U}=\Lambda$ and from \Cref{lm:pd} we know that there exists an $\alpha_{P_U}$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_{P_U})$.
+
 \end{proof}
 
-\begin{theorem}[Change of Basis]
-Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
+\begin{lemma}[Change of Basis]\label{lm:cb}
+Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ as in \Cref{assmp:lsa} and let $U$ be chosen according to \Cref{lm:hur}.
+Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, then
 \begin{align}
@@
 \left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
-\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
+\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
 %\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
@@
 where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
-\end{theorem}
+\end{lemma}
 \begin{proof}
-From proof of \Cref{pdrate} we know that
+We have
 \begin{align*}
-e_t&=(I-\alpha A_t)e_{t-1}+\zeta_t\\
-U^{-1}e_t&=(I-\alpha U^{-1}A_t U) U^{-1}e_{t-1}+ U^{-1}\zeta_t\\
-z_t&=(I-\alpha \Lambda_t) z_{t-1}+ H_t,
+e_t&=(I-\alpha A_t)e_{t-1}+\alpha\zeta_t\\
+U^{-1}e_t&=(I-\alpha U^{-1}A_t U) U^{-1}e_{t-1}+ \alpha U^{-1}\zeta_t\\
+z_t&=(I-\alpha \Lambda_t) z_{t-1}+\alpha H_t,
 \end{align*}
 where $z_t=\gamma_t-\gamma^*$, $\Lambda_t=U^{-1}A_t U$ and $H_t=U^{-1}\zeta_t$.
-we know that $\EE{\norm{H_t}^2}\leq \norm{U^{-1}}^2\EE{\norm{\zeta}}$ and $\EE{\norm{\Lambda_t H_t}}=\EE{\norm{U^{-1}A_t UU^-1\zeta_t}}=\EE{\norm{U^{-1}A_t \zeta_t}}\leq \norm{U^-1}\EE{\norm{A_t\zeta_t}}=\norm{U^{-1}}\sigma_2^2$.
-We follow through the steps of \Cref{pdrate} to obtain
+Note that the error recursion in $z_t$ might involve complex computations (depending on whether $U$ has complex entries or not), and hence \eqref{eq:lsacmplx} and \Cref{assmp:lsacmplx} are useful in analyzing $z_t$.
+We know that $\EE{\norm{H_t}^2}\leq \norm{U^{-1}}^2\EE{\norm{\zeta_t}}$ and $\EE{\norm{\Lambda_t H_t}}=\EE{\norm{U^{-1}A_t UU^{-1}\zeta_t}}=\EE{\norm{U^{-1}A_t \zeta_t}}\leq \norm{U^{-1}}\EE{\norm{A_t\zeta_t}}=\norm{U^{-1}}\sigma_2^2$.
+We follow through the steps of \Cref{th:pdrate} to obtain
 \begin{align}
@@
 
-Let $\F_t = \sigma( A_1,\dots,A_t, \zeta_1,\dots,\zeta_t ) $, with $\F_0$ the $\sigma$-field that holds all random variables.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
-Let $t>i\ge 1$, $x,y\in \C^d$ be $\F_{i}$-measurable random vectors. Then,
-\begin{align*}
-\E[x^\dag F_{t,i+1}y|\F_i]=x^\dag (I-\alpha H)^{t-i} y\,.
-\end{align*}
-\end{lemma}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{proof}
-By the definition of $F_{t,i+1}$,
-and because $F_{t-1,i+1} = (I-\alpha A_{t-1}) \dots (I-\alpha A_{i+1})$ is $\F_{t-1}$-measurable,
-as are $x$ and $y$,
-\begin{align*}
-\EE{x^\dag F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha A_t) | \F_{t-1} } F_{t-1,i+1} y\\
-&=x^\dag  (I-\alpha A_P)  F_{t-1,i+1} y\,.
-\end{align*}
-By the tower-rule for conditional expectations and our measurability assumptions,
-\begin{align*}
-\EE{x^\dag F_{t,i+1} y | \F_{t-2} }
-&=x^\dag  (I-\alpha A_P)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
-&= x^\dag (I-\alpha A_P)^2 F_{t-2,i+1} y\,.
-\end{align*}
-Continuing this way we get
-\begin{align*}
-\EE{x^\dag F_{t,i+1} y | \F_{t-j} }
-= x^\dag (I-\alpha A_P)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
-\end{align*}
-Specifically, for $j=t-i$ we get
-\begin{align*}
-\EE{x^\dag F_{t,i+1} y | \F_{i} }  = x^\dag (I-\alpha A_P)^{t-i} y\,.
-\end{align*}
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}\label{noisecancel}
-Let $t>i\ge 1$ and let $x$ be an $\F_{i-1}$-measurable random vector. Then,
-$\E[x^\dag F_{t,i+1}\zeta_{i}]=0$.
-\end{lemma}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{proof}
-By \cref{lem:genunroll},
-\begin{align*}
-\EE{x^\dag F_{t,i+1} \zeta_i | \F_{i} }  = x^\dag (I-\alpha A_P)^{t-i} \zeta_i\,.
-\end{align*}
-Using the tower rule,
-\begin{align*}
-\EE{x^\dag F_{t,i+1} \zeta_i | \F_{i-1} }
-= x^\dag (I-\alpha A_P)^{t-i}\EE{ \zeta_i | \F_{i-1} }= 0\,.
-\end{align*}
-\end{proof}
-
-\begin{lemma}\label{lem:unroll}
-For all $t>i\ge 0$, $\E \ip{e_i,F_{t,i+1} e_i}=\E\ip{e_i,(I-\alpha A_P)^{t-i} e_i}$.
-\end{lemma}
-\begin{proof}
-The lemma follows directly from \cref{lem:genunroll}. Indeed,
-$\theta_i$ depends only on $A_1,\dots,A_{i},b_1,\dots,b_{i}$, $\theta_i$ and so is $e_i$ $\F_i$-measurable.
-Hence, the lemma is applicable and implies that
-\begin{align*}
-\EE{ \ip{e_i, F_{t,i+1} e_i} | \F_i } =
-\EE{ \ip{e_i, (I-\alpha A_P)^{t-i} e_i} | \F_i }\,.
-\end{align*}
-Taking expectation of both sides gives the desired result.
-\end{proof}
-
-\begin{lemma}\label{innerproduct}
-Let $i>j \ge 0$ and let $x\in \R^d$ be an $\F_j$-measurable random vector.
-Then,
-\begin{align*}
-\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rho_\alpha)^{i-j}\E\norm{x}^2\,.
-\end{align*}
-\end{lemma}
-\begin{proof}
-Note that
-$S_t\doteq \EE{ (I-\alpha A_t)^\dag (I-\alpha A_t) | \F_{t-1} }
-= I - \alpha (A_P^\dag + A_P) + \alpha^2 \EE{ A_t^\dag A_t | \F_{t-1} }$.
-Since $(b_t,A_t)_t$ is an independent sequence, $\EE{ A_t^\dag A_t|\F_{t-1}} = \EE{ A_1^\dag A_1 }$.
-Now, using the definition of $\rhos{P}$ from \eqref{contract},
-$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (A_P^\dag + A_P - \alpha \EE{A_1^\top A_1}) x
-= 1-\alpha \rhos{P}$.
-Hence,
-\begin{align*}
-&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
-&= \EE{x^\dag F_{i-1,j+1}^\top (I-\alpha A_i)^\dag (I-\alpha A_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
-&=(x F_{i-1,j+1})^\dag \, S_i \, F_{i-1,j+1} x\\
-&\le (1-\alpha \rhos{P}) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
-& \le (1-\alpha \rhos{P})^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
-& \quad \vdots \\
-& \le (1-\alpha \rho_\alpha)^{i-j}\, \norm{x}^2\,.
-%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
-%&=(1-\alpha\rhos{P}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
-\end{align*}
-\end{proof}
+\paragraph{Proof of \Cref{th:rate}}
+Follows by sustuting $\theta_t=U\gamma_t$ in \Cref{lm:cb}.
 
@@
 \begin{proof}
-By \cref{lem:genunroll},
+By \Cref{lem:genunroll},
 \begin{align*}
@@
 \begin{proof}
-The lemma follows directly from \cref{lem:genunroll}. Indeed,
+The lemma follows directly from \Cref{lem:genunroll}. Indeed,
 $\theta_i$ depends only on $H_1,\dots,H_{i},g_1,\dots,g_{i}$, $\theta_i$ and so is $e_i$ $\F_i$-measurable.
@@
 &=\E \ip{e_i,\big[F_{j,i+1} e_i+\alpha\textstyle\sum_{k=i+1}^j F_{j,k+1}\zeta_{k}\big]}\\
-&=\E\ip{e_i,F_{j,i+1} e_i}  \text{(from \cref{noisecancel})}\\
-&=\E\ip{e_i, (I-\alpha J)^{j-i} e_i} \text{(from \cref{lem:unroll})}
+&=\E\ip{e_i,F_{j,i+1} e_i}  \text{(from \Cref{noisecancel})}\\
+&=\E\ip{e_i, (I-\alpha J)^{j-i} e_i} \text{(from \Cref{lem:unroll})}
 \end{align*}
@@
 \end{align*}
-Expanding $e_i$ using \eqref{eq:etft} and then using \cref{innerproduct} and \Cref{cmplxassmp}
+Expanding $e_i$ using \eqref{eq:etft} and then using \Cref{innerproduct} and \Cref{cmplxassmp}
 \begin{align*}

### 2017-05-17 not.tex
latest

@@
 \section{Notations and Definitions}\label{sec:def}
-We denote the sets real and complex numbers by $\R$ and $\C$ respectively. For $x=a+ib\in \C$, we denote its real and imaginary parts by $\re{x}=a$ and $\im{x}=b$ respectively.
-For any $x\in \C$ we denote its modulus $\md{x}=\sqrt{\re{x}^2+\im{x}^2}$ and its complex conjugate by $\bar{x}=a-ib$.
-We denote $d$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively.
-We denote the transpose of $x\in \C^{d}$ by $x^\top$ and the conjugate transpose by $x^*={\bar{x}}^\top$.
+We denote the sets real and complex numbers by $\R$ and $\C$ respectively.
+For $x\in \C$ we denote its modulus and complex conjugate by $\md{x}$ and $\bar{x}$ respectively.
+We denote $d$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively, and use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively.
+We denote the transpose of $x$ by $x^\top$ and the conjugate transpose by $x^*={\bar{x}}^\top$.
 We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^* y$.
-%We denote the transpose of $x\in \C^{d}, A\in \C^{\dcd}$ by $x^\top$ and $A^\top$ respectively.
-%Further, by $x^*$ denotes the conjugate transpose, i.e., $x^*=\bar{x}^\top$ (the same convention carries over to matrices as well).
-
- %Also, by $A_{ij}$ we denote the entry of matrix $A$ in the $i^{th}$ row and $j^{th}$ column.
-We use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively. The notation for (conjugate) transpose carries over from vectors.
-The norm of the matrix $A$ in given by $\norm{A}\eqdef \sup_{x\in \C^d:\norm{x}=1} \norm{Ax}$.
-For any $x\in\C^d$, we denote the general quadratic norm with respect to a hermitian matrix $C$ by $\norm{x}^2_C\eqdef x^*\, C \,x$.
+For $x\in\C^d$, we denote the general quadratic norm with respect to a Hermitian matrix $C$ by $\norm{x}^2_C\eqdef x^*\, C \,x$.
+The norm of the matrix $A$ is given by $\norm{A}\eqdef \sup_{x\in \C^d:\norm{x}=1} \norm{Ax}$.
 We use $\cond{A}=\norm{A}\norm{A^{-1}}$ to denote the condition number of the matrix $A$.
-We denote the identity matrix in $\C^{\dcd}$ by $\I$ and the set of invertible $\dcd$ complex matrices by $\gln$.
-\begin{comment}
-We use $A\succeq 0$ to denote that the
-square matrix $A$ is hermitian and positive semidefinite (HPD):
-$A = A^*$, $\inf_x x^* A x\ge 0$.
-For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
-We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
-We also use $\preceq$ and $\prec$ analogously.
-\end{comment}
+We denote the identity matrix in $\C^{\dcd}$ by $\I$ and the set of invertible $\dcd$ complex matrices by $\gld$.
 For a positive real number $B>0$, we denote by $\C^{d}_B=\{b\in \C^d\mid \norm{b}\leq B\}$ and by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ the set of complex vectors and matrices whose norms are bounded $B$.
-%We denote by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$ and  $\R^{\dcd}_B=\{A\in \R^{\dcd}\mid \md{A_{ij}}\leq B,\,\forall i,j=1,\ldots,n\}$.
-%We use $\P$ to denote the set of probability distribution over $\C^{\dcd}$.
-%We denote the set of distributions over $\C_B^{\dcd}$ by $\P$.
-Let $P$ be a distribution over $C^{\dcd}$, we use $A\sim P$ to denote $A$ is distributed according to $P$.
-For $U\in \gln$, we denote by $P_U$ the distribution of $U^{-1}\,A\,U$, where $A\sim P$.
+We use $\left(\cdot\right)\sim P$ to denote the fact that $\left(\cdot\right)$ is distributed according to probability distribution $P$.
 $\E$ denotes mathematical expectation.
-\begin{comment}
- We use $A\sim P$ to denote the fact that the random variable $A$ is distributed according to distribution $P$.
-Let the random matrix $A\sim P$, and $U\in \gln$, we use $P_U$ to denote the distribution of the random matrix $U^{-1}A U$.
-\end{comment}
 \begin{definition}\label{def:dist}
@@
 Let $P=(P^V,P^M)$ as in \Cref{def:dist}; $b\sim P^V$ and $A\sim \P^M$ be random variables distributed according to $P^V$ and $P^M$.
-For $U\in \gln$ define $P_U=(P_U^V,P_U^M)$ to be the distributions of $U^{-1}b$ and $U^{-1}AU$.
+For $U\in \gld$ define $P_U\eqdef(P_U^V,P_U^M)$, to be the distributions of $U^{-1}b$ and $U^{-1}AU$.
 \end{definition}
@@
 \begin{definition}\label{distpd}
-We call a distribution $P$ to be H/PD/SPD if $A_P$ is H/PD/SPD.
+We call the distribution $P$ in \Cref{def:dist} to be H/PD/SPD if $A_P$ is H/PD/SPD.
 \end{definition}

### 2017-05-17 prob.tex
latest

@@
-\section{Problem Setup}
-We consider the linear stochastic approximation (LSA) algorithm with constant step-size (CS) and Polyak-Ruppert (PR) averaging of the iterates given as below:
+\section{Problem Setup}\label{sec:prob}
+We consider linear stochastic approximation algorithm (LSAs) with constant step-size (CS) and Polyak-Ruppert (PR) averaging of the iterates given as below:
 \begin{subequations}\label{eq:lsa}
@@
 \label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
-\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{s=0}^{t}\theta_i,
+\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_i,
 \end{align}
 \end{subequations}
-where\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^d$ and $A_t\in \R^{\dcd}, b_t\in \R^d$.
-However, it serves well for our analysis to consider these qunatites to be complex valued.} $\thh_t, \theta_t \in \C^{d}$.
-The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\dcd}$ incrementally, in discrete time steps $t=1,2,\dots$
-based on data $b_t\in \C^d$, $A_t\in \C^{\dcd}$. Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
+where $\thh_t, \theta_t \in \R^{d}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \R^{\dcd}$ incrementally, in discrete time steps $t=1,2,\dots$
+based on data $b_t\in \R^d$, $A_t\in \R^{\dcd}$. Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
 initial value $\theta_0$.
@@
 Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$ can be efficiently computed in $O(d)$ time.
-This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp} and linear prediction \cite{bach,bachaistats} will be presented in a later sections (\Cref{sec:rl,sec:lp}).
-The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
+This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp} will be presented in \Cref{sec:related}.
+The significance of efficient computation of the matrix-vector product is that an update of the algorithm
 can then be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $d$ is in the range of thousands or more.\par
 We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa}.
-In what follows, we make use of the $\sigma$-field given by $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.
+In what follows, we make use of the $\sigma$-fields $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$, and $\F_0$ that holds all random variables.
 \begin{assumption}\label{assmp:lsa}
 \begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \label{dist} $(b_t, A_t)\sim (P^b,P^A)$ is an $i.i.d$ sequence, where $P^b$ is a distribution over $\C^d$ and $P^A$ is a Hurwitz distribution over $\C^{\dcd}$.
+\item \label{dist} $(b_t, A_t)\sim (P^b,P^A), t\geq 0$ is an $i.i.d$ sequence, where $P^b$ is a distribution over $\R^d$ and $P^A$ is a distribution over $\R^{\dcd}$. We assume that $P$ is Hurwitz.
 \item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following
@@
 %For the rest of the paper we denote the errors of the internal and the output variables of \eqref{eq:lsa} at time $t$ by $e_t\eqdef \theta_t-\ts$ and $\eh_t\eqdef \thh_t-\ts$ respectively.
-\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ $\EE{\norm{\thh_t-\ts}_C^2}$, the MSE with respect a hermitian matrix $C$ is desirable, we cannot exploit the use of $C$ unless it is connected to $P$ in a special way.
+\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ $\EEP{\norm{\thh_t-\ts}_C^2}$, the MSE with respect a hermitian matrix $C$ is desirable, we cannot exploit the use of $C$ unless it is connected to $P$ in a special way.
 For more discussion, see \Cref{sec:related}.}
@@
 \begin{comment}
-\footnote{ We note here that $\EE{\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$.
+\footnote{We note here that $\EE{P\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$.
 However, unless $C$ has some special structure, it is not possible to exploit this generality in our bounds.
-Further, when data is bounded $\norm{C}^2$ is also bounded and it is straightforward to see that $\EE{\norm{\thh_t-\ts}}^2_C\leq \norm{C}^2\EE{\norm{\thh_t-\ts}^2}$.
+Further, when data is bounded $\norm{C}^2$ is also bounded and it is straightforward to see that $\EEP{\norm{\thh_t-\ts}}^2_C\leq \norm{C}^2\EEP{\norm{\thh_t-\ts}^2}$.
 So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in the linear regression setting, where its usage is relevant.}
@@
 Another aspect of interest is to understand the contribution of the \emph{bias} (caused by the initial condition $\theta_0$) and the \emph{variance} (caused by $\sigma^2_P,\sigma^2_b$) to the MSE.\par
-\begin{comment}
-\textbf{Challenges:}
-The idea of constant step size and iterate averaging is not new and goes back to \cite{polyak-judisky}, where it was first shown that $L_t\ra 0$ as $t\ra\infty$.
-Recently \cite{bach} revisited the idea in connection to the stochastic gradient descent (SGD) algorithm (which is also an LSA of the form in \eqref{eq:lsa}) to solve linear regression.
-Much similar to the spirit of \cite{bach}, our aim here is to understand the aspects that effect the behaviour of $L_t$, such as:
-\begin{itemize}
-\item The variances $\sigma_P^2$ and $\sigma_b^2$ of the noise terms $M_t$ and $b_t$ in the update equation.
-\item The bias $\norm{\theta_0-\ts}$ due to the initial condition $\theta_0$.
-\item The effect of the structural properties of the $A_P$, i.e., whether it is AS or PD or SPD.
-\end{itemize}
-Most results in the past including those that of \cite{bach} consider an SGD algorithm for quadratic loss minimization, where the matrix $A_P$ is SPD and the noise has additional structural properties (we discuss this briefly in \Cref{sec:lp}).
-As it turns out these nice properties are lost when $A_P$ is AS or PD (a property that is holds in RL settings).
-While from the standpoint of the tools and techniques used, we use techniques much similar to \cite{bach}, however, as mentioned, the fact that we no longer are dealing with gradient systems (that involve to SPD matrices) but general vector fields (that involve AS or PD matrices), poses important challenges.
-Many RL algorithms \cite{} are LSA algorithms that need only $O(n)$ updates.
-However, there is a lot of variability in their analyses.
-The analysis as well as the results in this paper are useful from the perspective that it presents a single unified way of understand the various LSA-RL algorithms.
-We now present and discuss the main results of this paper in the next section.
-The comparison with existing literature is done in two separate sections, one to compare our results with \cite{bach} in the linear predictions setting (\Cref{sec:lp}) and the other to compare our results with prior results in RL literature \cite{}.
 \todoch{Mention importance of forgetting the bias etc. Mention that we really don't have control over $\norm{U}$}
-\end{comment}

### 2017-05-17 related.tex
latest

@@
-\section{Related Work}
+\section{Related Work}\label{sec:related}
 We first discuss the related work outside of RL setting, followed by related work in the RL setting. In both cases, we highlight the new insights that follows from the results in this paper.\par
@@
 The problem class considered by them belongs to $\P_{\text{PSD},B}$ (see \Cref{def:pclass} ) and hence is weakly admissible with a uniform step-size choice.
-In order to produce a rate of $\frac{C}{t}$ (where $C>0$ is a problem independent constant) for the bias term \citet{bach} do not rely on admissibility, but instead use the fact that the MSE, $\EE{\norm{\thh_t-\ts}^2_{A_P}}$ for the linear prediction problem is measured in the general quadratic norm with respect to the symmetric positive definite matrix $A_P$ and the fact that the noise is `structured' i.e., the variance is bounded by $R\,A_P$ for some constant $R>0$.
+In order to produce a rate of $\frac{C}{t}$ (where $C>0$ is a problem independent constant) for the variance term \citet{bach} do not rely on admissibility, but instead use the fact that the MSE, $\EE{\norm{\thh_t-\ts}^2_{A_P}}$ for the linear prediction problem is measured in the general quadratic norm with respect to the symmetric positive definite matrix $A_P$ and the fact that the noise is `structured' i.e., the variance is bounded by $R\,A_P$ for some constant $R>0$.
 At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.

### 2017-05-17 uniform.tex
latest

@@
-\section{Uniform stepsizes}
+\section{Uniform stepsizes}\label{sec:uniform}
 If $\P$ is weakly admissible, then one can choose a step-size $\alpha_{\P}>0$ solely based on the knowledge of $\P$ and

### 2017-05-17 instance.tex
further updates

@@
 $\forall \alpha\in (0,\alpha_{P_U})$.
-\begin{theorem}[Lower Bound]
+\begin{theorem}[Lower Bound]\label{th:lb}
 There exists a distribution $P$ over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, such that $\alpha_P>0$, such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in (0,\alpha_P)$ and

### 2017-05-17 linal.tex
further updates

@@
 \item Define $\forall\, t\geq 0$ random vectors $\zeta_t\eqdef b_t-b-(A_t-A_P)\ts$.
-%\item Define constants $\EE{\norm{\zeta_t}^2}\eqdef\sigma_1^2=\sigma_A^2\norm{\ts}^2+\sigma_b^2$ and $\EE{\norm{M_t\zeta_t}}\eqdef \sigma_2^2=\sigma_A^2\norm{\ts}$.
-\item Define constants $\EE{\norm{\zeta_t}^2}\eqdef\sigma_1^2$ and $\EE{\norm{M_t\zeta_t}}\eqdef \sigma_2^2$.
+\item Define constants $\sigma_1^2\eqdef\sigma_A^2\norm{\ts}^2+\sigma_b^2$ and $\sigma_2^2\eqdef\sigma_A^2\norm{\ts}$.
+Note that $\EE{\norm{\zeta_t}^2}\eqdef\leq \sigma_1^2$ and $\EE{\norm{M_t\zeta_t}}\leq \sigma_2^2$.
+%\item Define constants $\EE{\norm{\zeta_t}^2}\eqdef\sigma_1^2$ and $\EE{\norm{M_t\zeta_t}}\eqdef \sigma_2^2$.
 \item Define $\forall\,i\geq j$, the random matrices $F_{i,j}=(I-\alpha A_i)\ldots (I-\alpha A_j)$ and $\forall,\,i<j$ $F_{i,j}=\I$.
@@
 \begin{lemma}[Change of Basis]\label{lm:cb}
-Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ as in \Cref{assmp:lsa} and let $U$ be chosen according to \Cref{lm:hur}.
-Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, then
+Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ as in \Cref{assmp:lsa} and let $U$ be chosen according to \Cref{lm:hur}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma_*\eqdef U^{-1}\ts$, then
 \begin{align}
-\EE{\norm{\gamma_t-\gamma^*}^2}
+\EE{\norm{\gamma_t-\gamma_*}^2}
 \leq
@@
 \left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
-%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
+%\left(\frac{\norm{\gamma_0-\gamma_*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma_*})}{t+1} \right)\,,
 \end{align}
@@
 \begin{proof}
-We have
-\begin{align*}
+Consider the modified error recursion in terms of $z_t\eqdef \gamma_t-\gamma_*$
+\begin{align}\label{eq:newerrrec}
+\begin{split}
 e_t&=(I-\alpha A_t)e_{t-1}+\alpha\zeta_t\\
@@
 z_t&=(I-\alpha \Lambda_t) z_{t-1}+\alpha H_t,
-\end{align*}
-where $z_t=\gamma_t-\gamma^*$, $\Lambda_t=U^{-1}A_t U$ and $H_t=U^{-1}\zeta_t$.
+\end{split}
+\end{align}
+where  $\Lambda_t=U^{-1}A_t U$ and $H_t=U^{-1}\zeta_t$.
 Note that the error recursion in $z_t$ might involve complex computations (depending on whether $U$ has complex entries or not), and hence \eqref{eq:lsacmplx} and \Cref{assmp:lsacmplx} are useful in analyzing $z_t$.
 We know that $\EE{\norm{H_t}^2}\leq \norm{U^{-1}}^2\EE{\norm{\zeta_t}}$ and $\EE{\norm{\Lambda_t H_t}}=\EE{\norm{U^{-1}A_t UU^{-1}\zeta_t}}=\EE{\norm{U^{-1}A_t \zeta_t}}\leq \norm{U^{-1}}\EE{\norm{A_t\zeta_t}}=\norm{U^{-1}}\sigma_2^2$.
-We follow through the steps of \Cref{th:pdrate} to obtain
+Now applying \Cref{th:pdrate} to $\hat{z}_t\eqdef \frac{1}{t+1}\sum_{s=0}^t z_t$, we have
 \begin{align}
@@
 Follows by sustuting $\theta_t=U\gamma_t$ in \Cref{lm:cb}.
+\begin{comment}
+\begin{theorem}[Lower Bound]
+There exists a distribution $P$ over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, such that $\alpha_P>0$, such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in (0,\alpha_P)$ and
+\begin{align*}
+\EE{\norm{\thh_t-\ts}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big( \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \big(\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}\big) \Big)
+\end{align*}
+\end{theorem}
+\end{comment}
+\paragraph{Proof of \Cref{th:lb}}
 
-\begin{comment}
-Let $\F_t = \sigma( H_1,\dots,H_t, \zeta_1,\dots,\zeta_t ) $, with $\F_0$ the $\sigma$-field that holds all random variables.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Product unroll lemma]\label{lem:genunroll}
-Let $t>i\ge 1$, $x,y\in \R^d$ be $\F_{i}$-measurable random vectors. Then,
+Consider the LSA with $(b_t,A_t)\sim P$ such that $b_t=(N_t,0)\top\in\R^2$ is a zero mean $i.i.d$ random variable with variance $\sigma^2_b$, and $A_t=A,\,\forall t\geq 0$, where $A=A_P=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$.
+Note that in this example $\ts=0$.
+By choosing $\alpha<\frac2{\lambda_{\max}}$, in this case it is straightforward to write the expression for $\eh_t$ explicitly as below:
 \begin{align*}
-\E[x^\top F_{t,i+1}y|\F_i]=x^\top (I-\alpha H)^{t-i} y\,.
-\end{align*}
-\end{lemma}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{proof}
-By the definition of $F_{t,i+1}$,
-and because $F_{t-1,i+1} = (I-\alpha H_{t-1}) \dots (I-\alpha H_{i+1})$ is $\F_{t-1}$-measurable,
-as are $x$ and $y$,
-\begin{align*}
-\EE{x^\top F_{t,i+1} y | \F_{t-1} } &= x^\top \EE{ (I-\alpha H_t) | \F_{t-1} } F_{t-1,i+1} y\\
-&=x^\top  (I-\alpha H)  F_{t-1,i+1} y\,.
-\end{align*}
-By the tower-rule for conditional expectations and our measurability assumptions,
-\begin{align*}
-\EE{x^\top F_{t,i+1} y | \F_{t-2} }
-&=x^\top  (I-\alpha H)  \EE{F_{t-1,i+1} |\F_{t-2}} y\\
-&= x^\top (I-\alpha H)^2 F_{t-2,i+1} y\,.
-\end{align*}
-Continuing this way we get
-\begin{align*}
-\EE{x^\top F_{t,i+1} y | \F_{t-j} }
-= x^\top (I-\alpha H)^j F_{t-j,i+1} y\,, \quad j=1,2,\dots,t-i\,.
-\end{align*}
-Specifically, for $j=t-i$ we get
-\begin{align*}
-\EE{x^\top F_{t,i+1} y | \F_{i} }  = x^\top (I-\alpha H)^{t-i} y\,.
-\end{align*}
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}\label{noisecancel}
-Let $t>i\ge 1$ and let $x$ be an $\F_{i-1}$-measurable random vector. Then,
-$\E[x^\top F_{t,i+1}\zeta_{i}]=0$.
-\end{lemma}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{proof}
-By \Cref{lem:genunroll},
-\begin{align*}
-\EE{x^\top F_{t,i+1} \zeta_i | \F_{i} }  = x^\top (I-\alpha H)^{t-i} \zeta_i\,.
-\end{align*}
-Using the tower rule,
-\begin{align*}
-\EE{x^\top F_{t,i+1} \zeta_i | \F_{i-1} }
-= x^\top (I-\alpha H)^{t-i}\EE{ \zeta_i | \F_{i-1} }= 0\,.
-\end{align*}
-\end{proof}
-
-\begin{lemma}\label{lem:unroll}
-For all $t>i\ge 0$, $\E \ip{He_i,H F_{t,i+1} e_i}=\E\ip{He_i,H(I-\alpha H)^{t-i} e_i}$.
-\end{lemma}
-\begin{proof}
-The lemma follows directly from \Cref{lem:genunroll}. Indeed,
-$\theta_i$ depends only on $H_1,\dots,H_{i},g_1,\dots,g_{i}$, $\theta_i$ and so is $e_i$ $\F_i$-measurable.
-Hence, the lemma is applicable and implies that
-\begin{align*}
-\EE{ \ip{He_i, HF_{t,i+1} e_i} | \F_i } =
-\EE{ \ip{He_i, H (I-\alpha H)^{t-i} e_i} | \F_i }\,.
-\end{align*}
-Taking expectation of both sides gives the desired result.
-\end{proof}
-
-\begin{lemma}\label{innerproduct}
-Let $i>j \ge 0$ and let $x\in \R^d$ be an $\F_j$-measurable random vector.
-Then,
-\begin{align*}
-\E\ip{F_{i,j+1}x,F_{i,j+1}x}\leq (1-\alpha \rho_\alpha)^{i-j}\E\norm{x}^2\,.
-\end{align*}
-\end{lemma}
-\begin{proof}
-Note that
-$S_t\doteq \EE{ (I-\alpha H_t)^\top (I-\alpha H_t) | \F_{t-1} }
-= I - \alpha (H^\top + H) + \alpha^2 \EE{ H_t^\top H_t | \F_{t-1} }$.
-Since $(g_t,H_t)_t$ is an independent sequence, $\EE{ H_t^\top H_t|\F_{t-1}} = \EE{ H_1^\top H_1 }$.
-Now, using the definition of $\rho_\alpha$ from \eqref{eq:rhodef},
-$\sup_{x:\norm{x}= 1} x^\top S_t x = 1 - \alpha \inf_{x:\norm{x}=1} x^\top (H^\top + H - \alpha \EE{H_1^\top H_1}) x
-= 1-\alpha \rho_\alpha$.
-Hence,
-\begin{align*}
-&\EE{\ip{F_{i,j+1}x,F_{i,j+1}x}|\F_{i-1} }\\
-&= \EE{x^\top F_{i-1,j+1}^\top (I-\alpha H_i)^\top (I-\alpha H_i) F_{i-1,j+1} x\,|\,\F_{i-1}}\\
-&=(x F_{i-1,j+1})^\top \, S_i \, F_{i-1,j+1} x\\
-&\le (1-\alpha \rho_\alpha) \, \ip{ F_{i-1,j+1} x, F_{i-1,j+1} x} \\
-& \le (1-\alpha \rho_\alpha)^2\, \ip{ F_{i-2,j+1} x, F_{i-2,j+1} x} \\
-& \quad \vdots \\
-& \le (1-\alpha \rho_\alpha)^{i-j}\, \norm{x}^2\,.
-%&= \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_i)^\top (I-\alpha H_i)\ldots (I-\alpha H_{j+1})x|\F_{i-1}]\\
-%&=(1-\alpha\rho_{\alpha}) \E[x^\top (I-\alpha H_{j+1})\ldots (I-\alpha H_{i-1})^\top (I-\alpha H_{i-1})\ldots (I-\alpha H_{j+1})x]
-\end{align*}
-\end{proof}
-
-\begin{align*}
-U^{-1}e_t&=(I-\alpha U^{-1}A_t U) U^{-1}e_{t-1}+ U^{-1}(b_t -b -(A_t-A)\ts)\\
-e_t&=(I-\alpha J_t) e_{t-1}+ H_t
+\eh_t&=\frac{1}{t+1}\sum_{s=0}^t e_t = \frac{1}{t+1}\sum_{s=0}^t (I-\alpha A_P)^{t-s} e_0 + \sum_{s=1}^t \sum_{i=s}^t (I-\alpha A_P)^{i-s} b_s\\
+&=\frac{1}{t+1}(\alpha A_P)^{-1}\left[\left(I-(I-\alpha A_P)^{t+1}\right)e_0 + \sum_{s=1}^t \left(I-(I-\alpha A_P)^{t+1-s}\right) b_s\right]
 \end{align*}
@@
 \begin{align*}
-e_t
-& = (I-\alpha J_t) (I-\alpha J_{t-1}) e_{t-2}\\ &+ \alpha (I-\alpha J_t) \zeta_{t-1} +\alpha H_t \\
-& \quad \vdots\\
-& = (I-\alpha J_t) \cdots (I-\alpha J_1) e_0\\ &+ \alpha (I-\alpha J_t) \cdots (I-\alpha J_2) \zeta_1 \\
-& + \alpha (I-\alpha J_t) \cdots (I-\alpha J_3) \zeta_2\\
-&  \quad \vdots \\
-&+ \alpha H_t\,,
+\EE{\norm{\eh_t}^2}&\stackrel{(a)}{=}\frac{1}{(t+1)^2}\left[\norm{(\alpha A_P)^{-1}\left(I-(I-\alpha A_P)^{t+1}\right)e_0}^2 + \sum_{s=1}^t \norm{(\alpha A_P)^{-1}\left(I-(I-\alpha A_P)^{t+1-s}\right)b_s}^2\right]
+%&\leq\frac{1}{(t+1)^2}\alpha^2\norm{A_P^{-1}}^2\norm{e_0}^2
 \end{align*}
-which can be written compactly as
-\begin{align}
-\label{eq:etft}
-e_t = F_{t,1} e_0 + \alpha (F_{t,2} \zeta_1 + \dots + F_{t,t+1} H_t )\,,
-\end{align}
+
 \begin{align*}
-\zh_t=\frac{1}{t+1}{\sum}_{i=0}^{t}e_i
-=\frac{1}{t+1}&\Big\{{\sum}_{i=0}^{t} F_{i,1} e_0 \\
-&+ \alpha \sum_{i=1}^{t} \left(\sum_{k=i}^{t} F_{k,i+1} \right)\zeta_i \Big\} ,
+\EE{\norm{\eh_t}^2}\geq \EE{\eh^2_t(1)}=\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}\big[\left(1-(1-\alpha \lambda_{\min})^{t+1}\right)^2 \theta^2_0(1)+ \frac{1}{(t+1)^2}\sum_{s=1}^t\left(1-(1-\alpha \lambda_{\min})^{t+1-s}\right)^2 b^2_s(1) \big]
 \end{align*}
-where in the second sum we flipped the order of sums and swapped the names of the variables that the sum runs over.
-
-It follows that \todoc{We should rather use $C$ instead of $H$ here?}
-\begin{align*}
-\E[\norm{\zh_t}^2]&=\E\ip{\zh_t,\zh_t}
-=\frac{1}{(t+1)^2} \sum_{i,j=0}^t \E\ip{e_i,e_j}\,.
-\end{align*}
-Hence, we see that it suffices to bound $\EE{\ip{ e_i,  e_j }}$.
-There are two cases depending on whether $i=j$. When $i< j$,
-\begin{align*}
-\E\ip{e_i,e_j}
-&=\E \ip{e_i,\big[F_{j,i+1} e_i+\alpha\textstyle\sum_{k=i+1}^j F_{j,k+1}\zeta_{k}\big]}\\
-&=\E\ip{e_i,F_{j,i+1} e_i}  \text{(from \Cref{noisecancel})}\\
-&=\E\ip{e_i, (I-\alpha J)^{j-i} e_i} \text{(from \Cref{lem:unroll})}
-\end{align*}
-and therefore
-\begin{align*}
-\label{inter}
-\sum_{i=0}^{t-1}\sum_{j=i+1}^t \E\ip{e_i,e_j}
-&=\frac1{\alpha\rhod{P}} {\sum}_{i=0}^{t-1}\E\ip{e_i,e_i}\\
-&\leq \frac2{\alpha\rhod{P}}{\sum}_{i=0}^{t}\E\ip{e_i,e_i}\,.
-\end{align*}
-Since $\sum_{i,j}\cdot{} = \sum_{i=j}\cdot{} + 2 \sum_i \sum_{j>i} \cdot{}$,
-\begin{align*}
-{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{e_i,e_j}&= \left(1+\frac2{\alpha\rhod{P}}\right){\sum}_{i=0}^{t}\E\ip{e_i,e_i}\,.
-\end{align*}
-Expanding $e_i$ using \eqref{eq:etft} and then using \Cref{innerproduct} and \Cref{cmplxassmp}
-\begin{align*}
-\E\ip{e_i,e_i}&=\E\ip{F_{i,1}e_0,F_{i,1}e_0}+\alpha^2{\sum}_{j=1}^i\E\ip{ F_{i,j+1}\zeta_j, F_{i,j+1}\zeta_j}+\alpha\sum_{j=1}^i  \E\ip{F_{i,1} e_0, F_{i,j+1}\zeta_j}\\
-&\leq (1-\alpha\rhos{P})^i\norm{e_0}^2+ \alpha^2\frac{{\sigma}_1^2}{\alpha \rhos{P}}+ \alpha \frac{{\sigma}^2_2 \norm{e_0}}{\alpha\rhos{P}}\,,
-\end{align*}
-and so
-\begin{align*}
-{\sum}_{i=0}^{t}{\sum}_{j=0}^{t} \E\ip{e_i,e_j}
-&\leq \left(1+\frac2{\alpha\rhod{P} }\right)\, \frac1{\alpha\rhos{P}}\, (t(\alpha^2{\sigma}_1^2+\alpha {\sigma}^2_2\norm{e_0}) +\norm{e_0}^2)\,.
-\end{align*}
-Putting things together,
-\begin{align}
-\E[\norm{\zh_t}^2]
-\leq \left(1+\frac2{\alpha\rhod{P}}\right)\, \frac1{\alpha\rhos{P}}\, \,
-\left(\frac{\norm{e_0}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+\alpha {\sigma}^2_2\norm{e_0}}{t+1} \right)\,.
-\end{align}
-
-\end{comment}
+Note that in this example, $\rhos{P}=\rhod{P}=2\lambda_{\min} -\alpha \lambda_{\min}^2=\lambda_{\min}(2-\alpha \lambda_{\min})$, and $\norm{\ts}=0$ and $\sigma^2_A=0$.
+Further, the result follows by noting the fact that noting the fact that $\norm{b_t}^2=b_t(1)^2$ and $\norm{\theta_t}^2=\theta_t(1)^2$.
+%\begin{align}
+%\EE{\norm{\eh_t}^2}\geq \frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)
+%\end{align}

### 2017-05-18 concl.tex
latest

\section{Conclusion}
We presented a finite time performance analysis of LSAs with CS-PR and showed that the MSE decays at a rate $O(\frac{1}{t})$. Our results extended the analysis of \citet{bach} for SGD with CS-PR for the problem of linear regression with quadratic loss and $i.i.d$ sampling to general LSAs with CS-PR. Our results were directly applicable to TD class of algorithms that are LSAs. In particular, new performance bounds for TDC follows from our results. Our results also improved the rates in the case of the GTD class of algorithms. We presented conditions under which a constant step-size can be chosen uniformly for a give class of data distributions. We showed a negative result in that not all data distributions `admit' such a constant step-size. This is a negative result from the perspective of TD algorithms in RL. We also argued heuristically that a constant step-size can be obtained in an automatic manner for a given LSA and a given data distribution. Thus, LSAs with CS-PR technique $i)$ achieves comparable or better error rates $ii)$ alongside automatic tuning of the constant step-size can perform comparably well in practice.

### 2017-05-18 intro.tex
latest

@@
 Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster learning, and at the same time, step-sizes that are too large might result in instability of \eqref{eq:lsaintro}.
-A useful choice in RL has been the diminishing step-sizes \cite{gtd2,gtdmp,borkarbook,konda-tsitsiklis}, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$.
-An alternate technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{i=0}^t \theta_i$.
+A useful choice in RL has been the diminishing step-sizes \cite{gtd2,gtdmp,konda-tsitsiklis}, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$.
+An alternate technique, which we call LSA with constant step-size and Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{i=0}^t \theta_i$.
 In LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
@@
 A ``remarkable'' result by \citet{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the mean squared (regression) error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
-In this paper, we study LSA with CS-PR (thereby extending prior work by \citet{bach} from SGD to general LSA) where data $(b_t,A_t),t\geq 0$ is $i.i.d$ with some underlying distribution $P$ such that the variance is bounded.
+In this paper, we study LSAs with CS-PR (thereby extending prior work by \citet{bach} from SGD to general LSAs) where the data sequence $(b_t,A_t),t\geq 0$ is $i.i.d$ with some underlying distribution $P$ such that the variance is bounded.
 An additional restriction we place on $P$ is that the matrix $\EEP{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts.
@@
 \begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{Instance Dependent Bounds} For a given $P$, we  measure the performance of LSA with CS-PR in terms of the mean square error (MSE) given by $\EEP{\norm{\thh_t-\ts}^2}$.
-We show there exists an $\alpha_P>0$ such that the mean squared error (MSE) of LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$.
-The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
-\item \textbf{Uniform Step-Sizes:} It is of interest to know, for a given class $\P$ of distributions, whether it is possible to ensure uniform performance $\forall P\in \P$.
+\item \textbf{Instance Dependent Bounds:} For a given $P$, we  measure the performance of a given LSA with CS-PR in terms of the mean square error (MSE) given by $\EEP{\norm{\thh_t-\ts}^2}$.
+We show that there exists an $\alpha_P>0$ such that the mean squared error (MSE) %of a given LSA with CS-PR
+is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$.
+%The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
+\item \textbf{Uniform Step-Sizes:} It is of interest to know that for a given LSA with CS-PR and a given class $\P$ of distributions, whether it is possible to ensure uniform performance $\forall P\in \P$.
 We also show (via an example) that there is a class $\P$ that does not `admit' a constant step-size $\alpha_{\P}$ that guarantees uniform performance $\forall P\in \P$.
@@
 \item \textbf{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion.
-This way we argue that the performance LSA with CS-PR is comparable to other popular step-size choices used for LSAs arising in RL.
+This way we argue that the performance of LSA with CS-PR is comparable to other popular step-size choices used for LSAs arising in RL.
 \end{itemize}
-\paragraph{Implications:}
-\paragraph{Organization:}
+\paragraph{Implications:} In addition to TD(0), our results directly can be applied to a host of \emph{off-policy} TD algorithms such as GTD/GTD2/TDC with CS-PR.
+In the case of TDC only asymptotic bounds were known \cite{gtd2} and our results are a first step towards proving finite time bounds.
+Also, our results show that the GTD class of algorithms guarantee a $O(\frac{1}{t})$ rate for MSE (without use of projections) as opposed to the previous results (using projections) \cite{gtdmp} that provide a $O(\frac{1}{\sqrt{t}})$ rate for this class.
+\paragraph{Organization:} The paper is organized as follows

### 2017-05-18 linal.tex
latest

@@
 \begin{align*}
-\EE{\norm{\eh_t}^2}\geq \EE{\eh^2_t(1)}=\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}\big[\left(1-(1-\alpha \lambda_{\min})^{t+1}\right)^2 \theta^2_0(1)+ \frac{1}{(t+1)^2}\sum_{s=1}^t\left(1-(1-\alpha \lambda_{\min})^{t+1-s}\right)^2 b^2_s(1) \big]
+\EE{\norm{\eh_t}^2}\geq \EE{\eh^2_t(1)}\stackrel{(b)}{=}\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}\big[\left(1-(1-\alpha \lambda_{\min})^{t+1}\right)^2 \theta^2_0(1)+ \frac{1}{(t+1)^2}\sum_{s=1}^t\left(1-(1-\alpha \lambda_{\min})^{t+1-s}\right)^2 b^2_s(1) \big]
 \end{align*}
+Here $(a)$ and $(b)$ follows from the $i.i.d$ assumption.
 Note that in this example, $\rhos{P}=\rhod{P}=2\lambda_{\min} -\alpha \lambda_{\min}^2=\lambda_{\min}(2-\alpha \lambda_{\min})$, and $\norm{\ts}=0$ and $\sigma^2_A=0$.
@@
 %\end{align}
+
+\paragraph{Proof of \Cref{lm:notwad}}
+Fix an arbitrary $\alpha>0$. We show that there exists $P\in \P$ such that $\rho_\alpha(P)<0$.
+For $\epsilon \in (0,1/2)$ let $P=(P^V,P^M)$ be the distribution such that $P^M$ is supported on $\{-I,I\}$ and takes on the value of $I$ with probability $1/2+\epsilon$.
+Then $A_P = 2\epsilon I \succ 0$, hence $P\in \P_1$.
+Further, $Q_P = I$.
+Hence, $\rhos{P} = 4\epsilon-\alpha$. Hence, if $\epsilon<\alpha/4$, $\rho_\alpha(P)<0$.
+
+\paragraph{Proof of \Cref{lm:ppsdbwd}}
+Since $\P_{\text{PSD},B}$ is supported on the set of positive semi-definite matrices, we know for any $A\in \R^{\dcd}$ that is PSD, we can consider the SVD of $A$: $A = U \Lambda U^\top$ where $U$ is orthonormal and $\Lambda$ is diagonal with
+nonnegative elements. Note that $\Lambda \preceq B\, \I$ and thus $\Lambda^2 \preceq B \Lambda$.
+Then for any $x\in \R^d$, $x^\top A^\top A x = x^\top U \Lambda^2 U^\top x \le B x^\top U \Lambda U^\top x = B x^\top A x$.
+Taking expectations we find that $x^\top C_P x \le B x^\top A_P x$.
+Hence, $\rhos{P_{\text{PSD},B}} = 2 x^\top A_P x - \alpha x^\top C_P x \ge (2- \alpha B ) \,x^\top A_P x $.
+Thus, for any $\alpha<2/B$, $\rho_\alpha(P)>0$.
+
+\paragraph{Proof of \Cref{lm:ppsdbna}}
+Consider the case when the smallest eigenvalue of $A_P$ is $0$.

### 2017-05-18 uniform.tex
latest

@@
 \item  $\P_{\text{PSD},B}$ be such that  $\P^M_{\text{PSD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PSD} \}\,$ , where PSD denotes the set of positive semi-definite matrices.
-\item  $\P_{\text{PD},B}$ be such that  $\P^M_{\text{SD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PD} \}\,$ , where PD denotes the set of positive definite matrices.
+%\item  $\P_{\text{PD},B}$ be such that  $\P^M_{\text{SD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PD} \}\,$ , where PD denotes the set of positive definite matrices.
 \end{itemize}
@@
 \end{proposition}
-\begin{proposition}
+\begin{proposition}\label{lm:ppsdbwd}
 For any $B>0$, the set $\P_{\text{PSD},B}$ is weakly admissible
@@
 \end{proposition}
-\begin{proposition}
+\begin{proposition}\label{lm:ppsdbna}
 For any $B>0$, the set $\P_{\text{PSD},B}$ is not admissible.
@@
 }
-Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
-However, when we impose additional structure in the data in the form of positive semi-definiteness,  we then observe weak admissibility in \Cref{wad}, and at the same time, it is not enough to guarantee admissibility in \Cref{notad}.
-Finally, when we impose structure in the form of positive definiteness we observe admissibility.
+Thus \Cref{lm:notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
+However, when we impose additional structure in the data in the form of positive semi-definiteness,  we then observe weak admissibility in \Cref{lm:ppsdwd}, and at the same time, it is not enough to guarantee admissibility in \Cref{lm:ppsdna}.
 \begin{comment}

### 2017-05-18 not.tex
up to main results some polishing

@@
+%!TEX root =  flsa.tex
 \section{Notations and Definitions}\label{sec:def}
@@
 For $x\in \C$ we denote its modulus and complex conjugate by $\md{x}$ and $\bar{x}$ respectively.
-We denote $d$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively, and use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries respectively.
-We denote the transpose of $x$ by $x^\top$ and the conjugate transpose by $x^*={\bar{x}}^\top$.
+We denote $d$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively, and use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries, respectively.
+We denote the transpose of $C$ by $C^\top$ and the conjugate transpose by $C^*={\bar{C}}^\top$ (and of course the same notation applies to vectors, as well).
 We will use $\ip{\cdot,\cdot}$ to denote the inner products: $\ip{x,y}=x^* y$.
-For $x\in\C^d$, we denote the general quadratic norm with respect to a Hermitian matrix $C$ by $\norm{x}^2_C\eqdef x^*\, C \,x$.
+\todoc{Usually, people define this as $y^* x \ne x^* y$.
+It does not make a difference in terms of the math, but there could be slight differences.
+And we will need to be consistent.}
+We use $\norm{x} = \ip{x,x}^{1/2}$ to denote the $2$-norm.
+For $x\in\C^d$, we denote the general quadratic norm with respect to a positive definite (see below) Hermitian matrix $C$ (i.e., $C=C^*$) by $\norm{x}^2_C\eqdef x^*\, C \,x$.
 The norm of the matrix $A$ is given by $\norm{A}\eqdef \sup_{x\in \C^d:\norm{x}=1} \norm{Ax}$.
-We use $\cond{A}=\norm{A}\norm{A^{-1}}$ to denote the condition number of the matrix $A$.
+We use $\cond{A}=\normsm{A}\normsm{A^{-1}}$ to denote the condition number of matrix $A$.
 We denote the identity matrix in $\C^{\dcd}$ by $\I$ and the set of invertible $\dcd$ complex matrices by $\gld$.
-For a positive real number $B>0$, we denote by $\C^{d}_B=\{b\in \C^d\mid \norm{b}\leq B\}$ and by $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ the set of complex vectors and matrices whose norms are bounded $B$.
-We use $\left(\cdot\right)\sim P$ to denote the fact that $\left(\cdot\right)$ is distributed according to probability distribution $P$.
+For a positive real number $B>0$, we define $\C^{d}_B=\{b\in \C^d\mid \norm{b}\leq B\}$ and $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ to be the balls in $\C^d$ and $\C^{d\times d}$ respectively of radius $B$.
+We use $Z\sim P$ to denote the fact that $Z$ (which can be a number, or vector, or matrix) is distributed according to probability distribution $P$;
 $\E$ denotes mathematical expectation.
+
+Let us now state some definitions that will be useful for presenting our main results.
 \begin{definition}\label{def:dist}
-Let $P=(P^V,P^M)$ denote a $2$-tuple of probability distributions; $P^V$ over $\C^{d}$ and $P^M$ over $\C^{\dcd}$. Define
+For a probability distribution $P$ over $\C^d \times \C^{d\times d}$, we let $P^V$ and $P^M$ 
+denote the respective marginals of $P$ over $\C^d$ and $\C^{d\times d}$. \todoc{Changed this.}
+By \emph{abusing notation} we will often write $P = (P^V,P^M)$ to mean that $P$ is a distribution with the given marginals.
+%Let $P=(P^V,P^M)$ denote a $2$-tuple of probability distributions; $P^V$ over $\C^{d}$ and $P^M$ over $\C^{\dcd}$. 
+Define
 \begin{align*}
-A_P&=\int M dP^M(M),\quad C_P=\int M^* M dP^M(M), \quad b_P=\int V dP^V(V)\\
-\rhod{P}&\eqdef {\inf}_{x\in\C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha A_P^* A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha C_P\right)x}
+A_P&=\int M\, dP^M(M),\quad C_P=\int M^* M \,dP^M(M), \quad b_P=\int v\, dP^V(v)\,,\\
+\rhod{P}&\eqdef {\inf}_{x\in\C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha A_P^* A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^d\colon\norm{x}=1}\ip{x,\left((A_P+A_P^*)-\alpha C_P\right)x}\,.
 \end{align*}
 \end{definition}
-Note that $\rhod{P}>\rhos{P}$.
+Note that $\rhod{P}\ge \rhos{P}$.  \todoc{Strict inequality?? How about $P^M$ concentrating on zero? I changed the strict inequality to non-strict.} 
+\todoc{Explain why we use subindex $d$ and $s$.}
+\todoc{Since $\rhod$ depends on $P^M$ only, why not make it a function of $P^M$ only? Or at least add a remark?}
 \begin{definition}\label{def:simdist}
-Let $P=(P^V,P^M)$ as in \Cref{def:dist}; $b\sim P^V$ and $A\sim \P^M$ be random variables distributed according to $P^V$ and $P^M$.
-For $U\in \gld$ define $P_U\eqdef(P_U^V,P_U^M)$, to be the distributions of $U^{-1}b$ and $U^{-1}AU$.
+Let $P=(P^V,P^M)$ as in \cref{def:dist}; $b\sim P^V$ and $A\sim \P^M$ be random variables distributed according to $P^V$ and $P^M$.
+For $U\in \gld$ define $P_U$ to be the  distribution of $(U^{-1}b,U^{-1}AU)$.
+We also let
+$(P_U^V,P_U^M)$ denote the corresponding marginals. \todoc{I hope this works out.}
 \end{definition}
 \begin{definition}
-We call a matrix $A\in \C^{\dcd}$ to be \emph{Hurwitz} (H) if all eigenvalues of $A$ have positive real parts.
-We call a matrix $A\in \C^{\dcd}$ to be \emph{Positive Definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$.
-We call a matrix $A\in \C^{\dcd}$ to be \emph{Symmetric Positive Definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
+We call a matrix $A\in \C^{\dcd}$  \emph{Hurwitz} (H) if all eigenvalues of $A$ have positive real parts.
+We call a matrix $A\in \C^{\dcd}$ \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$.
+\todoc{I think usually this is defined as $\ip{Ax,x}>0$. Same issue as with $\ip{\cdot,\cdot}$.}
+We call a matrix $A\in \R^{\dcd}$ to be \emph{symmetric positive definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD. \todoc{Add note that SPD implies real.}
 \end{definition}
@@
 \begin{example}
-The matrices $\begin{bmatrix}0.1 &-1\\ 1 & 0.1\end{bmatrix}$, $\begin{bmatrix} 0.1 & 0.1 \\ 0 & 0.1\end{bmatrix}$ and $\begin{bmatrix}0.1 &0 \\ 0 & 0.1\end{bmatrix}$ are examples of H, PD and SPD respectively.
+The matrices $\begin{bmatrix}0.1 &-1\\ 1 & 0.1\end{bmatrix}$, $\begin{bmatrix} 0.1 & 0.1 \\ 0 & 0.1\end{bmatrix}$ and $\begin{bmatrix}0.1 &0 \\ 0 & 0.1\end{bmatrix}$ are examples of H, PD and SPD matrices, respectively, and they show that while SPD implies PD, which implies H, the reverse implications do not hold.
 \end{example}
@@
 \begin{definition}
-Call a set of distributions $\P=(\P^V,\P^M)$ over $\C^{d}\times \C^{\dcd}$
+Call a set of distributions $\P$ over $\C^{d}\times \C^{\dcd}$
 \emph{weakly admissible} if there exists $\alpha_{\P}>0$ such that
-$\rhos{P}>0,\forall \alpha\in(0,\alpha_{\P})$ holds for all $P\in \P$.
+$\rhos{P}>0$ holds for all $P\in \P$ and $\alpha\in(0,\alpha_{\P})$.
 \end{definition}
 \begin{definition}
-Call a set of distributions $\P=(\P^V,\P^M)$ over $\C^{d}\times \C^{\dcd}$ \emph{admissible}
-if there exists some $\alpha_{\P}>0$ such that $\inf_{P\in \P} \rhos{P}>0,\,\forall \alpha\in(0,\alpha_{\P})$.
+Call a set of distributions $\P$ over $\C^{d}\times \C^{\dcd}$ \emph{admissible}
+if there exists some $\alpha_{\P}>0$ such that $\inf_{P\in \P} \rhos{P}>0$ holds for all $\alpha\in(0,\alpha_{\P})$.
 The value of $\alpha_{\P}$ is called a witness.

### 2017-05-18 prob.tex
up to main results some polishing

@@
+%!TEX root =  flsa.tex
 \section{Problem Setup}\label{sec:prob}
@@
 \begin{align}
-\label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
-\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_i,
+\label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1})\,,\\
+\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\,\theta_i\,.
 \end{align}
 \end{subequations}
-where $\thh_t, \theta_t \in \R^{d}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \R^{\dcd}$ incrementally, in discrete time steps $t=1,2,\dots$
+The algorithm updates a pair of parameters $\theta_t,\tb_t\in \R^{d}$ incrementally, in discrete time steps $t=1,2,\dots$
 based on data $b_t\in \R^d$, $A_t\in \R^{\dcd}$. Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
@@
 The iterate $\theta_t$ is treated as an internal state of the algorithm, while $\thh_t$ is the output at time step $t$.
-The update of $\theta_t$ alone is considered a form of constant step-size LSA.\par
+The update of $\theta_t$ alone is considered a form of constant step-size LSA.
+
 Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(d)$ storage. 
-Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$ can be efficiently computed in $O(d)$ time.
-This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp} will be presented in \Cref{sec:related}.
+Moreover, sometimes $A_t$ will have a special form and then the matrix-vector product $A_t \theta_{t-1}$ can also be computed in $O(d)$ time. This happens for example when $A_t$ is rank one.
+When the data $(b_t,A_t)$ is sparse, further speedups are possible.
+Examples of this kind arise in reinforcement learning \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}, 
+for details see \cref{sec:related}.
 The significance of efficient computation of the matrix-vector product is that an update of the algorithm
-can then be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $d$ is in the range of thousands or more.\par
-We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa}.
-In what follows, we make use of the $\sigma$-fields $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$, and $\F_0$ that holds all random variables.
+can then be implemented in $O(d)$ time (or less, in case of sparse data) and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $d$ is in the range of thousands, or millions, or more, as may be required by modern applications (e.g., \citep{LiMaTaBo16})
+
+In what follows, for $t\ge1$ we make use of the $\sigma$-fields $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$; $\F_0$ holds all random variables.
+\todoc{A bit unusual as a choice.}
+We are interested in the behaviour of \eqref{eq:lsa} under the following assumption:
 \begin{assumption}\label{assmp:lsa}
 \begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \label{dist} $(b_t, A_t)\sim (P^b,P^A), t\geq 0$ is an $i.i.d$ sequence, where $P^b$ is a distribution over $\R^d$ and $P^A$ is a distribution over $\R^{\dcd}$. We assume that $P$ is Hurwitz.
+\item \label{dist} $(b_t, A_t)\sim P$, $t\geq 0$ is an $i.i.d$ sequence.
+We let $A_P$ be the expectation of $A_t$, $b_P$ be the expectation of $b_t$, as in \cref{def:dist}.
+%where $P^b$ is a distribution over $\R^d$ and $P^A$ is a distribution over $\R^{\dcd}$. 
+\todoc{I got rid off $P^b$ and $P^A$. There were $P^V$ and $P^M$ previously.. Confusing.}
+We assume that $P$ is Hurwitz.
 \item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following
-\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_{A_P}, \, \E[N_t^* N_t]=\sigma^2_{b_P}.\end{align*}
-\item $A_P$ is invertible and there exists a $\ts=A^{-1}_Pb_P$.
+\begin{align*}
+	\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_{A_P}\,, \qquad 
+	\E[N_t^* N_t]=\sigma^2_{b_P}\,.
+\end{align*}
+with some $\sigma^2_{A_P}$ and $\sigma^2_{b_P}$. \todoc{Could not these just be $\sigma^2_A$, $\sigma^2_b$, or with $M$ and $V$?}
+\item $A_P$ is invertible and thus the vector $\ts=A^{-1}_Pb_P$ is well-defined. \todoc{Bertsekas looked at the case when $\ts$ is well-defined but $A_P$ is not invertible. Future work..?}
 \end{enumerate}
@@
 %For the rest of the paper we denote the errors of the internal and the output variables of \eqref{eq:lsa} at time $t$ by $e_t\eqdef \theta_t-\ts$ and $\eh_t\eqdef \thh_t-\ts$ respectively.
-\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ $\EEP{\norm{\thh_t-\ts}_C^2}$, the MSE with respect a hermitian matrix $C$ is desirable, we cannot exploit the use of $C$ unless it is connected to $P$ in a special way.
-For more discussion, see \Cref{sec:related}.}
- by $\EE{\norm{\thh_t-\ts}^2}$.
+\paragraph{Performance Metric:}  
+We are interested in the behavior of the mean squared error (MSE) at time $t$ given%
+\footnote{ $\EEP{\norm{\thh_t-\ts}_C^2}$, the MSE with respect to a PD Hermitian matrix $C$ is desirable.
+However, we cannot exploit the use of $C$ unless it is connected to $P$ in a special way. 
+For more discussion, see \cref{sec:related}.}
+ by $\EE{\normsm{\thh_t-\ts}^2}$. \todoc{Move up footnote. Polish..}
 \begin{comment}

### 2017-05-18 uniform.tex
up to main results some polishing

@@
-\section{Uniform stepsizes}\label{sec:uniform}
+\section{Uniform bounds}\label{sec:uniform}
 If $\P$ is weakly admissible, then one can choose a step-size $\alpha_{\P}>0$ solely based on the knowledge of $\P$ and

### 2017-05-18 abstract.tex
latest
iid fix

@@
 For LSAs with a constant step-size, and PR averaging, we provide bounds for the mean squared error (MSE) after $t$ iterations.
-We assume that data is $i.i.d$ with finite variance (underlying distribution being $P$) and that the expected dynamics is Hurwitz.
+We assume that data is \iid with finite variance (underlying distribution being $P$) and that the expected dynamics is Hurwitz.
 For a given LSA with PR averaging, and data distribution $P$ satisfying the said assumptions, we show that there exists a range of constant step-sizes such that its MSE decays as $O(\frac{1}{t})$.\par
@@
 We also suggest a heuristic step-size tuning algorithm to choose a constant step-size of a given LSA for a given data distribution $P$.
-We compare our results with related work and also discuss the implication of our results in the context of LSA-TD algorithms.
+We compare our results with related work and also discuss the implication of our results in the context of TD algorithms that are LSAs.
 \end{abstract}

### 2017-05-18 concl.tex
iid fix

@@
 We presented a finite time performance analysis of LSAs with CS-PR and showed that the MSE decays at a rate $O(\frac{1}{t})$.
-Our results extended the analysis of \citet{bach} for SGD with CS-PR for the problem of linear regression with quadratic loss and $i.i.d$ sampling to general LSAs with CS-PR.
+Our results extended the analysis of \citet{bach} for SGD with CS-PR for the problem of linear regression with quadratic loss and \iid sampling to general LSAs with CS-PR.
 Our results were directly applicable to TD class of algorithms that are LSAs.

### 2017-05-18 intro.tex
iid fix

@@
+%!TEX root =  flsa.tex
 \section{Introduction}
@@
 Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection.
-\paragraph{Motivation and Contribution:} Recently, \citet{bach} considered SGD\footnote{SGD is an LSA of the form in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and $i.i.d$ sampling.
+\paragraph{Motivation and Contribution:} Recently, \citet{bach} considered SGD\footnote{SGD is an LSA of the form in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and \iid sampling.
 A ``remarkable'' result by \citet{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the mean squared (regression) error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
-In this paper, we study LSAs with CS-PR (thereby extending prior work by \citet{bach} from SGD to general LSAs) where the data sequence $(b_t,A_t),t\geq 0$ is $i.i.d$ with some underlying distribution $P$ such that the variance is bounded.
+In this paper, we study LSAs with CS-PR (thereby extending prior work by \citet{bach} from SGD to general LSAs) where the data sequence $(b_t,A_t),t\geq 0$ is \iid with some underlying distribution $P$ such that the variance is bounded.
 An additional restriction we place on $P$ is that the matrix $\EEP{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts.

### 2017-05-18 linal.tex
iid fix

@@
 \begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \label{dist} $(b_t, A_t)\sim (P^b,P^A), t\geq 0$ is an $i.i.d$ sequence, where $P^b$ is a distribution over $\C^d$ and $P^A$ is a distribution over $\C^{\dcd}$.
+\item \label{dist} $(b_t, A_t)\sim (P^b,P^A), t\geq 0$ is an \iid sequence, where $P^b$ is a distribution over $\C^d$ and $P^A$ is a distribution over $\C^{\dcd}$.
 We assume that $P$ is positive definite.
@@
 
-Consider the LSA with $(b_t,A_t)\sim P$ such that $b_t=(N_t,0)\top\in\R^2$ is a zero mean $i.i.d$ random variable with variance $\sigma^2_b$, and $A_t=A,\,\forall t\geq 0$, where $A=A_P=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$.
+Consider the LSA with $(b_t,A_t)\sim P$ such that $b_t=(N_t,0)\top\in\R^2$ is a zero mean \iid random variable with variance $\sigma^2_b$, and $A_t=A,\,\forall t\geq 0$, where $A=A_P=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$.
 Note that in this example $\ts=0$.
@@
 \end{align*}
-Here $(a)$ and $(b)$ follows from the $i.i.d$ assumption.
+Here $(a)$ and $(b)$ follows from the \iid assumption.
 Note that in this example, $\rhos{P}=\rhod{P}=2\lambda_{\min} -\alpha \lambda_{\min}^2=\lambda_{\min}(2-\alpha \lambda_{\min})$, and $\norm{\ts}=0$ and $\sigma^2_A=0$.

### 2017-05-18 intro.tex
intro

--- 2017-05-17 intro.tex
@@
+%!TEX root =  flsa.tex
 \section{Introduction}
-We consider linear stochastic approximation algorithms (LSA) of the general form given by
+Linear stochastic approximation algorithms (LSAs) 
+of the form
+%We consider linear stochastic approximation algorithms (LSA) of the general form given by
 \begin{align}\label{eq:lsaintro}
@@
 \end{align}
-where the iterates, $\theta_t\in \R^d$, and step-sizes, $\alpha_t>0$, are algorithmic quantities, and $(b_t,A_t), t\geq 0\in \R^d\times \R^{\dcd}$ is the noisy data sequence presented to the algorithm.
-LSAs are widely applied in machine learning and reinforcement learning (RL), where the aim is to compute an appropriate $\ts\in \R^d$ using only the noisy data sequence and $O(d)$ updates per iteration.
-Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear regression with quadratic loss \cite{bach,bachaistats}, and the \emph{temporal difference} (TD) class of learning algorithms in RL \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.\par
+with $(\alpha_t)_t$ a positive step-size sequence chosen by the user and 
+$(b_t,A_t)\in \R^d\times \R^{\dcd}$,  $t\geq 0$, a sequence of identically distributed random variables is widely used in
+machine learning, and in particular in reinforcement learning (RL), to compute the solution of the equation 
+$\E[b_t] - \E[A_t] \theta = 0$, where $\E$ stands for mathematical expectation.
+%where the iterates, $\theta_t\in \R^d$, and step-sizes, $\alpha_t>0$, are algorithmic quantities, and $(b_t,A_t), t\geq 0\in \R^d\times \R^{\dcd}$ is the noisy data sequence presented to the algorithm.
+%LSAs are widely applied in machine learning and reinforcement learning (RL), where the aim is to compute an appropriate $\ts\in \R^d$ using only the noisy data sequence and $O(d)$ updates per iteration.
+Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear regression with quadratic loss \cite{bach,bachaistats}, and the \emph{temporal difference} (TD) class of learning algorithms in RL \cite{sutton,konda-tsitsiklis,KoTsi03:LSA,gtd,gtd2,gtdmp}.
+\todoc{konda-tsitsiklis reference resolved to tsitsiklis-van-roy. is this what you want?? I also added KoTsi03:LSA, maybe that's what you wanted.}
 %An additional feature in these class of applications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par
 
-The choice of the step-sizes $\alpha_t$ is critical for the performance of LSAs such as \eqref{eq:lsaintro}.
-Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster learning, and at the same time, step-sizes that are too large might result in instability of \eqref{eq:lsaintro}.
-A useful choice in RL has been the diminishing step-sizes \cite{gtd2,gtdmp,borkarbook,konda-tsitsiklis}, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$.
-An alternate technique, which we call LSA with constant step-size Polyak-Ruppert (CS-PR) averaging, is to run \eqref{eq:lsaintro} as a constant step-size LSA by choosing $\alpha_t=\alpha>0,\forall t\geq 0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{i=0}^t \theta_i$.
-In LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
-The idea of averaging in stochastic approximation to improve noise rejection is a technique that goes back to  \citet{ruppert} and \citet{polyak-judisky}.
-Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection.
-\paragraph{Motivation and Contribution:} Recently, \citet{bach} considered SGD\footnote{SGD is an LSA of the form in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and $i.i.d$ sampling.
-A ``remarkable'' result by \citet{bach} is that there exists a constant step-size such that for any problem where the noisy data is bounded by a known constant, the mean squared (regression) error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data.\par\todoc{Dimension?}
-In this paper, we study LSA with CS-PR (thereby extending prior work by \citet{bach} from SGD to general LSA) where data $(b_t,A_t),t\geq 0$ is $i.i.d$ with some underlying distribution $P$ such that the variance is bounded.
-An additional restriction we place on $P$ is that the matrix $\EEP{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts.
-Our study is motivated by the problem of \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using TD class  of algorithms \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.
-Here, we wish to understand the effectiveness of the CS-PR technique for the TD class of LSA algorithms used in RL.
-We now list our contributions in this paper as follows:
+The choice of the step-size sequence $(\alpha_t)_t$ is critical for the performance of LSAs such as \eqref{eq:lsaintro}.
+Informally speaking, smaller step-sizes are better for noise rejection and larger step-sizes lead to faster forgetting of initial conditions (smaller bias).
+At the same time, step-sizes that are too large might result in instability of \eqref{eq:lsaintro} even when $(A_t)_t$ has favourable properties.
+A useful choice has been the diminishing step-sizes \cite{gtd2,gtdmp,konda-tsitsiklis}, where $\alpha_t\ra 0$ such that $\sum_{t\geq 0} \alpha_t=\infty$.
+Here, $\alpha_t\to0$ circumvents the need for guessing the magnitude of step-sizes that stabilize the updates, while the second condition ensures that initial conditions are forgotten.
+An alternate idea, which we call LSA with constant step-size and Polyak-Ruppert averaging (LSA with CS-PR, in short), is to run \eqref{eq:lsaintro} by choosing $\alpha_t=\alpha>0$ $\forall t\geq 0$ with some $\alpha>0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{i=0}^t \theta_i$.
+Thus, in LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
+The idea is that the constant step-size leads to faster forgetting of initial conditions, while the averaging on the top
+reduces noise.
+This idea goes back to  \citet{ruppert} and \citet{polyak-judisky} who considered it in the context of stochastic approximation that LSA is a special case of. 
+%Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection.
+\paragraph{Motivation and Contribution:} Recently, \citet{bach} considered stochastic gradient descent (SGD)\footnote{SGD is an LSA of the form in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and \iid sampling.
+They showed that one can calculate a constant step-size from only a bound on the magnitude of the noisy data so that the leading term as $t\to\infty$
+\todoc{There is a subtlety here, because, if I believe you, 
+\citet{bach} sells their result as a non-asymptotic result. How can they do this?
+Can we be more explicit where in the paper $C_P'/t^2$ shows up -- see the notation below? We must be very explicit here.
+The reviewers may not like this.
+}
+ in the mean-squared (regression) error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data, and is in particular independent of the eigenspectrum of $\E[A_t]$, a property which is not shared by other step-size tunings and variations of the basic SGD method.
+\todoc{Dimension?}
+
+In this paper, we study LSAs with CS-PR (thereby extending prior work by \citet{bach} from SGD to general LSAs) \todoc{Do our results imply theirs?
+Do we simplify improve their bounds, while extending the scope?
+Even if the answer is no, we need to add remarks on these questions.}
+in an effort to understand the effectiveness of the CS-PR technique beyond SGD.
+Of particular interest is whether a similar result to that  \citet{bach} holds
+for the TD class of LSA algorithms used in RL.
+For simplicity, we still consider the \iid case.
+Our restrictions on the common distribution is that the ``noise variance'' should be bounded (as we consider squared errors), and that the matrix $\E[A_t]$ must be Hurwitz, i.e., all its eigenvalues have positive real parts.
+This latter assumption can be shown to be necessary for the stability of the LSA iterations.
+\todoc{Add citation.
+Or remove/comment out the sentence.}
+One setting that fits our assumption is \emph{policy evaluation} \cite{dann} using linear value function approximation from experience replay \cite{lin} in a batch setting \cite{lange} in RL using the TD class of algorithms \cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.
+
+Our \textbf{main contributions} are as follows:
 %While asymptotically diminishing step-sizes are common in RL literature \cite{ilstd,gtdmp,korda-prashanth}, the constant step-size choice has also been used in the past \cite{gtd2}.
 \begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{Instance Dependent Bounds} For a given $P$, we  measure the performance of LSA with CS-PR in terms of the mean square error (MSE) given by $\EEP{\norm{\thh_t-\ts}^2}$.
-We show there exists an $\alpha_P>0$ such that the mean squared error (MSE) of LSA with CS-PR is at most $O(\frac{1}{t})$, whenever $\alpha\in (0,\alpha_P)$.
-The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
-\item \textbf{Uniform Step-Sizes:} It is of interest to know, for a given class $\P$ of distributions, whether it is possible to ensure uniform performance $\forall P\in \P$.
-We also show (via an example) that there is a class $\P$ that does not `admit' a constant step-size $\alpha_{\P}$ that guarantees uniform performance $\forall P\in \P$.
-This is a negative result in the context of RL.
-However, we also argue that the subclass (which we denote by $\P_{\text{SPD},B}$) related to SGD for linear regression does `admit' a uniform constant step-size, thereby recovering a part of the result by \citet{bach}.
-\item \textbf{Automatic Step-Size:} We argue informally that for any given $P$, a constant step-size can be chosen in an `automated' fashion.
-This way we argue that the performance LSA with CS-PR is comparable to other popular step-size choices used for LSAs arising in RL.
+\item \textbf{Finite-time Instance Dependent Bounds}: For a given $P$, we  measure the performance of a given LSA with CS-PR in terms of the mean square error (MSE) given by $\EEP{\normsm{\thh_t-\ts}^2}$.
+For the first time in the literature,
+we show that (under our stated assumptions) there exists an $\alpha_P>0$ such that 
+for any $\alpha\in (0,\alpha_P)$,
+the MSE %of a given LSA with CS-PR
+is at most $\frac{C_P}{t}+\frac{C_P'}{t^2}$ with some positive constants $C_P,C_P'$ that we explicitly compute from $P$.
+%The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
+\item \textbf{Uniform Bounds}:
+It is of major interest to know whether for a given class $\P$ of distributions 
+one can choose some step-size $\alpha$ 
+such that $C_P$ from above is uniformly bounded (i.e., replicating the result of \citet{bach}).%
+\footnote{Of course, the term $C_P'/t^2$ needs to be controlled, as well.
+Just like \citet{bach}, here we focus on $C_P$, which is justified if one considers the MSE as $t\to\infty$.
+Further justification is that we actually find a negative result.
+See above.}
+%It is of interest to know that for a given LSA with CS-PR and a given class $\P$ of distributions, whether it is possible to ensure uniform performance $\forall P\in \P$. 
+We show via an example that in general this is not possible.
+%that there is a class $\P$ that does not `admit' a constant step-size $\alpha_{\P}$ that guarantees uniform performance $\forall P\in \P$. 
+In particular, the example applies to RL, hence, we get a negative result for RL, which states that from only bounds on the data one cannot choose a step-size to guarantee that $C_P$ of CS-PR is uniformly bounded over $\P$.
+We also define a subclass  $\P_{\text{SPD},B}$ of problems, related to SGD for linear regression, that does `admit' a uniform constant step-size, thereby recovering a part of the result by \citet{bach}.
+\todoc{Why only part?
+We may need to explain the difference.} Our study thus sheds light on what structural assumptions are key to achieve a uniform bound for CS-PR.
+\item \textbf{Automatic Step-Size}: 
+The above negative result implies that in RL one needs to choose the constant step-size based on properties of the instance $P$ to avoid the explosion of the MSE.
+To circumvent this, we propose a natural step-size tuning method to guarantee instance-dependent boundedness.
+We experimentally evaluate the proposed method and find that it is indeed able to achieve its goal on a set of synthetic examples
+where no constant step-size is available to prevent exploding MSE.
 \end{itemize}
-\paragraph{Implications:}
-\paragraph{Organization:}
+%\paragraph{Implications:} 
+In addition to TD($0$), our results directly can be applied to a host of \emph{off-policy} TD algorithms such as GTD/GTD2/TDC with CS-PR.
+\todoc{How about TD($\lambda$)?
+Will we discuss this somewhere?
+Maybe in the organization section this can be mentioned.}
+In the case of TDC only asymptotic bounds were known \cite{gtd2} and our results are a first step 
+\todoc{Why only a step?
+What is missing?} towards proving finite time bounds.
+Also, our results show that the GTD class of algorithms guarantee a $O(\frac{1}{t})$ rate for MSE (without use of projections), improving on a previous result by \citet{gtdmp} that guaranteed a $O(\frac{1}{\sqrt{t}})$ rate for this class for the projected version of the algorithm.
+\todoc{And projections are problematic on their own.
+Will we discuss this somewhere.
+Or should the projection issue be a footnote here?}
+\todoc{How about Prashanth's paper? Should we mention it somewhere? The reviewers will push back if we don't.}
+\paragraph{Organization:} The paper is organized as follows \todoc{Do not forget to fill this in.}

### 2017-05-18 not.tex
intro

@@
 \section{Notations and Definitions}\label{sec:def}
-We denote the sets real and complex numbers by $\R$ and $\C$ respectively.
-For $x\in \C$ we denote its modulus and complex conjugate by $\md{x}$ and $\bar{x}$ respectively.
-We denote $d$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$ respectively, and use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries, respectively.
+We denote the sets of real and complex numbers by $\R$ and $\C$, respectively.
+For $x\in \C$ we denote its modulus and complex conjugate by $\md{x}$ and $\bar{x}$, respectively.
+We denote $d$-dimensional vector spaces over $\R$ and $\C$ by $\R^{d}$ and $\C^{d}$, respectively, and use $\R^{\dcd}$ and $\C^{\dcd}$ to denote $\dcd$ matrices with real and complex entries, respectively.
 We denote the transpose of $C$ by $C^\top$ and the conjugate transpose by $C^*={\bar{C}}^\top$ (and of course the same notation applies to vectors, as well).
@@
 We denote the identity matrix in $\C^{\dcd}$ by $\I$ and the set of invertible $\dcd$ complex matrices by $\gld$.
-For a positive real number $B>0$, we define $\C^{d}_B=\{b\in \C^d\mid \norm{b}\leq B\}$ and $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ to be the balls in $\C^d$ and $\C^{d\times d}$ respectively of radius $B$.
+For a positive real number $B>0$, we define $\C^{d}_B=\{b\in \C^d\mid \norm{b}\leq B\}$ and $\C^{\dcd}_B=\{A\in \C^{\dcd}\mid \norm{A}\leq B\}$ to be the balls in $\C^d$ and $\C^{d\times d}$, respectively, of radius $B$.
 We use $Z\sim P$ to denote the fact that $Z$ (which can be a number, or vector, or matrix) is distributed according to probability distribution $P$;

### 2017-05-18 instance.tex
main results

--- 2017-05-17 instance.tex
@@
-\section{Main Results}\label{sec:mainresults}
-In this section, we derive instance dependent bounds that are valid for a given problem $P$ (satisfying \Cref{assmp:lsa}) and in the \Cref{sec:uniform}, we address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions (problems).
-Here, we only present the main results followed by a discussion and the detailed proofs can be found in \Cref{sec:proofs}.
+%!TEX root =  flsa.tex
+\section{Main Results and Discussion}\label{sec:mainresults}
+In this section, we derive instance dependent bounds that are valid for a given problem $P$ (satisfying \Cref{assmp:lsa}) and in the \cref{sec:uniform}, we address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions (problems).
+Here, we only present the main results followed by a discussion.
+The detailed proofs can be found in \cref{sec:proofs}.
+\todoc{Usually it is expected that one gives an outline of the proof technique.
+Especially in a theory paper.
+At least verbally we could explain the tools used, how the proof is similar and/or different to previous proofs.}
 In what follows, for the sake of brevity, we drop the subscript $P$ in the quantities $\EEP{\cdot}$, $\sigma^2_{A_P}$ and $\sigma^2_{b_P}$.
+\todoc{I dropped $P$ from $\EEP{\cdot}$ before.
+Either put there back, or remove $\EEP{\cdot}$ from here.}
 %\begin{lemma}\label{lm:simtran}[Similarity Transformation]
@@
 %\end{lemma}
+We start with a lemma, which is needed to meaningfully state our main result:
 \begin{lemma}\label{lm:hur}
-Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, then there exists an $\alpha_{P_U}>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_{P_U})$.
+Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}. 
+Then there exists an $\alpha_{P_U}>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0$
+holds for all $\alpha \in (0,\alpha_{P_U})$. 
 \end{lemma}
@@
 \end{align*}
+\end{theorem}
 
-\end{theorem}
 \begin{theorem}\label{thm:simtran}[Change of Basis]
@@
 \begin{theorem}\label{th:rate}
-Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, then there exists $U\in\gln$ and $\alpha_{P_U}>0$ such that
+Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}.
+Then, for  $U\in\gln$ and $\alpha_{P_U}>0$ as in \cref{lm:hur},
+\todoc{Why not use $\alpha_P$? $U$ is dependent on $P$. If $\alpha_{P_U}$ does not have a specific definition, there is no point to denote this dependence on $U$. It just adds to the clutter.}
+for all $\alpha\in (0,\alpha_{P_U})$ and for all $t\ge 0$,
 \begin{align*}
-\EE{\norm{\thh_t-\ts}^2}
+\EE{\normsm{\thh_t-\ts}^2}
 \leq
-\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}
+\nu_P\,
 \left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,,
 \end{align*}
+where $\nu_P = \left(1+\tfrac2{\alpha\rhod{P_U}}\right)\tfrac{\cond{U}^2}{\alpha \rhos{P_U}}$.
 \end{theorem}
-$\forall \alpha\in (0,\alpha_{P_U})$.
-\begin{theorem}[Lower Bound]
-There exists a distribution $P$ over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, such that $\alpha_P>0$, such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in (0,\alpha_P)$ and
+
+\begin{theorem}[Lower Bound]\label{th:lb}
+There exists a distribution $P$ over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, such that
+there exists $\alpha_P>0$ so that $\rhos{P}>0$ and $\rhod{P}>0$ hold for all $\alpha\in (0,\alpha_P)$ and
 \begin{align*}
-\EE{\norm{\thh_t-\ts}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big( \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \big(\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}\big) \Big)
+\EE{\norm{\thh_t-\ts}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big\{ \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \big(\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}\big) \Big\}
 \end{align*}
@@
 \end{comment}
-
-\subsection{Result Discussion}
+In what follows, we discuss the specific details of these results. 
+\vspace*{0.5em}
+%\subsection{Result Discussion}
 \begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{Role of $U$:} When the given problem $P$ is positive definite, it is straightforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$.
+\item \textbf{Role of $U$:} When the given problem $P$ is positive definite, it is straightforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$ by considering the definitions of $\rhos{P}$ and $\rhod{P}$.
 However, the problem $P$ being Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}). 
-This can be remedied by a similarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$.
-To see this, let $\gamma_t\eqdef U^{-1 }\theta_t$ be a basis transformation, and let us re-write \eqref{conststep} in terms of $\gamma_t$ by pre-multiplying by $U^{-1}$ as shown below:
+Following  \cite{lihong}, we remedy this 
+by applying a similarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$.
+To see how this works, let $\gamma_t\eqdef U^{-1 }\theta_t$.
+Premultiplying
+ \eqref{conststep} by $U^{-1}$ and plugging in $\gamma_t$, we get
 \begin{align*}
-U^{-1}\theta_t&=U^{-1}\theta_{t-1}+\alpha(U^{-1} b_t- U^{-1}A_t U U^{-1}\theta_{t-1}),\\
+%U^{-1}\theta_t&=U^{-1}\theta_{t-1}+\alpha(U^{-1} b_t- U^{-1}A_t U U^{-1}\theta_{t-1}),\\
 \gamma_t&=\gamma_{t-1}+\alpha(U^{-1} b_t- J_t\gamma_{t-1}),
 \end{align*}
-This completes the transformation of a Hurwitz problem $P$ to a positive definite problem $P_U$.
-One can revert to $\theta_t$ by using $\theta_t=U\gamma_t$, and this explains the reason for the condition number $\cond{U}$ in \Cref{th:rate}.
-We would like to add that when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}.
-We would like to mention that such similarity transformations have been in the past for analyzing LSA algorithms in RL \cite{lihong} (see \Cref{sec:related} for more details).
+where $J_t = U^{-1}A_t U$.
+Now, since any Hurwitz matrix is similar to a PD matrix with an appropriate similarity transformation $U$, choosing $U$ this way we get that $\E[J_t]$ is PD, hence $P_U$ is PD.
+%This completes the transformation of a Hurwitz problem $P$ to a positive definite problem $P_U$. 
+The condition number $\cond{U}$ appears in our bound because $\theta_t$ is recovered from $\gamma_t$ using
+$\theta_t=U\gamma_t$. Of course, when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}. 
 \item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} When $P$ is positive definite, we can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{cross} terms of$\EE{\ip{e_i,e_j}},i\neq j$.

### 2017-05-18 linal.tex
main results

@@
+%!TEX root =  flsa.tex
 \appendix
@@
 \paragraph{Proof of \Cref{lm:hur}}
-\begin{lemma}
+\begin{lemma} %\label{lm:hur}
 Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}, then there exists an $\alpha_{P_U}>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_P)$.

### 2017-05-18 prob.tex
iid fix
main results

@@
 \begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \label{dist} $(b_t, A_t)\sim P$, $t\geq 0$ is an $i.i.d$ sequence.
+\item \label{dist} $(b_t, A_t)\sim P$, $t\geq 0$ is an \iid sequence.
 We let $A_P$ be the expectation of $A_t$, $b_P$ be the expectation of $b_t$, as in \cref{def:dist}.
@@
 We assume that $P$ is Hurwitz.
-\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following
+\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following \todoc{This would be the place to state if we make assumptions above the correlations between $M_t$ and $N_t$.}
 \begin{align*}
@@
 We are interested in the behavior of the mean squared error (MSE) at time $t$ given%
-\footnote{ $\EEP{\norm{\thh_t-\ts}_C^2}$, the MSE with respect to a PD Hermitian matrix $C$ is desirable.
-However, we cannot exploit the use of $C$ unless it is connected to $P$ in a special way. 
-For more discussion, see \cref{sec:related}.}
+\footnote{More generally, one can be interested in $\EEP{\normsm{\thh_t-\ts}_C^2}$, the MSE with respect to a PD Hermitian matrix $C$.
+Since in general it is not possible to exploit the presence of $C$ unless it is connected to $P$ in a special way,
+here we restrict ourselves to $C = \I$. For more discussion, including the discussion of the case of SGD for linear least-squares when $C$ and $P$ are favourably connected see \cref{sec:related}.}
  by $\EE{\normsm{\thh_t-\ts}^2}$. \todoc{Move up footnote. Polish..}
 \begin{comment}
-\footnote{We note here that $\EE{P\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$.
+\footnote{We note here that $\E[\normsm{\thh_t-\ts}_C^2]$, which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$.
 However, unless $C$ has some special structure, it is not possible to exploit this generality in our bounds.

### 2017-05-18 instance.tex
discussion

@@
 \leq
-\nu_P\,
-\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,,
+\nu\,
+\left\{\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{v^2}{t+1} \right\}\,,
 \end{align*}
-where $\nu_P = \left(1+\tfrac2{\alpha\rhod{P_U}}\right)\tfrac{\cond{U}^2}{\alpha \rhos{P_U}}$.
+where $\nu = \left(1+\tfrac2{\alpha\rhod{P_U}}\right)\tfrac{\cond{U}^2}{\alpha \rhos{P_U}}$ and
+$v^2 = 
+\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}$.
 \end{theorem}
-
+Note that $\nu$ depends on $P_U$ and $\alpha$, while $v^2$ in addition also depends on $\theta_0$. The dependence,  when it is essential, will be shown as a subscript.
 \begin{theorem}[Lower Bound]\label{th:lb}
@@
 there exists $\alpha_P>0$ so that $\rhos{P}>0$ and $\rhod{P}>0$ hold for all $\alpha\in (0,\alpha_P)$ and
+for any $t\ge 0$, \todoc{$t\ge 1$?}
 \begin{align*}
-\EE{\norm{\thh_t-\ts}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big\{ \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \big(\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}\big) \Big\}
+\EE{\normsm{\thh_t-\ts}^2} 
+&\geq \frac{1}{\alpha^2 \, \rhod{P}\rhos{P}} \,\left\{ \frac{\beta_t \norm{\theta_0-\ts}^2}{(t+1)^2} 
++ \frac{v^2\sum_{s=1}^t \beta_{t-s}  }{(t+1)^2} \right\}\,,
+%&\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big\{ \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\
+%&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \big(\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}\big) \Big\}\,.
 \end{align*}
-
+where $\beta_{t} =  \big(1-(1-\alpha \rhos{P})^t\big)$ and $v^2$ is as in \cref{th:rate}.
 \begin{comment}
@@
 \end{theorem}
-
+Note that $\beta_t \to 1$ as $t\to\infty$. Hence, the lower bound essentially matches the upper bound.
 \begin{comment}
@@
 In what follows, we discuss the specific details of these results.
-\vspace*{0.5em}
+
+%\vspace*{0.5em}
 %\subsection{Result Discussion}
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{Role of $U$:} When the given problem $P$ is positive definite, it is straightforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$ by considering the definitions of $\rhos{P}$ and $\rhod{P}$.
+%\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\textbf{Role of $U$}: When the given problem $P$ is positive definite, it is straightforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$ by considering the definitions of $\rhos{P}$ and $\rhod{P}$.
 However, the problem $P$ being Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}). 
@@
 by applying a similarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$.
-To see how this works, let $\gamma_t\eqdef U^{-1 }\theta_t$.
-Premultiplying
+This can be done because any Hurwitz matrix is similar to a PD matrix.
+Taking $U$ to be this similarity matrix, we transform  \eqref{conststep} by premultiplying it by $U^{-1}$ and introducing a new variable $\gamma_t\eqdef U^{-1 }\theta_t$.
+\if0
+To see how this works, let $\gamma_t\eqdef U^{-1 }\theta_t$. Premultiplying 
  \eqref{conststep} by $U^{-1}$ and plugging in $\gamma_t$, we get
@@
 Now, since any Hurwitz matrix is similar to a PD matrix with an appropriate similarity transformation $U$, choosing $U$ this way we get that $\E[J_t]$ is PD, hence $P_U$ is PD.
+\fi
 %This completes the transformation of a Hurwitz problem $P$ to a positive definite problem $P_U$. 
@@
 $\theta_t=U\gamma_t$. Of course, when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}. 
-\item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} When $P$ is positive definite, we can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{cross} terms of$\EE{\ip{e_i,e_j}},i\neq j$.
-The growth of the diagonal terms and the cross terms depends on the spectral norm of the random matrix $H_t=I-\alpha A_t$ and the deterministic matrix $H_P=I-\alpha A_P$ respectively.
+More generally, one can always take $U$ in the result that leads to the smallest bound.
+
+\textbf{Role of $\rhos{P}$ and $\rhod{P}$}: 
+When $P$ is positive definite, we can expand the MSE as 
+\begin{align}
+\label{eq:exp}
+	\EE{\norm{\eh_t}^2}=\tfrac{1}{(t+1)^2}\, \ip{ \textstyle\sum_{s=0}^t e_s,\textstyle\sum_{s=0}^t e_s}\,,
+\end{align} 
+where $\eh_t = \thh_t-\ts$ and $e_t = \theta_t-\ts$.
+The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_s,e_s}}$ and \emph{cross} terms of $\EE{\ip{e_s,e_q}}$, $s\neq q$.
+The growth of the diagonal terms and the cross terms depends on the spectral norm of the random matrices $H_t=I-\alpha A_t$ and that of the deterministic matrix $H_P=I-\alpha A_P$, respectively.
 These are given by
@@
 \begin{align}\label{eq:spectralrand}
-\EE{\norm{I-\alpha A_t}^2}=1-\alpha \rhos{P}, {\norm{I-\alpha A_P}^2}=1-\alpha \rhod{P}
+\EE{\norm{I-\alpha A_t}^2}=1-\alpha \rhos{P}, \quad  {\norm{I-\alpha A_P}^2}=1-\alpha \rhod{P}\,,
 \end{align}
-Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:dist} to the spectral norms matrices $H_t$ and $H_P$.
-For the MSE to be bounded, we need the spectral norms to be less than unity, which in turn translates to the requirement that $\rhos{P}>0$ and $\rhod{P}>0$.
-If $P$ is Hurwitz, we can argue on similar lines by first transforming it into a positive definite problem $P_U$ and replacing $\rhos{P}$ and $\rhod{P}$ by $\rhos{P_U}$ and $\rhod{P_U}$, and introducing $\cond{U}$ to account for the forward ($\gamma=U^{-1}\theta$) and reverse ($\theta=U\gamma$) transformations using $U^{-1}$ and $U$ respectively.
-\item \textbf{Constants} {$\alpha$, $\rhos{P}$ and $\rhod{P}$} do not affect the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
-\item \textbf{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms.
-The first term is the bias term given by $\B=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}$.
-The \emph{bias} term that captures the rate at which the initial condition $\norm{\theta_0-\ts}^2$ is forgotten.
-The second term is the \emph{variance} given by $\V=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}} \frac{\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} $.
-The variance term that captures the rate at which noise is rejected.
-\item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of step-size, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term.
-This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower.
-However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:rate} where $\alpha$ is seen to multiply the variances $\sigma^2_A$ and $\sigma^2_b$.
-In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other.
-For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.
-\item \textbf{Lower Bound} in \Cref{th:lb} shows that the factor $\rhos{P}\rhod{P}$ appearing in the upper bound in \Cref{th:rate} is inevitable for general LSA with CS-PR.
-However, there are specific examples such as SGD for the linear regression problem considered by \citet{bach} where this factor can be avoided (for further remarks see \Cref{sec:related}).
-\end{itemize}
+justifying the appearance of $\rhos{P}$ and $\rhod{P}$.
+%Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:dist} to the spectral norms matrices $H_t$ and $H_P$. 
+For the MSE to be bounded, we need the spectral norms to be less than unity, implying the conditions $\rhos{P}>0$ and $\rhod{P}>0$.
+If $P$ is Hurwitz, we can argue on similar lines by first transforming $P$ into a positive definite problem $P_U$ and replacing $\rhos{P}$ and $\rhod{P}$ by $\rhos{P_U}$ and $\rhod{P_U}$, and introducing $\cond{U}$ to account for the forward ($\gamma=U^{-1}\theta$) and reverse ($\theta=U\gamma$) transformations using $U^{-1}$ and $U$ respectively.
+
+\textbf{Constants} {$\alpha$, $\rhos{P}$ and $\rhod{P}$} do not affect the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.%
+\todoc{Why is this a great thing? Refer to other works etc.}
+
+\textbf{Bias and Variance}: The MSE at time $t$ is bounded by a sum of two terms.
+The first \emph{bias} term is given by $\B=\nu_{P_U,\alpha} \,\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}$, bounding how fast the initial error $\norm{\theta_0-\ts}^2$ is forgotten.
+The second \emph{variance} term is  given by $\V=\nu_{P_U,\alpha}\, \frac{\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} $ and captures the rate at which noise is rejected.
+
+\textbf{Behaviour for extreme values of $\alpha$}: 
+As $\alpha\to 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ there.
+This is unavoidable (see also \cref{th:lb}) and is due to the slow forgetting of initial conditions for small $\alpha$.
+Small step-sizes are however useful to suppress noise, as seen from that in our bound $\alpha$ is seen to multiply the variances $\sigma^2_A$ and $\sigma^2_b$.
+In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms are trading off the two types of errors.
+For larger values of $\alpha$ with $\alpha_P$ chosen so that $\rhos{P}\ra 0$ as $\alpha\ra \alpha_{P}$  (or $\alpha_{P_U}$ as the case may be), the bounds blow up again.
+
+\textbf{The lower bound} of \Cref{th:lb} shows that the upper bound of \cref{th:rate} is tight in a number of ways.
+In particular, the coefficients of both the $1/t$ and $1/t^2$ terms inside $\{ \cdot \}$ are essentially matched.
+Further, we also see that
+the $(\rhos{P}\rhod{P})^{-1}$ appearing in $\nu_{P_u,\alpha}$ cannot be removed from the upper bound. 
+Note however that there are specific examples, such as SGD for linear least-squares,
+where this latter factor can in fact be avoided (for further remarks see \Cref{sec:related}).
+
 \input{illustrate}

### 2017-05-18 intro.tex
discussion

@@
 \citet{bach} sells their result as a non-asymptotic result. How can they do this?
-Can we be more explicit where in the paper $C_P'/t^2$ shows up -- see the notation below? We must be very explicit here.
+Can we be more explicit where in the paper $C_{P',\alpha}/t^2$ shows up -- see the notation below? We must be very explicit here.
 The reviewers may not like this.
@@
 the MSE %of a given LSA with CS-PR
-is at most $\frac{C_P}{t}+\frac{C_P'}{t^2}$ with some positive constants $C_P,C_P'$ that we explicitly compute from $P$.
+is at most $\frac{C_{P,\alpha}}{t}+\frac{C_{P',\alpha}}{t^2}$ with some positive constants $C_{P,\alpha},C_{P',\alpha}$ that we explicitly compute from $P$.
 %The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
@@
 one can choose some step-size $\alpha$ 
-such that $C_P$ from above is uniformly bounded (i.e., replicating the result of \citet{bach}).%
-\footnote{Of course, the term $C_P'/t^2$ needs to be controlled, as well.
-Just like \citet{bach}, here we focus on $C_P$, which is justified if one considers the MSE as $t\to\infty$.
+such that $C_{P,\alpha}$ from above is uniformly bounded (i.e., replicating the result of \citet{bach}).%
+\footnote{Of course, the term $C_{P',\alpha}/t^2$ needs to be controlled, as well.
+Just like \citet{bach}, here we focus on $C_{P,\alpha}$, which is justified if one considers the MSE as $t\to\infty$.
 Further justification is that we actually find a negative result.
@@
 %that there is a class $\P$ that does not `admit' a constant step-size $\alpha_{\P}$ that guarantees uniform performance $\forall P\in \P$. 
-In particular, the example applies to RL, hence, we get a negative result for RL, which states that from only bounds on the data one cannot choose a step-size to guarantee that $C_P$ of CS-PR is uniformly bounded over $\P$.
+In particular, the example applies to RL, hence, we get a negative result for RL, which states that from only bounds on the data one cannot choose a step-size $\alpha$ to guarantee that $C_{P,\alpha}$ of CS-PR is uniformly bounded over $\P$.
 We also define a subclass  $\P_{\text{SPD},B}$ of problems, related to SGD for linear regression, that does `admit' a uniform constant step-size, thereby recovering a part of the result by \citet{bach}.
@@
 where no constant step-size is available to prevent exploding MSE.
+Further, it is empirically demonstrated that the method is able to find a step-size, which is close to the theoretically recommended stepsize which results in the fastest convergence.
+\todoc{True??}
 \end{itemize}

### 2017-05-18 prob.tex
discussion

@@
 We assume that $P$ is Hurwitz.
-\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following \todoc{This would be the place to state if we make assumptions above the correlations between $M_t$ and $N_t$.}
+\item \label{matvar} The martingale difference sequences\footnote{That is, $\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$ and $M_t,N_t$ are $\F_t$ measurable, $t\ge 0$.
+\todoc[inline]{This means, $\F_0$ should be the trivial $\sigma$-algebra.}} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following \todoc{This would be the place to state if we make assumptions above the correlations between $M_t$ and $N_t$.}
 \begin{align*}

### 2017-05-18 illustrate.tex
discussion
up to sec 5

@@
-\subsection{Illustrative Examples}
-In this section, we present some illustrative examples and make some useful remarks about the same.
+%!TEX root =  flsa.tex
+%\subsection{Illustrative Examples}
+%In this section, we present some illustrative examples and make some useful remarks about the same.
+\if0
 \begin{example}\label{ex:pdas}
-Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix},0<\epsilon<0.1$ are not positive definite.
+Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$,
+$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix},0<\epsilon<0.1$ are not positive definite.
 However choosing $U_1=\begin{bmatrix} \frac{1}{\sqrt{2}}& -\frac{1}{\sqrt{2}}\\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}} \end{bmatrix}$, $U_2=\begin{bmatrix} \frac{\sqrt{2}}{1-i\sqrt{3}} & \frac{\sqrt{2}}{1+i\sqrt{3}} \\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}}\end{bmatrix}$ and $U_3=\begin{bmatrix} 1 & 0\\ 0 &\frac{1}{10} \end{bmatrix}$, we can see that matrices $J_i=U_i^{-1}A_i U_i,i=1,2,3$ are positive definite.
@@
 The above example \Cref{ex:pdas} illustrates the use of appropriate $U_i,i=1,2,3$ matrices to transform Hurwitz problems into positive definite problems.
-In particular, problem $P^2$ involves the use of a complex matrix $U_2$.
+In particular, the second problem involves the use of a complex matrix $U_2$.
+\fi
+\todoc{I commented out the Hurwitz example. I did not find it very useful.}
+The following example can shed further light on the structure of the bound:%
+\todoc{I am wondering whether this could replace Theorem 2?}
 \begin{example}\label{ex:geo}
-Consider the one dimensional case where $A_t=\rho$ for some $0<\rho<1$, $b_t$ a bounded $i.i.d$ zero mean random variable with variance $\sigma^2$, and $\ts=0$.
-We know that $e_t=(1-\alpha\rho)^t e_0+\sum_{s=1}^t (1-\alpha \rho)^{t-s}\alpha b_s$ and the MSE is given by
+Consider the one dimensional case where $A_t=\rho$ for some $0<\rho<1$, $b_t$ a bounded \iid zero mean random variable with variance $\sigma^2$, and $\ts=0$. We can easily calculate 
+that $e_t=(1-\alpha\rho)^t e_0+\sum_{s=1}^t (1-\alpha \rho)^{t-s}\alpha b_s$ and thus the MSE becomes \todoc{Check what happens with $t=0$!}
 \begin{align*}
-\EE{\norm{\thh_t}^2}&=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2\\
-&\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)
+\EE{\normsm{\thh_t}^2}&=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2\\
+&\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)\,.
 \end{align*}
 \end{example}
-The above example helps to interpret the error dynamics of \Cref{eq:lsa} in terms of sum of a geometric series with common factor $\alpha\rho$.
+%The above example helps to interpret the error dynamics of \Cref{eq:lsa} in terms of sum of a geometric series with common factor $\alpha\rho$.
 \begin{comment}

### 2017-05-19 concl.tex
up to conclusion

@@
 We presented a finite time performance analysis of LSAs with CS-PR and showed that the MSE decays at a rate $O(\frac{1}{t})$.
-Our results extended the analysis of \citet{bach} for SGD with CS-PR for the problem of linear regression with quadratic loss and \iid sampling to general LSAs with CS-PR.
+Our results extended the analysis of \citet{bach} for SGD with CS-PR for the problem of linear least-squares estimation and \iid sampling to general LSAs with CS-PR.
 Our results were directly applicable to TD class of algorithms that are LSAs.

### 2017-05-19 instance.tex
up to conclusion

@@
 Let $P$ be a distribution over $\R^d\times \R^{\dcd}$ satisfying \Cref{assmp:lsa}.
+\todoc{Is it important that the distribution is over real-valued stuff?
+Would the results work for complexed-valued data?
+I would think so.}
 Then there exists an $\alpha_{P_U}>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0$
@@
 \textbf{Role of $U$}: When the given problem $P$ is positive definite, it is straightforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$ by considering the definitions of $\rhos{P}$ and $\rhod{P}$.
-However, the problem $P$ being Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}). 
+However, the problem $P$ being Hurwitz does not guarantee positive definiteness. % (see \Cref{ex:pdas}). 
+\todoc{I commented out the examples.}
 Following  \cite{lihong}, we remedy this 

### 2017-05-19 intro.tex
up to conclusion

@@
 %!TEX root =  flsa.tex
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
 Linear stochastic approximation algorithms (LSAs) 
@@
 %LSAs are widely applied in machine learning and reinforcement learning (RL), where the aim is to compute an appropriate $\ts\in \R^d$ using only the noisy data sequence and $O(d)$ updates per iteration.
-Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear regression with quadratic loss \cite{bach,bachaistats}, and the \emph{temporal difference} (TD) class of learning algorithms in RL \cite{sutton,konda-tsitsiklis,KoTsi03:LSA,gtd,gtd2,gtdmp}.
+Some examples of LSAs include the stochastic gradient descent algorithm (SGD) for the problem of linear least-squares estimation  (LSE) \cite{bach,bachaistats}, and the \emph{temporal difference} (TD) class of learning algorithms in RL \cite{sutton,konda-tsitsiklis,KoTsi03:LSA,gtd,gtd2,gtdmp}.
 \todoc{konda-tsitsiklis reference resolved to tsitsiklis-van-roy. is this what you want?? I also added KoTsi03:LSA, maybe that's what you wanted.}
@@
 %Intuitively, the CS-PR technique can be seen to address the learning vs noise rejection trade-off, where the constant step-size takes care of the learning  and the averaging of the iterates takes care of the noise rejection.
-\paragraph{Motivation and Contribution:} Recently, \citet{bach} considered stochastic gradient descent (SGD)\footnote{SGD is an LSA of the form in \eqref{eq:lsaintro}.} with CS-PR for the linear regression problem with quadratic loss and \iid sampling.
+\paragraph{Motivation and Contribution:} Recently, \citet{bach} considered stochastic gradient descent (SGD)\footnote{SGD is an LSA of the form in \eqref{eq:lsaintro}.} with CS-PR for LSE and \iid sampling.
 They showed that one can calculate a constant step-size from only a bound on the magnitude of the noisy data so that the leading term as $t\to\infty$
-\todoc{There is a subtlety here, because, if I believe you, 
-\citet{bach} sells their result as a non-asymptotic result. How can they do this?
-Can we be more explicit where in the paper $C_{P',\alpha}/t^2$ shows up -- see the notation below? We must be very explicit here.
-The reviewers may not like this.
-}
- in the mean-squared (regression) error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data, and is in particular independent of the eigenspectrum of $\E[A_t]$, a property which is not shared by other step-size tunings and variations of the basic SGD method.
+ in the mean-squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data, and is in particular independent of the eigenspectrum of $\E[A_t]$, a property which is not shared by other step-size tunings and variations of the basic SGD method.%
+ \footnote{See \cref{sec:related} for further discussion of the nature of these results.}
 \todoc{Dimension?}
@@
 In particular, the example applies to RL, hence, we get a negative result for RL, which states that from only bounds on the data one cannot choose a step-size $\alpha$ to guarantee that $C_{P,\alpha}$ of CS-PR is uniformly bounded over $\P$.
-We also define a subclass  $\P_{\text{SPD},B}$ of problems, related to SGD for linear regression, that does `admit' a uniform constant step-size, thereby recovering a part of the result by \citet{bach}.
-\todoc{Why only part?
-We may need to explain the difference.} Our study thus sheds light on what structural assumptions are key to achieve a uniform bound for CS-PR.
+We also define a subclass  $\P_{\text{SPD},B}$ of problems, related to SGD for LSE, that does `admit' a uniform constant step-size, thereby recovering a part of the result by \citet{bach}.
+Our results in particular shed light on the precise structural assumptions that are needed 
+to achieve a uniform bound for CS-PR. 
+For further details, see \cref{sec:related}.
 \item \textbf{Automatic Step-Size}: 

### 2017-05-19 prob.tex
up to conclusion

--- 2017-05-18 prob.tex
@@
 
-In what follows, for $t\ge1$ we make use of the $\sigma$-fields $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$; $\F_0$ holds all random variables.
-\todoc{A bit unusual as a choice.}
+In what follows, for $t\ge1$ we make use of the $\sigma$-fields $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$; $\F_{-1}$ is the trivial $\sigma$ algebra.
+\todoc{You wrote $\F_0$ holds all random variables, an unusual choice.
+And it does not work, I think and the index should be $-1$.
+The martingale property breaks otherwise with the first time step.}
 We are interested in the behaviour of \eqref{eq:lsa} under the following assumption:
@@
 We assume that $P$ is Hurwitz.
-\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following \todoc{This would be the place to state if we make assumptions above the correlations between $M_t$ and $N_t$.}
+\item \label{matvar} The martingale difference sequences\footnote{That is, $\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$ and $M_t,N_t$ are $\F_t$ measurable, $t\ge 0$.} $M_t\eqdef A_t-A_{P}$ and $N_t\eqdef b_t-b_{P}$ associated with $A_t$ and $b_t$ satisfy the following
+\todoc{So there is conditioning for $M_t$, and no conditions for $N_t$? Interesting. Maybe comment on this
+that this is not a mistake.
+Also, why write $N_t^* N_t$ instead of $\norm{N_t}^2$???}
+\todoc{This would be the place to state if we make assumptions above the correlations between $M_t$ and $N_t$.}
 \begin{align*}
@@
 Further, when data is bounded $\norm{C}^2$ is also bounded and it is straightforward to see that $\EEP{\norm{\thh_t-\ts}}^2_C\leq \norm{C}^2\EEP{\norm{\thh_t-\ts}^2}$.
-So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in the linear regression setting, where its usage is relevant.}
+So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in LSE, where its usage is relevant.}
 \end{comment}

### 2017-05-19 related.tex
further updates
iid fix
main results
up to conclusion

--- 2017-05-15 related.tex
@@
-\section{Related Work}
-We first discuss the related work outside of RL setting, followed by related work in the RL setting. In both cases, we highlight the new insights that follows from the results in this paper.\par
-\paragraph{SGD for Linear Regression} As mentioned in \Cref{sec:intro}, we were motivated by the ``remarkable'' results obtained by \citet{bach}, where the authors analyze the performance of SGD for the problem of linear regression with quadratic loss.
-The problem class considered by them belongs to $\P_{\text{PSD},B}$ (see \Cref{def:pclass} ) and hence is weakly admissible with a uniform step-size choice.
-In order to produce a rate of $\frac{C}{t}$ (where $C>0$ is a problem independent constant) for the bias term \citet{bach} do not rely on admissibility, but instead use the fact that the MSE, $\EE{\norm{\thh_t-\ts}^2_{A_P}}$ for the linear prediction problem is measured in the general quadratic norm with respect to the symmetric positive definite matrix $A_P$ and the fact that the noise is `structured' i.e., the variance is bounded by $R\,A_P$ for some constant $R>0$.
-At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.
-\paragraph{Our Results:} Analysis of LSA with CS-PR goes back to the work by \citet{polyak-judisky}, wherein they considered the additive noise setting i.e., $A_t=A$ for some deterministic Hurwitz matrix $A\in \R^{\dcd}$.
-A key improvement in our paper is that we consider the `multiplicative' noise case, i.e., $A_t$ is a random matrix.
+%!TEX root =  flsa.tex
+\section{Related Work}\label{sec:related}
+We first discuss the related work outside of RL setting, followed by related work in the RL setting. In both cases, we highlight the  insights that follows from the results in this paper.
+
+\textbf{SGD for LSE}: As mentioned in the previous section, distributions underlying
+% in \Cref{sec:intro}, we were motivated by the results obtained by \citet{bach}, where the authors analyze the performance of 
+SGD for LSE with bounded data 
+is a subset of $\P_{\text{PSD},B}$ and hence is weakly admissible under a fixed constant step-size choice. 
+However, we also noted that $\P_{\text{PSD},B}$ is not admissible. This seems to be at odds with the result of \citet{bach}
+who prove that the variance term of SGD with CS-PR with an appropriate constant is bounded by
+$\frac{C}{t}$ where $C>0$ only depends on $B$. The apparent contradiction is resolved by noting that 
+{\em (i)} in SGD the natural loss is $\EE{\normsm{\thh_t-\ts}^2_{A_P}}$ with $A_P$ SPD, and 
+{\em (ii)} the noise is ``structured'', i.e., the variance \todoc{What exactly is this?} 
+is bounded by $R\,A_P$ for some constant $R>0$. 
+Nevertheless, the bias term of their result
+decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is thus dependent on $P$ (in particular, it is dependent on the condition number of $A_P$).
+\todoc{There is a subtlety here, because, if I believe you, 
+\citet{bach} sells their result as a non-asymptotic result. How can they do this?
+Can we be more explicit where in the paper $C_{P',\alpha}/t^2$ shows up -- see the notation below? We must be very explicit here.
+The reviewers may not like this.
+}
+
+\textbf{Additive vs.
+multiplicative noise}: Analysis of LSA with CS-PR goes back to the work by \citet{polyak-judisky}, wherein they considered the additive noise setting i.e., $A_t=A$ for some deterministic Hurwitz matrix $A\in \R^{\dcd}$.
+A key improvement in our paper is that we consider the `multiplicative' noise case, i.e., $A_t$ is non-constant random matrix.
 To tackle the multiplicative noise we use newer analysis introduced by \citet{bach}.
-However, since the general LSA setting (with Hurwitz assumption) does not enjoy special structures of the SGD setting of \citet{bach}, we make use of Jordan decomposition and similarity transformations in a critical way to prove our results.
-\paragraph{Notation for RL:} We are presented with data in the form of an $i.i.d$ sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
+However, since the general LSA setting (with Hurwitz assumption) does not enjoy special structures of the SGD setting of \citet{bach}, we make use of Jordan decomposition and similarity transformations in a critical way to prove our results, thus diverging from the line of analysis of
+\citet{bach}.
+
+\textbf{Results for RL}: We are presented with data in the form of an \iid sequence $(\phi_t,\phi'_t,r_t)\in \R^d\times\R^d\times \R$.
 For a fixed constant $\gamma \in (0,1)$ define  $\Delta_t\eqdef \phi_t\phi_t^\top-\gamma \phi_t\phi_t^{'\top}$, $C_t=\phi_t\phi_t^\top$ and $b_t=\phi_r r_t$.
@@
 The general setting where $\mu_t>0$ is known as \emph{off-policy}.}.
-The various TD class of algorithms that can be cast as LSAs are given in \Cref{tb:tdalgo}
+The various TD class of algorithms that can be cast as LSAs are given in \cref{tb:tdalgo}.
 \FloatBarrier
@@
 Algorithm& Update & Remark\\ \hline
-TD(0)& $\begin{aligned}\theta_t&=\theta_{t-1}+\alpha_t (b_t -\Delta_t\theta_{t-1})\end{aligned}$ & \cite{korda-prashanth}:$\alpha_t=O(\frac{1}{t})^\beta,\beta\in(0,1)$; PR-avg; $\EE{\norm{\eh_t}}= O(\frac{1}{\sqrt{t}}).$ \\ \hline
-GTD/GTD2& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}- Q_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t A^\top_t y_{t-1})\end{aligned}$
-& \cite{gtd2}: $\beta_t=\eta \alpha_t,\sum_{t\geq 0}\alpha_t=\infty, \sum_{t\geq 0}\alpha^2_t<\infty$; $e_t\ra 0$ as $t\ra\infty$ w.p.1. \cite{gtdmp}:
-$\alpha_t=\beta_t=O(\frac{1}{\sqrt{t}})$; Projection+PR; $\,e_t =O(\frac{1}{{t}^{\frac{1}{4}}})$ w.h.p.\\\hline
-TDC& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t b_t-\mu_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$ & \cite{gtd2}: $\beta_t=\eta \alpha_t,\sum_{t\geq 0}\alpha_t=\infty, \sum_{t\geq 0}\alpha^2_t<\infty$; $e_t\ra 0$ as $t\ra\infty$ w.p.1.
-\\\hline
+TD(0) 
+	& $\begin{aligned}\theta_t&=\theta_{t-1}+\alpha_t (b_t -\Delta_t\theta_{t-1})\end{aligned}$ 
+	& \cite{korda-prashanth}: $\alpha_t=O(\frac{1}{t})^\beta$, $\beta\in(0,1)$; PR-avg, ``on-policy''; $\EE{\norm{\eh_t}}= O(\frac{1}{\sqrt{t}}).$ \\ \hline
+GTD/GTD2
+	& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}- Q_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t A^\top_t y_{t-1})\end{aligned}$
+	& \cite{gtd2}: $\beta_t=\eta \alpha_t$, $\sum_{t\geq 0}\alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$; $e_t\ra 0$ as $t\ra\infty$ w.p.1. 
+	\cite{gtdmp}: $\alpha_t=\beta_t=O(\frac{1}{\sqrt{t}})$; Projection+PR; $\,\norm{e_t} =O(t^{-\frac{1}{4}})$ w.h.p.\\\hline
+TDC
+	& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t b_t-\mu_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$
+	& \cite{gtd2}: $\beta_t=\eta \alpha_t$, $\sum_{t\geq 0}\alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$; $e_t\ra 0$ as $t\ra\infty$ w.p.1. \\\hline
 \end{tabular}
 }
+\caption{Rates for TD algorithms available in the literature \cite{korda-prashanth,gtdmp,gtd2}. }
 \label{tb:tdalgo}
-\caption{Some popular TD algorithms with rates known in literature \cite{korda-prashanth,gtd2,gtdmp}. }
 \end{table}
@@
 \end{comment}
-\paragraph{Related Work in RL:} The \emph{temporal difference} learning algorithm or simply known as TD(0), is the most basic of the class of TD algorithms.
-An important shortcoming of TD(0) was its instability in the \emph{off-policy} case, and was successfully mitigated by the \emph{gradient temporal difference} learning GTD algorithm\cite{gtd2}.
-The GTD was first proposed by \citet{gtd} and its variants namely GTD2 and TDC were proposed later by \citet{gtd2}.
-However, the initial convergence analysis for GTD/GTD2/TDC were only asymptotic in nature \cite{gtd,gtd2} with diminishing step-sizes.\par
+%\paragraph{Related Work in RL:} 
+The TD(0) algorithm is the most basic of the class of TD algorithms. 
+An important shortcoming of TD(0) was its instability in the \emph{off-policy} case, which 
+was successfully mitigated by the \emph{gradient temporal difference} learning GTD algorithm \cite{gtd2}. 
+GTD was proposed by \citet{gtd}; its variants, namely GTD2 and TDC, were proposed later by \citet{gtd2}. 
+The initial convergence analysis for GTD/GTD2/TDC was only asymptotic in nature \cite{gtd,gtd2} with diminishing step-sizes.
+
 The most relevant to our results are those by \citet{korda-prashanth} in TD(0) and by \citet{gtdmp} in GTD.
 For the TD(0) case,  diminishing step-sizes $\alpha_t=O(\frac{1}{t})^\beta,\beta \in(0,1)$ with PR averaging is showed to exhibit a rate of $O(\frac{1}{t})$ decay for the MSE when $\beta\ra 1$ \cite{korda-prashanth}.
-In the case of GTD/GTD2 diminishing step-sizes $\alpha_t=O(\frac{1}{\sqrt{t}})$, projection of iterates and PR-averaging leads to a rate of $O(\frac{1}{\sqrt{t}})$ is achieved for the prediction error $\norm{A_P\thh_t-b_P}^2$ with high probability \cite{gtdmp}.
-\citet{gtdmp} also suggest a new version of GTD based on stochastic mirror prox ideas, called the GTD-Mirror-Prox which also achieves $O(\frac{1}{\sqrt{t}})$ rate for $\norm{A_P\thh_t-b_P}^2$ with high probability under similar step-size choice that was used by them for the GTD.
-\paragraph{Our Results} \Cref{th:rate} applies to TD(0)/TDC/GTD/GTD2/GTDMP with CS-PR since all the TD class of algorithms satisfy the Hurwitz assumption.
+In the case of GTD/GTD2 diminishing step-sizes $\alpha_t=O(\frac{1}{\sqrt{t}})$, projection of iterates and PR-averaging leads to a rate of $O(\frac{1}{\sqrt{t}})$ 
+for the prediction error $\norm{A_P\thh_t-b_P}^2$ with high probability \cite{gtdmp}. 
+\citet{gtdmp} also suggest a new version of GTD based on stochastic mirror prox ideas, called the GTD-Mirror-Prox, 
+which also shown to achieve an $O(\frac{1}{\sqrt{t}})$ rate for $\norm{A_P\thh_t-b_P}^2$ with high probability under similar step-size choice that was used by them for the GTD.
+
+All previous results on these RL algorithms assume that \Cref{assmp:lsa} holds (the Hurwitz assumption is satisfied by definition for on-policy TD(0), while it holds by design for the others).
+\todoc{Are we sure?
+How about TDC? How about GTD-Mirror-Prox?}
+Thus, \Cref{th:rate} applies to all of TD(0)/TDC/GTD/GTD2/GTDMP with CS-PR in all cases considered in the literature.
 Finite time performance of TDC was unkown in prior literature and our work provides a first step by understanding TDC with CS-PR.
-Further, our results show that the error in the GTD class of algorithms decay at $O(\frac{1}{t})$ rate (even without use of projection or mirror maps) instead of $O(\frac{1}{\sqrt{t}})$,  which is an improvement.
+\todoc{I think this could be considered a major result if the Hurwitz condition can be shown to hold.}
+Further, our results show that the error in the GTD class of algorithms decay at the $O(\frac{1}{t})$ rate (even without use of projection or mirror maps) instead of $O(\frac{1}{\sqrt{t}})$, a major improvement on previously published results.
 In comparison to the TD(0) results by \citet{korda-prashanth}, \Cref{th:rate} is better in that it provides the bias/variance decomposition.
-Further, their results are specific to the TD(0) where $A_P$ is known to be positive definite, and hence cannot be extended in a straightforward manner to GTD class or TDC.
-However, it is important to note that \citet{korda-prashanth} handle the Markov noise case which is not dealt with in this paper.
+Further, their results are specific to TD(0) where $A_P$ is known to be positive definite, and it is unclear whether it extends in a straightforward manner to the GTD class or TDC.
+\todoc{Would not the similarity transform idea work?
+Why not?
+A comment is needed.}
+However, it is important to note that \citet{korda-prashanth} handle the Markov noise case which is not dealt with in this paper. \todoc{How about others? We should mention if others also assume iid.}
 \begin{comment}

### 2017-05-19 stepsizes.tex
up to conclusion

@@
+%!TEX root =  flsa.tex
 \section{Automatic Tuning of Step-Sizes}
@@
 We now present some heuristic arguments in favour of a constant step-size over asymptotically diminishing step-sizes in \eqref{eq:lsaintro}.
-It has been observed that when the step-sizes of form $\alpha_t=\frac{c}{t}$ or $\alpha_t=\frac{c}{c+t}$ (for some $c>0$) are used, the MSE, $\EE{\norm{\theta_t-\ts}^2}$, is not robust to the choice of $c>0$ \cite{korda-prashanth,bach-moulines}.
-In particular only a $O(\frac{1}{t^{\mu c}/2})$ decay can be achieved for the MSE, where $\mu$ is the smallest positive part of the $A_P$ \cite{bach-moulines}.
+It has been observed that when the step-sizes of form $\alpha_t=\frac{c}{t}$ or $\alpha_t=\frac{c}{c+t}$ (for some $c>0$) are used, the MSE, $\EE{\normsm{\theta_t-\ts}^2}$, is not robust to the choice of $c>0$ \cite{korda-prashanth,bach-moulines}.
+In particular only a $O(\frac{1}{t^{{\mu c}/2}})$ decay can be achieved for the MSE, where $\mu$ is the smallest positive part of the eigenvalues of \todoc{I added: ``eigenvalues of''} $A_P$ \cite{bach-moulines}.
 Note that, in the case of LSA with CS-PR, \Cref{th:rate} guarantees a $O(\frac{1}{t})$ rate of decay for the MSE and the problem dependent quantities affect only the constants and not the exponent.
@@
 Independently, \citet{dann} also observe in their experiments that a constant step-size is better than diminishing step-sizes.\par
-We would like to remind that in \Cref{sec:uniform} we showed that weak admissibility might not hold for all problem classes, and hence a uniform choice for the constant step-size might not be possible, However, motivated by \Cref{th:rate} and also by the usage of constant step-size in practice \cite{dann,gtd2,gtdmp}, we suggest a heuristic algorithm to tune the constant step-size (see \Cref{alg:tuning}).
+We would like to remind that in \Cref{sec:uniform} we showed that weak admissibility might not hold for all problem classes, and hence a uniform choice for the constant step-size might not be possible, However, motivated by \Cref{th:rate} and also by the usage of constant step-size in practice \cite{dann,gtd2,gtdmp}, we suggest a natural algorithm to tune the constant step-size, shown as \Cref{alg:tuning}:
 \begin{algorithm}[H]
@@
 \STATE{$\theta_t=\theta_{t-1}+\alpha (b_t-A_t\theta_{t-1}), \thh_t=\thh_{t-1}+\frac{1}{t+1}(\theta_t-\thh_{t-1})$}
-\IF{$IsUnstable(\norm{\theta_t},\ldots,\norm{\theta_{(t-kT)\wedge 0}})=1$}
+\IF{$IsUnstable(\normsm{\thh_t},\ldots,\normsm{\thh_{(t-kT)\wedge 0}})=\mathrm{True}$}
 \STATE{$\alpha=\alpha/2$}
@@
 \end{algorithm}
-In \Cref{alg:tuning} $T>0$ is a time epoch and $k$ is a given integer and $\alpha_{\max}>0$ is the maximum step-size that is allowable.
-It is easy to show using a Gronwall argument that in \Cref{alg:tuning} $\norm{\theta_t}\leq (1-e^{\beta t})$, where ($\beta>0)$) $\beta<0$ implies (in)stability of iterates.
-Using this fact, we observe that the sequence $r_i=\frac{\norm{\theta_{(t-kT+iT)\wedge 0}}}{\norm{\theta_{(t-kT+(i-1)T)\wedge 0}}},i=1,\ldots,k$ should be `roughly' (making allowance for the persistent noise) decreasing and converge to $1$.
-The $IsUnstable()$ routine in \Cref{alg:tuning} can take this sequence of $r_i$s as input, and in the case of instability, the step-size can be scaled down by a factor of $2$ (or any constant factor).
+In \Cref{alg:tuning}, $T>0$ is a time epoch and $k$ is a given integer and $\alpha_{\max}>0$ is the maximum step-size that is allowable.
+From the Gronwall-Bellman lemma it follows that in \Cref{alg:tuning} $\norm{\theta_t}\leq C(1+e^{\beta t})$ with some $C>0$, where the sign of $\beta$ determines whether the iterates are bounded.
+Using this fact, we observe that the sequence $r_i=\frac{\normsm{\thh_{(t-kT+iT)\wedge 0}}}{\normsm{\thh_{(t-kT+(i-1)T)\wedge 0}}},i=1,\ldots,k$ should be ``roughly'' (making allowance for the persistent noise) decreasing and converge to $1$ when the step-size is large enough so that the iterates stay bounded and eventually converge.
+The idea is that the $IsUnstable()$ routine in \Cref{alg:tuning} calculates $\{r_i\}_i$ based on its input and returns true when any of these is larger than a preset constant $c>1$.
+By choosing a larger the constant $c$, the probability of false detection of a run-away event decreases rapidly, while still controlling for the probability of altogether missing a run-away event.

### 2017-05-19 uniform.tex
up to conclusion

@@
+%!TEX root =  flsa.tex
 \section{Uniform bounds}\label{sec:uniform}
-If $\P$ is weakly admissible, then one can choose a step-size $\alpha_{\P}>0$ solely based on the knowledge of $\P$ and
-conclude that no matter the joint distribution $P$ of $(b_t,A_t)$, as long as the distribution of $P$ is an element of $\P$, the conclusions of \Cref{th:rate} hold.
-Further, if $\P$ is admissible then the error bound stated in  \cref{maintheorem} becomes independent of the instance.
+If $\P$ is weakly admissible, \todoc{Maybe moving the definitions concerning admissibility here would be better.
+The text could be shortened, etc.
+Ideally, everything is defined just before its used for the first time, except some things which are needed everywhere.}
+then one can choose some step-size $\alpha_{\P}>0$ solely based on the knowledge of $\P$ and
+conclude that for any $P\in \P$, the MSE will be bounded as shown in \Cref{th:rate}. 
+When $\P$ is not weakly admissible but rich enough to include the examples showing \cref{th:lb}, 
+no fixed step-size can guarantee bounded MSE for all $P\in \P$.
+On the other hand, if $\P$ is admissible then the error bound stated in  \cref{th:rate} becomes independent of the instance,
+while when $\P$ is not admissible, but ``sufficiently rich'', this does not hold.
 Hence, an interesting question to investigate is whether a given set $\P$ is (weakly) admissible.
-We consider some reasonable classes of distributions and test whether or not they are (weakly) admissible.
-\begin{definition}\label{def:pclass}
-\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
-\item For $B>0$, let $\P_B$ be the set of distributions over $\C_B^{d}\times \C_B^{\dcd}$ where for every $P\in \P_B$ is positive definite.
-\item  $\P_{\text{PSD},B}$ be such that  $\P^M_{\text{PSD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PSD} \}\,$ , where PSD denotes the set of positive semi-definite matrices.
-%\item  $\P_{\text{PD},B}$ be such that  $\P^M_{\text{SD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PD} \}\,$ , where PD denotes the set of positive definite matrices.
-\end{itemize}
-\end{definition}
-Note that all distributions are a subset of $\P_B$ and in most applications the knowledge of a suitable $B>0$ is available,and by appropriately normalizing the vectors and matrices involved, we may as well assume that $B=1$ without any loss of generality.
-\begin{proposition}\label{notwad}
-The set $\P_B$ is not weakly admissible.
+
+A reasonable assumption is that $(b_t,A_t)\in \C_B^{d}\times \C_B^{\dcd}$ with some $B>0$ (i.e., the data  is bounded with bound $B$) and that $A_P$ is positive definite for $P\in \P$.
+Call the set of such distributions will $\P_B$.
+\todoc{By allowing complex valued data, \cref{lm:notwad}, a negative result, is weakened. In fact, for precision, we probably should make it clear what distributions need to be in $\P_B$ for
+\cref{lm:notwad} to hold.}
+Is positive definiteness and boundedness sufficient for weak admissibility? The answer is no:
+% $\P_{\text{PD},B}$ be such that  $\P^M_{\text{SD},B}=\{ P^M \in \P^M_B \,:\,  \supp(P^M)\subset \text{PD} \}\,$ , where PD denotes the set of positive definite matrices.
+\begin{proposition}\label{lm:notwad}
+For any fixed $B>0$,
+the set $\P_B$ is not weakly admissible.
+% i.e.,
+%positive definiteness alone is insufficient for weak admissibility.
 \end{proposition}
+Consider now the strict subset of $\P_B$ that contains distributions $P$ such that for any $A$ in the support of $P$, $A$ is PSD.
+Call the resulting set of distributions $\P_{\text{PSD},B}$.
+Note that the distribution of data originating from linear least-squares estimation with SGD is of this type.
+Is $\P_{\text{PSD},B}$ weakly admissible? The answer is yes in this case:
 \begin{proposition}\label{lm:ppsdbwd}
 For any $B>0$, the set $\P_{\text{PSD},B}$ is weakly admissible
-and in particular any $0<\alpha < 2/B$ witnesses this.
+and in particular any $0<\alpha < 2/B$ witnesses this.%
+\todoc{So this altogether is weaker than what Bach founds. They prove strong admissibility.
+I am guessing that for this they use that the loss is special and also that the noise is special.
+This would be a good place to somehow incorporate this with a (short?) proof.
+}
 \end{proposition}
+However, admissibility does not hold for the same set:
 \begin{proposition}\label{lm:ppsdbna}
@@
 \end{comment}
-\todoc[inline]{So this altogether is weaker than what Bach founds. They prove strong admissibility.
-I am guessing that for this they use that the loss is special and also that the noise is special.
-This would be a good place to somehow incorporate this with a (short?) proof.
-}
-Thus \Cref{lm:notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
-However, when we impose additional structure in the data in the form of positive semi-definiteness,  we then observe weak admissibility in \Cref{lm:ppsdwd}, and at the same time, it is not enough to guarantee admissibility in \Cref{lm:ppsdna}.
+% However, when we impose additional structure in the data in the form of positive semi-definiteness,  we then observe weak admissibility in \Cref{lm:ppsdwd}, and at the same time, it is not enough to guarantee admissibility in \Cref{lm:ppsdna}.
 \begin{comment}

### 2017-05-19 concl.tex
conclusions

@@
+%!TEX root =  flsa.tex
 \section{Conclusion}
@@
 Our results also improved the rates in the case of the GTD class of algorithms.
-We presented conditions under which a constant step-size can be chosen uniformly for a give class of data distributions.
+We presented conditions under which a constant step-size can be chosen uniformly for a given class of data distributions.
 We showed a negative result in that not all data distributions `admit' such a constant step-size.
 This is a negative result from the perspective of TD algorithms in RL.
-We also argued heuristically that a constant step-size can be obtained in an automatic manner for a given LSA and a given data distribution.
+We also argued that a constant step-size can be obtained in an automatic manner for a given LSA and a given data distribution.
 Thus, LSAs with CS-PR technique $i)$ achieves comparable or better error rates $ii)$ alongside automatic tuning of the constant step-size can perform comparably well in practice.

### 2017-05-19 illustrate.tex
space

@@
 that $e_t=(1-\alpha\rho)^t e_0+\sum_{s=1}^t (1-\alpha \rho)^{t-s}\alpha b_s$ and thus the MSE becomes \todoc{Check what happens with $t=0$!}
-\begin{align*}
-\EE{\normsm{\thh_t}^2}&=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2\\
-&\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)\,.
-\end{align*}
+$%\begin{align*}
+\E[(\thh_t)^2]=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2
+\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)
+$.
+%\,.
+%\end{align*}
 \end{example}

### 2017-05-19 related.tex
space

@@
 In the case of GTD/GTD2 diminishing step-sizes $\alpha_t=O(\frac{1}{\sqrt{t}})$, projection of iterates and PR-averaging leads to a rate of $O(\frac{1}{\sqrt{t}})$ 
-for the prediction error $\norm{A_P\thh_t-b_P}^2$ with high probability \cite{gtdmp}. 
+for the prediction error $\normsm{A_P\thh_t-b_P}^2$ with high probability \cite{gtdmp}. 
 \citet{gtdmp} also suggest a new version of GTD based on stochastic mirror prox ideas, called the GTD-Mirror-Prox, 
-which also shown to achieve an $O(\frac{1}{\sqrt{t}})$ rate for $\norm{A_P\thh_t-b_P}^2$ with high probability under similar step-size choice that was used by them for the GTD.
+which also shown to achieve an $O(\frac{1}{\sqrt{t}})$ rate for $\normsm{A_P\thh_t-b_P}^2$ with high probability under similar step-size choice that was used by them for the GTD.
 

### 2017-05-19 instance.tex
space

@@
 \textbf{Bias and Variance}: The MSE at time $t$ is bounded by a sum of two terms.
-The first \emph{bias} term is given by $\B=\nu_{P_U,\alpha} \,\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}$, bounding how fast the initial error $\norm{\theta_0-\ts}^2$ is forgotten.
-The second \emph{variance} term is  given by $\V=\nu_{P_U,\alpha}\, \frac{\alpha^2(\sigma_A^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_A^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} $ and captures the rate at which noise is rejected.
+The first \emph{bias} term is given by $\B=\nu \,\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}$, bounding how fast the initial error $\norm{\theta_0-\ts}^2$ is forgotten.
+The second \emph{variance} term is  given by $\V=\nu\, \frac{v^2}{t+1} $ and captures the rate at which noise is rejected.
 
@@
 Further, we also see that
-the $(\rhos{P}\rhod{P})^{-1}$ appearing in $\nu_{P_u,\alpha}$ cannot be removed from the upper bound. 
+the $(\rhos{P}\rhod{P})^{-1}$ appearing in $\nu = \nu_{P_u,\alpha}$ cannot be removed from the upper bound. 
 Note however that there are specific examples, such as SGD for linear least-squares,

### 2017-05-19 instance.tex
updated experiments

@@
 %\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
+\textbf{Role of $U$}: $U$ is helpful in transforming the recursion in $\theta_t$ to $\gamma_t=U^{-1}\theta_t$, which helps in ensuring $\rhos{P_U}>0$.
+Such similarity transformation have also been considered in analysis of RL algorithms \cite{lihong}.
+The condition number $\cond{U}$ appears in our bound because $\theta_t$ is recovered from $\gamma_t$ using
+$\theta_t=U\gamma_t$. Of course, when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}.
+More generally, one can always take $U$ in the result that leads to the smallest bound.
+\begin{comment}
 \textbf{Role of $U$}: When the given problem $P$ is positive definite, it is straightforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$ by considering the definitions of $\rhos{P}$ and $\rhod{P}$.
@@
 More generally, one can always take $U$ in the result that leads to the smallest bound.
-
+\end{comment}
 \textbf{Role of $\rhos{P}$ and $\rhod{P}$}: 

### 2017-05-19 intro.tex
updated experiments

@@
 Our restrictions on the common distribution is that the ``noise variance'' should be bounded (as we consider squared errors), and that the matrix $\E[A_t]$ must be Hurwitz, i.e., all its eigenvalues have positive real parts.
-This latter assumption can be shown to be necessary for the stability of the LSA iterations.
+%This latter assumption can be shown to be necessary for the stability of the LSA iterations.
 \todoc{Add citation.
@@
 where no constant step-size is available to prevent exploding MSE.
-Further, it is empirically demonstrated that the method is able to find a step-size, which is close to the theoretically recommended stepsize which results in the fastest convergence.
+%Further, it is empirically demonstrated that the method is able to find a step-size, which is close to the theoretically recommended stepsize which results in the fastest convergence.
 \todoc{True??}
@@
 %\paragraph{Implications:} 
-In addition to TD($0$), our results directly can be applied to a host of \emph{off-policy} TD algorithms such as GTD/GTD2/TDC with CS-PR.
+In addition to TD($0$), our results directly can be applied to other \emph{off-policy} TD algorithms such as GTD/GTD2 with CS-PR.
 \todoc{How about TD($\lambda$)?
@@
 Maybe in the organization section this can be mentioned.}
-In the case of TDC only asymptotic bounds were known \cite{gtd2} and our results are a first step 
-\todoc{Why only a step?
-What is missing?} towards proving finite time bounds.
-Also, our results show that the GTD class of algorithms guarantee a $O(\frac{1}{t})$ rate for MSE (without use of projections), improving on a previous result by \citet{gtdmp} that guaranteed a $O(\frac{1}{\sqrt{t}})$ rate for this class for the projected version of the algorithm.
+%In the case of TDC only asymptotic bounds were known \cite{gtd2} and our results are a first step
+%\todoc{Why only a step? What is missing?} towards proving finite time bounds.
+In particular, our results show that the GTD class of algorithms guarantee a $O(\frac{1}{t})$ rate for MSE (without use of projections), improving on a previous result by \citet{gtdmp} that guaranteed a $O(\frac{1}{\sqrt{t}})$ rate for this class for the projected version of the algorithm.
 \todoc{And projections are problematic on their own.
@@
 \todoc{How about Prashanth's paper? Should we mention it somewhere? The reviewers will push back if we don't.}
-\paragraph{Organization:} The paper is organized as follows \todoc{Do not forget to fill this in.}
+%\paragraph{Organization:} The paper is organized as follows \todoc{Do not forget to fill this in.}

### 2017-05-19 related.tex
updated experiments

@@
 	\cite{gtdmp}: $\alpha_t=\beta_t=O(\frac{1}{\sqrt{t}})$; Projection+PR; $\,\norm{e_t} =O(t^{-\frac{1}{4}})$ w.h.p.\\\hline
-TDC
-	& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t b_t-\mu_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$
-	& \cite{gtd2}: $\beta_t=\eta \alpha_t$, $\sum_{t\geq 0}\alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$; $e_t\ra 0$ as $t\ra\infty$ w.p.1. \\\hline
+%TDC
+%	& $\begin{aligned} y_{t}&=y_{t-1}+\beta_t(\mu_t b_t -\mu_t \Delta_t \theta_{t-1}-C_t y_{t-1})\\ \theta_t &=\theta_{t-1}+\alpha_t (\mu_t b_t-\mu_t\Delta_t\theta_{t-1}+ (\Delta^\top_t-C_t)y_{t-1})\end{aligned}$
+	%& \cite{gtd2}: $\beta_t=\eta \alpha_t$, $\sum_{t\geq 0}\alpha_t=\infty$, $\sum_{t\geq 0}\alpha^2_t<\infty$; $e_t\ra 0$ as $t\ra\infty$ w.p.1. \\\hline
 \end{tabular}
@@
 How about TDC? How about GTD-Mirror-Prox?}
-Thus, \Cref{th:rate} applies to all of TD(0)/TDC/GTD/GTD2/GTDMP with CS-PR in all cases considered in the literature.
-Finite time performance of TDC was unkown in prior literature and our work provides a first step by understanding TDC with CS-PR.
+Thus, \Cref{th:rate} applies to all of TD(0)/GTD/GTD2/ with CS-PR in all cases considered in the literature.
+%Finite time performance of TDC was unkown in prior literature and our work provides a first step by understanding TDC with CS-PR.
 \todoc{I think this could be considered a major result if the Hurwitz condition can be shown to hold.}
-Further, our results show that the error in the GTD class of algorithms decay at the $O(\frac{1}{t})$ rate (even without use of projection or mirror maps) instead of $O(\frac{1}{\sqrt{t}})$, a major improvement on previously published results.
+In particular, our results show that the error in the GTD class of algorithms decay at the $O(\frac{1}{t})$ rate (even without use of projection or mirror maps) instead of $O(\frac{1}{\sqrt{t}})$, a major improvement on previously published results.
 In comparison to the TD(0) results by \citet{korda-prashanth}, \Cref{th:rate} is better in that it provides the bias/variance decomposition.

### 2017-05-19 tune.tex
updated experiments

\FloatBarrier
\begin{figure}[htp]
\resizebox{\columnwidth}{!}{
\begin{tabular}{cc}
\begin{tikzpicture}[scale=0.7]
\begin{axis}[
xlabel=$\sigma_A$,
ylabel=$\alpha_P$, legend pos= north east
]
\addplot[only marks, blue, mark=*, mark options={blue}, error bars/.cd,y dir=both,  y explicit] table [x index=0, y index=1, y error index=2]{./exp/hur};
\addplot[only marks, red, mark=square, mark options={red}, error bars/.cd,y dir=both,  y explicit] table [x index=0, y index=1, y error index=2]{./exp/hur_act};
\addlegendentry{\Cref{alg:tuning}}
\addlegendentry{closed form computation}
\end{axis}
\end{tikzpicture}

&
\begin{tikzpicture}[scale=0.7]
\begin{axis}[
xlabel=$t$,
ylabel=$\normsm{\thh_t-\ts}^2$, legend pos=north east
]

\addplot[only marks,mark=square,red,each nth point=25] plot file {./exp/hur1};
\addplot[only marks,mark=diamond,blue,each nth point=25] plot file {./exp/hur2};
\addplot[only marks,mark=star,green,each nth point=25] plot file {./exp/hur3};
\addplot[only marks,mark=o,black,each nth point=25] plot file {./exp/hur4};
\addplot[only marks,mark=triangle,magenta,each nth point=25] plot file {./exp/hur5};
\addlegendentry{$\sigma_A=0$}
\addlegendentry{$\sigma_A=2$}
\addlegendentry{$\sigma_A=5$}
\addlegendentry{$\sigma_A=10$}
\addlegendentry{$\sigma_A=20$}

\end{axis}
\end{tikzpicture}

\end{tabular}
}
\caption{The left plot shows comparison of the constant step-size $\alpha_P$ as function of $\sigma_A$ found by \Cref{alg:tuning} versus the constant step-size computed in closed form.  The right plot shows the performance of LSA with CS-PR (with the step-size choosen by \Cref{alg:tuning}) for various $\sigma_A$ values.}
\end{figure}

### 2017-05-19 instance.tex
uploaded version

@@
 there exists $\alpha_P>0$ so that $\rhos{P}>0$ and $\rhod{P}>0$ hold for all $\alpha\in (0,\alpha_P)$ and
-for any $t\ge 0$, \todoc{$t\ge 1$?}
+for any $t\ge 1$, \todoc{$t\ge 1$?}
 \begin{align*}
@@
 Such similarity transformation have also been considered in analysis of RL algorithms \cite{lihong}.
-The condition number $\cond{U}$ appears in our bound because $\theta_t$ is recovered from $\gamma_t$ using
-$\theta_t=U\gamma_t$. Of course, when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}.
-More generally, one can always take $U$ in the result that leads to the smallest bound.
+%The condition number $\cond{U}$ appears in our bound because $\theta_t$ is recovered from $\gamma_t$ using $\theta_t=U\gamma_t$.
+Of course, when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}.
+More generally, one can always take $U$ in the result that leads to the smallest bound.\par
 \begin{comment}
@@
 This can be done because any Hurwitz matrix is similar to a PD matrix.
-Taking $U$ to be this similarity matrix, we transform  \eqref{conststep} by premultiplying it by $U^{-1}$ and introducing a new variable $\gamma_t\eqdef U^{-1 }\theta_t$.
+Taking $U$ to be this similarity matrix, we transform  \eqref{conststep} by pre-multiplying it by $U^{-1}$ and introducing a new variable $\gamma_t\eqdef U^{-1 }\theta_t$.
 \if0
-To see how this works, let $\gamma_t\eqdef U^{-1 }\theta_t$. Premultiplying 
+To see how this works, let $\gamma_t\eqdef U^{-1 }\theta_t$. Pre-multiplying 
  \eqref{conststep} by $U^{-1}$ and plugging in $\gamma_t$, we get
@@
 In such a scenario, we can guess that whether or not the algorithm forgets the bias $\norm{\theta_0-\ts}$, depends on whether the matrix product is contracting.
-One way to characterize this contracting property is to look at the spectral radius of the random matrix, which is given by
+One way to characterize this contracting property is to look at the spectral radius of the random matrix, which is given by $\EE{\norm{I-\alpha A_t}^2}=1-\alpha \rhos{P},  {\norm{I-\alpha A_P}^2}=1-\alpha \rhod{P}\,,$
 \end{comment}
+\begin{comment}
 \begin{align}\label{eq:spectralrand}
@@
 \end{align}
+\end{comment}
 justifying the appearance of $\rhos{P}$ and $\rhod{P}$.
@@
 
-\textbf{Constants} {$\alpha$, $\rhos{P}$ and $\rhod{P}$} do not affect the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.%
+\textbf{Constants} {$\alpha$, $\rhos{P}$ and $\rhod{P}$} do not affect the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
+This property is not enjoyed by all step-size schemes, for instance, step-sizes that diminish at $O(\frac{c}{t})$ are know to exhibit $O(\frac{1}{t^{\mu c/2}})$ ($\mu$ is the smallest real part of eigenvalue of $A_P$), and hence the exponent of the rates are not robust to the choice of $c>0$ \cite{bach-moulines,korda-prashanth}.
 \todoc{Why is this a great thing? Refer to other works etc.}
@@
 
-\input{illustrate}
+%\input{illustrate}

### 2017-05-19 intro.tex
uploaded version

@@
 An alternate idea, which we call LSA with constant step-size and Polyak-Ruppert averaging (LSA with CS-PR, in short), is to run \eqref{eq:lsaintro} by choosing $\alpha_t=\alpha>0$ $\forall t\geq 0$ with some $\alpha>0$, and output the average $\thh_t\eqdef\frac{1}{t+1}\sum_{i=0}^t \theta_i$.
-Thus, in LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm.
+Thus, in LSA with CS-PR, $\theta_t$ is an internal variable and $\thh_t$ is the output of the algorithm (see \Cref{sec:prob} for a formal definition of LSA with CS-PR).
 The idea is that the constant step-size leads to faster forgetting of initial conditions, while the averaging on the top
@@
 They showed that one can calculate a constant step-size from only a bound on the magnitude of the noisy data so that the leading term as $t\to\infty$
- in the mean-squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data, and is in particular independent of the eigenspectrum of $\E[A_t]$, a property which is not shared by other step-size tunings and variations of the basic SGD method.%
+ in the mean-squared prediction error after $t$ updates is at most $\frac{C}{t}$ with a constant $C>0$ that depends \emph{only} on the bound on the data, the dimension $d$ and is in particular independent of the eigenspectrum of $\E[A_t]$, a property which is not shared by other step-size tunings and variations of the basic SGD method.%
  \footnote{See \cref{sec:related} for further discussion of the nature of these results.}
@@
 
-In this paper, we study LSAs with CS-PR (thereby extending prior work by \citet{bach} from SGD to general LSAs) \todoc{Do our results imply theirs?
+In this paper, we study LSAs with CS-PR (thereby extending the scope of prior work by \citet{bach} from SGD to general LSAs) \todoc{Do our results imply theirs?
 Do we simplify improve their bounds, while extending the scope?
@@
 in an effort to understand the effectiveness of the CS-PR technique beyond SGD.
+Our analysis for the case of general LSA does not use specific structures, and hence cannot recover entirely, the results of \citet{bach} who use the problem specific structures in their analysis.
 Of particular interest is whether a similar result to that  \citet{bach} holds
@@
 \begin{itemize}[leftmargin=*]%, before = \leavevmode\vspace{-\baselineskip}]
-\item \textbf{Finite-time Instance Dependent Bounds}: For a given $P$, we  measure the performance of a given LSA with CS-PR in terms of the mean square error (MSE) given by $\EEP{\normsm{\thh_t-\ts}^2}$.
+\item \textbf{Finite-time Instance Dependent Bounds} (\Cref{sec:mainresults}): For a given $P$, we  measure the performance of a given LSA with CS-PR in terms of the mean square error (MSE) given by $\EEP{\normsm{\thh_t-\ts}^2}$.
 For the first time in the literature,
@@
 %The MSE can further be broken down to a \emph{bias} term (arising due to the initial condition $\theta_0$) that decays at a rate $O(\frac{1}{t^2})$ and a \emph{variance} term (due to the noise) that decays at at rate $O(\frac{1}{t})$.
-\item \textbf{Uniform Bounds}:
+\item \textbf{Uniform Bounds} (\Cref{sec:uniform}):
 It is of major interest to know whether for a given class $\P$ of distributions 
@@
 For further details, see \cref{sec:related}.
-\item \textbf{Automatic Step-Size}: 
+\item \textbf{Automatic Step-Size} (\Cref{sec:stepsizes}):
 The above negative result implies that in RL one needs to choose the constant step-size based on properties of the instance $P$ to avoid the explosion of the MSE.
@@
 where no constant step-size is available to prevent exploding MSE.
-%Further, it is empirically demonstrated that the method is able to find a step-size, which is close to the theoretically recommended stepsize which results in the fastest convergence.
+%Further, it is empirically demonstrated that the method is able to find a step-size, which is close to the theoretically recommended step-size which results in the fastest convergence.
 \todoc{True??}
@@
 %\paragraph{Implications:} 
-In addition to TD($0$), our results directly can be applied to other \emph{off-policy} TD algorithms such as GTD/GTD2 with CS-PR.
+In addition to TD($0$), our results directly can be applied to other \emph{off-policy} TD algorithms such as GTD/GTD2 with CS-PR (\Cref{sec:related}).
 \todoc{How about TD($\lambda$)?
@@
 %\todoc{Why only a step? What is missing?} towards proving finite time bounds.
-In particular, our results show that the GTD class of algorithms guarantee a $O(\frac{1}{t})$ rate for MSE (without use of projections), improving on a previous result by \citet{gtdmp} that guaranteed a $O(\frac{1}{\sqrt{t}})$ rate for this class for the projected version of the algorithm.
+In particular, our results show that the GTD class of algorithms guarantee a $O(\frac{1}{t})$ rate for MSE (without use of projections), improving on a previous result by \citet{gtdmp} that guaranteed a $O(\frac{1}{\sqrt{t}})$ rate for this class for the projected version\footnote{Projections can be problematic since they assume knowledge of $\norm{\ts}$, which is not available in practice.} of the algorithm.
 \todoc{And projections are problematic on their own.

### 2017-05-19 not.tex
main results
up to conclusion
uploaded version

@@
 
-Let us now state some definitions that will be useful for presenting our main results.
+Let us now state some definitions that will be useful for presenting our main results. \todoc{Actually, we should remove all definitions not needed by the main body.}
 \begin{definition}\label{def:dist}
@@
 \end{definition}
-Note that $\rhod{P}\ge \rhos{P}$.  \todoc{Strict inequality?? How about $P^M$ concentrating on zero? I changed the strict inequality to non-strict.} 
+Note that $\rhod{P}\ge \rhos{P}$.
+Here, subscripts $s$ and $d$ stand for \emph{stochastic} and \emph{deterministic} respectively.
+\todoc{Strict inequality??
+How about $P^M$ concentrating on zero?
+I changed the strict inequality to non-strict.}
 \todoc{Explain why we use subindex $d$ and $s$.}
@@
 We call a matrix $A\in \C^{\dcd}$ \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{d}$.
+If $\inf_x \ip{x,Ax}\ge 0$ then $A$ is \emph{positive semi-definite} (PSD).
 \todoc{I think usually this is defined as $\ip{Ax,x}>0$. Same issue as with $\ip{\cdot,\cdot}$.}
@@
 \end{definition}
+Note that SPD implies that the underlying matrix is real.
 \begin{definition}\label{distpd}
@@
 \end{definition}
+Though $\rhos{P}$ and $\rhod{P}$ depend only on $P^M$, we use $P$ instead of $P^M$ to avoid notational clutter.
 \begin{example}

### 2017-05-19 prob.tex
uploaded version

@@
 The update of $\theta_t$ alone is considered a form of constant step-size LSA.
-
+Sometimes $A_t$ will have a special form and then the matrix-vector product $A_t \theta_{t-1}$ can also be computed in $O(d)$ time, a scenario common in reinforcement learning\cite{sutton,konda-tsitsiklis,gtd,gtd2,gtdmp}.
+This makes the algorithm particularly attractive in large-scale computations when $d$ is in the range of thousands, or millions, or more, as may be required by modern applications (e.g., \citep{LiMaTaBo16})
+\begin{comment}
 Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(d)$ storage. 
@@
 can then be implemented in $O(d)$ time (or less, in case of sparse data) and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $d$ is in the range of thousands, or millions, or more, as may be required by modern applications (e.g., \citep{LiMaTaBo16})
-
+\end{comment}
 In what follows, for $t\ge1$ we make use of the $\sigma$-fields $\F_{t-1}\eqdef\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$; $\F_{-1}$ is the trivial $\sigma$ algebra.
@@
 	\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_{A_P}\,, \qquad 
-	\E[N_t^* N_t]=\sigma^2_{b_P}\,.
+\E\left[\norm{N_t}^2\mid\F_{t-1}\right]\leq\sigma^2_{b_P}\,.
 \end{align*}
-with some $\sigma^2_{A_P}$ and $\sigma^2_{b_P}$. \todoc{Could not these just be $\sigma^2_A$, $\sigma^2_b$, or with $M$ and $V$?}
+with some $\sigma^2_{A_P}$ and $\sigma^2_{b_P}$. Further, we assume $\EE{M_t N_t}=0$ \todoc{Could not these just be $\sigma^2_A$, $\sigma^2_b$, or with $M$ and $V$?}
 \item $A_P$ is invertible and thus the vector $\ts=A^{-1}_Pb_P$ is well-defined. \todoc{Bertsekas looked at the case when $\ts$ is well-defined but $A_P$ is not invertible. Future work..?}
@@
 \paragraph{Performance Metric:}  
-We are interested in the behavior of the mean squared error (MSE) at time $t$ given%
-\footnote{More generally, one can be interested in $\EEP{\normsm{\thh_t-\ts}_C^2}$, the MSE with respect to a PD Hermitian matrix $C$.
+We are interested in the behavior of the mean squared error (MSE) at time $t$ given by $\EE{\normsm{\thh_t-\ts}^2}$.
+{More generally, one can be interested in $\EEP{\normsm{\thh_t-\ts}_C^2}$, the MSE with respect to a PD Hermitian matrix $C$.
 Since in general it is not possible to exploit the presence of $C$ unless it is connected to $P$ in a special way,
 here we restrict ourselves to $C = \I$. For more discussion, including the discussion of the case of SGD for linear least-squares when $C$ and $P$ are favourably connected see \cref{sec:related}.}
- by $\EE{\normsm{\thh_t-\ts}^2}$. \todoc{Move up footnote. Polish..}
+\todoc{Move up footnote. Polish..}
 \begin{comment}

### 2017-05-19 related.tex
uploaded version

@@
 However, we also noted that $\P_{\text{PSD},B}$ is not admissible. This seems to be at odds with the result of \citet{bach}
-who prove that the variance term of SGD with CS-PR with an appropriate constant is bounded by
+who prove that the MSE of SGD with CS-PR with an appropriate constant is bounded by
 $\frac{C}{t}$ where $C>0$ only depends on $B$. The apparent contradiction is resolved by noting that 
 {\em (i)} in SGD the natural loss is $\EE{\normsm{\thh_t-\ts}^2_{A_P}}$ with $A_P$ SPD, and 
-{\em (ii)} the noise is ``structured'', i.e., the variance \todoc{What exactly is this?} 
-is bounded by $R\,A_P$ for some constant $R>0$. 
-Nevertheless, the bias term of their result
-decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is thus dependent on $P$ (in particular, it is dependent on the condition number of $A_P$).
+{\em (ii)} the noise (arising due to the residual error) is ``structured'', i.e., its variance \todoc{What exactly is this?}
+is bounded by $R\,A_P$ for some constant $R>0$ (see $\mathcal{A}3$, \cite{bach}).
+%Nevertheless, the bias term of their result decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is thus dependent on $P$ (in particular, it is dependent on the condition number of $A_P$).
 \todoc{There is a subtlety here, because, if I believe you, 
@@
 The various TD class of algorithms that can be cast as LSAs are given in \cref{tb:tdalgo}.
-\FloatBarrier
-\begin{table}[h]
+\begin{table}
 \resizebox{\columnwidth}{!}{
@@
 Thus, \Cref{th:rate} applies to all of TD(0)/GTD/GTD2/ with CS-PR in all cases considered in the literature.
-%Finite time performance of TDC was unkown in prior literature and our work provides a first step by understanding TDC with CS-PR.
+%Finite time performance of TDC was unknown in prior literature and our work provides a first step by understanding TDC with CS-PR.
 \todoc{I think this could be considered a major result if the Hurwitz condition can be shown to hold.}
@@
 In comparison to the TD(0) results by \citet{korda-prashanth}, \Cref{th:rate} is better in that it provides the bias/variance decomposition.
-Further, their results are specific to TD(0) where $A_P$ is known to be positive definite, and it is unclear whether it extends in a straightforward manner to the GTD class or TDC.
+%Further, their results are specific to TD(0) where $A_P$ is known to be positive definite, and it is unclear whether it extends in a straightforward manner to the GTD class or TDC.
 \todoc{Would not the similarity transform idea work?
@@
 A comment is needed.}
-However, it is important to note that \citet{korda-prashanth} handle the Markov noise case which is not dealt with in this paper. \todoc{How about others? We should mention if others also assume iid.}
+While the $i.i.d$ assumption is made in much of prior work \cite{gtd2,gtdmp}, however, it is important to note that \citet{korda-prashanth} handle the Markov noise case which is not dealt with in this paper.
+\todoc{How about others?
+We should mention if others also assume iid.}
 \begin{comment}

### 2017-05-19 tune.tex
uploaded version

@@
-\FloatBarrier
-\begin{figure}[htp]
+\begin{figure}
 \resizebox{\columnwidth}{!}{
 \begin{tabular}{cc}
-\begin{tikzpicture}[scale=0.7]
+\begin{tikzpicture}[scale=0.5]
 \begin{axis}[
 xlabel=$\sigma_A$,
-ylabel=$\alpha_P$, legend pos= north east
+ylabel=$\alpha_P$, legend pos= north east,
+title= Tuned Vs Hand Computed Step-Size
 ]
@@
 \addlegendentry{\Cref{alg:tuning}}
-\addlegendentry{closed form computation}
+\addlegendentry{Hand Computed}
 \end{axis}
@@
 &
-\begin{tikzpicture}[scale=0.7]
+\begin{tikzpicture}[scale=0.5]
 \begin{axis}[
 xlabel=$t$,
-ylabel=$\normsm{\thh_t-\ts}^2$, legend pos=north east
+ylabel=$\EE{\normsm{\thh_t-\ts}^2}$, legend pos=north east,
+title= MSE corresponding to step-size tuned by \Cref{alg:tuning}
 ]
-
 \addplot[only marks,mark=square,red,each nth point=25] plot file {./exp/hur1};
@@
 \end{tikzpicture}
-
 \end{tabular}
@@
 \caption{The left plot shows comparison of the constant step-size $\alpha_P$ as function of $\sigma_A$ found by \Cref{alg:tuning} versus the constant step-size computed in closed form.
-The right plot shows the performance of LSA with CS-PR (with the step-size choosen by \Cref{alg:tuning}) for various $\sigma_A$ values.}
+The right plot shows the performance of LSA with CS-PR (with the step-size choosen by \Cref{alg:tuning}) for various $\sigma_A$ values.
+The errors were insignificant and hence error bars are not shown in the right plot.}
+\label{fig:tune}
 \end{figure}

### 2017-05-21 exp/construn.m
updates

@@ -1,0 +1,1 @@
+function [errx]=construn(A,sigma,alpha,RunL)
iters=RunL;
tol=0.1;
k=10;
d=2;
I=eye(d);
last=1;
x=zeros(d,iters);
x_mean=zeros(d,iters);
x(:,1)=rand(d,1);
x(:,1)=zeros(d,1);
%x(:,1)=x(:,1)/norm(x(:,1));
x_mean(:,1)=x(:,1);
errx(1)=norm(x(1))^2;
b=A*[1;1];
%b=[1;1];
for i=2:iters
n=0*(rand-0.5);
At=A+sigma*(rand(d)-0.5);
x(:,i)=x(:,i-1)+alpha*(n+b -At*x(:,i-1));
x_mean(:,i)=mean(x(:,1:i),2);
errx(i)=norm(x_mean(:,i)-inv(A)*b)^2;
end;

### 2017-05-21 exp/cron.m

@@ -1,0 +1,1 @@
+clear all;
close all;
iters=20;
runs=100;
errx=zeros(runs,iters);
for i=1:runs
[errx(i,:),erry(i,:),errz(i,:)]=flsa(iters);

end;
mx=mean(errx);
my=mean(erry);
mz=mean(errz);

fptr=fopen('as','w');
for i=1:iters
fprintf(fptr,'%e\t%e\n',i,mx(i));
end
fclose(fptr);


fptr=fopen('pd','w');
for i=1:iters
fprintf(fptr,'%e\t%e\n',i,my(i));
end
fclose(fptr);

fptr=fopen('spd','w');
for i=1:iters
fprintf(fptr,'%e\t%e\n',i,mz(i));
end
fclose(fptr);

### 2017-05-21 exp/cronconst.m

@@ -1,0 +1,1 @@
+clear all;
close all;
alpha=load('hur_data');
iters=2;
sigma=[0 2 5 10 20];
RunL=1000;
err=zeros(iters,RunL);
A=[1 -10;10 1]; 
for j=1:5
fptr=fopen(strcat('hur',num2str(j)),'w');
err=zeros(iters,RunL);
  for i=1:iters
      [err(i,:)]=construn(A,sigma(j),alpha(j,2),RunL);
  end;
err_mean=mean(err);
err_std=std(err);
  for i=1:RunL
%    fprintf(fptr,'%d\t%e\t%e\n',i,err_mean(i),err_std(i));
    fprintf(fptr,'%d\t%e\n',i,err_mean(i));
  end;
  fclose(fptr);
  
end;


clear all;
close all;
alpha=load('pd_data');
iters=2;
sigma=[0 2 5 10 20];
RunL=10000;
errx=zeros(iters,RunL);
A=[1 10;0 1]; %U as below
for j=1:5
fptr=fopen(strcat('pd',num2str(j)),'w');
  for i=1:iters
  err=zeros(iters,RunL);
      [err(i,:)]=construn(A,sigma(j),alpha(j,2),RunL);
  end;
err_mean=mean(err);
err_std=std(err);
  for i=1:RunL
    fprintf(fptr,'%d\t%e\t%e\n',i,err_mean(i),err_std(i));
  end;
  fclose(fptr);
  
end;

### 2017-05-21 exp/cronstab.m

@@ -1,0 +1,1 @@
+close all;
clear all;
iters=20;
sigma=[0 2 5 10 20];
A=[10 -100; 100 10]; %No U required
A=[1 -10; 10 1]; %No U required
%A=[1 0; 0 1]; %No U required
%A=[10 100;0 10]; %U as below
U=[1 0; 0 0.1]
fptr=fopen('hur','w');
fptrd=fopen('hur_data','w');
fp=fopen('hur_last','w');

fprintf(fptr,'X Y Y_error\n');
for j=1:length(sigma)

  for i=1:iters
      [alpha(i),last(i)]=stab(A,sigma(j));
  end;

fprintf(fptr,'%d\t%e\t%e\n',sigma(j),mean(alpha),std(alpha));
fprintf(fptrd,'%d\t%e\t%e\n',sigma(j),mean(alpha),std(alpha));
fprintf(fp,'%d\t%d\n',sigma(j),mean(last));
end;
fclose(fptr);
fclose(fp);
fclose(fptrd);




close all;
clear all;
iters=20;
sigma=[0 2 5 10 20];
A=[1 10;0 1]; %U as below
U=[1 0; 0 0.1];
fptr=fopen('pd','w');
fptrd=fopen('pd_data','w');
fp=fopen('pd_last','w');
fprintf(fptr,'X Y Y_error\n');
for j=1:length(sigma)

  for i=1:iters
      [alpha(i),last(i)]=stab(A,sigma(j));
  end;

fprintf(fptr,'%d\t%e\t%e\n',sigma(j),mean(alpha),std(alpha));
fprintf(fptrd,'%d\t%e\t%e\n',sigma(j),mean(alpha),std(alpha));
fprintf(fp,'%d\t%d\n',sigma(j),mean(last));

end;
fclose(fptr);
fclose(fptrd);
fclose(fp);

### 2017-05-21 exp/flsa.m

@@ -1,0 +1,1 @@
+function [errx,erry,errz]=flsa(iters)
sigma=0.5;
alpha=0.05;
n=2;
I=eye(n);
A=[10 100; 0 10];
B=[10 9.99; 0 10];
C=[10 0; 0 10];

U=[1 0; 0 0.2];
x=zeros(n,iters);
x(:,1)=rand(n,1);
%x(:,1)=rand(n,1);
x(:,1)=x(:,1)/norm(x(:,1));

y=zeros(n,iters);
y(:,1)=ones(n,1);
y(:,1)=x(:,1);

z=zeros(n,iters);
z(:,1)=ones(n,1);
z(:,1)=x(:,1);

%y(:,1)=rand(n,1);
%y(:,1)=y(:,1)/norm(y(:,1));

errx(1)=norm(x(1))^2;
erry(1)=norm(y(1))^2;
errz(1)=norm(z(1))^2;
for i=2:iters
n=sigma*(rand-0.5);

x(:,i)=(I-alpha*A)*x(:,i-1)+n;
y(:,i)=(I-alpha*B)*y(:,i-1)+n;
z(:,i)=(I-alpha*C)*z(:,i-1)+n;
x_mean(i)=mean(x(:,i));
y_mean(i)=mean(y(:,i));
z_mean(i)=mean(z(:,i));
errx(i)=norm(x_mean(i))^2;
erry(i)=norm(y_mean(i))^2;
errz(i)=norm(z_mean(i))^2;
end;

### 2017-05-21 exp/stab.m

@@ -1,0 +1,1 @@
+function [alpha,last]=stab(A,sigma)
iters=1000;
alpha=1;
tol=0.025;
k=10;
d=2;
I=eye(d);
last=1;
x=zeros(d,iters);
x_mean=zeros(d,iters);
x(:,1)=rand(d,1);
x(:,1)=x(:,1)/norm(x(:,1));
x_mean(:,1)=x(:,1);
errx(1)=norm(x(1))^2;
b=A*[1;1];
for i=2:iters
n=0.5*(rand-0.5);
At=A+sigma*(rand(d)-0.5);
x(:,i)=x(:,i-1)+alpha*(n+b -At*x(:,i-1));
  if(mod(i,k)==0) %Peridoically check every k iterations
      r_1=(norm(x(:,i-k/2))/norm(x(:,i-k+1)));
      r_2=(norm(x(:,i))/norm(x(:,i-k/2)));
    if( r_2/r_1> (1+tol))
     alpha=alpha/2;
     last=i;
     x(:,i)=zeros(d,1);
    end;
  end;
x_mean(:,i)=mean(x(:,1:i),2);
errx(i)=norm(x(:,i))^2;

end;

### 2017-07-26 issues.tex
issues-with-flstd

\documentclass{article}
\input{pack}
% if you need to pass options to natbib, use, e.g.:
 \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
%\usepackage{varioref}
%\usepackage{thmtools}
\usepackage{aliascnt}
\newcommand{\tcite}[1]{\citeauthor{#1}~\citeyear{#1}}
\usepackage{titlecaps}
\usepackage[capitalize]{cleveref}
\usepackage{amsmath,amssymb,graphicx}
\crefname{assumption}{Assumption}{Assumption}

% BIBLIOGRAHY -----------------------------------------------------------------------------
% Option 1: bibtex natbib
%\usepackage[numbers]{natbib}
%\usepackage[colorlinks=true,linkcolor=blue,citecolor=purple,pagebackref=true]{hyperref}
% Option 2: biblatex
%\usepackage[backend=bibtex,natbib=true,backref=false]{biblatex}
%\usepackage[backend=bibtex,style=ieee,natbib=true,backref=false]{biblatex}
%\addbibresource{ref.bib}
% Adding back references to bib entries:
%\DefineBibliographyStrings{english}{%
%backrefpage = {page},% originally "cited on page"
%backrefpages = {pages},% originally "cited on pages"
%}
% Title, authors, etc. -------------------------------------------------------------------------

\title{Issues with Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
%  David S.~Hippocampus\thanks{Use footnote for providing further
%    information about author (webpage, alternative
%    address)---\emph{not} for acknowledging funding agencies.} \\
%  Department of Computer Science\\
%  Cranberry-Lemon University\\
%  Pittsburgh, PA 15213 \\
%  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle
In what follows we describe the issues with \citet{flstd}. The complete proofs of \cite{flstd} are available in the arxiv version in \cite{flstda}.
\section{Bounded Linear Space Assumption}
\citet{flstda} assume that the linear space is bounded (see assumption \textbf{$(A3)$} of \cite{flstd} and \cite{flstda}) and they invoke it in Lemma~7 (see pages 15 and 16, \cite{flstda}) to show that $f_i(\theta_{i-1})$ is a bounded random variable. We suspect the following shortcomings with assumption \textbf{$(A3)$}
\begin{itemize}
\item Is it not strong to assume the boundedness of the iterates? What if the iterates of the algorithm are not even stable? (we do know that the algorithm diverges for incorrect choice of step-sizes)
\item For the sake of an argument let us also additionally assume that the iterates are bounded by a projection step (which  \citet{flstda} don't mention explicitly). Then would the recursion unfold the same way as it does in Equation (26) of \cite{flstda} (or Equation (13) of \cite{flstd})? How are we to ensure decay/contraction (as shown in Equation (27) of \cite{flstda}) at each step (since projection might be non-contracting)?
\item The recursion eventually (see Equation (26) of \cite{flstda}) has only \emph{additive noise} and not \emph{multiplicative noise} as analyzed in our submission. The additive noise is captured in the form of the $H_\beta^2$ term in Equation (28) of \cite{flstd}. This reduction to additive noise is possible due to \textbf{$(A3)$}.
\end{itemize}
\section{Relation between Covariance Matrix and $\bar{A}_T$ matrix}
\citet{flstda} assume positive definiteness of \textbf{$(A4)$} covariance matrix $\frac{1}{T}\Phi^\top_T \Phi_T$. However, unlike regression problems, in reinforcement learning problems what appears in the recursion (see Equation (6)) is not the covariance matrix, but a different matrix $\bar{A}_T=\frac{1}{T}\sum_{i=1}^T \phi(s_i)(\phi(s_i)-\beta\phi(s_i'))^\top$ defined in pages 2, 4 (below Equation (5)), 8 and 16 of \cite{flstda}. Usually, without a sampling assumption known as the `on-policy' case (see \cite{gtd} for a discussion on `on-policy' vs `off-policy') the eigenvalues of $\bar{A}_T$ cannot be guaranteed to have all positive real parts. While \citet{flstda} mention the `on-policy' sampling in the introduction, there is no explicit sampling assumption in the list of assumption. This is worrisome for the following reasons
\begin{itemize}
\item \textbf{$(A4)$} deals only with the covariance matrix and it is not clear how that alone is sufficient to say that $\bar{A}_n-(1-\beta)\mu$ is positive definite (see below Equation (26) page 16 of \cite{flstda}). 
\item For the sake of the argument, let us suppose that $(1-\beta)\mu$ is the smallest eigenvalue of $\bar{A}_n$. However, the problem here is that $\bar{A}_n$ is not symmetric in general and it is typical in reinforcement learning applications to say a non-symmetric matrix $M$ is positive definite if $x^\top M x>0$ for all real valued vectors $x$. While it follows from this notion of positive definiteness of $M$ that all the eigenvalues of $M$ have positive real parts, there could still be eigenvalues with complex conjugate parts. Hence to assume that the smallest eigenvalue of $\bar{A}_n$ to be $(1-\beta)\mu$ might be problematic (because there could be other complex eigenvalues with smaller magnitude). While it might be true in practice that the smallest eigenvalue of $\bar{A}_n$, is greater than $(1-\beta)\mu$ to our knowledge there is no proof for this fact in literature.
\end{itemize}
\section{Blow Up}
We would like to note that the rate expression in Corollary~4 of \cite{flstda} (or Corollary~2 of \cite{flstd}) contains a constant $C$. The authors do mention that the sampling error blows up as $\alpha \rightarrow 1$, however, it also looks like even constant $C=\sum_{n=1}^\infty \exp(-\mu cn^{1-\alpha})$ blows up as $\alpha\rightarrow 1$, in which case it is not clear whether a $1/\sqrt{n}$ rate is achieved.
\bibliographystyle{plainnat}
\bibliography{issuesref}
%\newpage
%\input{linal}

%$f_m(\theta_{m-1})=(\theta^\top \phi(s_{i_m})-(r_{i_m}+\beta\theta^\top \phi(s_{i_{m+1}})))$
\end{document}

### 2017-07-27 instance.tex
issues fixed

@@
 \textbf{Constants} {$\alpha$, $\rhos{P}$ and $\rhod{P}$} do not affect the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
-This property is not enjoyed by all step-size schemes, for instance, step-sizes that diminish at $O(\frac{c}{t})$ are know to exhibit $O(\frac{1}{t^{\mu c/2}})$ ($\mu$ is the smallest real part of eigenvalue of $A_P$), and hence the exponent of the rates are not robust to the choice of $c>0$ \cite{bach-moulines,korda-prashanth}.
+This property is not enjoyed by all step-size schemes, for instance, step-sizes that diminish at $O(\frac{c}{t})$ are known to exhibit $O(\frac{1}{t^{\mu c/2}})$ ($\mu$ is the smallest real part of eigenvalue of $A_P$), and hence the exponent of the rates are not robust to the choice of $c>0$ \cite{bach-moulines,korda-prashanth}.
 \todoc{Why is this a great thing? Refer to other works etc.}

### 2017-07-27 issues.tex
issues fixed

@@
 In what follows we describe the issues with \citet{flstd}. The complete proofs of \cite{flstd} are available in the arxiv version in \cite{flstda}.
+
+\section{Problems with the Proof}
+
+\subsection{Expected Error Bound}
+
+The proof of Theorem 1 can be found in section A.2, starting on page 16 of \cite{flstda}.
+The proof up to Eq. (26) is correct (in the definition of $M_{n+1}$, $F_n$ should be $F$).
+However, after this display, we are told that by (A4), which ensures that $\frac1T \sum_{t=1}^T \phi_t \phi_t^\top \succ \mu I$ for some $\mu>0$, $\bar A_n = \frac1n \sum_{i=1}^n \phi_i (\phi_i - \beta \phi_i')^\top$ is such that $\bar A_n - (1-\beta) \mu I $ is positive definite.
+Here, $n>0$ is an arbitrary index and we use the abbreviation $\phi_t = \phi(s_{i_t})$ and $\phi_t' = \phi(s_{i_t}')$.
+\begin{quote}
+In general, (A4) does \emph{not} imply that  $\bar A_n - (1-\beta) \mu I $ is positive definite!
+\end{quote}
+Take for example $n=1$. We would need that $\phi_1 (\phi_1 - \beta \phi_1')^\top - (1-\beta) \mu I$ is positive definite.
+It is easy to construct examples where this is not true.
+(Note that the matrices involved are \emph{not} symmetric.
+Unfortunately, none of the two papers defines what is meant by positive definite in this case.
+We assume that the definition used is that a square, real-valued matrix $A$ is called positive definite if $x^\top A x\ge 0$ for any $x$ real-valued vector of appropriate dimension.)
+
+In fact, we don't see whether this holds even when $n=T$ and we in fact suspect that this claim is false in full generality.
+
+\section{Problems with the Proof: High Probability Bound}
+
+The proof of the high probability bound starts on page 14, in section A.1.
+The first problem happens in the display on the bottom of this page in the proof of Lemma 6.
+Here, we are told that for $A = \phi \phi^\top - \beta \phi (\phi')^\top$ (we are dropping indices to remove clutter),
+\begin{align*}
+A^\top A = \norm{\phi}^2 \phi \phi^\top - \beta ( 2 - \norm{\phi}^2\beta ) \phi' (\phi')^\top\,,
+\end{align*}
+where $\norm{x}$ denotes the $2$-norm of $x$.
+However, using $A = \phi( \phi-\beta \phi')^\top$, a direct calculation gives:
+\begin{align*}
+A^\top A 
+& = ( \phi-\beta \phi') \phi^\top \phi ( \phi-\beta \phi')^\top \\
+& = \norm{\phi}^2  ( \phi-\beta \phi') ( \phi-\beta \phi')^\top \\
+& = \norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}\,,
+\end{align*}
+which does not match the previous display.
+The terms that do not match are the linear-in-$\beta$ terms.
+In the first display we have $ -2 \beta \phi' (\phi')^\top$, while in the bottom we have $-\beta (\phi' \phi^\top + \phi (\phi')^\top)$.
+
+The first equality of their display states (in equivalent form) that 
+\begin{align*}
+A^\top A = \norm{\phi}^2 ( \phi \phi^\top - 2\beta \phi (\phi')^\top +\beta^2 \phi' (\phi')^\top )\,.
+\end{align*}
+We see that while this is closer to the correct result, here the mistake is that $-2\beta \phi  (\phi')^\top$ is replacing $-\beta(\phi (\phi')^\top + \phi' \phi^\top)$.
+
+We also have a hard time following the rest of the proof of this lemma.
+In fact, we have a simple proof for a result similar to stated in this lemma.
+This works by observing that one needs to 
+bound the Lipschitz factor of $\theta \mapsto \bar T_n \dots \bar T_{t+1} ( (I-\gamma_t A_t)\theta + b_t )$, where
+$A_t = \phi_t (\phi_t -\beta \phi_t')^\top$, $b_t = r_{i_t} \phi_t$ and
+ $\bar T_t(\theta) = (I-\gamma_t \bar A_T) \theta + \bar b_T$ (we assume that the stepsize sequence is deterministic).
+ Clearly, this Lipschitz factor can be bounded by $\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 \norm{I-\gamma_t A_t}_2$
+ (the composition of Lipschitz maps is Lipschitz with a factor that is the product of the individual maps' Lipschitz factors). 
+Invoking (A3) then indeed gives some version of Lemma 7.
+By making an appropriate assumption (similar to those used in our paper), 
+$\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 $ can be controlled.
+
 \section{Bounded Linear Space Assumption}
-\citet{flstda} assume that the linear space is bounded (see assumption \textbf{$(A3)$} of \cite{flstd} and \cite{flstda}) and they invoke it in Lemma~7 (see pages 15 and 16, \cite{flstda}) to show that $f_i(\theta_{i-1})$ is a bounded random variable.
+\citet{flstda} assume that the parameter vector stays such that the value function $\Phi \theta$ (with their notation) will be bounded in $L^\infty$-norm (see assumption \textbf{$(A3)$} of \cite{flstd} and \cite{flstda}).
+This assumption is critical in establishing Lemma~7 (see pages 15 and 16, \cite{flstda}).
 We suspect the following shortcomings with assumption \textbf{$(A3)$}
 \begin{itemize}
-\item Is it not strong to assume the boundedness of the iterates? What if the iterates of the algorithm are not even stable? (we do know that the algorithm diverges for incorrect choice of step-sizes)
-\item For the sake of an argument let us also additionally assume that the iterates are bounded by a projection step (which  \citet{flstda} don't mention explicitly).
+\item The assumption is stated in a somewhat sloppy fashion.
+We take the authors meant to say that $\sup_n \norm{\Phi \theta_n}_\infty<+\infty$ holds almost surely.
+This seems like a strong assumption!
+Would not one of the main question be whether this assumption even holds?
+We see in our paper that for some cases, the iterates can blow up if the step-size is inappropriately selected!
+\item One possibility would be to modify the algorithm by adding a projection step to guarantee boundedness.
+It is still unclear to us whether this alone would ensure convergence of the error to zero.
+In any case, the analysis is invalidated if a projection step is present (basically, the algebraic identities will all fail to hold).
+%For the sake of an argument let us also additionally assume that the iterates are bounded by a projection step (which  \citet{flstda} don't mention explicitly).
 Then would the recursion unfold the same way as it does in Equation (26) of \cite{flstda} (or Equation (13) of \cite{flstd})?
 How are we to ensure decay/contraction (as shown in Equation (27) of \cite{flstda}) at each step (since projection might be non-contracting)?
-\item The recursion eventually (see Equation (26) of \cite{flstda}) has only \emph{additive noise} and not \emph{multiplicative noise} as analyzed in our submission.
+%\item The recursion eventually (see Equation (26) of \cite{flstda}) has only \emph{additive noise} and not \emph{multiplicative noise} as analyzed in our submission.
 The additive noise is captured in the form of the $H_\beta^2$ term in Equation (28) of \cite{flstd}.
@@
 While \citet{flstda} mention the `on-policy' sampling in the introduction, there is no explicit sampling assumption in the list of assumption.
+In fact, we doubt that the proposed algorithm will converge without extra assumption (as discussed above).
+\if0
 This is worrisome for the following reasons
 \begin{itemize}
-\item \textbf{$(A4)$} deals only with the covariance matrix and it is not clear how that alone is sufficient to say that $\bar{A}_n-(1-\beta)\mu$ is positive definite (see below Equation (26) page 16 of \cite{flstda}).
+\item \textbf{$(A4)$} deals only with the covariance matrix and it is not clear how that alone is sufficient to say that $\bar{A}_n-(1-\beta)\mu$ is positive definite (see below Equation (26) page 16 of \cite{flstda}; the problem mentioned earlier).
 \item For the sake of the argument, let us suppose that $(1-\beta)\mu$ is the smallest eigenvalue of $\bar{A}_n$.
@@
 \end{itemize}
-\section{Blow Up}
+\fi
+\section{Blow Up of the Bound}
 We would like to note that the rate expression in Corollary~4 of \cite{flstda} (or Corollary~2 of \cite{flstd}) contains a constant $C$.
-The authors do mention that the sampling error blows up as $\alpha \rightarrow 1$, however, it also looks like even constant $C=\sum_{n=1}^\infty \exp(-\mu cn^{1-\alpha})$ blows up as $\alpha\rightarrow 1$, in which case it is not clear whether a $1/\sqrt{n}$ rate is achieved.
+The authors do mention that the sampling error blows up as $\alpha \rightarrow 1$.
+However, it also looks like that even the constant $C=\sum_{n=1}^\infty \exp(-\mu cn^{1-\alpha})$ will blow up as $\alpha\rightarrow 1$, in which case it is not clear whether a $1/\sqrt{n}$ rate is achieved (even if we disregard all the other problems).
+
+\section{Doubt about the Effectiveness of the Algorithm}
+In Corollary~4 of \cite{flstda} (or Corollary~2 of \cite{flstd}), we learn that the value of $c$ governing the stepsize of the primary update must be in a small range (it must be between $1.33$ and $2$).
+This means, that effectively, the stepsize $\gamma_n$ behaves as $1/n^\alpha$ ($c$ has very little effect).
+At least when $\alpha=1$, we know that stepsizes like this make the bias decrease slowly and averaging remains ineffective.
+This seems to be at odds with the suggestion after this result that $\alpha\to 1$ is a desirable choice.
+In fact, we would be inclined to choose $\alpha=1/2$, i.e., its lowest value.
+This is because then the bias is decreased relatively quickly, while the variance will be controlled by the additional averaging.
+However, given all the problems with this it remains to be seen whether this is indeed a reasonable choice and under exactly what conditions this is reasonable.
 \bibliographystyle{plainnat}

### 2017-07-27 rebuttal.txt
rebuttal, too long..

We thank the reviewers for their comments.
We would like to start with answering what appears to be the main issue raised by Reviewer #3.
1) Missing citation to Prashanth, Korda and Munos' ECML'14 paper.
In fact, Prashanth and Korda later published arxiv.org/pdf/1411.3224.pdf (ICML'15), which we thought was the "state-of-the-art" and we cited this paper. Now, strangely this paper does not cite the ECML'14 paper and so we failed to find the ECML'14 paper and hence failed to cite it. Now, as to the actual contributions, unfortunately, the ECML'14 paper seems to have many problems. We believe that the stated results in it are incorrect: They miss essential conditions, after which their result would not be different from ours. This can be seen by looking at the arxiv version which has the proofs. The factual errors are many: for the expected error bound, on page 16 in Section A.2, the derivation up to (26) is correct. However, then it is claimed that \bar A_n - (1-beta) mu I is positive definite, which is easily seen to be false, e.g, when n = 1 (in which case \bar A_n is rank 1, asymmetric, has little to do with mu, the minimum eigenvalue of the empirical feature covariance matrix). This is not a minor problem: Fixing the proof requires arguments similar to ours. The high-probability bound is similarly incorrect: On page 14, the display in the bottom of the page is just incorrect. Again, fixing the proof requires extra conditions, leading to an analysis similar to ours. As a sidenote we remark that even the ICML'15 seems to have problems: We have contacted the first author, Prashanth, who confirmed the problems.

2) Work of Bach and Moulines: We think this is a precursor to and is superseded by the papers by Défossez and Bach [2] (see page 5) and that of Dieuleveut, Flammarion and Bach [3] (see page 2), which we cite and compare to (page 6, lines 226-241). In fact, we think that this comparison is truly enlightening as it brings out in what way the LMS update is a very special case of LSA.

2) Knowledge of minimum eigenvalue: Indeed, the stepsize dependence on a quantity that is most likely unknown in an application is highly problematic! Our intention if anything was to make this very clear. In fact, here is where our lower bound becomes crucial: We are showing that this is not a weakness of the analysis, but a weakness of the approach. We think that shining light on this dependence is in fact what makes the paper uniquely valuable (however unhappy one may be about this dependence!). We realize that their is a tendency in the community to prefer constructive and positive results, but we think that negative results like ours are also extremely useful: there is much to be learned from knowing what does not work (and why). Gaining this level of understanding is only possible by a careful theoretical analysis like ours. The importance of our result is further elevated by that the topic of RL/TD/.. is rapidly gaining popularity, and that numerous works attempt to transport results from supervised learning to this setting, overlooking the extra difficulties (e.g., see the papers by Prashanth et al.). Our paper may be valuable for both people solely interested in RL, but also for the fast growing community of researchers coming to RL with a supervised learning background.

3-4) Differences to Prashanth-Korda and Prashanth-Korda-Munos? Why is our bound better? Also: line 275

5) TD/GTD can of course be used with data-resampling, which will not involve Markov noise. Since the LSTD equation is not derived from a loss, we are puzzled about the comment about SGD. Mirror-prox (=GTD) comes closest to what we think SGD may look like in this setting.

Difference to state of the art: See above.
Upper bound no guidelines: Lower bound shows there will be no guidelines. The adaptive algorithm is meant to address this, but further work is needed on this for sure. We think even without this further work, the work should be interesting for the community who cares about stochastic-approximation in the RL setting.

### 2017-07-28 issues.tex
updated issues

@@
 
-\title{Issues with Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control}
+\title{Finite Time Bounds for Temporal Difference Learning with Function Approximation: Roadblocks and Errors in current state-of-the-art results}
 
@@
 \maketitle
-In what follows we describe the issues with \citet{flstd}. The complete proofs of \cite{flstd} are available in the arxiv version in \cite{flstda}.
+\begin{abstract}
+Learning the value function from the samples of a given policy is an important problem in reinforcement learning (RL).
+This problem is also known as the `on-policy' learning setting, where in samples in the form of state transitions are presented to the learner.
+It is well known that the temporal difference (TD(0)) algorithm has been used for this problem.
+At each iteration, TD(0) corrects its iterates by making an incremental step in the direction of the \emph{temporal-difference} feedback obtained from the data presented to the learner.
+In the limit, the TD algorithm converges to the solution of the projected Bellman equation.
+It is of interest to derive finite time bounds (probability as well as in expectation) for the error i.e., difference between the TD estimate at time $t$ and the final TD solution.
+This technical report discusses the possible roadblocks in attempting a finite time error analysis for TD by closely examining two important (\cite{flstda,lstdicml}) \emph{state-of-the-art} results in literature that turn out to be erroneous.
+\end{abstract}
+\section{Introduction}
+This report focusses on the specific technical details related to the TD(0) reinforcement learning algorithm.
+As a result, the report is not self contained and its scope is limited to an audience who are interested in RL and TD(0).
+Further, this report will also borrow notation freely from other sources \cite{flstda,lstdicmla}; the aim here is to quickly get to the heart of the technicalities (we request the reader to look at \cite{flstda,lstdicmla} for detailed notation).\par
+Finite time error analysis of TD(0) has attracted the interest of researchers in the recent past.
+The TD(0) algorithm has its share of similarities and differences with the stochastic gradient descent (SGD) algorithm for linear regression.
+The similarities are that both TD and SGD are stochastic approximation algorithms, they make use of noisy feedback to update their parameters and are incremental algorithms that use step-sizes.
+However, the critical difference is that TD(0) is not a gradient descent algorithm i.e., the temporal difference error does not correspond to a gradient of any function.
+We believe this non-gradient nature of the TD(0) manifests itself at various places when an analysis is attempted.\par
+For the purpose of our discussion use \cite{flstda} and \cite{lstdicmla} which are arxiv versions (containing detailed proofs) of \cite{flstd} and \cite{lstdicml} respectively.
+Unfortunately, both state-of-the-art results \cite{flstd,lstdicml} have issues in the form of bugs in the proof.
+In what follows, we re-use the notions from \cite{flstda,lstdicmla}.
+%While the arguments presented below are about technical issues in the analysis, we don't provide any proofs as such and hence the presentation turns out to be a commentary of the prior results.\par
+%The report is organized as follows.
+In \Cref{sec:flstd} we discuss the issues with the approach in \cite{flstda} and in \Cref{sec:lstdicml} the roadblocks as seen in \cite{lstdicmla} are presented.
+Both sections have other subsections which deal with the various separate difficulties.
 
-\section{Problems with the Proof}
+\section{Expected Error Bound}
 
-\subsection{Expected Error Bound}
-
+\subsection{Bugs in ``Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control"}
 The proof of Theorem 1 can be found in section A.2, starting on page 16 of \cite{flstda}.
@@
 We assume that the definition used is that a square, real-valued matrix $A$ is called positive definite if $x^\top A x\ge 0$ for any $x$ real-valued vector of appropriate dimension.)
-
 In fact, we don't see whether this holds even when $n=T$ and we in fact suspect that this claim is false in full generality.
+\subsection{Bugs in ``On TD(0) with function approximation: Concentration bounds and a centered variant with exponential convergence''}
+In page 10 of \cite{lstdicmla} the expression for $\mathbf{E}[a_{j+1}^\top a_{j+1}]$ contains quadratic probability terms.
 
 \section{Problems with the Proof: High Probability Bound}
-
+\subsection{Bugs in ``Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control"}
 The proof of the high probability bound starts on page 14, in section A.1.
@@
 We see that while this is closer to the correct result, here the mistake is that $-2\beta \phi  (\phi')^\top$ is replacing $-\beta(\phi (\phi')^\top + \phi' \phi^\top)$.
-
+\begin{comment}
 We also have a hard time following the rest of the proof of this lemma.
@@
 $\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 $ can be controlled.
+\end{comment}
+\subsection{Bugs in ``On TD(0) with function approximation: Concentration bounds and a centered variant with exponential convergence''}
+In page 14 the expression below "A is a possibly random matrix..." is not entirely justified (personal communication with one of the authors).
 
 \section{Bounded Linear Space Assumption}
+\subsection{Bugs in ``Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control"}
 \citet{flstda} assume that the parameter vector stays such that the value function $\Phi \theta$ (with their notation) will be bounded in $L^\infty$-norm (see assumption \textbf{$(A3)$} of \cite{flstd} and \cite{flstda}).
@@
 Would not one of the main question be whether this assumption even holds?
-We see in our paper that for some cases, the iterates can blow up if the step-size is inappropriately selected!
+% We see in our paper that for some cases, the iterates can blow up if the step-size is inappropriately selected!
 \item One possibility would be to modify the algorithm by adding a projection step to guarantee boundedness.
-It is still unclear to us whether this alone would ensure convergence of the error to zero.
-In any case, the analysis is invalidated if a projection step is present (basically, the algebraic identities will all fail to hold).
+It is still unclear  whether this alone would ensure convergence of the error to zero.
+In any case, the expected error bound analysis is invalidated if a projection step is present (basically, the algebraic identities will all fail to hold).
+
 %For the sake of an argument let us also additionally assume that the iterates are bounded by a projection step (which  \citet{flstda} don't mention explicitly).
@@
 \end{itemize}
+\subsection{Bugs in ``On TD(0) with function approximation: Concentration bounds and a centered variant with exponential convergence''}
+\citet{gugan} mention that (citing a personal communication with \citet{lstdicmla}) \cite{lstdicmla} assume implicitly the projection step in all the high probability bounds in \cite{lstdicmla}.
+While this implicit projection in itself does not affect the high probability bound proofs directly, it is important to note that the set to which the iterates needs to be projected has to contain the TD(0) solution, a condition that cannot be guaranteed.
+
 \section{Relation between Covariance Matrix and $\bar{A}_T$ matrix}

### 2017-07-29 rebuttal.txt
updated rebuttal

@@
 We think even without this further work, the work should be interesting for the community who cares about stochastic-approximation in the RL setting.
+
+Rev2: We can push the adaptive stepsize section to the appendix and provide proof sketch in the main body.
+
+Rev 1: Markov noise is absent in the data-resampling setting (quite common in RL [4,10,15,16]), where the noise is i.i.d.
+Novelties are analysis of general LSA schemes with multiplicative noise for constant stepsize+ PR avg (is absent in prior literature; only SGD is known); a negative result: problem dependent stepsize in RL (with matching lower bound).
+In concluding that there is no novelty in the paper, we note that the review here fails to justify his position, which is unwarranted.

### 2017-07-29 rebuttal.txt
updated rebuttal

--- 2017-07-27 rebuttal.txt
@@
 We think even without this further work, the work should be interesting for the community who cares about stochastic-approximation in the RL setting.
+
+Reviewer 2: We can push the adaptive stepsize section to the appendix and provide proof sketch in the main body.
+
+Reviewer 1: 
+1) Markov noise is absent in the data-resampling setting (quite common in RL [4,10,15,16]), where the noise is i.i.d. 
+2) Novelties are significant- i) Analysis of general LSA schemes with multiplicative noise for constant stepsize+ PR avg is new in literature.
+Prior results were known only for general LSA with additive noise and for SGD (a special case of LSA) ii) Matrices in general LSA schemes are not symmetric and prior results do not apply directly; the similarity transformation that we employ in the paper is needed to handle the non-symmetric case; this is a major step iii) We show a negative result: for general LSAs the stepsize cannot be chosen in a problem independent manner; we provide a lower bound to show that the negative result is not due to a weakness in our analysis iv) Better rates for GTD follows immediately due to our results.
+Our paper obviates the projection step in GTD; this is a major gain.
+We would like to mention that we have discussed the significance of our contributions in the paper in the introduction as well as section 6.
+In concluding that there is no novelty in the paper, we are afraid that the reviewer here fails to justify his position which is unwarranted.

### 2017-07-29 rebuttal.txt
rebuttal rewording

--- 2017-07-27 rebuttal.txt
@@
 We thank the reviewers for their comments.
-We would like to start with answering what appears to be the main issue raised by Reviewer #3.
+Reviewer 1:
+Naturally, we are disappointed by the negative recommendation, which we think is perhaps the result of overlooking some facts:
+1) Two justifications for the iid setting: The noise is iid in the data-resampling setting, which is quite common in RL [4,10,15,16].
+Understanding the differences between the standard supervised setting and the RL setting should start with a setting with the least number of differences.
+We see that already the iid noise setting presents surprises (see below).
+2) Novelties are significant:
+i) Analysis of general LSA schemes with multiplicative noise for constant stepsize combined with PR averaging is new in literature.
+Prior results were known only for general LSA with additive noise and for SGD (a special case of LSA).
+The new analysis required significant departure from the previous analysis and produces different results.
+ii) Matrices in general LSA schemes lack symmetry, hence prior results do not apply directly; the similarity transformation that we employ in the paper is needed to handle the non-symmetric case; this is a major step
+iii) We show a negative result: for general LSAs the stepsize cannot be chosen in a problem independent manner; we provide a lower bound to show that the negative result is not due to a weakness in our analysis
+iv) Our results give better than previous rates for GTD.
+This way, our paper obviates a highly problematic projection step in GTD introduced by[10]; this is a major gain.
+The projection step is problematic because the range of dual variables cannot be known a priori in GTD.
+We would like to mention that we have explicitly discussed the significance of our contributions in the paper in the introduction as well as section 6.
+Hence, we think that the reviewer's position is unwarranted and lacks justification.
+
+Reviewer 2:
+We will be happy to push the adaptive stepsize section to the appendix and provide a proof sketch in the main body.
+
+Reviewer 3:
 1) Missing citation to Prashanth, Korda and Munos' ECML'14 paper.
 In fact, Prashanth and Korda later published arxiv.org/pdf/1411.3224.pdf (ICML'15), which we thought was the "state-of-the-art" and we cited this paper.
-Now, strangely this paper does not cite the ECML'14 paper and so we failed to find the ECML'14 paper and hence failed to cite it.
+Strangely, this paper does not cite the ECML'14 paper and so we failed to find the ECML'14 paper and hence failed to cite it.
 Now, as to the actual contributions, unfortunately, the ECML'14 paper seems to have many problems.
-We believe that the stated results in it are incorrect: They miss essential conditions, after which their result would not be different from ours.
-This can be seen by looking at the arxiv version which has the proofs.
-The factual errors are many: for the expected error bound, on page 16 in Section A.2, the derivation up to (26) is correct.
-However, then it is claimed that \bar A_n - (1-beta) mu I is positive definite, which is easily seen to be false, e.g, when n = 1 (in which case \bar A_n is rank 1, asymmetric, has little to do with mu, the minimum eigenvalue of the empirical feature covariance matrix).
+We believe that the stated results in it are incorrect and fixing the bugs would lead to an analysis and results identical to ours.
+The factual errors are severe: for the expected error bound, on page 16 in Section A.2 of the arxiv version, after Eq.
+(26) it is claimed that \bar A_n - (1-beta) mu I is positive definite.
+This is easily seen to be false, e.g, when n = 1, in which case \bar A_n is rank 1, asymmetric and has nothing to do with mu, the minimum eigenvalue of the empirical feature covariance matrix.
 This is not a minor problem: Fixing the proof requires arguments similar to ours.
@@
 Again, fixing the proof requires extra conditions, leading to an analysis similar to ours.
-As a sidenote we remark that even the ICML'15 seems to have problems: We have contacted the first author, Prashanth, who confirmed the problems.
+As a sidenote, we remark that even the ICML'15 seems to have problems: (Earlier) we have contacted the first author, Prashanth, who confirmed these problems.
 
-2) Work of Bach and Moulines: We think this is a precursor to and is superseded by the papers by Défossez and Bach [2] (see page 5) and that of Dieuleveut, Flammarion and Bach [3] (see page 2), which we cite and compare to (page 6, lines 226-241).
-In fact, we think that this comparison is truly enlightening as it brings out in what way the LMS update is a very special case of LSA.
+2) Work of Bach and Moulines: We think this is a precursor to and is superseded by the papers by Défossez and Bach [2] (see page 5 in [2]) and that of Dieuleveut, Flammarion and Bach [3] (see page 2 in [3]), which we cite and compare to (page 6, lines 226-241).
+In fact, we think that this comparison is truly enlightening as it brings out in what way the LMS update is a uniquely special case of LSA.
 
@@
 Our intention if anything was to make this very clear.
-In fact, here is where our lower bound becomes crucial: We are showing that this is not a weakness of the analysis, but a weakness of the approach.
-We think that shining light on this dependence is in fact what makes the paper uniquely valuable (however unhappy one may be about this dependence!).
-We realize that their is a tendency in the community to prefer constructive and positive results, but we think that negative results like ours are also extremely useful: there is much to be learned from knowing what does not work (and why).
+In fact, here is where our lower bound becomes crucial: We are showing that the dependence is not an artifact of the analysis, but a weakness of the approach.
+Shining light on this dependence makes the paper uniquely valuable.
+We realize that there is a tendency in the community to prefer constructive and positive results, but we think that negative results like ours are also extremely useful: there is much to be learned from knowing what does not work (and why).
 Gaining this level of understanding is only possible by a careful theoretical analysis like ours.
@@
 
-3-4) Differences to Prashanth-Korda and Prashanth-Korda-Munos? Why is our bound better? Also: line 275
+3-4) Differences to Prashanth-Korda and Prashanth-Korda-Munos? See above. Why is our bound better? Also: line 275
 
-5) TD/GTD can of course be used with data-resampling, which will not involve Markov noise.
+5) TD/GTD can of course be used with data-resampling, which will not involve Markov noise (as recognized by many earlier, e.g., [4,10,15,16]).
 Since the LSTD equation is not derived from a loss, we are puzzled about the comment about SGD.
@@
 Difference to state of the art: See above.
-Upper bound no guidelines: Lower bound shows there will be no guidelines.
-The adaptive algorithm is meant to address this, but further work is needed on this for sure.
-We think even without this further work, the work should be interesting for the community who cares about stochastic-approximation in the RL setting.
+Upper bound gives no guideline: Lower bound shows there will be no guidelines; an adaptive algorithm is necessary (contrasting LMS).
+Our adaptive algorithm is meant to address this, but further work is needed on this for sure.
+We think even without this further work, the work should be interesting for the community as outlined above (see also the response to Reviewer 1).

### 2017-07-29 issues.tex
updated issues
rebuttal? issues..

--- 2017-07-27 issues.tex
@@
 \documentclass{article}
+\usepackage[a4paper]{geometry}
 \input{pack}
+\newcommand{\Prob}[1]{\mathbb{P}\left(#1\right)}
 % if you need to pass options to natbib, use, e.g.:
- \PassOptionsToPackage{numbers, compress}{natbib}
+ \PassOptionsToPackage{compress}{natbib}
+ \usepackage{natbib}
 % before loading nips_2017
@@
 
-\usepackage{nips_2017}
+%\usepackage{nips_2017}
 
@@
 
-\title{Issues with Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control}
+\title{Finite Time Bounds for Temporal Difference Learning with Function Approximation: Bugs in state-of-the-art results}
 
@@
 \author{
+Chandrashekar Lakshmi Narayanan
+\and
+Csaba Szepesv\'ari 
 %  David S.~Hippocampus\thanks{Use footnote for providing further
@@
 \maketitle
-In what follows we describe the issues with \citet{flstd}. The complete proofs of \cite{flstd} are available in the arxiv version in \cite{flstda}.
+\begin{abstract}
+This short report lists some problems with the recent papers by \cite{flstd,lstdicml}, whose longer version containing the proofs are available on arxiv \citep{flstda,lstdicmla}.
+\end{abstract}
+\section{Introduction}
+This short report lists some problems with the proofs of the claims in two recent papers  by \cite{flstd,lstdicml}.
+As the problems seem serious, our conclusion is that to obtain results similar to those claimed in these two works, the assumptions of the stated claims need to be considerably strengthened and the form of the results will also need to be adjusted in significant ways.
 
-\section{Problems with the Proof}
+The report is not self contained and its scope is limited to an audience who are interested in RL and TD(0) and specifically the above-mentioned papers.
+Furthermore, since \cite{flstd,lstdicml} omitted the proofs, we will instead discuss their longer version in what follows, which have identical statements but include the proofs.
+The longer version of the paper by \cite{flstd} is \citep{flstda}, while the longer version
+of \cite{lstdicml} is \citep{lstdicmla}.
+We will  borrow the notation directly from these works.
+%While the arguments presented below are about technical issues in the analysis, we don't provide any proofs as such and hence the presentation turns out to be a commentary of the prior results.\par
+%The report is organized as follows.
+In \Cref{sec:flstd} we discuss the issues with the approach in \cite{flstda} and in \Cref{sec:lstdicml} the roadblocks as seen in \cite{lstdicmla} are presented.
+Both sections have other subsections which deal with the various separate difficulties.
 
-\subsection{Expected Error Bound}
+\section{Expected Error Bound}
 
-The proof of Theorem 1 can be found in section A.2, starting on page 16 of \cite{flstda}.
+\subsection{Bugs in \cite{flstda}}
+One of the main results of \citet{flstda} is Theorem 1, which states a bound on the expected error.
+The proof of this theorem can be found in section A.2, starting on page 16.
 The proof up to Eq. (26) is correct (in the definition of $M_{n+1}$, $F_n$ should be $F$).
-However, after this display, we are told that by (A4), which ensures that $\frac1T \sum_{t=1}^T \phi_t \phi_t^\top \succ \mu I$ for some $\mu>0$, $\bar A_n = \frac1n \sum_{i=1}^n \phi_i (\phi_i - \beta \phi_i')^\top$ is such that $\bar A_n - (1-\beta) \mu I $ is positive definite.
-Here, $n>0$ is an arbitrary index and we use the abbreviation $\phi_t = \phi(s_{i_t})$ and $\phi_t' = \phi(s_{i_t}')$.
+However, after this display, we are told that by (A4), which ensures that $\frac1T \sum_{i=1}^T \phi(s_i) \phi_t(s_i)^\top \succ \mu I$ for some $\mu>0$, $\bar A_n = \frac1n \sum_{t=1}^n \phi_t (\phi_t - \beta \phi_t')^\top$ is such that $\bar A_n - (1-\beta) \mu I $ is positive definite.
+Here, $n>0$ is an arbitrary index and for $t\ge 1$ we use the abbreviation $\phi_t = \phi(s_{i_t})$ and $\phi_t' = \phi(s_{i_t}')$, where $i_t\in \{1,\dots,T\}$ is a random index.
 \begin{quote}
-In general, (A4) does \emph{not} imply that  $\bar A_n - (1-\beta) \mu I $ is positive definite!
+In general, (A4) does \emph{not} imply that  $\bar A_n - (1-\beta) \mu I $ is positive definite.
 \end{quote}
 Take for example $n=1$. We would need that $\phi_1 (\phi_1 - \beta \phi_1')^\top - (1-\beta) \mu I$ is positive definite.
-It is easy to construct examples where this is not true.
+It is easy to construct examples where this is not true: Nothing prevents, for example, $\phi_1 = \beta \phi_1'$, in which case $\bar A_1 - (1-\beta)\mu I = -(1-\beta)\mu I$ is \emph{negative} definite.
 (Note that the matrices involved are \emph{not} symmetric.
@@
 We assume that the definition used is that a square, real-valued matrix $A$ is called positive definite if $x^\top A x\ge 0$ for any $x$ real-valued vector of appropriate dimension.)
-
-In fact, we don't see whether this holds even when $n=T$ and we in fact suspect that this claim is false in full generality.
+In fact, we don't see why the claimed relationship would hold even when 
+$\bar A_n$ is replaced by $\hat A_T \doteq  \frac1T \sum_{i=1}^T \phi( s_i) (\phi(s_i) - \beta \phi(s_i'))^\top$,
+and we in fact suspect that this claim is false in full generality. But at minimum, a proof would be required and the whole subsequent argument will need to be changed.
+\subsection{Bugs in  \cite{lstdicmla}}
+In page 10 of \cite{lstdicmla} the expression for $\mathbf{E}[a_{j+1}^\top a_{j+1}]$ contains terms that involve the product of $P$ and $P^\top$.
+This cannot be correct, as here we can take the expectation first over the next state, which will bring in a \emph{single} instance of $P$.
+To remove clutter, drop the $j$ subindex, and set $A = \phi(\phi- \beta \phi')^\top$, where $\phi  = \phi(s_{I})$ and $\phi' = \phi(s_{I}')$.
+The incriminated expression from Eq. (15) of the paper is
+\begin{align}
+\EE{ A - \tfrac{\gamma}{2} A^\top A  } = \Phi^\top (I - \beta \Psi P - \tfrac{\gamma}{2}(\Delta - \beta P^\top (2I - \beta \Delta)\Psi P) ) \Phi\,.
+\end{align}
+Here, $\Phi$ is the $S\times d$ matrix whose $i$th row is $\phi^\top(s_i)$,
+$\Psi$ is the $T \times T$ diagonal matrix whose $i$th diagonal entry is $\Prob{I = i}$, while $\Delta$ is another diagonal matrix whose $i$th entry is $\norm{\phi(s_i)}_2^2$.
+A direct calculation gives that
+\begin{align}
+A^\top A 
+& = ( \phi-\beta \phi') \phi^\top \phi ( \phi-\beta \phi')^\top  \nonumber \\
+& = \norm{\phi}^2  ( \phi-\beta \phi') ( \phi-\beta \phi')^\top \nonumber \\
+& = \norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}\, \nonumber \\
+& = \norm{\phi}^2  \phi \phi^\top-\beta \norm{\phi}^2 \phi' \phi^\top -\beta \norm{\phi}^2  \phi (\phi')^\top + \beta^2 
+\norm{\phi}^2 \phi' (\phi')^\top\,.
+\end{align}
+%where we abbreviate the 2-norm with $\norm{\cdot}$.
+Hence, 
+\begin{align*}
+\EE{ A^\top A \,|\, I } = 
+\norm{\phi}^2  \phi \phi^\top-\beta \norm{\phi}^2 \EE{ \phi' \,|\, I} \phi^\top -\beta \norm{\phi}^2  \phi (\EE{ \phi' \,|\, I})^\top + \beta^2 
+\norm{\phi}^2 \EE{ \phi' (\phi')^\top\,|\, I } \,.
+\end{align*}
+Now, $\EE{ \phi' (\phi')^\top\,|\, I=i } = \phi(
+Notice that $P \Phi$ is an $S\times d$ matrix whose $i$th row is $\EE{ \phi' \,|\, I=i}^\top$.
+Hence, $ \EE{ \phi' \,|\, I} = (P \Phi)_{I:\cdot}$.
+Now, $\EE{\norm{\phi}^2 \phi \phi^\top} = \Phi \Phi^\top \Psi \Phi \Phi^\top$.
+Hence,
+\begin{align*}
+\EE{ A^\top A } = 
+\EE{\norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}}\,.
+\end{align*}
 
 \section{Problems with the Proof: High Probability Bound}
-
+\subsection{Bugs in ``Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control"}
 The proof of the high probability bound starts on page 14, in section A.1.

### 2017-07-29 rebuttal.txt
updated rebuttal

@@
 
-3-4) Differences to Prashanth-Korda and Prashanth-Korda-Munos? See above. Why is our bound better? Also: line 275
+3) The word "results" in l.266 needs to be changed to the iterate averaged case (relevant to our work) in Theorem 2 [Korda-Prashanth 2015] which does not provide a bias/variance terms separately ($K_1^{IA}(n)$).
+
+4) Going by Theorem 1 of fast LSTD it should be $\mu c$.
+However, the idea was to say that 1/n or c/c+n stepsize are bad in general and in l.
+275 we were citing [Bach-Moulines 2011, see page 4 case $\alpha=1$], where it is $\mu c/2$.
+The point to be noted is that $\mu c$ or $\mu/2$ in the exponent is not desirable.
 

### 2017-07-29 rebuttal.txt
updated rebuttal

--- 2017-07-29 rebuttal.txt
@@
 Understanding the differences between the standard supervised setting and the RL setting should start with a setting with the least number of differences.
-We see that already the iid noise setting presents surprises (see below).
-2) Novelties are significant:
+We see that already the iid noise setting presents surprises (see below)
+2) Novelties are many:
 i) Analysis of general LSA schemes with multiplicative noise for constant stepsize combined with PR averaging is new in literature.
-Prior results were known only for general LSA with additive noise and for SGD (a special case of LSA).
-The new analysis required significant departure from the previous analysis and produces different results.
+Prior results were known only for general LSA with additive noise and for SGD (a special case).
+New analysis required significant departure from the prior work and produces different results.
 ii) Matrices in general LSA schemes lack symmetry, hence prior results do not apply directly; the similarity transformation that we employ in the paper is needed to handle the non-symmetric case; this is a major step
@@
 The projection step is problematic because the range of dual variables cannot be known a priori in GTD.
-We would like to mention that we have explicitly discussed the significance of our contributions in the paper in the introduction as well as section 6.
+We would like to mention that we have discussed the significance of our contributions in the introduction as well as section 6.
 Hence, we think that the reviewer's position is unwarranted and lacks justification.
@@
 The importance of our result is further elevated by that the topic of RL/TD/..
-is rapidly gaining popularity, and that numerous works attempt to transport results from supervised learning to this setting, overlooking the extra difficulties (e.g., see the papers by Prashanth et al.).
+is rapidly gaining popularity, and that numerous works attempt to transport results from supervised learning to this setting, overlooking the extra difficulties (e.g., see papers by Prashanth et al.).
 Our paper may be valuable for both people solely interested in RL, but also for the fast growing community of researchers coming to RL with a supervised learning background.
 
-3-4) Differences to Prashanth-Korda and Prashanth-Korda-Munos? See above. Why is our bound better? Also: line 275
+3) l.266 should have cited Theorem 2 [6] (PR avg case) which has no bias/variance decomposition (see $K_1^{IA}(n)$).
+
+4) In l. 275 we cite [11], (page 4, $\alpha=1$); point is both $\mu c$ or $\mu/2$ in the exponent are undesirable.
 

### 2017-07-30 issues.tex
issues in flux

@@
 \subsection{Bugs in  \cite{lstdicmla}}
+(In this paper, $s_t$ is a sequence of states obtained whole following a fixed policy in an MDP.)
 In page 10 of \cite{lstdicmla} the expression for $\mathbf{E}[a_{j+1}^\top a_{j+1}]$ contains terms that involve the product of $P$ and $P^\top$.
 This cannot be correct, as here we can take the expectation first over the next state, which will bring in a \emph{single} instance of $P$.
-To remove clutter, drop the $j$ subindex, and set $A = \phi(\phi- \beta \phi')^\top$, where $\phi  = \phi(s_{I})$ and $\phi' = \phi(s_{I}')$.
+To remove clutter, drop the $j$ subindex, and set $A = \phi(\phi- \beta \phi')^\top$, where $\phi  = \phi(s_{j})$ and $\phi' = \phi(s_{j+1})$.
 The incriminated expression from Eq. (15) of the paper is
@@
 \EE{ A - \tfrac{\gamma}{2} A^\top A  } = \Phi^\top (I - \beta \Psi P - \tfrac{\gamma}{2}(\Delta - \beta P^\top (2I - \beta \Delta)\Psi P) ) \Phi\,.
+\label{eq:bad1}
 \end{align}
-Here, $\Phi$ is the $S\times d$ matrix whose $i$th row is $\phi^\top(s_i)$,
-$\Psi$ is the $T \times T$ diagonal matrix whose $i$th diagonal entry is $\Prob{I = i}$, while $\Delta$ is another diagonal matrix whose $i$th entry is $\norm{\phi(s_i)}_2^2$.
+Here, $\Phi$ is the $S\times d$ matrix whose $s$th row is $\phi^\top(s)$ ($s\in \{1, \dots, S\}$),
+$\Psi$ is the $S \times S$ diagonal matrix whose $i$th diagonal entry is $\Prob{s_t = i}$,  \todoc{Do they assume stationarity?}
+while $\Delta$ is another $S\times S$ diagonal matrix whose $s$th entry is $\norm{\phi(s)}_2^2$.
 A direct calculation gives that
@@
 %where we abbreviate the 2-norm with $\norm{\cdot}$.
-Hence, 
+The expectation of each terms are as follows:
 \begin{align*}
-\EE{ A^\top A \,|\, I } = 
-\norm{\phi}^2  \phi \phi^\top-\beta \norm{\phi}^2 \EE{ \phi' \,|\, I} \phi^\top -\beta \norm{\phi}^2  \phi (\EE{ \phi' \,|\, I})^\top + \beta^2 
-\norm{\phi}^2 \EE{ \phi' (\phi')^\top\,|\, I } \,.
+\EE{\norm{\phi}^2  \phi \phi^\top } 
+& = \sum_s \Prob{s_t = s}\norm{\phi(s)}^2 \phi(s) \phi(s)^\top = \Phi^\top \Delta \Psi \Phi\,,\\
+\beta \EE{\norm{\phi}^2 \phi' \phi^\top } 
+& = \beta \sum_s \Prob{s_t = s}\norm{\phi(s)}^2 \left\{ \sum_{s'}  P(s'|s) \phi(s')  \right\} \phi(s)^\top 
+   = \beta (P \Phi)^\top \Delta \Psi \Phi\,,\\
+\beta \EE{\norm{\phi}^2 \phi (\phi')^\top } 
+& = \beta \EE{\norm{\phi}^2 \phi' \phi^\top }^\top
+= \beta \left\{(P \Phi)^\top \Delta \Psi \Phi\right\}^\top 
+=  \beta (\Phi^\top \Delta \Psi P \Phi) \,,\\
+\EE{ \beta^2  \norm{\phi}^2 \phi' (\phi')^\top } 
+& =\beta^2 \sum_s P(s_t=s) \norm{\phi(s)}^2 \left\{ \sum_{s'} P(s'|s) \phi(s') \phi(s')^\top \right\}\,.
 \end{align*}
-Now, $\EE{ \phi' (\phi')^\top\,|\, I=i } = \phi(
-Notice that $P \Phi$ is an $S\times d$ matrix whose $i$th row is $\EE{ \phi' \,|\, I=i}^\top$.
-Hence, $ \EE{ \phi' \,|\, I} = (P \Phi)_{I:\cdot}$.
-Now, $\EE{\norm{\phi}^2 \phi \phi^\top} = \Phi \Phi^\top \Psi \Phi \Phi^\top$.
-Hence,
+Further,
 \begin{align*}
-\EE{ A^\top A } = 
-\EE{\norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}}\,.
+\EE{ A }  = \Phi^\top (I - \beta P) \Phi\,.
+\end{align*}
+Putting together things we see the mismatch with \eqref{eq:bad1}.
+To see this even more clearly, assume that $ \norm{\phi(s)}^2 = 1$ for any $s\in \{1,\dots,S\}$.
+Then, $\Delta = I$, and by stationarity, $\Prob{s_{t+1} = s } = \Prob{s_t=s}$, hence,
+\begin{align*}
+\EE{ \beta^2  \norm{\phi}^2 \phi' (\phi')^\top } & =
+\beta^2 \EE{ \phi (\phi)^\top } =\beta^2 \Phi^\top \Psi \Phi\, .
+\end{align*}
+Thus, 
+\begin{align*}
+\EE{ A^\top A } & = 
+\Phi^\top  \Psi \Phi -  \beta \Phi^\top P^\top  \Psi \Phi - \beta \Phi^\top  \Psi P \Phi + \beta^2 \Phi^\top \Psi \Phi \\
+& = 
+\Phi^\top  \left(\Psi -  \beta  (P^\top  \Psi +   \Psi P) + \beta^2 \Psi \right) \Phi \\
+\end{align*}
+and hence
+\begin{align*}
+\EE{ A - \frac{\gamma}{2} A^\top A } = 
+ \Phi^\top (I - \beta P) \Phi - \frac{\gamma}{2} 
+ \Phi^\top  \Psi \Phi + \frac{\gamma\beta }{2} \left( \Phi^\top P^\top  \Psi \Phi + (\Phi^\top  \Psi P \Phi) \right)
+ - \frac{\gamma\beta^2}{2}\Phi^\top \Psi \Phi\,,
+\end{align*}
+while \eqref{eq:bad1} gives
+\begin{align*}
+\EE{ A - \tfrac{\gamma}{2} A^\top A  } 
+& = \Phi^\top (I - \beta \Psi P - \tfrac{\gamma}{2}(I - \beta P^\top (2I - \beta I)\Psi P) ) \Phi\,.
 \end{align*}

### 2017-07-30 issues.tex
edits to issues.tex

@@
 
-\subsection{Bugs in \cite{flstda}}
+\subsection{Bugs in the paper by \cite{flstda}}
 One of the main results of \citet{flstda} is Theorem 1, which states a bound on the expected error.
@@
 and we in fact suspect that this claim is false in full generality. But at minimum, a proof would be required and the whole subsequent argument will need to be changed.
-\subsection{Bugs in  \cite{lstdicmla}}
+
+\subsection{Bugs in the paper by  \cite{lstdicmla}}
 (In this paper, $s_t$ is a sequence of states obtained whole following a fixed policy in an MDP.)
@@
 \begin{align*}
-\EE{ A }  = \Phi^\top (I - \beta P) \Phi\,.
+\EE{ A }  = \Phi^\top  \Psi (I - \beta P) \Phi\,.
 \end{align*}
@@
 \begin{align*}
-\EE{ A - \frac{\gamma}{2} A^\top A } = 
- \Phi^\top (I - \beta P) \Phi - \frac{\gamma}{2} 
- \Phi^\top  \Psi \Phi + \frac{\gamma\beta }{2} \left( \Phi^\top P^\top  \Psi \Phi + (\Phi^\top  \Psi P \Phi) \right)
- - \frac{\gamma\beta^2}{2}\Phi^\top \Psi \Phi\,,
+\EE{ A - \frac{\gamma}{2} A^\top A } & = 
+ \Phi^\top \Psi(I - \beta P) \Phi - \tfrac{\gamma}{2} 
+ \Phi^\top  \left(\Psi -  \beta  (P^\top  \Psi +   \Psi P) + \beta^2 \Psi \right) \Phi \\
+& = 
+\Phi^\top  \left\{ \Psi (I - \beta P)- \tfrac{\gamma}{2} 
+\left(\Psi -  \beta  (P^\top  \Psi +   \Psi P) + \beta^2 \Psi \right) \right\} \Phi \,.
 \end{align*}
@@
 \EE{ A - \tfrac{\gamma}{2} A^\top A  } 
-& = \Phi^\top (I - \beta \Psi P - \tfrac{\gamma}{2}(I - \beta P^\top (2I - \beta I)\Psi P) ) \Phi\,.
+& = \Phi^\top (I - \beta \Psi P - \tfrac{\gamma}{2}(I - \beta P^\top (2I - \beta I)\Psi P) ) \Phi \\
+& = \Phi^\top \left\{I - \beta \Psi P - \tfrac{\gamma}{2}(I - 2 \beta P^\top\Psi P  +  \beta^2 P^\top \Psi P) \right\} \Phi\,.
 \end{align*}
+Choosing $\Phi = I$, we find that the two expressions are equal if and only if
+\begin{align*}
+ \Psi (I - \beta P)- \tfrac{\gamma}{2} \left( \Psi -  \beta  (P^\top  \Psi +   \Psi P) + \beta^2 \Psi \right)
+= 
+I - \beta \Psi P - \tfrac{\gamma}{2}(I - 2 \beta P^\top\Psi P  +  \beta^2 P^\top \Psi P)\,,
+\end{align*}
+which implies, e.g., that $\Psi=I$ (by choosing $\gamma=0$), which is not possible since the diagonal elements of $\Psi$ must some to one.
+Even if we correct the first identity to $\Psi$, we see that we must have
+\begin{align*}
+ \Psi -  \beta  (P^\top  \Psi +   \Psi P) + \beta^2 \Psi 
+ = I - 2 \beta P^\top\Psi P  +  \beta^2 P^\top \Psi P\,,
+\end{align*}
+which again, means that $\Psi = I$, and also that $P^\top \Psi + \Psi P = P^\top \Psi P$ and that $\Psi = P^\top \Psi P$.
+The first equality is always false, and the others are false except (perhaps) in some very special cases.
 
 \section{Problems with the Proof: High Probability Bound}
-\subsection{Bugs in ``Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control"}
+\subsection{Bugs in the paper by \cite{flstda}}
 The proof of the high probability bound starts on page 14, in section A.1.
@@
 where $\norm{x}$ denotes the $2$-norm of $x$.
-However, using $A = \phi( \phi-\beta \phi')^\top$, a direct calculation gives:
+However, using $A = \phi( \phi-\beta \phi')^\top$, a direct calculation, as before, gives:
 \begin{align*}
@@
 \end{comment}
-\subsection{Bugs in ``On TD(0) with function approximation: Concentration bounds and a centered variant with exponential convergence''}
+\subsection{Bugs in the paper by  \cite{lstdicmla}}
 In page 14 the expression below "A is a possibly random matrix..." is not entirely justified (personal communication with one of the authors).
@@
 \section{Bounded Linear Space Assumption}
-\subsection{Bugs in ``Fast LSTD using stochastic approximation: Finite time analysis and application to traffic control"}
+\subsection{Bugs in the paper by \citet{flstda}}
 \citet{flstda} assume that the parameter vector stays such that the value function $\Phi \theta$ (with their notation) will be bounded in $L^\infty$-norm (see assumption \textbf{$(A3)$} of \cite{flstd} and \cite{flstda}).
@@
 \end{itemize}
-\subsection{Bugs in ``On TD(0) with function approximation: Concentration bounds and a centered variant with exponential convergence''}
+\subsection{Bugs in the paper by \citet{lstdicmla}}
 \citet{gugan} mention that (citing a personal communication with \citet{lstdicmla}) \cite{lstdicmla} assume implicitly the projection step in all the high probability bounds in \cite{lstdicmla}.
@@
 
-\section{Doubt about the Effectiveness of the Algorithm}
+\section{Doubt about the Effectiveness of the Algorithms}
 In Corollary~4 of \cite{flstda} (or Corollary~2 of \cite{flstd}), we learn that the value of $c$ governing the stepsize of the primary update must be in a small range (it must be between $1.33$ and $2$).

### 2017-07-30 issues.tex
pass on issues

@@
 \subsection{Bugs in the paper by  \cite{lstdicmla}}
-In page 14 the expression below "A is a possibly random matrix..." is not entirely justified (personal communication with one of the authors).
+In page 14 the expression below ``$A$ is a possibly random matrix...'' is not justified (personal communication with one of the authors confirmed this).
+In particular, the claim here is that if $A$ is a random matrix with $\norm{A}_2\le C$ with $C$ a deterministic constant then for any $\theta$ deterministic vector,
+\begin{align*}
+\EE{ \theta^\top A^\top \EE{ \epsilon_n | \mathcal{F}_n } A \theta \,|\, s_0 } \le C^2 \theta^\top \EE{ \epsilon_n|s_0 } \theta\,.
+\end{align*}
+Recall that here $\epsilon_n$ is a matrix of appropriate dimensions, the ``mixing-error term''
+and is actually $\mathcal{F}_n$ measurable ($\epsilon_n = \EE{ a_n | \mathcal{F}_n } - \mathbb{E}_{\Psi}[a_n]$). 
+When the Markov chain is started from its stationary state 
+(which is not ruled out by the conditions of the theorem under question), $\EE{ \epsilon_n|s_0 } = 0$.
+If the above inequality was true, we would get 
+\begin{align*}
+\EE{ A^\top \epsilon_n  A  \,|\, s_0 } = 0\,.
+\end{align*}
+However, letting $B = \EE{\epsilon_n|\mathcal{F}_n}$ and, for example,
+$A = C B/\norm{ B }_2$, we have 
+\begin{align*}
+A^\top \EE{ \epsilon_n | \mathcal{F}_n } A  = \frac{C^2}{\norm{ B }_2^2} B^\top B B
+\end{align*}
+and it is easy to construct examples where the expectation of this is nonzero.
 
-\section{Bounded Linear Space Assumption}
-\subsection{Bugs in the paper by \citet{flstda}}
-\citet{flstda} assume that the parameter vector stays such that the value function $\Phi \theta$ (with their notation) will be bounded in $L^\infty$-norm (see assumption \textbf{$(A3)$} of \cite{flstd} and \cite{flstda}).
-This assumption is critical in establishing Lemma~7 (see pages 15 and 16, \cite{flstda}).
-We suspect the following shortcomings with assumption \textbf{$(A3)$}
+\section{Issues with the Setup}
+\subsection{Boundedness of iterates: \citet{flstda}}
+\citet{flstda} assume that the parameter vector stays such that the value function $\Phi \theta$  will be bounded in $L^\infty$-norm (see assumption (A3) of \cite{flstd} and \cite{flstda}).
+This assumption is critical in establishing Lemma~7 (see pages 15 and 16, \cite{flstda}), in an argument that is similar to the proof of McDiarmid's inequality.
+We suspect the following shortcomings with assumption (A3):
 \begin{itemize}
@@
 We take the authors meant to say that $\sup_n \norm{\Phi \theta_n}_\infty<+\infty$ holds almost surely.
-This seems like a strong assumption!
-Would not one of the main question be whether this assumption even holds?
+This seems like a strong assumption:  ensuring this will most likely further restrict the step-size sequences that can be used.
+The step-size sequences that give the best rate under (A3) may include step-size sequences which in fact lead to $\Prob{\limsup_{n\to\infty}\norm{\theta_n}=\infty}>0$.
+Without proving that this is not the case, the results of the paper have limited utility.
 % We see in our paper that for some cases, the iterates can blow up if the step-size is inappropriately selected!
@@
 It is still unclear  whether this alone would ensure convergence of the error to zero.
-In any case, the expected error bound analysis is invalidated if a projection step is present (basically, the algebraic identities will all fail to hold).
+In any case, the expected error bound analysis is invalidated if a projection step is present (basically, the algebraic identities will all fail to hold) and a new proof will be required.
 
@@
 \end{itemize}
-\subsection{Bugs in the paper by \citet{lstdicmla}}
-\citet{gugan} mention that (citing a personal communication with \citet{lstdicmla}) \cite{lstdicmla} assume implicitly the projection step in all the high probability bounds in \cite{lstdicmla}.
-While this implicit projection in itself does not affect the high probability bound proofs directly, it is important to note that the set to which the iterates needs to be projected has to contain the TD(0) solution, a condition that cannot be guaranteed.
+\subsection{Boundedness of iterates:  \citet{lstdicmla}}
+\citet{gugan} mention that (citing a personal communication with \citet{lstdicmla}) 
+that \cite{lstdicmla} assume implicitly a projection step in all the high probability bounds. 
+While this implicit projection in itself does not affect the high probability bound proofs directly, 
+the algebraic steps are invalidated.
+Furthermore, the set that the iterates are projected to should contain the TD(0) solution.
+How to ensure this (without knowing $A,b$) remains to be seen.
 
-\section{Relation between Covariance Matrix and $\bar{A}_T$ matrix}
+\subsection{Relation between Covariance Matrix and $\bar{A}_T$ matrix}
 \citet{flstda} assume positive definiteness of \textbf{$(A4)$} covariance matrix $\frac{1}{T}\Phi^\top_T \Phi_T$.
@@
 \fi
-\section{Blow Up of the Bound}
+\subsection{Blow Up of the Bound}
 We would like to note that the rate expression in Corollary~4 of \cite{flstda} (or Corollary~2 of \cite{flstd}) contains a constant $C$.
-The authors do mention that the sampling error blows up as $\alpha \rightarrow 1$.
-However, it also looks like that even the constant $C=\sum_{n=1}^\infty \exp(-\mu cn^{1-\alpha})$ will blow up as $\alpha\rightarrow 1$, in which case it is not clear whether a $1/\sqrt{n}$ rate is achieved (even if we disregard all the other problems).
+The authors do mention that the sampling error (a.k.a.
+variance) blows up as $\alpha \rightarrow 1$.
+However, it also looks like that even the constant $C=\sum_{n=1}^\infty \exp(-\mu cn^{1-\alpha})$ (appearing in the bound of the bias) will blow up as $\alpha\rightarrow 1$, in which case it seems that the claim that the $1/\sqrt{n}$ rate can be achieved in the limit will not hold.
 
-\section{Doubt about the Effectiveness of the Algorithms}
+\subsection{Doubt about the Effectiveness of the Algorithms}
 In Corollary~4 of \cite{flstda} (or Corollary~2 of \cite{flstd}), we learn that the value of $c$ governing the stepsize of the primary update must be in a small range (it must be between $1.33$ and $2$).
@@
 This is because then the bias is decreased relatively quickly, while the variance will be controlled by the additional averaging.
-However, given all the problems with this it remains to be seen whether this is indeed a reasonable choice and under exactly what conditions this is reasonable.
+However, given all the problems with this, it remains to be seen whether this is indeed a reasonable choice and under exactly what conditions this is reasonable.
 \bibliographystyle{plainnat}

### 2017-07-30 issues.tex
pass on issues

@@
 \begin{abstract}
-This short report lists some problems with the recent papers by \cite{flstd,lstdicml}, whose longer version containing the proofs are available on arxiv \citep{flstda,lstdicmla}.
+In mathematics, results build on previous results. Thus, it is important to keep the literature free of erroneous claims.
+This short report lists some problems with the proofs and claims in the recent papers by \cite{flstd,lstdicml}, whose longer version containing the proofs are available on arxiv \citep{flstda,lstdicmla}.
+In particular, it follows that the results in these papers, if true, would need completely new proofs, and thus should not be used as stated by the authors.
 \end{abstract}

### 2017-07-30 issues.tex
pass on issues

@@
 \begin{abstract}
-In mathematics, results build on previous results. Thus, it is important to keep the literature free of erroneous claims.
+In all branches of mathematics, including learning theory,
+results build on previous results. Thus, it is important to keep the literature free of erroneous claims.
 This short report lists some problems with the proofs and claims in the recent papers by \cite{flstd,lstdicml}, whose longer version containing the proofs are available on arxiv \citep{flstda,lstdicmla}.
-In particular, it follows that the results in these papers, if true, would need completely new proofs, and thus should not be used as stated by the authors.
+In particular, it follows that the results in these papers, if true, would need completely new proofs, and thus should not be used in the form stated by the authors.
 \end{abstract}

### 2017-07-30 rebuttal.txt
issues in flux
rebuttal final touches

@@
 We would like to mention that we have discussed the significance of our contributions in the introduction as well as section 6.
-Hence, we think that the reviewer's position is unwarranted and lacks justification.
+Hence, we think that the negative position is unwarranted and lacks justification.
 
@@
 Now, as to the actual contributions, unfortunately, the ECML'14 paper seems to have many problems.
-We believe that the stated results in it are incorrect and fixing the bugs would lead to an analysis and results identical to ours.
+In fact, a report by Lakshmi Narayanan and Szepesvári available at https://goo.gl/q9A4vn describes some of these problems.
 The factual errors are severe: for the expected error bound, on page 16 in Section A.2 of the arxiv version, after Eq.
@@
 Again, fixing the proof requires extra conditions, leading to an analysis similar to ours.
-As a sidenote, we remark that even the ICML'15 seems to have problems: (Earlier) we have contacted the first author, Prashanth, who confirmed these problems.
+As a sidenote, per the report above, it appears that the ICML'15 paper seems to have similar problems.
 

### 2017-07-30 issues.tex
no date
updated issues

@@
 
-\title{Finite Time Bounds for Temporal Difference Learning with Function Approximation: Bugs in state-of-the-art results}
+\title{Finite Time Bounds for Temporal Difference Learning with Function Approximation: Problems with some ``state-of-the-art'' results}
 
@@
 }
+\date{}
 
@@
 \end{align*}
-which implies, e.g., that $\Psi=I$ (by choosing $\gamma=0$), which is not possible since the diagonal elements of $\Psi$ must some to one.
+which implies, e.g., that $\Psi=I$ (by choosing $\gamma=0$), which is not possible since the diagonal elements of $\Psi$ must sum to one.
 Even if we correct the first identity to $\Psi$, we see that we must have
@@
 and it is easy to construct examples where the expectation of this is nonzero.
+
+Also in page 14, we find the following inequality
+\begin{align*}
+(d+2)\left( e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\norm{\theta_0}^2_2+ e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\left(\sum_{k=1}^n \gamma_k e^{-(1+\beta)\sum_{j=1}^{k-1} \gamma_j}\right)^2+\norm{\theta^*}_2^2\right)\norm{\E(\epsilon_n|s_0)}_2\\
+\leq (d+2)\frac{\norm{\theta_0}_2^2+1+\norm{\theta^*}_2^2}{(1-\beta)^2}e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\norm{\E(\epsilon_n|s_0)}_2,
+\end{align*}
+where it is not clear as to how the $(1-\beta)^2$ factor appears in the denominator (also confirmed by a personal communication with one of the authors).
 

### 2017-07-30 issues.tex
update issues

@@
 and we in fact suspect that this claim is false in full generality. But at minimum, a proof would be required and the whole subsequent argument will need to be changed.
+\subsection{Bugs in the paper by  \cite{lstdicmla}}
+In page 14 the expression below ``$A$ is a possibly random matrix...'' is not justified (personal communication with one of the authors confirmed this).
+In particular, the claim here is that if $A$ is a random matrix with $\norm{A}_2\le C$ with $C$ a deterministic constant then for any $\theta$ deterministic vector,
+\begin{align*}
+\EE{ \theta^\top A^\top \EE{ \epsilon_n | \mathcal{F}_n } A \theta \,|\, s_0 } \le C^2 \theta^\top \EE{ \epsilon_n|s_0 } \theta\,.
+\end{align*}
+Recall that here $\epsilon_n$ is a matrix of appropriate dimensions, the ``mixing-error term''
+and is actually $\mathcal{F}_n$ measurable ($\epsilon_n = \EE{ a_n | \mathcal{F}_n } - \mathbb{E}_{\Psi}[a_n]$). 
+When the Markov chain is started from its stationary state 
+(which is not ruled out by the conditions of the theorem under question), $\EE{ \epsilon_n|s_0 } = 0$.
+If the above inequality was true, we would get 
+\begin{align*}
+\EE{ A^\top \epsilon_n  A  \,|\, s_0 } = 0\,.
+\end{align*}
+However, letting $B = \EE{\epsilon_n|\mathcal{F}_n}$ and, for example,
+$A = C B/\norm{ B }_2$, we have 
+\begin{align*}
+A^\top \EE{ \epsilon_n | \mathcal{F}_n } A  = \frac{C^2}{\norm{ B }_2^2} B^\top B B
+\end{align*}
+and it is easy to construct examples where the expectation of this is nonzero.
 
+Also in page 14, we find the following inequality
+\begin{align*}
+(d+2)\left( e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\norm{\theta_0}^2_2+ e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\left(\sum_{k=1}^n \gamma_k e^{-(1+\beta)\sum_{j=1}^{k-1} \gamma_j}\right)^2+\norm{\theta^*}_2^2\right)\norm{\E(\epsilon_n|s_0)}_2\\
+\leq (d+2)\frac{\norm{\theta_0}_2^2+1+\norm{\theta^*}_2^2}{(1-\beta)^2}e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\norm{\E(\epsilon_n|s_0)}_2,
+\end{align*}
+where it is not clear as to how the $(1-\beta)^2$ factor appears in the denominator (also confirmed by a personal communication with one of the authors).
+
+\section{Problems with the Proof: High Probability Bound}
+\subsection{Bugs in the paper by \cite{flstda}}
+The proof of the high probability bound starts on page 14, in section A.1.
+The first problem happens in the display on the bottom of this page in the proof of Lemma 6.
+Here, we are told that for $A = \phi \phi^\top - \beta \phi (\phi')^\top$ (we are dropping indices to remove clutter),
+\begin{align*}
+A^\top A = \norm{\phi}^2 \phi \phi^\top - \beta ( 2 - \norm{\phi}^2\beta ) \phi' (\phi')^\top\,,
+\end{align*}
+where $\norm{x}$ denotes the $2$-norm of $x$.
+However, using $A = \phi( \phi-\beta \phi')^\top$, a direct calculation, as before, gives:
+\begin{align*}
+A^\top A 
+& = ( \phi-\beta \phi') \phi^\top \phi ( \phi-\beta \phi')^\top \\
+& = \norm{\phi}^2  ( \phi-\beta \phi') ( \phi-\beta \phi')^\top \\
+& = \norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}\,,
+\end{align*}
+which does not match the previous display.
+The terms that do not match are the linear-in-$\beta$ terms.
+In the first display we have $ -2 \beta \phi' (\phi')^\top$, while in the bottom we have $-\beta (\phi' \phi^\top + \phi (\phi')^\top)$.
+
+The first equality of their display states (in equivalent form) that 
+\begin{align*}
+A^\top A = \norm{\phi}^2 ( \phi \phi^\top - 2\beta \phi (\phi')^\top +\beta^2 \phi' (\phi')^\top )\,.
+\end{align*}
+We see that while this is closer to the correct result, here the mistake is that $-2\beta \phi  (\phi')^\top$ is replacing $-\beta(\phi (\phi')^\top + \phi' \phi^\top)$.
+\begin{comment}
+We also have a hard time following the rest of the proof of this lemma.
+In fact, we have a simple proof for a result similar to stated in this lemma.
+This works by observing that one needs to 
+bound the Lipschitz factor of $\theta \mapsto \bar T_n \dots \bar T_{t+1} ( (I-\gamma_t A_t)\theta + b_t )$, where
+$A_t = \phi_t (\phi_t -\beta \phi_t')^\top$, $b_t = r_{i_t} \phi_t$ and
+ $\bar T_t(\theta) = (I-\gamma_t \bar A_T) \theta + \bar b_T$ (we assume that the stepsize sequence is deterministic).
+ Clearly, this Lipschitz factor can be bounded by $\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 \norm{I-\gamma_t A_t}_2$
+ (the composition of Lipschitz maps is Lipschitz with a factor that is the product of the individual maps' Lipschitz factors). 
+Invoking (A3) then indeed gives some version of Lemma 7.
+By making an appropriate assumption (similar to those used in our paper), 
+$\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 $ can be controlled.
+\end{comment}
 \subsection{Bugs in the paper by  \cite{lstdicmla}}
@@
 
-\section{Problems with the Proof: High Probability Bound}
-\subsection{Bugs in the paper by \cite{flstda}}
-The proof of the high probability bound starts on page 14, in section A.1.
-The first problem happens in the display on the bottom of this page in the proof of Lemma 6.
-Here, we are told that for $A = \phi \phi^\top - \beta \phi (\phi')^\top$ (we are dropping indices to remove clutter),
-\begin{align*}
-A^\top A = \norm{\phi}^2 \phi \phi^\top - \beta ( 2 - \norm{\phi}^2\beta ) \phi' (\phi')^\top\,,
-\end{align*}
-where $\norm{x}$ denotes the $2$-norm of $x$.
-However, using $A = \phi( \phi-\beta \phi')^\top$, a direct calculation, as before, gives:
-\begin{align*}
-A^\top A 
-& = ( \phi-\beta \phi') \phi^\top \phi ( \phi-\beta \phi')^\top \\
-& = \norm{\phi}^2  ( \phi-\beta \phi') ( \phi-\beta \phi')^\top \\
-& = \norm{\phi}^2  \left\{ \phi \phi^\top-\beta (\phi' \phi^\top + \phi (\phi')^\top) + \beta^2 \phi' (\phi')^\top \right\}\,,
-\end{align*}
-which does not match the previous display.
-The terms that do not match are the linear-in-$\beta$ terms.
-In the first display we have $ -2 \beta \phi' (\phi')^\top$, while in the bottom we have $-\beta (\phi' \phi^\top + \phi (\phi')^\top)$.
-
-The first equality of their display states (in equivalent form) that 
-\begin{align*}
-A^\top A = \norm{\phi}^2 ( \phi \phi^\top - 2\beta \phi (\phi')^\top +\beta^2 \phi' (\phi')^\top )\,.
-\end{align*}
-We see that while this is closer to the correct result, here the mistake is that $-2\beta \phi  (\phi')^\top$ is replacing $-\beta(\phi (\phi')^\top + \phi' \phi^\top)$.
-\begin{comment}
-We also have a hard time following the rest of the proof of this lemma.
-In fact, we have a simple proof for a result similar to stated in this lemma.
-This works by observing that one needs to 
-bound the Lipschitz factor of $\theta \mapsto \bar T_n \dots \bar T_{t+1} ( (I-\gamma_t A_t)\theta + b_t )$, where
-$A_t = \phi_t (\phi_t -\beta \phi_t')^\top$, $b_t = r_{i_t} \phi_t$ and
- $\bar T_t(\theta) = (I-\gamma_t \bar A_T) \theta + \bar b_T$ (we assume that the stepsize sequence is deterministic).
- Clearly, this Lipschitz factor can be bounded by $\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 \norm{I-\gamma_t A_t}_2$
- (the composition of Lipschitz maps is Lipschitz with a factor that is the product of the individual maps' Lipschitz factors). 
-Invoking (A3) then indeed gives some version of Lemma 7.
-By making an appropriate assumption (similar to those used in our paper), 
-$\prod_{s=t+1}^n \norm{I-\gamma_s \bar A_T}_2 $ can be controlled.
-\end{comment}
-\subsection{Bugs in the paper by  \cite{lstdicmla}}
-In page 14 the expression below ``$A$ is a possibly random matrix...'' is not justified (personal communication with one of the authors confirmed this).
-In particular, the claim here is that if $A$ is a random matrix with $\norm{A}_2\le C$ with $C$ a deterministic constant then for any $\theta$ deterministic vector,
-\begin{align*}
-\EE{ \theta^\top A^\top \EE{ \epsilon_n | \mathcal{F}_n } A \theta \,|\, s_0 } \le C^2 \theta^\top \EE{ \epsilon_n|s_0 } \theta\,.
-\end{align*}
-Recall that here $\epsilon_n$ is a matrix of appropriate dimensions, the ``mixing-error term''
-and is actually $\mathcal{F}_n$ measurable ($\epsilon_n = \EE{ a_n | \mathcal{F}_n } - \mathbb{E}_{\Psi}[a_n]$). 
-When the Markov chain is started from its stationary state 
-(which is not ruled out by the conditions of the theorem under question), $\EE{ \epsilon_n|s_0 } = 0$.
-If the above inequality was true, we would get 
-\begin{align*}
-\EE{ A^\top \epsilon_n  A  \,|\, s_0 } = 0\,.
-\end{align*}
-However, letting $B = \EE{\epsilon_n|\mathcal{F}_n}$ and, for example,
-$A = C B/\norm{ B }_2$, we have 
-\begin{align*}
-A^\top \EE{ \epsilon_n | \mathcal{F}_n } A  = \frac{C^2}{\norm{ B }_2^2} B^\top B B
-\end{align*}
-and it is easy to construct examples where the expectation of this is nonzero.
-
-Also in page 14, we find the following inequality
-\begin{align*}
-(d+2)\left( e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\norm{\theta_0}^2_2+ e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\left(\sum_{k=1}^n \gamma_k e^{-(1+\beta)\sum_{j=1}^{k-1} \gamma_j}\right)^2+\norm{\theta^*}_2^2\right)\norm{\E(\epsilon_n|s_0)}_2\\
-\leq (d+2)\frac{\norm{\theta_0}_2^2+1+\norm{\theta^*}_2^2}{(1-\beta)^2}e^{2(1+\beta)\sum_{k=1}^n \gamma_k}\norm{\E(\epsilon_n|s_0)}_2,
-\end{align*}
-where it is not clear as to how the $(1-\beta)^2$ factor appears in the denominator (also confirmed by a personal communication with one of the authors).
-
 \section{Issues with the Setup}
